### 2024-01-24 latex_files/cover_letter.tex
first-commit-(moved-from-private-repo)

\documentclass[10pt]{letter}
\usepackage[letterpaper,margin=2.2cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\usepackage{graphicx}
\usepackage{scrlayer}
\usepackage{microtype}

\DeclareNewLayer[
  foreground,
  textarea,
  contents={\includegraphics[width=0.5\textwidth]{logos.png}}
]{letterimage}
\AddLayersToPageStyle{empty}{letterimage}

\address{Stewart Charles McDowall\\
Leiden University\\
Institute of Environmental Sciences (CML)\\
s.c.mcdowall@cml.leidenuniv.nl\\
P.O. Box 9518, 2300 RA Leiden\\
The Netherlands}

\begin{document}
\thispagestyle{empty}
\begin{letter}{Editors\\\textit{Resources, Conservation and Recycling}\\Elsevier}

\opening{Dear Editors,}

We are pleased to submit our original research article titled ``WasteAndMaterialFootprint: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}. This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.

Our submission provides a comprehensive description of the tool, which integrates seamlessly with the Brightway2 framework to offer a user-friendly and adaptable solution for LCA practitioners. By automating data analysis and aggregation processes, our tool expedites the identification of environmental hotspots within supply chains, thus facilitating more informed decision-making.

We believe our manuscript is an excellent fit for \textit{Resources, Conservation and Recycling} due to its potential contribution to advancing resource conservation and recycling methodologies. The case study presented, focusing on battery supply chains, underlines the tool's practical utility and potential impact on sustainable resource management practices.

The manuscript complies with the journal's requirements for an original research article, having a word count of [Insert Word Count], including [Insert Number of Figures] figures and [Insert Number of Tables] tables.

We have no conflicts of interest to declare and have ensured that this work is original, has not been published previously, and is not under consideration by any other publication. All authors have agreed to this submission and are collectively responsible for its content.

\vspace{1em}

We suggest the following reviewers for their expertise relevant to our manuscript's focus:

\begin{enumerate}
    \item {Reviewer 1 Name, Affiliation, Email Address}
    \item {Reviewer 2 Name, Affiliation, Email Address}
    \item {Reviewer 3 Name, Affiliation, Email Address}
\end{enumerate}
\vspace{1em}

Thank you for considering our manuscript for publication. We look forward to the opportunity to contribute to \textit{Resources, Conservation and Recycling} and eagerly await your response.

\vspace{1em}

\closing{Sincerely,}
% Signature block with names in a row using dummy images for alignment
\begin{center}
    \begin{minipage}[t]{0.25\textwidth}
      \centering
      \vspace{-2cm}
      \includegraphics[height=2cm]{signature_scm.png}\\ % Replace with your actual signature file name.
      Stewart Charles McDowall\\
    %   \textbf{Corresponding author}
    \end{minipage}%
    \begin{minipage}[t]{0.25\textwidth}
      \centering
      \vspace{-2cm}
      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
      Elizabeth Lanphere
    \end{minipage}%
    \begin{minipage}[t]{0.25\textwidth}
      \centering
      \vspace{-2cm}
      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
      Stefano Cucurachi
    \end{minipage}%
    \begin{minipage}[t]{0.25\textwidth}
      \centering
      \vspace{-2cm}
      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
      Carlos Felipe Rocha Blanco
    \end{minipage}
  \end{center}

\end{letter}
\end{document}

### 2024-01-24 latex_files/introduction.tex
first-commit-(moved-from-private-repo)

\cbox{Section word count: ~1700}
\cbox{Needs work to make it flow better, also shortened}

The development of a `circular economy' has become a critical area of focus in the imperative pursuit of achieving sustainability objectives and curtailing our environmental footprint within planetary boundaries \citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}. Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling) \citep{eu2022ecodesign, eu2022repair,eu2015reman}. In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions, wrought as an act of competition or outright hostility \citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.

While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use-- and end-of-life-- (EOL) phases, a significant proportion of these are often `hidden' in the supply chain and thus not reported directly in the final results \citep{laurenti2016wastefootprint,salviulo2021supplychain}. It has been found that these material footprints can be `highly representative of damage to human health and biodiversity' \citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage \citep{laurenti2023wastefootprint}. Thus, to reduce the negative externalities of consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible.

Thus, we were motivated to create a flexible and policy-driven approach to understanding a product’s demand on the waste management system both now and in the future to ensure that models of circularity and improvements in technology and policy for waste are included in LCA modeling.

\subsection{LCA}

Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes. LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'---, often identifying critical hotspots and guiding prioritisation of actions. The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe \citep{huijbregts2016recipe} and CML \citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows. These scores are then aggregated into a single score for each impact category, which can be compared across products and processes.

Several LCIA methods include, to some extent, waste generation \citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption \citep{arvidsson2020csi,foen2021ecofactors}. These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., kg-Si equivalents or Ümweltbelastungspunkte (UBP)).

\subsection{Material Demand in LCA}
In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs) \citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}. Given the increasing focus placed on this, as well as ideals of improving resource efficiency and developing the circular economy, it is essential to understand the material demands of products and processes. While LCA seeks to model the technosphere (the sum of all supply-chains), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.

A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess longterm global scarcity of minerals in life cycle assessment (LCA). This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations. CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies. The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.

While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories. Furthermore, the CSPs are not available for all elements, and the method does not allow for the quantification of material demands in terms of mass or volume.

\subsection{Waste in LCA}

Though often described simply as `a material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.  Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows. Thus, precise and detailed categorisation of these is essential to understand the `circularity' of an activity and its life cycle externalities. There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage \citep{laurenti2023wastefootprint}. Conventional LCAs consider waste as a `service' \citep{guinee2021wasteisnotaservice} and typically use generic waste processing models \citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.

A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining 'invisible' in traditional life cycle assessment (LCA) practices~\citep{laurenti2016wastefootprint}. This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators. Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment. The environmental significance of waste and its correlation with other indicators has been a subject of extensive research. For example, studies have shown that popular resource footprints can cover a significant portion of environmental impact variance in product rankings~\citep{steinmann2017resourcefootprints}. However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate with other impact assessment scores~\citep{laurent2012carbonfootprint}. The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.

Moreover, existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste. This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle. Addressing these hotspots is crucial for advancing towards circularity. Thus, there is a pressing need for more comprehensive methods that can effectively quantify waste impacts and contribute to a better understanding of a product's total environmental footprint.
There is currently a lack of a convenient and flexible method for calculating waste flows in Life Cycle Assessment (LCA).

\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity. In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health. The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting. The WMF tool presented in the current work, builds on the work of Laurenti et.al., providing a more flexible, transparent and user-friendly approach to quantifying waste flows in LCA. Moreover, the WMF tool is not limited to waste, but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.

\subsection{The WasteAndMaterialFootprint Tool}

To better assess waste and material flows in LCA, the authors have developed a Python program built on the Brightway2 framework and designed to track these exchanges by translating them into indicators and impact categories. In this study, we present the WasteAndMaterialFootprint tool that enables LCA practitioners to easily aggregate the mass and volume of any desired exchange, and to creation of flexible categories to differentiate between waste types and End-of-Life (EOL) handling using (in this case) the Ecoinvent 3.9 cutoff database.

While methods with similar aims exist, they either lack flexibility and specificity \citep{foen2021ecofactors} or are cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.

This tool provides a method for the calculation of waste footprint impact category results, differentiated by the type of waste handling. Furthermore, the tool facilitates rapid investigation and identification of waste hotspots, enabled by standard contribution analysis and Sankey diagram visualization tools. The authors consider this a crucial step in addressing the deficit of Life Cycle Assessment (LCA) methods that consider waste flows in the evaluation of a product or process' circular economy potential.

### 2024-01-24 latex_files/main.tex
first-commit-(moved-from-private-repo)

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
\documentclass[final,5p,authoryear]{elsarticle}
%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{lineno}
\usepackage{booktabs}
% ---------------------------------------------------------------------
%             FOR DRAFTING / REVIEW (DELETE BEFORE SUBMISSION)
% ---------------------------------------------------------------------
\usepackage{caption}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage[tt=false]{libertine}
\usepackage{xcolor}
\definecolor{blue}{RGB}{33,150,209}
\definecolor{red}{RGB}{255,0,0}
\usepackage{xurl}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{lineno}
\setlength{\linenumbersep}{5pt}
% Save the old \texttt command
\let\oldtexttt\texttt
% Redefine the \UrlFont to use a smaller typewriter font
\renewcommand{\UrlFont}{\ttfamily\small}
% Redefine the \texttt command to use a smaller font size and allow line breaks
\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}

\usepackage{tcolorbox}
\newcommand{\cbox}[1]{
    \begin{tcolorbox}[hbox, colback=red!5!white, colframe=red!65!black, boxrule=0.25pt, boxsep=2pt, left=2pt, right=2pt, top=1pt, bottom=1pt]
        \small\sffamily #1
    \end{tcolorbox}
}

% ------------------------------------------------------

\journal{Resources, Conservation and Recycling}

\begin{document}

\linenumbers

    % SPOTLIGHTS (5 bullet points, max 120 characters each)
    {\Large Spotlights}
    \vspace{1em}
    \begin{description}[style=nextline]
        \item[Bullet 1: Critical context and background information on the problem addressed] Understanding waste and material flows is essential for developing the circular economy and supply chain resilience
        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A program was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is flexible and easy to use.
        \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
    \end{description}

    \newpage
    % %% GRAPHICAL ABSTRACT
    % \begin{graphicalabstract}
    %     %\includegraphics{grabs}
    % \end{graphicalabstract}

    %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
    \begin{highlights}
        \item WasteAndMaterialFootprint, a new tool for quantifying waste and material flows in LCA.
        \item Assesses supply risks by calculating demand for critical materials.
        \item Simplifies quantification of user-specified waste and material categories.
        \item Rapidly identifies waste and material demand hotspots.
        \item Presents a case study of the battery supply chain.
    \end{highlights}

\begin{frontmatter}

    \title{WasteAndMaterialFootprint: A python package to quantify supply chain flows of waste and material in LCA databases}
    \author[1]{Stewart Charles McDowall\corref{cor1}}
    \ead{s.c.mcdowall@cml.leidenuniv.nl}
    \author[1]{Elizabeth Lanphere}
    \author[1]{Stefano Cucurachi}
    \author[1]{Carlos Felipe Blanco}

    \cortext[cor1]{Corresponding author:}

    \affiliation[1]{
        organization={Institute of Environmental Sciences (CML), Leiden University},
        addressline={P.O. Box 9518},
        city={Leiden},
        postcode={2300 RA},
        state={South Holland},
        country={The Netherlands}}

    \begin{abstract}
        The `circular economy' is a concept focused on the prevention of waste and the reuse of resources. Thus, identifying and quantifying waste and material flows are of fundamental importance. Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of products and services, those where the implementation of circular principles could be most effective.

        Introducing WasteAndMaterialFootprint, a Python tool extending the Brightway framework for flexible, user-defined quantification of supply-chain demands in current and future scenarios. This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

        A case study on battery supply chains demonstrates the tool's utility. WasteAndMaterialFootprint quantifies and compares demands and potential environmental impacts, aiding sustainable decision-making. It contributes to a more sustainable, circular economy by providing detailed material usage and waste generation analysis.

    \end{abstract}

    % KEYWORDS (max 6)
    \begin{keyword}
        circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
    \end{keyword}

\end{frontmatter}

%% main text
\section{Introduction}
\label{sec:introduction}
\input{introduction.tex}

\section{Methodology}
\label{sec:methodology}
\input{methodology.tex}

\section{Results}
\label{sec:results}
\input{results.tex}

\section{Discussion}
\label{sec:discussion}
\input{discussion.tex}
\cbox{Section word count: aim for 400?}
\begin{itemize}
    \item Summary of the results, 1 par
    \item -----------------------------
    \item Advantages of the WMF tool, 1 par
    \item -----------------------------
    \item Limitations, 1 par
    \item -----------------------------
    \item Future possibilities, 1 par
\end{itemize}

\section{Conclusions}
\cbox{Section word count: aim for 150?}

This paper extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply-chain risk and potential environmental damage.

\label{sec:conclusions}

% Back matter
\section*{Data availability} 
All data used in this analysis are publicly available online under the noted sources. The WasteAndMaterialFootprint tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/WasteAndMaterialFootprint}.
The full source code for the WasteAndMaterialFootprint tool is available at \url{https://www.github.com/Stew-McD/WasteAndMaterialFootprint}.
A user guide and comprehensive documentation are available at \url{https://wasteandmaterialfootprint.readthedocs.io}.

\section*{CRediT authorship contribution statement}
\cbox{Co-authors, please check this and change as necessary.}
\textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
\textbf{Elizabeth Lanphere:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
\textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
\textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.

\subsection*{Alternative CRediT statement}

\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit.pdf}

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Acknowledgements}
\cbox{Should I put FutuRaMa here?}

    Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No. 101058522 (project FutuRaM --- futuram.eu). The authors would like to thank the reviewers for their valuable comments and suggestions.

\section*{Supplementary material}
    The supplementary material contains the following:
    \begin{enumerate}
        \item List of waste and material categories
        \item Example output of the WasteAndMaterialFootprint tool
        \item Extended results of the case study
    \end{enumerate}

\bibliographystyle{elsarticle-harv}
\bibliography{references.bib}

% \appendix

% \section{Example output}
% \label{app:example_output}

\end{document}

### 2024-01-24 latex_files/methodology.tex
first-commit-(moved-from-private-repo)

\cbox{Section word count: 1600}

First, we identified all Ecoinvent technosphere exchanges that produce waste. We then further classified the waste into non-mutually exclusive categories such as its destination (i.e., dumped, incinerated, etc.), hazardousness, and form (solid vs. liquid). After cloning the technosphere exchanges as biosphere exchanges, we aggregated and quantified them, by waste category, into matching impact categories in the Life Cycle Impact Assessment.

In our simplified test case of several battery types, we were able to identify `waste hotspots' and distinguish the major sources of contribution to waste generation on a process level. One conspicuous result from the case study (and potential direction for further work) is that many waste flows are tied to processes lacking a clear EOL pathway. Further development of this tool could involve developing an algorithm using identifiers of each background waste process to predict where these uncategorised wastes land in their EOL management.

This section is divided into two parts. In \autoref{sec:method-WMF_tool}, we describe the WasteAndMaterialFootprint tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.

\subsection{The WasteAndMaterialFootprint tool}
\label{sec:method-WMF_tool}

The WasteAndMaterialFootprint tool is a Python package that allows one to calculate the waste and material footprint of any product or service inside of LCA databases. The tool is built on the \texttt{Brightway2} LCA framework~\citep{mutel2017brightway} and is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser} an open-source graphical user interface for LCA. The WMF tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023wmfpipy} and is open source under the CC-0 licence. The full source code for the WasteAndMaterialFootprint tool is indexed on Zenodo~\citep{mcdowall2023wmfzenodo} and under further development in the GitHub repository~\citep{mcdowall2024wmfgithub}. The tool is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases as well by changing the search criteria. Currently, it has been tested with all available system models of ecoinvent 3.5--3.10.

The program can be used directly from the command line, or imported as a Python module, in which case, the user can access the individual functions and modules. In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the ecoinvent database. The user can also customise the program to calculate the waste and material footprint of a custom database, or a prospective database based on future scenarios. The program is designed to be modular so that the user can easily customise the program to their needs.

The following lists outline the constituent modules of the WMF tool, with a brief description of their functions. More extensive details can be found in the user guide and documentation of the program~\citep{mcdowall2023wmfdocs}.

\subsubsection{Functional modules:}
\begin{itemize}
    \item \texttt{future\_scenarios}: Creates prospective LCA databases based on future scenarios.
    \item \texttt{explode\_database}: Responsible for expanding a Brightway2 database into detailed exchange lists.
    \item \texttt{search\_waste}: Provides functions for searching and categorising waste generation-related exchange data.
    \item \texttt{search\_material}: Provides functions for searching and categorising material demand-related exchange data.
    \item \texttt{make\_custom\_database}: Facilitates the creation of custom databases based on the waste and material search categories.
    \item \texttt{method\_editor}: Manages the custom LCIA methods for waste and material footprint calculations.
    \item \texttt{exchange\_editor}: Appends `pseudo-biosphere' exchanges to activities to match their waste generation and material demand exchanges in the technosphere.
    \item \texttt{verify\_database}: Performs verification of the manipulated databases.
\end{itemize}

\subsubsection{Configuration modules:}
\begin{itemize}
    \item \texttt{custom\_config}: Provides functions for managing the configuration of the WasteAndMaterialFootprint package.
    \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
    \item \texttt{queries\_waste}: Defines search parameters and categories for waste generation exchanges (user editable).
    \item \texttt{queries\_materials}: Defines search parameters and categories for material demand exchanges (user editable).
\end{itemize}

\autoref{fig:WMF_flowchart} shows the flowchart of the WasteAndMaterialFootprint tool. The subsequent subsections describe the computational framework and the modules in more detail.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/WMF_flowchart.pdf}
    \caption{Flowchart of the WasteAndMaterialFootprint tool \cbox{Figure needs to be updated with new names + premise}}

    \label{fig:WMF_flowchart}
\end{figure}

\subsubsection{Computational framework}
Developed in the Python programming language (version 3), the WMF tool extends the brightway2 LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}. Additionally, the \texttt{wurst} package is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}. Integration with \texttt{premise} package~\citep{sacchi2022premise} enables the user to easily create and manipulate prospective LCA databases.

\subsubsection{Generation of prospective LCA databases}
Future waste and material footprints can be projected using the \texttt{future\_scenarios} module, which uses \texttt{premise} to generate prospective scenario databases based on the configuration in \texttt{user\_settings}. These prospective databases can be custom-defined by the user or can be constructed with the future projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg}  that can be paired with a variety of mitigation scenarios.

\subsubsection{Database expansion}
The \texttt{explode\_database} module uses \texttt{wurst} to deconstruct LCA databases into a list of individual exchanges representing all of material and energy flows in the technosphere model. This dataset being converted into a pandas DataFrame and stored as a binary \texttt{.pickle} file for subsequent analysis.

\subsubsection{Waste and material flow identification and categorisation}

The \texttt{search\_waste} and \texttt{search\_material} modules apply user-defined search parameters from \texttt{queries\_waste} and \texttt{queries\_materials} to identify relevant waste and material flows in the list of technosphere exchanges generated by \texttt{explode\_database} and categorises them accordingly. The results of the search functions are stored in \texttt{.csv} files for subsequent use in the WMF tool's workflow.

In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods. The waste categories are:
\begin{itemize}[itemsep=0pt]
    \item digestion
    \item composting
    \item open burning
    \item incineration
    \item recycling
    \item landfill
    \item hazardous
    \item non-hazardous
    \item carbon dioxide (in prospective databases, carbon capture and storage is included)
    \item total
\end{itemize}

In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}. The CRM list is a list of 30 materials that are considered critical to the EU economy and are at risk of supply disruption. Further materials of interest were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas. The identity of the materials considered and their categorical groupings are easily customisable by the user. A full list of 59 materials included in the default configuration is provided in the supplementary material.

\subsubsection{Creation of custom `pseudo-biosphere' databases}
Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module. This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project.

\subsubsection{LCIA method management}
The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the WMF tool. This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange. The methods are stored in the Brightway2 project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods. Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods, changing the perspective from waste consumed by treatment to waste generated by the activity.

\subsection{Exchange editing}
The \texttt{exchange\_editor} module loads the \texttt{.csv} files created by the search functions and appends `pseudo-biosphere' exchanges to the matching activities in the LCA database. This is the most computationally intensive part of the WMF tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.

\subsubsection{Database Verification}
The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the WMF tool has processed the database correctly.

\subsection{Case Study Methodology}
\label{sec:method-case_study}

\subsubsection{Activities}
This case study investigated five types of Li-ion batteries, each represented by specific market activities:
\begin{itemize}[itemsep=0pt]
    \item Li-ion, NMC111, rechargeable, prismatic
    \item Li-ion, LiMn2O4, rechargeable, prismatic
    \item Li-ion, NCA, rechargeable, prismatic
    \item Li-ion, NMC811, rechargeable, prismatic
    \item Li-ion, LFP, rechargeable, prismatic
\end{itemize}

\subsubsection{Methods}
In addition to the Waste Footprint and Material Demand footprint methods created by the WMF tool, the following standard LCIA methods were applied for comparison:
\begin{itemize}[itemsep=0pt]
    \item ReCiPe 2016 v1.03, midpoint (I)
    \item EF v3.0 no LT
    \item EDIP 2003 no LT
    \item Crustal Scarcity
\end{itemize}

\subsubsection{Databases}
The primary source of life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff. Additionally, the WMF tool was used to create prospective database sets using the \mbox{REMIND} model with the following Representative Concentration Pathways (RCPs):
\begin{itemize}
    \item \textbf{SSP2 -base:}\\ representing ca. 3.5°C increase in global temperatures to 2100
    \item \textbf{SSP2-PkBudg500:}\\ meeting Paris climate goals, ca. 1.3°C increase to 2100
\end{itemize}

For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with the WMF tool over the time series: 2020, 2040, 2060, 2080, 2100.

\subsubsection{Calculations}
For each combination of activity, method, and database, a single score `LCIA' was calculated along with details of the top contributing processes. Additionally, for the Waste and Material Footprint methods, a contribution analysis was performed. This involved utilizing the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\cite{mutel2016brightway2analyzer}, an additional component of the Brightway2 LCA framework. This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their CPC codes. This provides insight into the products and sectors in the supply chain of the activity that carry the most responsibility for the final footprint.

### 2024-01-24 latex_files/results.tex
first-commit-(moved-from-private-repo)

\cbox{Section word count: 1400}

\subsection{WasteAndMaterialFootprint tool}\label{sec:results-wmf}
We have written an extension to the brightway2 LCA framework that calculates the waste footprint of a product or service.
It finds upstream waste flows in a supply chain, categorises waste flows into 14 types and finds hotspots in waste generation.
It explodes the database, identifies upstream waste exchanges, edits them and writes custom WasteFootprint methods. The waste flows then become pseudo-biosphere flows and the waste footprint can be calculated as an LCIA method. The WasteFootprint tool can be easily applied to calculate the `footprints' of other supply-chain flows such as water, gas, and critical raw materials.

In the context of environmental impact assessments, where a variety of indicators have been proposed and debated for their significance and communicability \citep{vanham2019footprints,ridoutt2013footprints}, the WasteAndMaterialFootprint (WMF) tool serves to re-simplify environmental decision-making. This tool leverages the findings that a limited set of indicators can cover a significant portion of environmental impact variance \citep{steinmann2017resourcefootprints}, enabling more streamlined and effective assessment processes.

\subsubsection{Waste exchanges}\label{sec:results-wmf-waste_exchanges}

In LCA waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows. Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}. There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill. In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments. In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the long-term impacts of disposal.

The purpose of the WMF tool is not to quantify the environmental impacts of waste treatment, but rather to quantify the waste flows themselves, even those that are finally consumed by waste treatment processes. It provides, thus, not an impact assessment in the traditional sense, but an accounting of the waste generated by a product or service, regardless of the end-of-life fate of these flows. By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others. In any case, waste avoidance is of critical importance and by quantifying and classifying the waste total attributed to a product or service in an LCA database, the WMF tool provides a practical means to identify hotspots and opportunities for waste reduction and efficiency.

The logic of screening for waste exchanges is based on a set of boolean search queries (`AND', `OR', and `NOT') that are applied in a list comprehension to the names of every exchange in the LCA database (see `search\_queries.py' for the full list). In this way, the search queries enable classification into categories (such as `hazardous solid' and `incineration liquid') and permit the identification of waste exchanges in addition to those directly connected to waste treatment processes. The search queries are tailored to the specific database and the user can easily modify them to suit their needs. In the default settings, there are a total of 18 waste classifications (9 categories, each separated into liquid and solid waste) For example, the identification of `non-hazardous solid' waste exchanges is based on the following search query; \texttt{AND = [`waste'], NOT = [`hazardous', `radioactive'], UNIT = [`kilogram']} (this can also be inferred and confirmed by comparison with the difference between the results of `total solid' and `hazardous solid').\ \autoref{tab:wf_results} presents a list of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100. Note that for prospective databases, a waste category for `carbon dioxide' has been created to account for the carbon capture and storage (CCS) that has been implemented in this model.

\begin{table}[ht]
    \centering
    \caption{WasteFootprint search results for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
    \begin{tabular}{llr}
        \toprule
        \textbf{Waste exchanges} & \textbf{Unit} & \textbf{Exchange count} \\
        \midrule
        digestion & kilogram & 4 \\
        composting & kilogram & 26 \\
        open burning & kilogram & 535 \\
        incineration & kilogram & 2171 \\
        recycling & kilogram & 137 \\
        landfill & kilogram & 1530 \\
        hazardous & kilogram & 1928 \\
        carbon dioxide & kilogram & 119 \\
        total & kilogram & 29524 \\
        digestion & cubic meter & 16 \\
        composting & cubic meter & 0 \\
        open burning & cubic meter & 0 \\
        incineration & cubic meter & 2 \\
        recycling & cubic meter & 0 \\
        landfill & cubic meter & 2 \\
        hazardous & cubic meter & 437 \\
        carbon dioxide & cubic meter & 0 \\
        total & cubic meter & 4360 \\
        \bottomrule
\end{tabular}
\end{table}

\subsubsection{Material exchanges}\label{sec:results-wmf-material_exchanges}

The logic for the identification of material exchanges with the WMF tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest. That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange. A useful feature of the WMF tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges. For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs. Similarly, the total demand for all critical raw materials (CRMs) can be easily calculated in the same manner.

As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the `crustal scarcity indicator' (which provides only an aggregated, abstracted endpoint)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}. In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere. In the WMF tool, however, the accounting for material demand is based on exchanges solely within the technosphere. This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials. Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the WMF tool can provide insight into the broader supply-chain impacts of the demand for this metal. If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the WMF results---supply chain pressure for one material can result in lessening of supply chain pressure for another. In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.

\subsection{Case study: Li-ion batteries}\label{sec:results-case_study}

As listed in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (as well as a variety of other indicators) for the unaltered inventories for five Li-ion batteries with the functional unit being 1~kg of battery. The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the WMF tool. This section includes some highlights of the results and the full results are available in the supplementary material.

\subsubsection{Temporal and scenario variation in waste and material footprints}\label{sec:results-case_study-total_footprints}

\autoref{fig:waste_total} shows the total solid waste footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model. The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced. The LiMn2O4 battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery. In each case there was a slight downward trend in the waste footprints between 2020 and 2100. This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is included in the models over this time. For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/total_waste.png}
    \caption{Total solid waste footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_total}
\end{figure}

The inclusion of carbon capture and storage in the prospective databases using the PkBudg500 RCP is evident in \autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/carbondioxide.png}
    \caption{Carbon dioxide waste (from carbon capture and storage) footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:carbondioxide}
\end{figure}

For the phosphate demand footprints that are depicted in \autoref{fig:phosphate}, the LFP (lithium iron phosphate) battery has a much larger footprint than the other batteries, consistent with its composition. In this case, the phosphate footprint of all batteries is shown to decrease over the period from 2020--2100, and the RCP scenarios are seen to converge between 2020 and 2040.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/phosphate.png}
    \caption{Phosphate material demand footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:phosphate}
\end{figure}

\subsubsection{Contribution of `top-processes' in the supply chain}\label{sec:results-case_study-topprocesses}

\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the cobalt footprint of the LiMn2O4 battery under the baseline scenario from 2020--2100. The total footprint is seen to almost triple, from 2.2~kg/kg in 2020 to 6.2~kg/kg in 2100. This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model. The fractional contributions of the top processes is remains relatively steady over the coming century in this case.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/top-processes.png}
    \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:top_contribution}
\end{figure}

\subsubsection{Contribution of industrial sectors in the supply chain}\label{sec:results-case_study-topsectors}

\autoref{fig:cpc_contribution} shows the contribution of industrial sectors (grouped by CPC classifications) to the total liquid waste footprint of the NCA battery under the PkBudg500 pathway.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/cpc_contribution.png
    }
    \caption{Contribution of industrial sectors to the liquid waste footprint of the NCA battery from 2020 to 2100 under the SSP2 scenario using the PkBudg500 RCP of the REMIND model.}\label{fig:cpc_contribution}
\end{figure}

### 2024-01-24 latex_files/supplementary_material.tex
first-commit-(moved-from-private-repo)

\documentclass{article}
\usepackage{booktabs} % For better quality table rules
\usepackage{caption} % For better caption format

\begin{document}

\section{List of Materials and Their Markets}

\begin{table}[ht]
\centering
\caption{A comprehensive list of various materials in the default configuration.}
\label{tab:materials}

\begin{tabular}{ll}
\toprule
\textbf{Market Name} & \textbf{Material group} \\
\midrule
market for aluminium & aluminium \\
market for antimony & antimony \\
market for bauxite & bauxite \\
market for beryllium & beryllium \\
market for bismuth & bismuth \\
market for cadmium & cadmium \\
market for calcium borates & borates \\
market for cement & cement \\
market for cerium & cerium \\
market for chromium & chromium \\
market for coal & coal \\
market for cobalt & cobalt \\
market for coke & coke \\
market for copper & copper \\
market for dysprosium & dysprosium \\
market for erbium & erbium \\
market for europium & europium \\
market for electricity & electricity \\
market for ferroniobium & niobium \\
market for fluorspar & fluorspar \\
market for gadolinium & gadolinium \\
market for gallium & gallium \\
market for gold & gold \\
market for graphite & graphite \\
market for hafnium & hafnium \\
market for helium & helium \\
market for holmium & holmium \\
market for hydrogen & hydrogen \\
market for indium & indium \\
market for latex & latex \\
market for lithium & lithium \\
market for magnesium & magnesium \\
market for natural gas & natural gas \\
market for nickel & nickel \\
market for palladium & palladium \\
market for petroleum & petroleum \\
market for phosphate & phosphate rock \\
market for platinum & platinum \\
market for rare earth & rare earth \\
market for rhodium & rhodium \\
market for sand & sand \\
market for selenium & selenium \\
market for scandium & scandium \\
market for silicon & silicon \\
market for silver & silver \\
market for sodium borates & borates \\
market for strontium & strontium \\
market for tantalum & tantalum \\
market for tellurium & tellurium \\
market for tin & tin \\
market for titanium & titanium \\
market for uranium & uranium \\
market for tungsten & tungsten \\
market for vanadium & vanadium \\
market for vegetable oil & vegetable oil \\
market for tap water & water \\
market for water & water \\
market for zinc & zinc \\
market for zirconium & zirconium \\
\bottomrule
\end{tabular}
\end{table}

\end{document}

### 2024-01-30 latex_files/discussion.tex
update-manuscript

--- 2024-01-24 latex_files/main.tex
@@
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,authoryear]{elsarticle}
-\documentclass[final,5p,authoryear]{elsarticle}
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-\usepackage{lineno}
-\usepackage{booktabs}
-% ---------------------------------------------------------------------
-%             FOR DRAFTING / REVIEW (DELETE BEFORE SUBMISSION)
-% ---------------------------------------------------------------------
-\usepackage{caption}
-\usepackage[T1]{fontenc}
-\usepackage[british]{babel}
-\usepackage{microtype}
-\usepackage[tt=false]{libertine}
-\usepackage{xcolor}
-\definecolor{blue}{RGB}{33,150,209}
-\definecolor{red}{RGB}{255,0,0}
-\usepackage{xurl}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{pdfpages}
-\usepackage{enumitem}
-\usepackage{lineno}
-\setlength{\linenumbersep}{5pt}
-% Save the old \texttt command
-\let\oldtexttt\texttt
-% Redefine the \UrlFont to use a smaller typewriter font
-\renewcommand{\UrlFont}{\ttfamily\small}
-% Redefine the \texttt command to use a smaller font size and allow line breaks
-\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
-
-\usepackage{tcolorbox}
-\newcommand{\cbox}[1]{
-    \begin{tcolorbox}[hbox, colback=red!5!white, colframe=red!65!black, boxrule=0.25pt, boxsep=2pt, left=2pt, right=2pt, top=1pt, bottom=1pt]
-        \small\sffamily #1
-    \end{tcolorbox}
-}
-
-% ------------------------------------------------------
-
-\journal{Resources, Conservation and Recycling}
-
-\begin{document}
-
-\linenumbers
-
-    % SPOTLIGHTS (5 bullet points, max 120 characters each)
-    {\Large Spotlights}
-    \vspace{1em}
-    \begin{description}[style=nextline]
-        \item[Bullet 1: Critical context and background information on the problem addressed] Understanding waste and material flows is essential for developing the circular economy and supply chain resilience
-        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A program was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
-        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is flexible and easy to use.
-        \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
-        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
-    \end{description}
-
-    \newpage
-    % %% GRAPHICAL ABSTRACT
-    % \begin{graphicalabstract}
-    %     %\includegraphics{grabs}
-    % \end{graphicalabstract}
-
-    %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
-    \begin{highlights}
-        \item WasteAndMaterialFootprint, a new tool for quantifying waste and material flows in LCA.
-        \item Assesses supply risks by calculating demand for critical materials.
-        \item Simplifies quantification of user-specified waste and material categories.
-        \item Rapidly identifies waste and material demand hotspots.
-        \item Presents a case study of the battery supply chain.
-    \end{highlights}
-
-\begin{frontmatter}
-
-    \title{WasteAndMaterialFootprint: A python package to quantify supply chain flows of waste and material in LCA databases}
-    \author[1]{Stewart Charles McDowall\corref{cor1}}
-    \ead{s.c.mcdowall@cml.leidenuniv.nl}
-    \author[1]{Elizabeth Lanphere}
-    \author[1]{Stefano Cucurachi}
-    \author[1]{Carlos Felipe Blanco}
-
-    \cortext[cor1]{Corresponding author:}
-
-    \affiliation[1]{
-        organization={Institute of Environmental Sciences (CML), Leiden University},
-        addressline={P.O. Box 9518},
-        city={Leiden},
-        postcode={2300 RA},
-        state={South Holland},
-        country={The Netherlands}}
-
-    \begin{abstract}
-        The `circular economy' is a concept focused on the prevention of waste and the reuse of resources.
-Thus, identifying and quantifying waste and material flows are of fundamental importance.
-Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of products and services, those where the implementation of circular principles could be most effective.
-
-        Introducing WasteAndMaterialFootprint, a Python tool extending the Brightway framework for flexible, user-defined quantification of supply-chain demands in current and future scenarios.
-This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
-
-        A case study on battery supply chains demonstrates the tool's utility.
-WasteAndMaterialFootprint quantifies and compares demands and potential environmental impacts, aiding sustainable decision-making.
-It contributes to a more sustainable, circular economy by providing detailed material usage and waste generation analysis.
-
-    \end{abstract}
-
-    % KEYWORDS (max 6)
-    \begin{keyword}
-        circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
-    \end{keyword}
-
-\end{frontmatter}
-
-%% main text
-\section{Introduction}
-\label{sec:introduction}
-\input{introduction.tex}
-
-\section{Methodology}
-\label{sec:methodology}
-\input{methodology.tex}
-
-\section{Results}
-\label{sec:results}
-\input{results.tex}
-
-\section{Discussion}
-\label{sec:discussion}
-\input{discussion.tex}
-\cbox{Section word count: aim for 400?}
 \begin{itemize}
@@
 \end{itemize}
-
-\section{Conclusions}
-\cbox{Section word count: aim for 150?}
-
-This paper extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply-chain risk and potential environmental damage.
-
-\label{sec:conclusions}
-
-% Back matter
-\section*{Data availability} 
-All data used in this analysis are publicly available online under the noted sources.
-The WasteAndMaterialFootprint tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/WasteAndMaterialFootprint}.
-The full source code for the WasteAndMaterialFootprint tool is available at \url{https://www.github.com/Stew-McD/WasteAndMaterialFootprint}.
-A user guide and comprehensive documentation are available at \url{https://wasteandmaterialfootprint.readthedocs.io}.
-
-\section*{CRediT authorship contribution statement}
-\cbox{Co-authors, please check this and change as necessary.}
-\textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-\textbf{Elizabeth Lanphere:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-\textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-\textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-
-\subsection*{Alternative CRediT statement}
-
-\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit.pdf}
-
-\section*{Declaration of competing interest}
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-\section*{Acknowledgements}
-\cbox{Should I put FutuRaMa here?}
-
-    Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
-101058522 (project FutuRaM --- futuram.eu).
-The authors would like to thank the reviewers for their valuable comments and suggestions.
-
-\section*{Supplementary material}
-    The supplementary material contains the following:
-    \begin{enumerate}
-        \item List of waste and material categories
-        \item Example output of the WasteAndMaterialFootprint tool
-        \item Extended results of the case study
-    \end{enumerate}
-
-\bibliographystyle{elsarticle-harv}
-\bibliography{references.bib}
-
-% \appendix
-
-% \section{Example output}
-% \label{app:example_output}
-
-\end{document}

### 2024-01-30 latex_files/introduction.tex
update-manuscript

@@
-\cbox{Section word count: ~1700}
-\cbox{Needs work to make it flow better, also shortened}
-
 The development of a `circular economy' has become a critical area of focus in the imperative pursuit of achieving sustainability objectives and curtailing our environmental footprint within planetary boundaries \citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
@@
 
-Thus, we were motivated to create a flexible and policy-driven approach to understanding a product’s demand on the waste management system both now and in the future to ensure that models of circularity and improvements in technology and policy for waste are included in LCA modeling.
+Thus, we were motivated to create a flexible and policy-driven approach to understanding a product`s demand on the waste management system both now and in the future to ensure that models of circularity and improvements in technology and policy for waste are included in LCA modeling.
 
@@
 Given the increasing focus placed on this, as well as ideals of improving resource efficiency and developing the circular economy, it is essential to understand the material demands of products and processes.
-While LCA seeks to model the technosphere (the sum of all supply-chains), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.
+While LCA seeks to model the technosphere (the sum of all supply chains), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.
 
-A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess longterm global scarcity of minerals in life cycle assessment (LCA).
+A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in life cycle assessment (LCA).
 This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations.
@@
 
-A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining 'invisible' in traditional life cycle assessment (LCA) practices~\citep{laurenti2016wastefootprint}.
+In LCA, waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows.
+Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
+There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill.
+In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
+In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the long-term impacts of disposal.
+
+A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional life cycle assessment (LCA) practices~\citep{laurenti2016wastefootprint}.
 This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
 Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment.
-The environmental significance of waste and its correlation with other indicators has been a subject of extensive research.
+The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
 For example, studies have shown that popular resource footprints can cover a significant portion of environmental impact variance in product rankings~\citep{steinmann2017resourcefootprints}.
@@
 The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting.
-The WMF tool presented in the current work, builds on the work of Laurenti et.al., providing a more flexible, transparent and user-friendly approach to quantifying waste flows in LCA.
-Moreover, the WMF tool is not limited to waste, but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.
+The WMF tool presented in the current work provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
+Moreover, the WMF tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.
 
@@
 To better assess waste and material flows in LCA, the authors have developed a Python program built on the Brightway2 framework and designed to track these exchanges by translating them into indicators and impact categories.
-In this study, we present the WasteAndMaterialFootprint tool that enables LCA practitioners to easily aggregate the mass and volume of any desired exchange, and to creation of flexible categories to differentiate between waste types and End-of-Life (EOL) handling using (in this case) the Ecoinvent 3.9 cutoff database.
+In this study, we present the WasteAndMaterialFootprint (WMF) tool that enables LCA practitioners to easily aggregate the mass and volume of any desired exchange, and to creation of flexible categories to differentiate between waste types and End-of-Life (EOL) handling using (in this case) the Ecoinvent 3.9 cutoff database.
 
-While methods with similar aims exist, they either lack flexibility and specificity \citep{foen2021ecofactors} or are cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.
+The purpose of the WMF tool is not to quantify the environmental impacts of waste treatment, but rather to quantify the waste flows themselves, even those that are finally consumed by waste treatment processes.
+It provides, thus, not an impact assessment in the traditional sense, but an accounting of the waste generated by a product or service, regardless of the end-of-life fate of these flows.
+By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
+In any case, waste avoidance is of critical importance and by quantifying and classifying the waste total attributed to a product or service in an LCA database, the WMF tool provides a practical means to identify hotspots and opportunities for waste reduction and efficiency.
+
+While methods with similar aims exist, they either lack flexibility and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.
 

### 2024-01-30 latex_files/main.tex
update-manuscript

@@
 %% \documentclass[final,5p,times,authoryear]{elsarticle}
-\documentclass[final,5p,authoryear]{elsarticle}
+\documentclass[review,3p,authoryear]{elsarticle}
 %% For including figures, graphicx.sty has been loaded in
@@
 \usepackage{lineno}
+\usepackage{float}
 \setlength{\linenumbersep}{5pt}
@@
     \begin{abstract}
+        \cbox{Abstract word count: 150, Limit: 150}
+
         The `circular economy' is a concept focused on the prevention of waste and the reuse of resources.
@@
 \end{frontmatter}
+\cbox{Total word count: ~6000, Limit: 5000 (can be easily condensed)}
 
@@
 \label{sec:introduction}
+\cbox{Section word count: 1650}
+\cbox{Needs work to make it flow better, also shortened}
 \input{introduction.tex}
@@
 \label{sec:methodology}
+\cbox{Section word count: 2000}
 \input{methodology.tex}
@@
 \section{Results}
+\cbox{Section word count: 1400}
 \label{sec:results}
@@
 \cbox{Section word count: aim for 400?}
-\begin{itemize}
-    \item Summary of the results, 1 par
-    \item -----------------------------
-    \item Advantages of the WMF tool, 1 par
-    \item -----------------------------
-    \item Limitations, 1 par
-    \item -----------------------------
-    \item Future possibilities, 1 par
-\end{itemize}
 
 \section{Conclusions}
-\cbox{Section word count: aim for 150?}
+\label{sec:conclusions}
+\cbox{Section word count: 250}
+We have written the WasteAndMaterialFootprint (WMF) tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
+It explodes the database, identifies upstream waste and material exchanges, edits them and writes matching custom WMF methods.
+These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
+The WMF tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.
+
+Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
+While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
+
+The WMF tool, in contrast, gives LCA practitioners a simple, flexible, and transparent way to calculate the supply chain waste and material footprints, delivering results in standard units as direct aggregations of the relevant demand inventories.
 
 This paper extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply-chain risk and potential environmental damage.
-
-\label{sec:conclusions}
 
@@
 \section*{Acknowledgements}
-\cbox{Should I put FutuRaMa here?}
-
     Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.

### 2024-01-30 latex_files/methodology.tex
update-manuscript

@@
-\cbox{Section word count: 1600}
-
-First, we identified all Ecoinvent technosphere exchanges that produce waste.
-We then further classified the waste into non-mutually exclusive categories such as its destination (i.e., dumped, incinerated, etc.), hazardousness, and form (solid vs.
-liquid).
-After cloning the technosphere exchanges as biosphere exchanges, we aggregated and quantified them, by waste category, into matching impact categories in the Life Cycle Impact Assessment.
-
-In our simplified test case of several battery types, we were able to identify `waste hotspots' and distinguish the major sources of contribution to waste generation on a process level.
-One conspicuous result from the case study (and potential direction for further work) is that many waste flows are tied to processes lacking a clear EOL pathway.
-Further development of this tool could involve developing an algorithm using identifiers of each background waste process to predict where these uncategorised wastes land in their EOL management.
-
 This section is divided into two parts.
@@
 
-\subsubsection{Functional modules:}
+\subsubsection{Functional modules}
 \begin{itemize}
@@
 
-\subsubsection{Configuration modules:}
+\subsubsection{Configuration modules}
 \begin{itemize}
@@
     \includegraphics[width=0.7\columnwidth]{figures/WMF_flowchart.pdf}
-    \caption{Flowchart of the WasteAndMaterialFootprint tool \cbox{Figure needs to be updated with new names + premise}}
-
-    \label{fig:WMF_flowchart}
+    \caption{Flowchart of the WasteAndMaterialFootprint tool \cbox{Figure needs to be updated with new names + premise}}\label{fig:WMF_flowchart}
 \end{figure}
@@
 
+\subsubsection{Waste exchanges}\label{sec:method-wmf-waste_exchanges}
+
 In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
@@
     \item non-hazardous
-    \item carbon dioxide (in prospective databases, carbon capture and storage is included)
+    \item carbon dioxide 
     \item total
@@
 
+The logic of screening for waste exchanges is based on a set of boolean search queries (`AND', `OR', and `NOT') that are applied in a list comprehension to the names of every exchange in the LCA database (see `search\_queries.py' for the full list).
+In this way, the search queries enable classification into categories (such as `hazardous solid' and `incineration liquid') and permit the identification of waste exchanges in addition to those directly connected to waste treatment processes.
+The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
+In the default settings, there are a total of 18 waste classifications (9 categories, each separated into liquid and solid waste) For example, the identification of `non-hazardous solid' waste exchanges is based on the following search query; \texttt{AND = [`waste'], NOT = [`hazardous', `radioactive'], UNIT = [`kilogram']} (this can also be inferred and confirmed by comparison with the difference between the results of `total solid' and `hazardous solid').\ \autoref{tab:wf_results} presents a list of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
+Note that the carbon dioxide waste category does not include emissions to the atmosphere.
+This category is based solely on the accounting of carbon capture and storage (CCS), which is included in many prospective databases as direct sequestration in reservoirs as well as solvent capture.
+
+\begin{table}[ht]
+    \centering
+    \caption{WasteFootprint search results for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
+    \begin{tabular}{llr}
+        \toprule
+        \textbf{Waste exchanges} & \textbf{Unit} & \textbf{Exchange count} \\
+        \midrule
+        digestion & kilogram & 4 \\
+        composting & kilogram & 26 \\
+        open burning & kilogram & 535 \\
+        incineration & kilogram & 2171 \\
+        recycling & kilogram & 137 \\
+        landfill & kilogram & 1530 \\
+        hazardous & kilogram & 1928 \\
+        carbon dioxide & kilogram & 119 \\
+        total & kilogram & 29524 \\
+        digestion & cubic meter & 16 \\
+        composting & cubic meter & 0 \\
+        open burning & cubic meter & 0 \\
+        incineration & cubic meter & 2 \\
+        recycling & cubic meter & 0 \\
+        landfill & cubic meter & 2 \\
+        hazardous & cubic meter & 437 \\
+        carbon dioxide & cubic meter & 0 \\
+        total & cubic meter & 4360 \\
+        \bottomrule
+\end{tabular}
+\end{table}
+
+\subsubsection{Material exchanges}\label{sec:method-wmf-material_exchanges}
 In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
@@
 
+The logic for the identification of material exchanges with the WMF tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
+That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
+A useful feature of the WMF tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
+For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
+Similarly, the total demand for all critical raw materials (CRMs) can be easily calculated in the same manner.
+
+As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the `crustal scarcity indicator' (which provides only an aggregated, abstracted endpoint)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
+In the WMF tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
+This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the WMF tool can provide insight into the broader supply-chain impacts of the demand for this metal.
+If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the WMF results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
+In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
+
 \subsubsection{Creation of custom `pseudo-biosphere' databases}
 Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module.
-This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project.
+This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `WasteAndMaterialFootprint'.
 
@@
 The methods are stored in the Brightway2 project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods.
-Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods, changing the perspective from waste consumed by treatment to waste generated by the activity.
+Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from waste consumed by treatment to waste generated by the activity.
 
@@
 
-\subsection{Case Study Methodology}
-\label{sec:method-case_study}
+\subsection{Case Study Methodology}\label{sec:method-case_study}
 

### 2024-01-30 latex_files/results.tex
update-figures
update-manuscript

@@
-\cbox{Section word count: 1400}
+\subsection{WasteAndMaterialFootprint tool}\label{sec:results-wmf}
 
-\subsection{WasteAndMaterialFootprint tool}\label{sec:results-wmf}
-We have written an extension to the brightway2 LCA framework that calculates the waste footprint of a product or service.
-It finds upstream waste flows in a supply chain, categorises waste flows into 14 types and finds hotspots in waste generation.
-It explodes the database, identifies upstream waste exchanges, edits them and writes custom WasteFootprint methods.
-The waste flows then become pseudo-biosphere flows and the waste footprint can be calculated as an LCIA method.
-The WasteFootprint tool can be easily applied to calculate the `footprints' of other supply-chain flows such as water, gas, and critical raw materials.
-
-In the context of environmental impact assessments, where a variety of indicators have been proposed and debated for their significance and communicability \citep{vanham2019footprints,ridoutt2013footprints}, the WasteAndMaterialFootprint (WMF) tool serves to re-simplify environmental decision-making.
-This tool leverages the findings that a limited set of indicators can cover a significant portion of environmental impact variance \citep{steinmann2017resourcefootprints}, enabling more streamlined and effective assessment processes.
-
-\subsubsection{Waste exchanges}\label{sec:results-wmf-waste_exchanges}
-
-In LCA waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows.
-Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
-There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill.
-In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the long-term impacts of disposal.
-
-The purpose of the WMF tool is not to quantify the environmental impacts of waste treatment, but rather to quantify the waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the waste generated by a product or service, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
-In any case, waste avoidance is of critical importance and by quantifying and classifying the waste total attributed to a product or service in an LCA database, the WMF tool provides a practical means to identify hotspots and opportunities for waste reduction and efficiency.
-
-The logic of screening for waste exchanges is based on a set of boolean search queries (`AND', `OR', and `NOT') that are applied in a list comprehension to the names of every exchange in the LCA database (see `search\_queries.py' for the full list).
-In this way, the search queries enable classification into categories (such as `hazardous solid' and `incineration liquid') and permit the identification of waste exchanges in addition to those directly connected to waste treatment processes.
-The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
-In the default settings, there are a total of 18 waste classifications (9 categories, each separated into liquid and solid waste) For example, the identification of `non-hazardous solid' waste exchanges is based on the following search query; \texttt{AND = [`waste'], NOT = [`hazardous', `radioactive'], UNIT = [`kilogram']} (this can also be inferred and confirmed by comparison with the difference between the results of `total solid' and `hazardous solid').\ \autoref{tab:wf_results} presents a list of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
-Note that for prospective databases, a waste category for `carbon dioxide' has been created to account for the carbon capture and storage (CCS) that has been implemented in this model.
-
-\begin{table}[ht]
-    \centering
-    \caption{WasteFootprint search results for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
-    \begin{tabular}{llr}
-        \toprule
-        \textbf{Waste exchanges} & \textbf{Unit} & \textbf{Exchange count} \\
-        \midrule
-        digestion & kilogram & 4 \\
-        composting & kilogram & 26 \\
-        open burning & kilogram & 535 \\
-        incineration & kilogram & 2171 \\
-        recycling & kilogram & 137 \\
-        landfill & kilogram & 1530 \\
-        hazardous & kilogram & 1928 \\
-        carbon dioxide & kilogram & 119 \\
-        total & kilogram & 29524 \\
-        digestion & cubic meter & 16 \\
-        composting & cubic meter & 0 \\
-        open burning & cubic meter & 0 \\
-        incineration & cubic meter & 2 \\
-        recycling & cubic meter & 0 \\
-        landfill & cubic meter & 2 \\
-        hazardous & cubic meter & 437 \\
-        carbon dioxide & cubic meter & 0 \\
-        total & cubic meter & 4360 \\
-        \bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Material exchanges}\label{sec:results-wmf-material_exchanges}
-
-The logic for the identification of material exchanges with the WMF tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
-That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
-A useful feature of the WMF tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
-For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
-Similarly, the total demand for all critical raw materials (CRMs) can be easily calculated in the same manner.
-
-As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the `crustal scarcity indicator' (which provides only an aggregated, abstracted endpoint)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
-In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
-In the WMF tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
-This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the WMF tool can provide insight into the broader supply-chain impacts of the demand for this metal.
-If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the WMF results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
+An example of the output from the application of the WMF tool has been included in the supplementary material. The manipulated ecoinvent databases 
+can be recreated using the code and instructions available in the package documentation~\citep{mcdowall2023wmfdocs}.
 
@@
 
-As listed in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (as well as a variety of other indicators) for the unaltered inventories for five Li-ion batteries with the functional unit being 1~kg of battery.
+As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (as well as a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the WMF tool.
 This section includes some highlights of the results and the full results are available in the supplementary material.
+Because the WMF methods are integrated into the brightway project as if they were LCIA methods, the results can be visualised in the same way.
+In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
 
@@
 
-\begin{figure}[ht!]
+\begin{figure}[H]
     \centering
@@
 
-The inclusion of carbon capture and storage in the prospective databases using the PkBudg500 RCP is evident in \autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline.
+The inclusion of carbon capture and storage (CCS) in the prospective databases using the PkBudg500 RCP is evident in \autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
+This result highlights the fact that (frequently) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
+The actual deployment of these technologies---approximately 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}---falls far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
     \centering
@@
 This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model.
-The fractional contributions of the top processes is remains relatively steady over the coming century in this case.
+The fractional contributions of the top processes remains relatively steady over the coming century in this case.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
     \centering
@@
 
-\subsubsection{Contribution of industrial sectors in the supply chain}\label{sec:results-case_study-topsectors}
+\subsubsection{Contribution of sectors in the supply chain}\label{sec:results-case_study-topsectors}
 
-\autoref{fig:cpc_contribution} shows the contribution of industrial sectors (grouped by CPC classifications) to the total liquid waste footprint of the NCA battery under the PkBudg500 pathway.
+\autoref{fig:cpc_contribution} shows the contribution of sectors (grouped by CPC) to the total natural gas footprint of the LFP battery under the PkBudg500 pathway.
+In this example, the relative contribution of the sector `47160: Electronic integrated circuits' is seen to decrease from 11\% in 2020 to 6\% in 2100, while over the same period the contribution of the sector `46430: Parts of primary cells, primary batteries and electrodes' increases from 29\% to 38\%.
+The method used to calculate these contributions involves traversing the supply chain branches to a certain level (max.
+4, in this case), cutting a specified point (5\% in this case), and grouping the value of the `leaves' by their CPC code.
+The results, therefore, will depend on how deeply the user would like to inspect the supply chain.
+Additionally, the utility of these results is dependent on how well the CPC codes define the processes in the supply chain for the particular case.
 
-\begin{figure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/cpc_contribution.png
+    \includegraphics[width=0.6\linewidth]{figures/cpc_contribution.png
     }
@@
 \end{figure}
+
+\subsubsection{Comparison with `similar' methods}\label{sec:results-case_study-methodcomparison}
+
+A comparison of the results from the WMF tool's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
+In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
+Such comparability was also observed for other fossil-fuel-related methods (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
+Correlation between standard LCIA methods and WMF methods, is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
+In standard LCIA methods, impact scores are derived exclusively from the magnitude of the exchanges between the biosphere and the technosphere, that is, extraction and emission.
+In the WMF methods, the footprint scores are based on an accounting of either waste generated or material demand, both of which are technosphere-technosphere exchanges in terms of LCA modelling.
+For the material demand methods especially, this distinction is critical.
+For example, application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the WMF method provides an aggregation of the exchanges with the market for that metal.
+The WMF method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply-chain pressures that are not captured by the standard LCIA methods.
+This makes the WMF methods more sensitive to the modeling choices (e.g., allocation) that are generally embedded in LCA databases.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figures/comparison_methods.png
+    }
+    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' with the `WMF - Coal (black)' in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+}\label{fig:comparison_methods}
+\end{figure}
+
+\subsubsection{Comparison with other studies}\label{sec:results-case_study-comparison}
+
+\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in ecoinvent version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
+The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
+\autoref{tab:results-case_study-comparison} presents a comparison of the results from the two studies, where possible.
+For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste, Laurenti et al.
+reported 95\% of the total waste, whereas the WMF tool reported only 3\%.
+The reason for this discrepancy is explained by the fact that in the method of Laurenti et al., a ``waste flow was regarded as hazardous when the hazardousness
+was clearly stated in its subsequent waste treatment activity''.
+The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
+In the WMF method, the source database is exploded into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+
+\begin{table}[H]
+\centering
+\caption{Comparison of the results from the WMF battery case study (database: ecoinvent 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: ecoinvent 3.5, activity:`Li-ion').}
+\label{tab:results-case_study-comparison}
+\begin{tabular}{lll}
+\toprule
+\textbf{Indicator} & \textbf{WMF} & \textbf{Laurenti et al.} \\
+\midrule
+Total solid waste (kg/kg)& 50.9 & 62.5 \\
+Total liquid waste (kg/kg) & 3.53 & 3.63 \\
+Hazardous waste (kg/kg) & 1.47 & 62.6 \\
+Recycled waste (kg/kg) & 1.59 & 1.98 \\
+\bottomrule
+\end{tabular}
+\end{table}

### 2024-01-31 latex_files/discussion.tex
update-manuscript

@@
-\begin{itemize}
-    \item Summary of the results, 1 par
-    \item -----------------------------
-    \item Advantages of the WMF tool, 1 par
-    \item -----------------------------
-    \item Limitations, 1 par
-    \item -----------------------------
-    \item Future possibilities, 1 par
-\end{itemize}
+Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
+While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
+
+The WMF tool, in contrast, gives LCA practitioners a simple, flexible, and transparent way to calculate the supply chain waste and material footprints, delivering results in standard units as direct aggregations of the relevant demand inventories.
+Once the databases have been processed with the WMF tool, the user can then easily apply the WMF methods to calculate the supply-chain waste and material footprints for any activity in the same way that they would calculate any LCIA indicator.
+
+The simple case study presented in this paper demonstrates both the utility and the limitations of the WMF tool.
+It was shown the WMF tool was able to calculate categorised aggregates and contribution analyses of both the waste generated and material demands in the present and future supply chains of the Li-ion batteries under consideration.
+Further, integration of the WMF methods as `pseudo-LCIA' methods allows the user to easily make use of the WMF tool in their preferred LCA software, be it code-based like \texttt{brightway2} or graphical like \texttt{ActivityBrowser}.
+
+One main limitation of the WMF tool is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated, which would be needed to thoroughly assess the environmental impacts of this waste.
+Currently, the user would need to manually explore the `waste inventory' produced by the application of the WMF to determine if, for example, the waste generated represents an actual loss of resources, or is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+A methodic classification of waste exchanges and the end-of-life fates will be facilitated by the more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
+
+Furthermore, the utility of the WMF tool in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is little adaption of the future background inventories.
+In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.

### 2024-01-31 latex_files/main.tex
update-manuscript

@@
 %% \documentclass[final,5p,times,authoryear]{elsarticle}
-\documentclass[review,3p,authoryear]{elsarticle}
+\documentclass[final,5p,authoryear]{elsarticle}
 %% For including figures, graphicx.sty has been loaded in
@@
 \usepackage{caption}
+\usepackage{subcaption}
 \usepackage[T1]{fontenc}
@@
 \input{discussion.tex}
-\cbox{Section word count: aim for 400?}
+\cbox{Section word count: 450}
 
@@
 We have written the WasteAndMaterialFootprint (WMF) tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
-It explodes the database, identifies upstream waste and material exchanges, edits them and writes matching custom WMF methods.
+It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom WMF methods.
 These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
 The WMF tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.
-
-Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
-While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
-
-The WMF tool, in contrast, gives LCA practitioners a simple, flexible, and transparent way to calculate the supply chain waste and material footprints, delivering results in standard units as direct aggregations of the relevant demand inventories.
 
@@
         \item Example output of the WasteAndMaterialFootprint tool
-        \item Extended results of the case study
+        \item List of identified waste and material exchanges
+        \item Code (python script) used for case study
+        \item Inventory and methods used in the case study
+        \item Complete tabulated results of the case study
+        \item Complete visualisations of the case study
     \end{enumerate}

### 2024-01-31 latex_files/methodology.tex
update-manuscript

@@
 This section is divided into two parts.
-In \autoref{sec:method-WMF_tool}, we describe the WasteAndMaterialFootprint tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
+In \autoref{sec:method-WMF_tool}, we describe the WasteAndMaterialFootprint (WMF) tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
 
-\subsection{The WasteAndMaterialFootprint tool}
+\subsection{The WasteAndMaterialFootprint (WMF) tool}
 \label{sec:method-WMF_tool}
 
-The WasteAndMaterialFootprint tool is a Python package that allows one to calculate the waste and material footprint of any product or service inside of LCA databases.
+The WMF tool is a Python package that allows one to calculate the waste and material footprint of any product or service inside of LCA databases.
+Metadata of the WMF tool is presented in \autoref{tab:methods-wmfmetadata}.
+
+\begin{table}[h]
+    \caption{WasteAndMaterialFootprint (WMF) tool metadata}
+    \label{tab:methods-wmfmetadata}
+    \centering
+    \begin{tabular}{ll}
+    \toprule
+    \textbf{Item} & \textbf{Details} \\
+    \midrule
+    Current version & 0.1.21 \\
+    DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
+    Code repository & \url{github.com/Stew-McD/WasteAndMaterialFootprint} \\
+    License & CC0--1.0 license \\
+    Versioning system & git \\
+    Language & Python \\
+    Documentation & \url{wasteandmaterialfootprint.readthedocs.io} \\
+    Main dependencies & brightway2, premise, wurst \\
+    Support email & s.c.mcdowall@cml.leidenuniv.nl \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
 The tool is built on the \texttt{Brightway2} LCA framework~\citep{mutel2017brightway} and is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser} an open-source graphical user interface for LCA.
 The WMF tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023wmfpipy} and is open source under the CC-0 licence.
-The full source code for the WasteAndMaterialFootprint tool is indexed on Zenodo~\citep{mcdowall2023wmfzenodo} and under further development in the GitHub repository~\citep{mcdowall2024wmfgithub}.
+The full source code for the WMF tool is indexed on Zenodo~\citep{mcdowall2023wmfzenodo} and under further development in the GitHub repository~\citep{mcdowall2024wmfgithub}.
 The tool is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases as well by changing the search criteria.
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/WMF_flowchart.pdf}
-    \caption{Flowchart of the WasteAndMaterialFootprint tool \cbox{Figure needs to be updated with new names + premise}}\label{fig:WMF_flowchart}
+    \includegraphics[height=0.8\textheight]{figures/WMF_flowchart.pdf}
+    \caption{Flowchart of the WasteAndMaterialFootprint (WMF) tool}\label{fig:WMF_flowchart}
 \end{figure}

### 2024-01-31 latex_files/results.tex
update-manuscript

@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/top-processes.png}
+    \includegraphics[width=0.6\linewidth]{figures/top-processes.png}
     \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:top_contribution}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/comparison_methods.png
-    }
-    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' with the `WMF - Coal (black)' in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
-}\label{fig:comparison_methods}
+    \begin{subfigure}[b]{0.8\linewidth}
+        \includegraphics[width=\linewidth]{figures/comparison_methods2.png}
+        \caption{}\label{fig:comparison_methods_a}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.8\linewidth}
+        \includegraphics[width=\linewidth]{figures/comparison_methods1.png}
+        \caption{}\label{fig:comparison_methods_b}
+    \end{subfigure}
+    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `WMF - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:comparison_methods}
 \end{figure}

### 2024-02-28 latex_files/cover_letter.tex
update name and readme
update

@@
 \begin{enumerate}
-    \item {Reviewer 1 Name, Affiliation, Email Address}
-    \item {Reviewer 2 Name, Affiliation, Email Address}
+    \item {Rafael Laurenti, IVL --- Swedish Environmental Research Institute, rafa@kth.se}
+    \item {Valentina Bisinella,  DTU --- Technical University of Denmark, valenb@dtu.dk}
     \item {Reviewer 3 Name, Affiliation, Email Address}
@@
       \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
-      Elizabeth Lanphere
+      Elizabeth Lanphear
     \end{minipage}%

### 2024-02-28 latex_files/main.tex
update name and readme
update

@@
     \ead{s.c.mcdowall@cml.leidenuniv.nl}
-    \author[1]{Elizabeth Lanphere}
+    \author[1]{Elizabeth Lanphear}
     \author[1]{Stefano Cucurachi}
@@
 
-        The `circular economy' is a concept focused on the prevention of waste and the reuse of resources.
+        The `circular economy' is a concept centered on `re-X' strategies---on the prevention of waste and the reuse of resources.
 Thus, identifying and quantifying waste and material flows are of fundamental importance.
@@
         A case study on battery supply chains demonstrates the tool's utility.
-WasteAndMaterialFootprint quantifies and compares demands and potential environmental impacts, aiding sustainable decision-making.
+WasteAndMaterialFootprint quantifies and compares inventory demands and potential environmental burdens, aiding sustainable decision-making.
 It contributes to a more sustainable, circular economy by providing detailed material usage and waste generation analysis.
@@
 \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-\textbf{Elizabeth Lanphere:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
+\textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
 \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.

### 2024-03-05 latex_files/cover_letter.tex
update abstract and introduction

@@
 
-We are pleased to submit our original research article titled ``WasteAndMaterialFootprint: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
+We are pleased to submit our original research article titled ``T-reX: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
 This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.

### 2024-03-05 latex_files/introduction.tex
update abstract and introduction

@@
 
-Thus, we were motivated to create a flexible and policy-driven approach to understanding a product`s demand on the waste management system both now and in the future to ensure that models of circularity and improvements in technology and policy for waste are included in LCA modeling.
-
-\subsection{LCA}
-
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
@@
 Several LCIA methods include, to some extent, waste generation \citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption \citep{arvidsson2020csi,foen2021ecofactors}.
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., kg-Si equivalents or Ümweltbelastungspunkte (UBP)).
+These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP)).
 
-\subsection{Material Demand in LCA}
 In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs) \citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
-Given the increasing focus placed on this, as well as ideals of improving resource efficiency and developing the circular economy, it is essential to understand the material demands of products and processes.
-While LCA seeks to model the technosphere (the sum of all supply chains), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.
+While LCA seeks to model the technosphere (a.k.a.
+the anthroposphere), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.
 
-A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in life cycle assessment (LCA).
+A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in LCA.
 This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations.
@@
 While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
-Furthermore, the CSPs are not available for all elements, and the method does not allow for the quantification of material demands in terms of mass or volume.
+Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
 
@@
 
-Though often described simply as `a material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.
+Though often described simply as a `material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.
 Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows.
@@
 
-A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional life cycle assessment (LCA) practices~\citep{laurenti2016wastefootprint}.
+A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional LCA practices~\citep{laurenti2016wastefootprint}.
 This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
@@
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing towards circularity.
-Thus, there is a pressing need for more comprehensive methods that can effectively quantify waste impacts and contribute to a better understanding of a product's total environmental footprint.
-There is currently a lack of a convenient and flexible method for calculating waste flows in Life Cycle Assessment (LCA).
-
+Addressing these hotspots is crucial for advancing towards circularity, however,  there is a lack of a convenient and flexible way to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a better understanding of a product's total environmental footprint.
 \cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity.
@@
 The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting.
-The WMF tool presented in the current work provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
-Moreover, the WMF tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.
+The T-reX tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
+Moreover, the T-reX tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.
 
-\subsection{The WasteAndMaterialFootprint Tool}
+\subsection{The T-reX Tool}
 
-To better assess waste and material flows in LCA, the authors have developed a Python program built on the Brightway2 framework and designed to track these exchanges by translating them into indicators and impact categories.
-In this study, we present the WasteAndMaterialFootprint (WMF) tool that enables LCA practitioners to easily aggregate the mass and volume of any desired exchange, and to creation of flexible categories to differentiate between waste types and End-of-Life (EOL) handling using (in this case) the Ecoinvent 3.9 cutoff database.
+To better assess waste and material flows in LCA, we have developed a Python program built on the Brightway framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into indicators and `pseudo' LCA impact (LCIA) categories.
+In this study, we present the T-reX tool that enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
 
-The purpose of the WMF tool is not to quantify the environmental impacts of waste treatment, but rather to quantify the waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the waste generated by a product or service, regardless of the end-of-life fate of these flows.
+While methods with similar aims exist, they lack customisability and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.
+
+The purpose of the T-reX tool is not to quantify the environmental impacts of material consumption and waste production, but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
+It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
 By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
-In any case, waste avoidance is of critical importance and by quantifying and classifying the waste total attributed to a product or service in an LCA database, the WMF tool provides a practical means to identify hotspots and opportunities for waste reduction and efficiency.
-
-While methods with similar aims exist, they either lack flexibility and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.
-
-This tool provides a method for the calculation of waste footprint impact category results, differentiated by the type of waste handling.
-Furthermore, the tool facilitates rapid investigation and identification of waste hotspots, enabled by standard contribution analysis and Sankey diagram visualization tools.
-The authors consider this a crucial step in addressing the deficit of Life Cycle Assessment (LCA) methods that consider waste flows in the evaluation of a product or process' circular economy potential.
+In any case, avoiding material consumption and generation of waste is of critical importance.
+By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and opportunities for waste reduction and material efficiency.

### 2024-03-05 latex_files/main.tex
update abstract and introduction

@@
     \begin{description}[style=nextline]
-        \item[Bullet 1: Critical context and background information on the problem addressed] Understanding waste and material flows is essential for developing the circular economy and supply chain resilience
-        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A program was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
-        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is flexible and easy to use.
+        \item[Bullet 1: Critical context and background information on the problem addressed] Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
+        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
+        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use.
         \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
@@
     \newpage
-    % %% GRAPHICAL ABSTRACT
-    % \begin{graphicalabstract}
-    %     %\includegraphics{grabs}
-    % \end{graphicalabstract}
+    %% GRAPHICAL ABSTRACT
+    \begin{graphicalabstract}
+        \includegraphics{grabs}
+    \end{graphicalabstract}
 
@@
     \begin{highlights}
-        \item WasteAndMaterialFootprint, a new tool for quantifying waste and material flows in LCA.
+        \item T-reX, a new tool for quantifying waste and material flows in LCA.
         \item Assesses supply risks by calculating demand for critical materials.
@@
 
-    \title{WasteAndMaterialFootprint: A python package to quantify supply chain flows of waste and material in LCA databases}
+    \title{T-reX: A python package to quantify supply chain flows of waste and material in LCA databases}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
     \begin{abstract}
-        \cbox{Abstract word count: 150, Limit: 150}
-
-        The `circular economy' is a concept centered on `re-X' strategies---on the prevention of waste and the reuse of resources.
+        \cbox{Abstract word count: 164, Limit: 150}
+        Minimising waste through the reuse of resources is the quintessential principle of the `circular economy', but relies on our ability to identify and quantify waste and material flows.
         Thus, identifying and quantifying waste and material flows are of fundamental importance.
@@
 
-        Introducing WasteAndMaterialFootprint, a Python tool extending the Brightway framework for flexible, user-defined quantification of supply-chain demands in current and future scenarios.
+        Introducing T-reX, a Python tool extending the Brightway framework for flexible quantification of user-defined supply-chain demands in current and future scenarios.
 This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
@@
         A case study on battery supply chains demonstrates the tool's utility.
-WasteAndMaterialFootprint quantifies and compares inventory demands and potential environmental burdens, aiding sustainable decision-making.
-It contributes to a more sustainable, circular economy by providing detailed material usage and waste generation analysis.
+T-reX quantifies and compares inventory demands and, thus, potential environmental burdens, aiding sustainable decision-making.
+It contributes to the development of the `circular economy' by providing detailed material usage and waste generation analysis.
 
@@
 \cbox{Section word count: 250}
-We have written the WasteAndMaterialFootprint (WMF) tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
+We have written the T-reX tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
 It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom WMF methods.
@@
 All data used in this analysis are publicly available online under the noted sources.
-The WasteAndMaterialFootprint tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/WasteAndMaterialFootprint}.
-The full source code for the WasteAndMaterialFootprint tool is available at \url{https://www.github.com/Stew-McD/WasteAndMaterialFootprint}.
-A user guide and comprehensive documentation are available at \url{https://wasteandmaterialfootprint.readthedocs.io}.
+The T-reX tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T-reX}.
+The full source code for the T-reX tool is available at \url{https://www.github.com/Stew-McD/T-reX}.
+A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.
 
@@
         \item List of waste and material categories
-        \item Example output of the WasteAndMaterialFootprint tool
+        \item Example output of the T-reX tool
         \item List of identified waste and material exchanges

### 2024-03-05 latex_files/methodology.tex
update abstract and introduction

@@
 This section is divided into two parts.
-In \autoref{sec:method-WMF_tool}, we describe the WasteAndMaterialFootprint (WMF) tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
+In \autoref{sec:method-WMF_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
 
-\subsection{The WasteAndMaterialFootprint (WMF) tool}
+\subsection{The T-reX tool}
 \label{sec:method-WMF_tool}
@@
 \begin{table}[h]
-    \caption{WasteAndMaterialFootprint (WMF) tool metadata}
+    \caption{T-reX tool metadata}
     \label{tab:methods-wmfmetadata}
@@
     DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
-    Code repository & \url{github.com/Stew-McD/WasteAndMaterialFootprint} \\
+    Code repository & \url{github.com/Stew-McD/T-reX} \\
     License & CC0--1.0 license \\
@@
     Language & Python \\
-    Documentation & \url{wasteandmaterialfootprint.readthedocs.io} \\
+    Documentation & \url{T-reX.readthedocs.io} \\
     Main dependencies & brightway2, premise, wurst \\
@@
 \begin{itemize}
-    \item \texttt{custom\_config}: Provides functions for managing the configuration of the WasteAndMaterialFootprint package.
+    \item \texttt{custom\_config}: Provides functions for managing the configuration of the T-reX package.
     \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
@@
 
-\autoref{fig:WMF_flowchart} shows the flowchart of the WasteAndMaterialFootprint tool. The subsequent subsections describe the computational framework and the modules in more detail.
+\autoref{fig:WMF_flowchart} shows the flowchart of the T-reX tool. The subsequent subsections describe the computational framework and the modules in more detail.
 
@@
     \includegraphics[height=0.8\textheight]{figures/WMF_flowchart.pdf}
-    \caption{Flowchart of the WasteAndMaterialFootprint (WMF) tool}\label{fig:WMF_flowchart}
+    \caption{Flowchart of the T-reX tool}\label{fig:WMF_flowchart}
 \end{figure}
@@
 Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module.
-This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `WasteAndMaterialFootprint'.
+This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `T-reX'.
 

### 2024-03-05 latex_files/discussion.tex
update-methods

--- 2024-01-30 latex_files/discussion.tex
@@
-\begin{itemize}
-    \item Summary of the results, 1 par
-    \item -----------------------------
-    \item Advantages of the WMF tool, 1 par
-    \item -----------------------------
-    \item Limitations, 1 par
-    \item -----------------------------
-    \item Future possibilities, 1 par
-\end{itemize}
+Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
+While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
+
+The T-reX tool, in contrast, gives LCA practitioners a simple, flexible, and transparent way to calculate the supply chain waste and material footprints, delivering results in standard units as direct aggregations of the relevant demand inventories.
+Once the databases have been processed with the T-reX tool, the user can then easily apply the T-reX methods to calculate the supply-chain waste and material footprints for any activity in the same way that they would calculate any LCIA indicator.
+
+The simple case study presented in this paper demonstrates both the utility and the limitations of the T-reX tool.
+It was shown the T-reX tool was able to calculate categorised aggregates and contribution analyses of both the waste generated and material demands in the present and future supply chains of the Li-ion batteries under consideration.
+Further, integration of the T-reX methods as `pseudo-LCIA' methods allows the user to easily make use of the T-reX tool in their preferred LCA software, be it code-based like \texttt{brightway2} or graphical like \texttt{ActivityBrowser}.
+
+One main limitation of the T-reX tool is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated, which would be needed to thoroughly assess the environmental impacts of this waste.
+Currently, the user would need to manually explore the `waste inventory' produced by the application of the T-reX to determine if, for example, the waste generated represents an actual loss of resources, or is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+A methodic classification of waste exchanges and the end-of-life fates will be facilitated by the more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
+
+Furthermore, the utility of the T-reX tool in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is little adaption of the future background inventories.
+In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.

### 2024-03-05 latex_files/introduction.tex
update-methods

@@
-The development of a `circular economy' has become a critical area of focus in the imperative pursuit of achieving sustainability objectives and curtailing our environmental footprint within planetary boundaries \citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling) \citep{eu2022ecodesign, eu2022repair,eu2015reman}.
-In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions, wrought as an act of competition or outright hostility \citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+The development of a `circular economy' has become a critical area of focus in the imperative pursuit of achieving sustainability objectives and curtailing our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{eu2022ecodesign, eu2022repair,eu2015reman}.
+In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions, wrought as an act of competition or outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
-While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use-- and end-of-life-- (EOL) phases, a significant proportion of these are often `hidden' in the supply chain and thus not reported directly in the final results \citep{laurenti2016wastefootprint,salviulo2021supplychain}.
-It has been found that these material footprints can be `highly representative of damage to human health and biodiversity' \citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage \citep{laurenti2023wastefootprint}.
+While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use-- and end-of-life-- (EOL) phases, a significant proportion of these are often `hidden' in the supply chain and thus not reported directly in the final results~\citep{laurenti2016wastefootprint,salviulo2021supplychain}.
+It has been found that these material footprints can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage~\citep{laurenti2023wastefootprint}.
 Thus, to reduce the negative externalities of consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible.
@@
 LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'---, often identifying critical hotspots and guiding prioritisation of actions.
-The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe \citep{huijbregts2016recipe} and CML \citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows.
+The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows.
 These scores are then aggregated into a single score for each impact category, which can be compared across products and processes.
 
-Several LCIA methods include, to some extent, waste generation \citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption \citep{arvidsson2020csi,foen2021ecofactors}.
+Several LCIA methods include, to some extent, waste generation~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption~\citep{arvidsson2020csi,foen2021ecofactors}.
 These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP)).
 
-In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs) \citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
+In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
 While LCA seeks to model the technosphere (a.k.a.
@@
 
-Though often described simply as a `material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.
+Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.
 Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows.
 Thus, precise and detailed categorisation of these is essential to understand the `circularity' of an activity and its life cycle externalities.
-There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage \citep{laurenti2023wastefootprint}.
-Conventional LCAs consider waste as a `service' \citep{guinee2021wasteisnotaservice} and typically use generic waste processing models \citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
+There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage~\citep{laurenti2023wastefootprint}.
+Conventional LCAs consider waste as a `service'~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
 
@@
 
-While methods with similar aims exist, they lack customisability and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.
+While methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
 

### 2024-03-05 latex_files/main.tex
update-methods

@@
 \label{sec:introduction}
-\cbox{Section word count: 1650}
-\cbox{Needs work to make it flow better, also shortened}
+\cbox{Section word count: 1450}
+\cbox{To be split into two (sub) sections: Introduction and Background, can be easily condensed by 200 words or so.}
 \input{introduction.tex}
@@
 \cbox{Section word count: 250}
-We have written the T-reX tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
-It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom WMF methods.
+We have written the T-reX tool, an extension to the brightway LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
+It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom `pseudo' LCIA methods.
 These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
-The WMF tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.
+The T-reX tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.
 

### 2024-03-05 latex_files/methodology.tex
update-methods

@@
 This section is divided into two parts.
-In \autoref{sec:method-WMF_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
+In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
 
 \subsection{The T-reX tool}
-\label{sec:method-WMF_tool}
+\label{sec:method-T-reX_tool}
 
-The WMF tool is a Python package that allows one to calculate the waste and material footprint of any product or service inside of LCA databases.
-Metadata of the WMF tool is presented in \autoref{tab:methods-wmfmetadata}.
+The T-reX tool is a Python package that allows one to calculate the waste and material inventory footprints of any product or service in LCA databases.
+Metadata of the T-reX tool is presented in the supplementary information [ADD LINK]
 
-\begin{table}[h]
-    \caption{T-reX tool metadata}
-    \label{tab:methods-wmfmetadata}
-    \centering
-    \begin{tabular}{ll}
-    \toprule
-    \textbf{Item} & \textbf{Details} \\
-    \midrule
-    Current version & 0.1.21 \\
-    DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
-    Code repository & \url{github.com/Stew-McD/T-reX} \\
-    License & CC0--1.0 license \\
-    Versioning system & git \\
-    Language & Python \\
-    Documentation & \url{T-reX.readthedocs.io} \\
-    Main dependencies & brightway2, premise, wurst \\
-    Support email & s.c.mcdowall@cml.leidenuniv.nl \\
-    \bottomrule
-    \end{tabular}
-\end{table}
-
-The tool is built on the \texttt{Brightway2} LCA framework~\citep{mutel2017brightway} and is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser} an open-source graphical user interface for LCA.
-The WMF tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023wmfpipy} and is open source under the CC-0 licence.
-The full source code for the WMF tool is indexed on Zenodo~\citep{mcdowall2023wmfzenodo} and under further development in the GitHub repository~\citep{mcdowall2024wmfgithub}.
+Developed in the Python programming language (version 3), the T-reX tool extends the brightway2 LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
+Additionally, the \texttt{wurst} package is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
+Integration with \texttt{premise} package~\citep{sacchi2022premise} enables the user to easily create and manipulate prospective LCA databases.
+The T-reX tool is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser} an open-source graphical user interface for LCA.
+The T-reX tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
+The full source code for the T-reX tool is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
 The tool is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases as well by changing the search criteria.
-Currently, it has been tested with all available system models of ecoinvent 3.5--3.10.
+Currently, it has been tested with all available system models of ecoinvent 3.5--3.10~\citep{ecoinvent2016version3}.
 
@@
 
-The following lists outline the constituent modules of the WMF tool, with a brief description of their functions.
-More extensive details can be found in the user guide and documentation of the program~\citep{mcdowall2023wmfdocs}.
+The supplementary information [ADD LINK] contains a list of the constituent modules of the T-reX tool, with a brief description of their functions.
+More extensive details can be found in the user guide and documentation of the program~\citep{mcdowall2023T-reXdocs}.
 
-\subsubsection{Functional modules}
-\begin{itemize}
-    \item \texttt{future\_scenarios}: Creates prospective LCA databases based on future scenarios.
-    \item \texttt{explode\_database}: Responsible for expanding a Brightway2 database into detailed exchange lists.
-    \item \texttt{search\_waste}: Provides functions for searching and categorising waste generation-related exchange data.
-    \item \texttt{search\_material}: Provides functions for searching and categorising material demand-related exchange data.
-    \item \texttt{make\_custom\_database}: Facilitates the creation of custom databases based on the waste and material search categories.
-    \item \texttt{method\_editor}: Manages the custom LCIA methods for waste and material footprint calculations.
-    \item \texttt{exchange\_editor}: Appends `pseudo-biosphere' exchanges to activities to match their waste generation and material demand exchanges in the technosphere.
-    \item \texttt{verify\_database}: Performs verification of the manipulated databases.
-\end{itemize}
-
-\subsubsection{Configuration modules}
-\begin{itemize}
-    \item \texttt{custom\_config}: Provides functions for managing the configuration of the T-reX package.
-    \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
-    \item \texttt{queries\_waste}: Defines search parameters and categories for waste generation exchanges (user editable).
-    \item \texttt{queries\_materials}: Defines search parameters and categories for material demand exchanges (user editable).
-\end{itemize}
-
-\autoref{fig:WMF_flowchart} shows the flowchart of the T-reX tool. The subsequent subsections describe the computational framework and the modules in more detail.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=0.8\textheight]{figures/WMF_flowchart.pdf}
-    \caption{Flowchart of the T-reX tool}\label{fig:WMF_flowchart}
-\end{figure}
-
-\subsubsection{Computational framework}
-Developed in the Python programming language (version 3), the WMF tool extends the brightway2 LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
-Additionally, the \texttt{wurst} package is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
-Integration with \texttt{premise} package~\citep{sacchi2022premise} enables the user to easily create and manipulate prospective LCA databases.
+% \autoref{fig:T-reX_flowchart} shows the flowchart of the T-reX tool. The subsequent subsections describe the computational framework and the modules in more detail.
 
@@
 The \texttt{search\_waste} and \texttt{search\_material} modules apply user-defined search parameters from \texttt{queries\_waste} and \texttt{queries\_materials} to identify relevant waste and material flows in the list of technosphere exchanges generated by \texttt{explode\_database} and categorises them accordingly.
-The results of the search functions are stored in \texttt{.csv} files for subsequent use in the WMF tool's workflow.
+The results of the search functions are stored in \texttt{.csv} files for subsequent use in the T-reX tool's workflow.
 
-\subsubsection{Waste exchanges}\label{sec:method-wmf-waste_exchanges}
+\subsubsection{Waste exchanges}\label{sec:method-T-reX-waste_exchanges}
 
@@
 
-\subsubsection{Material exchanges}\label{sec:method-wmf-material_exchanges}
+\subsubsection{Material exchanges}\label{sec:method-T-reX-material_exchanges}
 In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
@@
 
-The logic for the identification of material exchanges with the WMF tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
+The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
 That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
-A useful feature of the WMF tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
+A useful feature of the T-reX tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
 For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
@@
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
-In the WMF tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
+In the T-reX tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
 This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the WMF tool can provide insight into the broader supply-chain impacts of the demand for this metal.
-If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the WMF results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply-chain impacts of the demand for this metal.
+If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
 In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
@@
 \subsubsection{LCIA method management}
-The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the WMF tool.
+The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the T-reX tool.
 This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange.
@@
 The \texttt{exchange\_editor} module loads the \texttt{.csv} files created by the search functions and appends `pseudo-biosphere' exchanges to the matching activities in the LCA database.
-This is the most computationally intensive part of the WMF tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
+This is the most computationally intensive part of the T-reX tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
 
 \subsubsection{Database Verification}
-The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the WMF tool has processed the database correctly.
+The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the T-reX tool has processed the database correctly.
 
@@
 \subsubsection{Methods}
-In addition to the Waste Footprint and Material Demand footprint methods created by the WMF tool, the following standard LCIA methods were applied for comparison:
+In addition to the Waste Footprint and Material Demand footprint methods created by the T-reX tool, the following standard LCIA methods were applied for comparison:
 \begin{itemize}[itemsep=0pt]
@@
 The primary source of life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff.
-Additionally, the WMF tool was used to create prospective database sets using the \mbox{REMIND} model with the following Representative Concentration Pathways (RCPs):
+Additionally, the T-reX tool was used to create prospective database sets using the \mbox{REMIND} model with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}
@@
 
-For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with the WMF tool over the time series: 2020, 2040, 2060, 2080, 2100.
+For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with the T-reX tool over the time series: 2020, 2040, 2060, 2080, 2100.
 

### 2024-03-05 latex_files/results.tex
update abstract and introduction
update-methods

@@
-\subsection{WasteAndMaterialFootprint tool}\label{sec:results-wmf}
+\subsection{T-reX tool}\label{sec:results-T-reX}
 
-An example of the output from the application of the WMF tool has been included in the supplementary material. The manipulated ecoinvent databases 
-can be recreated using the code and instructions available in the package documentation~\citep{mcdowall2023wmfdocs}.
+An example of the output from the application of the T-reX tool has been included in the supplementary material. The manipulated ecoinvent databases 
+can be recreated using the code and instructions available in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (as well as a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at market.
-The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the WMF tool.
+The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the T-reX tool.
 This section includes some highlights of the results and the full results are available in the supplementary material.
-Because the WMF methods are integrated into the brightway project as if they were LCIA methods, the results can be visualised in the same way.
+Because the T-reX methods are integrated into the brightway project as if they were LCIA methods, the results can be visualised in the same way.
 In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
@@
 
-A comparison of the results from the WMF tool's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
+A comparison of the results from the T-reX tool's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
 In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
 Such comparability was also observed for other fossil-fuel-related methods (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
-Correlation between standard LCIA methods and WMF methods, is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
+Correlation between standard LCIA methods and T-reX methods, is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
 In standard LCIA methods, impact scores are derived exclusively from the magnitude of the exchanges between the biosphere and the technosphere, that is, extraction and emission.
-In the WMF methods, the footprint scores are based on an accounting of either waste generated or material demand, both of which are technosphere-technosphere exchanges in terms of LCA modelling.
+In the T-reX methods, the footprint scores are based on an accounting of either waste generated or material demand, both of which are technosphere-technosphere exchanges in terms of LCA modelling.
 For the material demand methods especially, this distinction is critical.
-For example, application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the WMF method provides an aggregation of the exchanges with the market for that metal.
-The WMF method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply-chain pressures that are not captured by the standard LCIA methods.
-This makes the WMF methods more sensitive to the modeling choices (e.g., allocation) that are generally embedded in LCA databases.
+For example, application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the T-reX method provides an aggregation of the exchanges with the market for that metal.
+The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply-chain pressures that are not captured by the standard LCIA methods.
+This makes the T-reX methods more sensitive to the modeling choices (e.g., allocation) that are generally embedded in LCA databases.
 
@@
     \end{subfigure}
-    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `WMF - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:comparison_methods}
+    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `T-reX - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:comparison_methods}
 \end{figure}
@@
 For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste, Laurenti et al.
-reported 95\% of the total waste, whereas the WMF tool reported only 3\%.
+reported 95\% of the total waste, whereas the T-reX tool reported only 3\%.
 The reason for this discrepancy is explained by the fact that in the method of Laurenti et al., a ``waste flow was regarded as hazardous when the hazardousness
@@
 The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
-In the WMF method, the source database is exploded into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+In the T-reX method, the source database is exploded into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
 
@@
 \centering
-\caption{Comparison of the results from the WMF battery case study (database: ecoinvent 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: ecoinvent 3.5, activity:`Li-ion').}
+\caption{Comparison of the results from the T-reX battery case study (database: ecoinvent 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: ecoinvent 3.5, activity:`Li-ion').}
 \label{tab:results-case_study-comparison}
@@
 \toprule
-\textbf{Indicator} & \textbf{WMF} & \textbf{Laurenti et al.} \\
+\textbf{Indicator} & \textbf{T-reX} & \textbf{Laurenti et al.} \\
 \midrule

### 2024-03-05 latex_files/supplementary_material.tex
update-manuscript
update-methods

@@
 \documentclass{article}
-\usepackage{booktabs} % For better quality table rules
-\usepackage{caption} % For better caption format
+\usepackage{booktabs}
+% ---------------------------------------------------------------------
+%             FOR DRAFTING / REVIEW (DELETE BEFORE SUBMISSION)
+% ---------------------------------------------------------------------
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage[british]{babel}
+\usepackage{microtype}
+\usepackage[tt=false]{libertine}
+\usepackage{xcolor}
+\definecolor{blue}{RGB}{33,150,209}
+\definecolor{red}{RGB}{255,0,0}
+\usepackage{xurl}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{lineno}
+\usepackage{float}
+\setlength{\linenumbersep}{5pt}
+% Save the old \texttt command
+\let\oldtexttt\texttt
+% Redefine the \UrlFont to use a smaller typewriter font
+\renewcommand{\UrlFont}{\ttfamily\small}
+% Redefine the \texttt command to use a smaller font size and allow line breaks
+\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
+
+\usepackage{longtable}
+% --------------------------------------------------------------------- %
+
+\usepackage{tcolorbox}
+\newcommand{\cbox}[1]{
+    \begin{tcolorbox}[hbox, colback=red!5!white, colframe=red!65!black, boxrule=0.25pt, boxsep=2pt, left=2pt, right=2pt, top=1pt, bottom=1pt]
+        \small\sffamily #1
+    \end{tcolorbox}
+}
+
+\usepackage{graphicx}
+\usepackage[margins=1cm]{geometry}
 
@@
 
+\title{Supplementary Material for the T-reX Tool}
+\author{Stewart Charles McDowall, Elizabeth Lanphear, Carlos Felipe Rocha Blanco, Stefano Cucurachi}
+\date{\today}
+
+\maketitle
+
+\section{Metadata of the T-reX tool}
+
+\begin{table}[h]
+    \caption{T-reX tool metadata}\label{tab:metadata}
+    \centering
+    \begin{tabular}{ll}
+    \toprule
+    \textbf{Item} & \textbf{Details} \\
+    \midrule
+    Current version & 0.1.21 \\
+    DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
+    Code repository & \url{github.com/Stew-McD/T-reX} \\
+    License & CC0--1.0 license \\
+    Versioning system & git \\
+    Language & Python \\
+    Documentation & \url{T-reX.readthedocs.io} \\
+    Main dependencies & brightway2, premise, wurst \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{List of modules in the T-reX python package}
+
+\subsection{Functional modules}
+\begin{itemize}
+    \item \texttt{future\_scenarios}: Creates prospective LCA databases based on future scenarios.
+    \item \texttt{explode\_database}: Responsible for expanding a Brightway2 database into detailed exchange lists.
+    \item \texttt{search\_waste}: Provides functions for searching and categorising waste generation-related exchange data.
+    \item \texttt{search\_material}: Provides functions for searching and categorising material demand-related exchange data.
+    \item \texttt{make\_custom\_database}: Facilitates the creation of custom databases based on the waste and material search categories.
+    \item \texttt{method\_editor}: Manages the custom LCIA methods for waste and material footprint calculations.
+    \item \texttt{exchange\_editor}: Appends `pseudo-biosphere' exchanges to activities to match their waste generation and material demand exchanges in the technosphere.
+    \item \texttt{verify\_database}: Performs verification of the manipulated databases.
+\end{itemize}
+
+\subsection{Configuration modules}
+\begin{itemize}
+    \item \texttt{custom\_config}: Provides functions for managing the configuration of the T-reX package.
+    \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
+    \item \texttt{queries\_waste}: Defines search parameters and categories for waste generation exchanges (user editable).
+    \item \texttt{queries\_materials}: Defines search parameters and categories for material demand exchanges (user editable).
+\end{itemize}
+
+\section{Flowchart of the T-reX tool}
+
+% \autoref{fig:flowchart} shows the computation flowchart of the T-reX tool.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=0.8\textheight]{figures/T-reX_flowchart.pdf}
+    \caption{Computational flowchart of the T-reX tool}\label{fig:flowchart}
+\end{figure}
+
 \section{List of Materials and Their Markets}
 
-\begin{table}[ht]
-\centering
-\caption{A comprehensive list of various materials in the default configuration.}
-\label{tab:materials}
-
-\begin{tabular}{ll}
+\begin{longtable}{ll}
+    \caption{A comprehensive list of various materials in the default configuration.} \label{tab:materials} \\
     \toprule
@@
     \midrule
+    \endfirsthead
+    \multicolumn{2}{c}%
+    {\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
+    \toprule
+    \textbf{Market Name} & \textbf{Material group} \\
+    \midrule
+    \endhead
+    \midrule \multicolumn{2}{r}{\textit{Continued on next page}} \\
+    \endfoot
+    \bottomrule
+    \endlastfoot
+
 market for aluminium & aluminium \\
@@
 \bottomrule
-\end{tabular}
-\end{table}
+\end{longtable}
 

### 2024-03-06 latex_files/discussion.tex
revise methodology

--- 2024-01-30 latex_files/discussion.tex
@@
-\begin{itemize}
-    \item Summary of the results, 1 par
-    \item -----------------------------
-    \item Advantages of the WMF tool, 1 par
-    \item -----------------------------
-    \item Limitations, 1 par
-    \item -----------------------------
-    \item Future possibilities, 1 par
-\end{itemize}
+Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
+While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
+
+Our contribution advances the state-of-the-art by giving LCA practitioners a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
+Once the databases have been processed with the T-reX tool, the user can  easily apply the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
+
+The simple case study of six Li-ion batteries presented in this paper demonstrated the the utility, flexibility and limitations of the T-reX tool.
+
+First, by adjusting the user configuration for future scenarios and waste/material categorisation we easily produced a set of customised versions of the ecoinvent database.
+Then, by applying the T-reX `pseudo-LCIA' methods, it was trivial to calculate categorised waste and material inventory footprints for a number of present and future supply chains.
+Additionally, visual exploration in \texttt{ActivityBrowser} was possible because the T-reX `pseudo-LCIA' methods are integrated in the \texttt{brightway} project in the same way as standard LCIA methods.
+
+One limitation of the T-reX tool is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated.
+This is the information would be needed to thoroughly assess the potential environmental impacts of this waste.
+Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if, for example, the waste generated represents an actual loss of resources, or is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+A methodic classification of waste exchanges and the end-of-life fates will be facilitated by the more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
+
+The utility of the T-reX tool in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is little adaption of the future background inventories.
+In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.

### 2024-03-06 latex_files/introduction.tex
revise methodology

@@
 It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
-In any case, avoiding material consumption and generation of waste is of critical importance.
+By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
+In any case, avoiding material consumption and generation of waste is of paramount importance.
 By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and opportunities for waste reduction and material efficiency.

### 2024-03-06 latex_files/main.tex
revise methodology

@@
 
-    \title{T-reX: A python package to quantify supply chain flows of waste and material in LCA databases}
+    \title{T-reX: A python package to quantify waste and material footprints in current and future LCA databases}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
     \begin{abstract}
-        \cbox{Abstract word count: 164, Limit: 150}
-        Minimising waste through the reuse of resources is the quintessential principle of the `circular economy', but relies on our ability to identify and quantify waste and material flows.
+        \cbox{Abstract word count: 158, Limit: 150}
+        Minimising material consumption and waste generation is the quintessential principle of the `circular economy'.
         Thus, identifying and quantifying waste and material flows are of fundamental importance.
@@
 
-        Introducing T-reX, a Python tool extending the Brightway framework for flexible quantification of user-defined supply-chain demands in current and future scenarios.
-This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
+        Introducing T-reX, a Python tool extending the `Brightway' LCA ecosystem, allowing facile quantification of user-defined supply-chain demands in current and prospective scenarios.
+T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
 
-        A case study on battery supply chains demonstrates the tool's utility.
-T-reX quantifies and compares inventory demands and, thus, potential environmental burdens, aiding sustainable decision-making.
-It contributes to the development of the `circular economy' by providing detailed material usage and waste generation analysis.
+        A case study of 6 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
 
@@
 \label{sec:methodology}
-\cbox{Section word count: 2000}
+\cbox{Section word count: 800}
 \input{methodology.tex}

### 2024-03-06 latex_files/methodology.tex
revise methodology

@@
 This section is divided into two parts.
-In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.
+In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of 6 Li-ion batteries.
 
-\subsection{The T-reX tool}
+\subsection{The T-reX tool}\label{sec:method-T-reX_tool}
+
+\subsubsection{Computational framework}
 \label{sec:method-T-reX_tool}
 
-The T-reX tool is a Python package that allows one to calculate the waste and material inventory footprints of any product or service in LCA databases.
-Metadata of the T-reX tool is presented in the supplementary information [ADD LINK]
-
-Developed in the Python programming language (version 3), the T-reX tool extends the brightway2 LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
-Additionally, the \texttt{wurst} package is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
-Integration with \texttt{premise} package~\citep{sacchi2022premise} enables the user to easily create and manipulate prospective LCA databases.
-The T-reX tool is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser} an open-source graphical user interface for LCA.
+Developed in the Python programming language, T-reX extends the brightway LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
+Additionally, the \texttt{wurst} package---which can expand entire databases into a list of exchanges---is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
+Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
+T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the 'pseudo' LCIA methods created by T-reX can be used in the accustomed way.
 The T-reX tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
 The full source code for the T-reX tool is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
-The tool is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases as well by changing the search criteria.
+T-reX is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
 Currently, it has been tested with all available system models of ecoinvent 3.5--3.10~\citep{ecoinvent2016version3}.
 
-The program can be used directly from the command line, or imported as a Python module, in which case, the user can access the individual functions and modules.
+T-reX can be used directly from the command line, or imported as a Python module, in which case, the user can access the individual functions and modules.
 In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the ecoinvent database.
-The user can also customise the program to calculate the waste and material footprint of a custom database, or a prospective database based on future scenarios.
-The program is designed to be modular so that the user can easily customise the program to their needs.
+The user can also customise T-reX to calculate the waste and material inventory footprints of a custom database, or a prospective database based on future scenarios by implementing \texttt{premise} with an included module.
 
-The supplementary information [ADD LINK] contains a list of the constituent modules of the T-reX tool, with a brief description of their functions.
-More extensive details can be found in the user guide and documentation of the program~\citep{mcdowall2023T-reXdocs}.
+The supplementary information [ADD LINK] contains the full metadata of T-reX tool, along with a list of the constituent modules of the T-reX tool, a description of their functions and a detailed computational workflow.
+Further details can be found in the user guide and documentation~\citep{mcdowall2023T-reXdocs}.
 
-% \autoref{fig:T-reX_flowchart} shows the flowchart of the T-reX tool. The subsequent subsections describe the computational framework and the modules in more detail.
+\subsubsection{Functionality and purpose}
 
-\subsubsection{Generation of prospective LCA databases}
-Future waste and material footprints can be projected using the \texttt{future\_scenarios} module, which uses \texttt{premise} to generate prospective scenario databases based on the configuration in \texttt{user\_settings}.
-These prospective databases can be custom-defined by the user or can be constructed with the future projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg}  that can be paired with a variety of mitigation scenarios.
+T-reX is a Python package that allows one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same was as standard LCIA methods.
 
-\subsubsection{Database expansion}
-The \texttt{explode\_database} module uses \texttt{wurst} to deconstruct LCA databases into a list of individual exchanges representing all of material and energy flows in the technosphere model.
-This dataset being converted into a pandas DataFrame and stored as a binary \texttt{.pickle} file for subsequent analysis.
+If desired, prospective databases can be custom-defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} that can be paired with a variety of mitigation scenarios.
 
-\subsubsection{Waste and material flow identification and categorisation}
-
-The \texttt{search\_waste} and \texttt{search\_material} modules apply user-defined search parameters from \texttt{queries\_waste} and \texttt{queries\_materials} to identify relevant waste and material flows in the list of technosphere exchanges generated by \texttt{explode\_database} and categorises them accordingly.
-The results of the search functions are stored in \texttt{.csv} files for subsequent use in the T-reX tool's workflow.
-
-\subsubsection{Waste exchanges}\label{sec:method-T-reX-waste_exchanges}
+Subsequent expansion of the databases into lists of exchanges allows relevent material and waste flows to be identified and categorised.
+The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
+The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand.
+The exchange editing function of T-reX then takes each list of exchanges and appends to the relevent activity a copy of the technosphere exchange as `pseudo-biosphere' exchange that matches the `pseudo LCIA' method.
 
 In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories are:
-\begin{itemize}[itemsep=0pt]
-    \item digestion
-    \item composting
-    \item open burning
-    \item incineration
-    \item recycling
-    \item landfill
-    \item hazardous
-    \item non-hazardous
-    \item carbon dioxide 
-    \item total
-\end{itemize}
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary information [ADD LINK].
+One advantage of T-reX is that is is able to identify of waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
+Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
 
-The logic of screening for waste exchanges is based on a set of boolean search queries (`AND', `OR', and `NOT') that are applied in a list comprehension to the names of every exchange in the LCA database (see `search\_queries.py' for the full list).
-In this way, the search queries enable classification into categories (such as `hazardous solid' and `incineration liquid') and permit the identification of waste exchanges in addition to those directly connected to waste treatment processes.
-The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
-In the default settings, there are a total of 18 waste classifications (9 categories, each separated into liquid and solid waste) For example, the identification of `non-hazardous solid' waste exchanges is based on the following search query; \texttt{AND = [`waste'], NOT = [`hazardous', `radioactive'], UNIT = [`kilogram']} (this can also be inferred and confirmed by comparison with the difference between the results of `total solid' and `hazardous solid').\ \autoref{tab:wf_results} presents a list of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
-Note that the carbon dioxide waste category does not include emissions to the atmosphere.
-This category is based solely on the accounting of carbon capture and storage (CCS), which is included in many prospective databases as direct sequestration in reservoirs as well as solvent capture.
-
-\begin{table}[ht]
-    \centering
-    \caption{WasteFootprint search results for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
-    \begin{tabular}{llr}
-        \toprule
-        \textbf{Waste exchanges} & \textbf{Unit} & \textbf{Exchange count} \\
-        \midrule
-        digestion & kilogram & 4 \\
-        composting & kilogram & 26 \\
-        open burning & kilogram & 535 \\
-        incineration & kilogram & 2171 \\
-        recycling & kilogram & 137 \\
-        landfill & kilogram & 1530 \\
-        hazardous & kilogram & 1928 \\
-        carbon dioxide & kilogram & 119 \\
-        total & kilogram & 29524 \\
-        digestion & cubic meter & 16 \\
-        composting & cubic meter & 0 \\
-        open burning & cubic meter & 0 \\
-        incineration & cubic meter & 2 \\
-        recycling & cubic meter & 0 \\
-        landfill & cubic meter & 2 \\
-        hazardous & cubic meter & 437 \\
-        carbon dioxide & cubic meter & 0 \\
-        total & cubic meter & 4360 \\
-        \bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Material exchanges}\label{sec:method-T-reX-material_exchanges}
 In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
 The CRM list is a list of 30 materials that are considered critical to the EU economy and are at risk of supply disruption.
-Further materials of interest were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
+Further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
 The identity of the materials considered and their categorical groupings are easily customisable by the user.
-A full list of 59 materials included in the default configuration is provided in the supplementary material.
+The full list of 59 materials included in the default configuration is provided in the supplementary material [ADD LINK].
 
 The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
-That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
 A useful feature of the T-reX tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
@@
 
-\subsubsection{Creation of custom `pseudo-biosphere' databases}
-Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module.
-This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `T-reX'.
+% \subsubsection{Theoretical basis}
 
-\subsubsection{LCIA method management}
-The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the T-reX tool.
-This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange.
-The methods are stored in the Brightway2 project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods.
-Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from waste consumed by treatment to waste generated by the activity.
+% XXX
 
-\subsection{Exchange editing}
-The \texttt{exchange\_editor} module loads the \texttt{.csv} files created by the search functions and appends `pseudo-biosphere' exchanges to the matching activities in the LCA database.
-This is the most computationally intensive part of the T-reX tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
+\subsubsection{T-reX tool workflow}
 
-\subsubsection{Database Verification}
-The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the T-reX tool has processed the database correctly.
+The workflow of the T-reX tool is divided into several modules, each of which performs a specific function.
+The modules are designed to be used in a specific order, but the user can also use them individually to perform specific tasks.
+The standard workflow is as follows:
+
+\begin{enumerate}
+    \item Configuration of waste and material exchange categorisation (optional)
+    \item Generation of prospective LCA databases (optional)
+    \item Database expansion---to create a list of all exchanges in the database
+    \item Identification and categorisation of exchanges
+    \item Creation of `pseudo-biosphere' databases
+    \item Creation of `pseudo-LCIA' methods to calculate waste and material inventory footprints
+    \item Exchange editing---whereby the technosphere exchange is mirrored as a `pseudo-biosphere' exchange
+    \item Database verification
+\end{enumerate}
+
+This workflow creates one or more manipulated LCA databases that can then be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+
+A generic flowchart of the T-reX tool workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material [ADD LINK] contains a more detailed computational flowchart.
+
+\begin{figure}[ht]
+    \centering
+    \caption{Flowchart of the T-reX tool workflow.}
+    \includegraphics[width=\textwidth]{figures/T-reX_workflow.pdf}
+    \label{fig:methods-flowchart}
+\end{figure}
 
@@
 
-\subsubsection{Activities}
-This case study investigated five types of Li-ion batteries, each represented by specific market activities:
+We investigated five types of Li-ion batteries, each represented by specific market activities:
 \begin{itemize}[itemsep=0pt]
@@
 
-\subsubsection{Methods}
-In addition to the Waste Footprint and Material Demand footprint methods created by the T-reX tool, the following standard LCIA methods were applied for comparison:
+In addition to the waste and material inventory footprint methods created by the T-reX tool, the following standard LCIA methods were applied for comparison:
+
 \begin{itemize}[itemsep=0pt]
@@
 
-\subsubsection{Databases}
 The primary source of life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff.
@@
 \begin{itemize}
-    \item \textbf{SSP2 -base:}\\ representing ca. 3.5°C increase in global temperatures to 2100
-    \item \textbf{SSP2-PkBudg500:}\\ meeting Paris climate goals, ca. 1.3°C increase to 2100
+    \item SSP2-base: representing an approximate 3.5°C increase in global temperatures to 2100
+    \item SSP2-PkBudg500: representing the achievement of Paris climate goals, ca. 1.3°C increase to 2100
 \end{itemize}
@@
 Additionally, for the Waste and Material Footprint methods, a contribution analysis was performed.
-This involved utilizing the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\cite{mutel2016brightway2analyzer}, an additional component of the Brightway2 LCA framework.
+This involved utilizing the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, an additional component of the Brightway2 LCA framework.
 This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their CPC codes.

### 2024-03-06 latex_files/supplementary_material.tex
add pdf of review notes
revise methodology

@@
 \usepackage{booktabs}
-% ---------------------------------------------------------------------
-%             FOR DRAFTING / REVIEW (DELETE BEFORE SUBMISSION)
-% ---------------------------------------------------------------------
 \usepackage{caption}
@@
 \usepackage{xcolor}
-\definecolor{blue}{RGB}{33,150,209}
-\definecolor{red}{RGB}{255,0,0}
 \usepackage{xurl}
@@
 \usepackage{float}
+\usepackage{natbib}
+\usepackage{longtable}
+\usepackage{geometry}
+\usepackage{tabularx}
+% Define colors
+\definecolor{blue}{RGB}{33,150,209}
+\definecolor{red}{RGB}{255,0,0}
+
+% Set geometry for the page
+\geometry{a4paper, margin=1in}
+
+% Line numbers settings
 \setlength{\linenumbersep}{5pt}
-% Save the old \texttt command
+\linenumbers
+
+% Modify URL and texttt fonts
+\renewcommand{\UrlFont}{\ttfamily\small}
 \let\oldtexttt\texttt
-% Redefine the \UrlFont to use a smaller typewriter font
-\renewcommand{\UrlFont}{\ttfamily\small}
-% Redefine the \texttt command to use a smaller font size and allow line breaks
 \renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
 
-\usepackage{longtable}
-% --------------------------------------------------------------------- %
-
+% Custom box command for drafting purposes
 \usepackage{tcolorbox}
@@
 
-\usepackage{graphicx}
-\usepackage[margins=1cm]{geometry}
-
 \begin{document}
 
-\title{Supplementary Material for the T-reX Tool}
-\author{Stewart Charles McDowall, Elizabeth Lanphear, Carlos Felipe Rocha Blanco, Stefano Cucurachi}
+\title{\textbf{Supplementary Information for the T-reX Manuscript}}
+\author{Stewart Charles McDowall, Elizabeth Lanphere, Carlos Felipe Blanco, Stefano Cucurachi \\
+\textit{Institute of Environmental Sciences (CML), Leiden University, Leiden, The Netherlands}}
 \date{\today}
@@
 \maketitle
+\tableofcontents
+\listoffigures
+\listoftables
+\clearpage
 
@@
 
-\section{Flowchart of the T-reX tool}
+\section{Detailled description of the T-reX tool workflow}
 
-% \autoref{fig:flowchart} shows the computation flowchart of the T-reX tool.
+\subsection{Generation of prospective LCA databases}
 
-\begin{figure}
+Future waste and material footprints can be projected using the \texttt{future\_scenarios} module, which uses \texttt{premise} to generate prospective scenario databases based on the configuration in \texttt{user\_settings}.
+These prospective databases can be custom-defined by the user or can be constructed with the future projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} that can be paired with a variety of mitigation scenarios.
+
+\subsection{Database expansion}
+The \texttt{explode\_database} module uses \texttt{wurst} to deconstruct LCA databases into a list of individual exchanges representing all of material and energy flows in the technosphere model.
+This dataset being converted into a pandas DataFrame and stored as a binary \texttt{.pickle} file for subsequent analysis.
+
+\subsection{Waste and material flow identification and categorisation}
+
+The \texttt{search\_waste} and \texttt{search\_material} modules apply user-defined search parameters from \texttt{queries\_waste} and \texttt{queries\_materials} to identify relevant waste and material flows in the list of technosphere exchanges generated by \texttt{explode\_database} and categorises them accordingly.
+The results of the search functions are stored in \texttt{.csv} files for subsequent use in the T-reX tool's workflow.
+
+\subsubsection{Waste exchanges}\label{sec:method-T-reX-waste_exchanges}
+
+The logic of screening for waste exchanges is based on a set of boolean search queries (`AND', `OR', and `NOT') that are applied in a list comprehension to the names of every exchange in the LCA database (see `search\_queries.py' for the full list).
+In this way, the search queries enable classification into categories (such as `hazardous solid' and `incineration liquid') and permit the identification of waste exchanges in addition to those directly connected to waste treatment processes.
+The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
+In the default settings, there are a total of 18 waste classifications (9 categories, each separated into liquid and solid waste) For example, the identification of `non-hazardous solid' waste exchanges is based on the following search query; \texttt{AND = [`waste'], NOT = [`hazardous', `radioactive'], UNIT = [`kilogram']} (this can also be inferred and confirmed by comparison with the difference between the results of `total solid' and `hazardous solid').
+
+The default waste categories and their search logic are listed in \autoref{tab:waste_categories}.
+
+\begin{table}[ht]
     \centering
-    \includegraphics[height=0.8\textheight]{figures/T-reX_flowchart.pdf}
-    \caption{Computational flowchart of the T-reX tool}\label{fig:flowchart}
-\end{figure}
+    \caption{Waste categories in the default configuration of the T-reX tool.}\label{tab:waste_categories}
+    \begin{tabularx}{\textwidth}{p{2.5cm}p{1cm}p{1.5cm}XX}
+    \toprule
+    \textbf{Waste Category} & \textbf{AND} & \textbf{AND +} & \textbf{OR} & \textbf{NOT} \\
+    \midrule
+    digestion & waste & digestion & & \\
+    composting & waste & composting & & \\
+    open burning & waste & burning & & \\
+    incineration & waste & incineration & & \\
+    recycling & waste & recycling & & \\
+    landfill & waste & & landfill, dumped, deposit & \\
+    hazardous & waste & & hazardous, radioactive & non-hazardous, non-radioactive \\
+    non-hazardous & waste & & & hazardous, radioactive \\
+    carbon dioxide & waste & & carbon dioxide storage, carbon dioxide, captured & methane \\
+    total & waste & & & \\
+    \bottomrule
+    \end{tabularx}
+\end{table}
 
-\section{List of Materials and Their Markets}
+As an example, \autoref{tab:wf_results} presents a list of the number of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
+Note that the carbon dioxide waste category does not include emissions to the atmosphere, which is a typical focus of LCIA studies.
+The carbon dioxide waste category is based solely on the accounting of carbon capture and storage (CCS), which is included in many prospective databases as direct sequestration in reservoirs as well as solvent capture.
+
+\begin{table}[ht]
+    \centering
+    \caption{T-reX waste search result counts for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
+    \begin{tabular}{llr}
+        \toprule
+        \textbf{Waste exchanges} & \textbf{Unit} & \textbf{Exchange count} \\
+        \midrule
+        digestion & kilogram & 4 \\
+        composting & kilogram & 26 \\
+        open burning & kilogram & 535 \\
+        incineration & kilogram & 2171 \\
+        recycling & kilogram & 137 \\
+        landfill & kilogram & 1530 \\
+        hazardous & kilogram & 1928 \\
+        carbon dioxide & kilogram & 119 \\
+        total & kilogram & 29524 \\
+        digestion & cubic meter & 16 \\
+        composting & cubic meter & 0 \\
+        open burning & cubic meter & 0 \\
+        incineration & cubic meter & 2 \\
+        recycling & cubic meter & 0 \\
+        landfill & cubic meter & 2 \\
+        hazardous & cubic meter & 437 \\
+        carbon dioxide & cubic meter & 0 \\
+        total & cubic meter & 4360 \\
+        \bottomrule
+\end{tabular}
+\end{table}
+
+\subsubsection{Material exchanges}\label{sec:method-T-reX-material_exchanges}
+In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
+The CRM list is a list of 30 materials that are considered critical to the EU economy and are at risk of supply disruption.
+Further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
+The identity of the materials considered and their categorical groupings are easily customisable by the user.
+A full list of 59 materials included in the default configuration is provided in \autoref{tab:list_materials}.
+
+The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
+That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
+A useful feature of the T-reX tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
+For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
+Similarly, the total demand for all critical raw materials (CRMs) can be easily calculated in the same manner.
+
+As discussed in the introduction of the paper to which this material is attached, there are some existing material demand methods in the standard LCIA method sets, including the `crustal scarcity indicator' (which provides only an aggregated, abstracted endpoint)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
+In the T-reX tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
+This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply-chain impacts of the demand for this metal.
+If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
+In the results of the Li-ion battery case study in the paper to which this material is attached, it was, indeed, the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
 
 \begin{longtable}{ll}
-    \caption{A comprehensive list of various materials in the default configuration.} \label{tab:materials} \\
+    \caption{A comprehensive list of various materials and their groupings in the default configuration.} \label{tab:list_materials} \\
     \toprule
@@
 
+\subsection{Creation of custom `pseudo-biosphere' databases}
+Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module.
+This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `T-reX'.
+
+\subsection{LCIA method management}
+The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the T-reX tool.
+This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange.
+The methods are stored in the Brightway2 project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods.
+Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from waste consumed by treatment to waste generated by the activity.
+
+\subsection{Exchange editing}
+The \texttt{exchange\_editor} module loads the \texttt{.csv} files created by the search functions and appends `pseudo-biosphere' exchanges to the matching activities in the LCA database.
+This is the most computationally intensive part of the T-reX tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
+
+\subsection{Database Verification}
+The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the T-reX tool has processed the database correctly.
+
+\section{Modular flowchart of the T-reX tool}
+
+\autoref{fig:flowchart} shows the logical flow through the individual modules of T-reX tool.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[height=0.8\textheight]{figures/T-reX_flowchart.pdf}
+    \caption{Computational flowchart of the T-reX tool.
+In green are the individual modules, in blue are their outputs.
+The final product is one or more manipulated databases with which one can calculate the material and waste inventory footprints using the `pseudo-LCIA' methods created by T-reX based on the search configuration.}\label{fig:flowchart}
+\end{figure}
+
+\clearpage
+
+\section{Example terminal output of the T-reX tool}
+
+Due to its length, an example of the terminal output of the T-reX tool is included as a separate file in the supplementary material, named `T-reX\_example\_output.html', which can be opened in an internet browser.
+The example output is from the execution of the T-reX tool on the `ecoinvent 3.9.1' database with the IAM model `REMIND' and the RCP `PkBudg500' in the years 2065 and 2100.
+The example output includes the execution of the main module which in turn executes the \texttt{future\_scenarios}, \texttt{explode\_database}, \texttt{search\_waste}, \texttt{search\_material}, \texttt{make\_custom\_database}, \texttt{method\_editor}, \texttt{exchange\_editor}, and \texttt{verify\_database} modules.
+
+Due to licensing restrictions, the manipulated ecoinvent databases produced by the T-reX tool are not included in the supplementary material.
+However, with a licence to the ecoinvent database, the user can easily reproduce the results of the T-reX tool by following the instructions in the documentation.
+
+\clearpage
+
+\bibliographystyle{elsarticle-harv}
+\bibliography{references.bib}
+
 \end{document}

### 2024-03-06 latex_files/main.tex
add sankeys, etc.

@@
 %% \usepackage{amsthm}
-\usepackage{lineno}
 \usepackage{booktabs}
@@
 
-\linenumbers
+% \linenumbers
 
@@
     \begin{abstract}
-        \cbox{Abstract word count: 158, Limit: 150}
+        \cbox{Abstract word count: 151, Limit: 150}
         Minimising material consumption and waste generation is the quintessential principle of the `circular economy'.
         Thus, identifying and quantifying waste and material flows are of fundamental importance.
-Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of products and services, those where the implementation of circular principles could be most effective.
+Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of activities, where the implementation of circular principles could be most effective.
 
-        Introducing T-reX, a Python tool extending the `Brightway' LCA ecosystem, allowing facile quantification of user-defined supply-chain demands in current and prospective scenarios.
+        Introducing T-reX, a Python tool extending the brightway ecosystem, allowing facile quantification of user-defined supply-chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
@@
 \label{sec:methodology}
-\cbox{Section word count: 800}
+\cbox{Section word count: 1600}
 \input{methodology.tex}
@@
 \section{Results}
-\cbox{Section word count: 1400}
+\cbox{Section word count: 1300}
 \label{sec:results}
@@
 \cbox{Section word count: 250}
-We have written the T-reX tool, an extension to the brightway LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database.
+We have created the T-reX tool, an extension to the brightway LCA ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database.
 It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom `pseudo' LCIA methods.
@@
     \begin{enumerate}
-        \item List of waste and material categories
-        \item Example output of the T-reX tool
-        \item List of identified waste and material exchanges
-        \item Code (python script) used for case study
-        \item Inventory and methods used in the case study
+        \item Metadata of the T-reX python package
+        \item Details of the modules and user configuration
+        \item Description of the computational workflow
+        \item Modular flowchart of the T-reX tool
+        \item Example of the terminal output of the T-reX tool
         \item Complete tabulated results of the case study

### 2024-03-06 latex_files/results.tex
add sankeys, etc.

@@
 
-\subsubsection{Temporal and scenario variation in waste and material footprints}\label{sec:results-case_study-total_footprints}
+\subsubsection{Sankey visualisation of flows in waste and material inventory footprints}\label{sec:results-case_study-sankey}
+\
+
+\autoref{fig:sankey} Sankey diagram of the total solid waste footprint flows for the NMC 811 battery in the case study
+\cbox{I won't write about these until I know what figures we want to show. We can explore them in ActivityBrowser and decide.}
+
+\textcolor{red}{[and decide which ones are most interesting (or will fit on a page.
+The most interesting are the ones that look like the flying spaghetti monster).
+I would probably make a new sankey (back to coding!), or edit the AB one in inkscape to make it understandable on the page.]}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
+    \caption{Sankey diagram of the total solid waste footprint flows for the NMC 811 battery in the case study.}\label{fig:sankey_waste}
+\end{figure}
+
+\autoref{fig:sankey_material} shows the Sankey diagram of the chromium inventory demand footprint flows for the LFP battery in the case study.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
+    \caption{Sankey diagram of the chromium inventory demand footprint flows for the LFP battery in the case study.}\label{fig:sankey_material}
+\end{figure}
+
+\subsubsection{Temporal and scenario variation in waste and material inventory footprints}\label{sec:results-case_study-total_footprints}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/total_waste.png}
+    \includegraphics[width=0.7\linewidth]{figures/total_waste.png}
     \caption{Total solid waste footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_total}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/carbondioxide.png}
+    \includegraphics[width=0.7\linewidth]{figures/carbondioxide.png}
     \caption{Carbon dioxide waste (from carbon capture and storage) footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:carbondioxide}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/phosphate.png}
+    \includegraphics[width=0.7\linewidth]{figures/phosphate.png}
     \caption{Phosphate material demand footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:phosphate}

### 2024-03-06 latex_files/main.tex
V2 almost

@@
 \end{frontmatter}
-\cbox{Total word count: ~6000, Limit: 5000 (can be easily condensed)}
+\cbox{Total word count: ~5000, Limit: 5000}
+
+% list of abbreviations
+\begin{table}[H]
+    \centering
+    \caption{List of abbreviations}
+    \label{tab:abbreviations}
+    \begin{tabular}{ll}
+    \toprule
+    CML & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
+    CRM & Critical Raw Material \\
+    CSI & Crustal Scarcity Indicator \\
+    CSP & Crustal Scarcity Potential \\
+    EDIP & Environmental Design of Industrial Products \\
+    EOL & End-of-Life \\
+    FOEN & Swiss Federal Office for the Environment \\
+    IAM & Integrated Assessment Model \\
+    LCA & Life Cycle Assessment \\
+    LCIA & Life Cycle Impact Assessment \\
+    LFP & Lithium Iron Phosphate \\
+    LiMn$_2$O$_4$ & Lithium Manganese Oxide \\
+    NCA & Nickel Cobalt Aluminum \\
+    NMC & Nickel Manganese Cobalt \\
+    PWF & Product Waste Footprint \\
+    RCP & Representative Concentration Pathway \\
+    ReCiPe & A standard LCIA method set \\
+    SSPs & Shared Socioeconomic Pathways \\
+    UBP & Umweltbelastungspünkte (Environmental Impact Points) \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
 
@@
 \label{sec:introduction}
-\cbox{Section word count: 1450}
+\cbox{Section word count: 1500}
 \cbox{To be split into two (sub) sections: Introduction and Background, can be easily condensed by 200 words or so.}

### 2024-03-06 latex_files/introduction.tex
V2 almost
updated the intro again

@@
-The development of a `circular economy' has become a critical area of focus in the imperative pursuit of achieving sustainability objectives and curtailing our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+% "As discussed in the introduction 1 there are some existing material demand methods in the standard LCIA method
+% 264 sets, including the ‘crustal scarcity indicator’ (which provides only an aggregated, abstracted endpoint) (Arvidsson
+% 265 et al., 2020) and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental
+% 266 units) (Hauschild and Potting, 2004). In these methods, the material demand is calculated based on the total mass that
+% 267 is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring
+% 268 these materials from the biosphere into the technosphere."
+\subsection{Background}
+
+The development of a `circular economy' has become a critical area of focus in the imperative pursuit of sustainability objectives and curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{eu2022ecodesign, eu2022repair,eu2015reman}.
@@
 
-While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use-- and end-of-life-- (EOL) phases, a significant proportion of these are often `hidden' in the supply chain and thus not reported directly in the final results~\citep{laurenti2016wastefootprint,salviulo2021supplychain}.
+While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use and end-of-life (EOL) phases, a significant proportion of these can be `hidden' in the supply chain and thus not reported directly in the final results~\citep{laurenti2016wastefootprint,salviulo2021supplychain}.
+The concept of the 'footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} being since popularised by the Carbon Footprint (CF) and more recently extended to include the Material Footprint (MF) and Waste Footprint (WF)~\cite{cucek2015environmentalfootprints}.
 It has been found that these material footprints can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage~\citep{laurenti2023wastefootprint}.
@@
 
-Several LCIA methods include, to some extent, waste generation~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption~\citep{arvidsson2020csi,foen2021ecofactors}.
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP)).
+Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
+These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP) in the case of the Swiss Eco-Factors).
+[ADD CITATION HERE]
 
-In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
-While LCA seeks to model the technosphere (a.k.a.
-the anthroposphere), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.
+\subsubsection{Waste in LCA}
 
-A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in LCA.
-This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations.
-CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
-The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.
+Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time.
+Moreover, from a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
+Returning from the abstract, the economic viability of the waste processing sector depends on precise knowledge (or reasonable predictions of material/waste flows), thus, as ~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be physically based'.
 
-While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
-Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
+Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of an activity and predicting its life cycle externalities, but there is a conspicuous knowledge gap regarding the waste footprints of human activities and their environmental impacts~\citep{laurenti2023wastefootprint}.
 
-\subsection{Waste in LCA}
-
-Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.
-Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows.
-Thus, precise and detailed categorisation of these is essential to understand the `circularity' of an activity and its life cycle externalities.
-There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage~\citep{laurenti2023wastefootprint}.
 Conventional LCAs consider waste as a `service'~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
-
-In LCA, waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows.
+wIn LCA, waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows.
 Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
@@
 This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
-For example, studies have shown that popular resource footprints can cover a significant portion of environmental impact variance in product rankings~\citep{steinmann2017resourcefootprints}.
-However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate with other impact assessment scores~\citep{laurent2012carbonfootprint}.
+Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment~\cite{bisinella2024wastelca, laurenti2023wastefootprint}.
+The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints}.
+However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
 The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
@@
 Addressing these hotspots is crucial for advancing towards circularity, however,  there is a lack of a convenient and flexible way to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a better understanding of a product's total environmental footprint.
+
+\cbox{Laurenti's latest. I'm not sure how to move it earlier...}
+
 \cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity.
@@
 The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting.
-The T-reX tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
-Moreover, the T-reX tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.
+The T-reX package presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
+Moreover, the T-reX tool is not limited to waste but can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
 
-\subsection{The T-reX Tool}
+\subsubsection{Material demand in LCA}
 
-To better assess waste and material flows in LCA, we have developed a Python program built on the Brightway framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into indicators and `pseudo' LCA impact (LCIA) categories.
-In this study, we present the T-reX tool that enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
+While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere), its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
 
+A relatively new (2020) LCA method, the CSI, was developed to assess long-term global scarcity of minerals.
+The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg silicon equivalents per kg element and derived from crustal concentrations.
+CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
+The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.
+
+While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
+Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
+
+\subsubsection{Introduction to the T-reX package}
+
+To facilitate quantification of waste and material flows in LCA, we have developed a Python program built on the Brightway framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX enables LCA practitioners to easily manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
 While methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
+
+\cbox{Prospective databases}
+
+Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
+The current utility of prospective databases (in general, and in particular for the waste sector) is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
+Indeed, despite the ever more critical need model and future waste management systems, \cite{bisinella2024wastelca} reports an alarming lack of
+
+\cbox{Purpose of the T-reX, conclusion}
 

### 2024-03-06 latex_files/main.tex
updated the intro again

--- 2024-03-06 latex_files/main.tex
@@
 
-        A case study of 6 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of 5 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
 \end{frontmatter}
-\cbox{Total word count: ~6000, Limit: 5000 (can be easily condensed)}
+\cbox{Total word count: ca. 5200, Limit: 5000}
+
+% list of abbreviations
+\begin{table}[H]
+    \centering
+    \caption{List of abbreviations}
+    \label{tab:abbreviations}
+    \begin{tabular}{ll}
+    \toprule
+    CF & Carbon Footprint \\
+    EF & Ecological Footprint \\
+    CML & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
+    CRM & Critical Raw Material \\
+    CSI & Crustal Scarcity Indicator \\
+    CSP & Crustal Scarcity Potential \\
+    EDIP & Environmental Design of Industrial Products \\
+    EOL & End-of-Life \\
+    FOEN & Swiss Federal Office for the Environment \\
+    IAM & Integrated Assessment Model \\
+    LCA & Life Cycle Assessment \\
+    LCI & Life Cycle Inventory \\
+    LCIA & Life Cycle Impact Assessment \\
+    LFP & Lithium Iron Phosphate \\
+    LiMn$_2$O$_4$ & Lithium Manganese Oxide \\
+    MF & Material Footprint \\
+    NCA & Nickel Cobalt Aluminum \\
+    NMC & Nickel Manganese Cobalt \\
+    PWF & Product Waste Footprint \\
+    RCP & Representative Concentration Pathway \\
+    ReCiPe & A standard LCIA method set \\
+    SSPs & Shared Socioeconomic Pathways \\
+    UBP & Umweltbelastungspünkte (Environmental Impact Points) \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
 
@@
 \label{sec:introduction}
-\cbox{Section word count: 1450}
+\cbox{Section word count: 1700}
 \cbox{To be split into two (sub) sections: Introduction and Background, can be easily condensed by 200 words or so.}

### 2024-03-06 latex_files/methodology.tex
add sankeys, etc.
V2 almost
updated the intro again

@@
 This section is divided into two parts.
-In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of 6 Li-ion batteries.
+In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of 5 Li-ion batteries.
 
@@
 \subsubsection{Computational framework}
-\label{sec:method-T-reX_tool}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
     \centering
-    \caption{Flowchart of the T-reX tool workflow.}
-    \includegraphics[width=\textwidth]{figures/T-reX_workflow.pdf}
+    \caption{Workflow of the T-reX tool.
+Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate the waste and material inventory footprints.}
+    \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}
     \label{fig:methods-flowchart}

### 2024-03-07 latex_files/discussion.tex
updated the intro again
update V2 of the manuscript

@@
-Given that both waste generation and material demand are often strongly associated with the environmental impacts of an activity, it is important that they are included in the LCA.
-While there are numerous examples of existing and proposed methods that attempt to provide endpoint LCIA scores through convoluted formulae or subjective weighting, there is little consensus on their application and their complexity and lack of transparency can make them difficult to use and interpret.
+Given that both waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints}, we consider it of high importance that they are included in their LCA.
+Although are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they generally contain convoluted formulae or subjective weighting, and their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}).
 
-Our contribution advances the state-of-the-art by giving LCA practitioners a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
-Once the databases have been processed with the T-reX tool, the user can  easily apply the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
+T-reX advances the state-of-the-art in LCA providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
+Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
 
-The simple case study of six Li-ion batteries presented in this paper demonstrated the the utility, flexibility and limitations of the T-reX tool.
+The simple case study of five Li-ion batteries presented in this paper demonstrated the  utility, flexibility as well as some limitations of T-reX.
 
-First, by adjusting the user configuration for future scenarios and waste/material categorisation we easily produced a set of customised versions of the ecoinvent database.
-Then, by applying the T-reX `pseudo-LCIA' methods, it was trivial to calculate categorised waste and material inventory footprints for a number of present and future supply chains.
-Additionally, visual exploration in \texttt{ActivityBrowser} was possible because the T-reX `pseudo-LCIA' methods are integrated in the \texttt{brightway} project in the same way as standard LCIA methods.
+First, by adjusting the user configuration for future scenarios and waste/material categorisation we produced a set of customised versions of current and prospective ecoinvent databases along with a `pseudo-biosphere' T-reX database.
+Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for a number of present and future supply chains were trivial to calculate.
+Additionally, visual exploration in \texttt{ActivityBrowser} was possible because the T-reX `pseudo-LCIA' methods are integrated within the \texttt{brightway} project in the same way as standard LCIA methods.
 
 One limitation of the T-reX tool is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated.
-This is the information would be needed to thoroughly assess the potential environmental impacts of this waste.
-Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if, for example, the waste generated represents an actual loss of resources, or is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
-A methodic classification of waste exchanges and the end-of-life fates will be facilitated by the more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
+This is the information that would be needed to thoroughly assess the potential environmental impacts of this waste.
+Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+A methodic classification of waste exchanges and the end-of-life fates is expected to be facilitated by the ever more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
 
 The utility of the T-reX tool in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
-As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is little adaption of the future background inventories.
-In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows.
-A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.
+As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adaption of the future background inventories.
+In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows\citep{bisinella2024wastelca}.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.

### 2024-03-07 latex_files/introduction.tex
update V2 of the manuscript

@@
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{eu2022ecodesign, eu2022repair,eu2015reman}.
-In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions, wrought as an act of competition or outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
 While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use and end-of-life (EOL) phases, a significant proportion of these can be `hidden' in the supply chain and thus not reported directly in the final results~\citep{laurenti2016wastefootprint,salviulo2021supplychain}.
-The concept of the 'footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} being since popularised by the Carbon Footprint (CF) and more recently extended to include the Material Footprint (MF) and Waste Footprint (WF)~\cite{cucek2015environmentalfootprints}.
+The concept of the 'footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} being since popularised by the Carbon Footprint (CF) and more recently extended to include the Material Footprint (MF) and Waste Footprint (WF)~\citep{cucek2015environmentalfootprints}.
 It has been found that these material footprints can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage~\citep{laurenti2023wastefootprint}.
@@
 LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'---, often identifying critical hotspots and guiding prioritisation of actions.
-The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows.
-These scores are then aggregated into a single score for each impact category, which can be compared across products and processes.
+The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
+These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
 
@@
 These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP) in the case of the Swiss Eco-Factors).
-[ADD CITATION HERE]
+[ADD CITATION HERE?]
 
@@
 Moreover, from a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-Returning from the abstract, the economic viability of the waste processing sector depends on precise knowledge (or reasonable predictions of material/waste flows), thus, as ~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be physically based'.
+Returning from the abstract, the economic viability of the waste processing sector depends on precise knowledge (or reasonable predictions) of material/waste flows, thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be `physically based'.
 
-Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of an activity and predicting its life cycle externalities, but there is a conspicuous knowledge gap regarding the waste footprints of human activities and their environmental impacts~\citep{laurenti2023wastefootprint}.
+Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and predicting its life cycle externalities, but there remains a conspicuous knowledge gap regarding the waste footprints of human activities and their environmental impacts~\citep{laurenti2023wastefootprint}.
 
 Conventional LCAs consider waste as a `service'~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
-wIn LCA, waste flows are not considered as fundamental biosphere exchanges, but rather as technosphere flows.
+In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy.
 Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
@@
 In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the long-term impacts of disposal.
+In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling the long-term impacts of disposal.
 
-A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional LCA practices~\citep{laurenti2016wastefootprint}.
+A considerable portion of a product's total waste is generated during earlier stages of its supply chain such as resource extraction, transportation, and manufacturing, thus, often remaining `invisible' to conventional LCA accounting practices~\citep{laurenti2016wastefootprint}.
 This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment~\cite{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints}.
-However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
+Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\cite{bisinella2024wastelca, laurenti2023wastefootprint}.
+The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints}[ADD ONE MORE].
+Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
 The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
 
-Moreover, existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste.
+Existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste.
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing towards circularity, however,  there is a lack of a convenient and flexible way to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a better understanding of a product's total environmental footprint.
+Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible ways to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
 
@@
 
-\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity.
+\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and `circularity'.
 In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health.
-The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting.
-The T-reX package presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
+The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting (as the authors themselves admit).
+The T-reX tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
 Moreover, the T-reX tool is not limited to waste but can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
@@
 
-In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere), its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
+In the context of a mineral-hungry renewable energy transition and constant geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
+While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
 
@@
 
-While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
+While useful for its stated purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
 Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
 
-\subsubsection{Introduction to the T-reX package}
+\subsubsection{Introduction to the T-reX tool}
 
-To facilitate quantification of waste and material flows in LCA, we have developed a Python program built on the Brightway framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into indicators and `pseudo' LCA impact (LCIA) categories.
+To facilitate the quantification of waste and material flows in LCA, we have developed a Python package built to extend the \texttt{brightway framework}~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
 T-reX enables LCA practitioners to easily manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
-While methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
+While some methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
 
@@
 
-Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-The current utility of prospective databases (in general, and in particular for the waste sector) is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Indeed, despite the ever more critical need model and future waste management systems, \cite{bisinella2024wastelca} reports an alarming lack of
+Integration with the \texttt{premise} python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
+Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
+Indeed, despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
 
@@
 It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others.
-In any case, avoiding material consumption and generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and opportunities for waste reduction and material efficiency.
+By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
+In any case, minimising primary material consumption and the generation of waste is of paramount importance.
+By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.

### 2024-03-07 latex_files/main.tex
update V2 of the manuscript

--- 2024-03-06 latex_files/main.tex
@@
 
-% \linenumbers
-
+\linenumbers
+% add something about the future stuffg
     % SPOTLIGHTS (5 bullet points, max 120 characters each)
@@
         \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use.
-        \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
+        \item[Bullet 4: The significance of the results to the environment, economics, or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
         \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
@@
     \begin{graphicalabstract}
+        \cbox{to be improved :D}
         \includegraphics{grabs}
@@
         \item Rapidly identifies waste and material demand hotspots.
-        \item Presents a case study of the battery supply chain.
+        \item Presents a case study of five Li-ion battery supply chains`'.
     \end{highlights}
@@
 
-    \title{T-reX: A python package to quantify waste and material footprints in current and future LCA databases}
+    \title{T-reX: A tool to quantify waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
     \begin{abstract}
-        \cbox{Abstract word count: 151, Limit: 150}
-        Minimising material consumption and waste generation is the quintessential principle of the `circular economy'.
+        \cbox{Abstract word count: 153, Limit: 150}
+        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
         Thus, identifying and quantifying waste and material flows are of fundamental importance.
-Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of activities, where the implementation of circular principles could be most effective.
+Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of activities, where the implementation of circular principles could be most effective.
 
-        Introducing T-reX, a Python tool extending the brightway ecosystem, allowing facile quantification of user-defined supply-chain demands in current and prospective scenarios.
+        Introducing T-reX, a Python tool extending the brightway ecosystem, allowing facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
 
-        A case study of 6 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of 5 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
 \end{frontmatter}
-\cbox{Total word count: ~6000, Limit: 5000 (can be easily condensed)}
+\cbox{Total word count: ca. 5200, Limit: 5000}
+
+% list of abbreviations
+\begin{table}[H]extension
+    \centering
+    \caption{List of abbreviations}
+    \label{tab:abbreviations}
+    \begin{tabular}{ll}
+    \toprule
+    CF & Carbon Footprint \\
+    EF & Ecological Footprint \\
+    CML & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
+    CRM & Critical Raw Material \\
+    CPC & Cooperative Patent Classification \\
+    CSI & Crustal Scarcity Indicator \\
+    CSP & Crustal Scarcity Potential \\
+    EDIP & Environmental Design of Industrial Products \\
+    EOL & End-of-Life \\
+    FOEN & Swiss Federal Office for the Environment \\
+    IAM & Integrated Assessment Model \\
+    LCA & Life Cycle Assessment \\
+    LCI & Life Cycle Inventory \\
+    LCIA & Life Cycle Impact Assessment \\
+    LFP & Lithium Iron Phosphate \\
+    LiMn$_2$O$_4$ & Lithium Manganese Oxide \\
+    MF & Material Footprint \\
+    NCA & Nickel Cobalt Aluminum \\
+    NMC & Nickel Manganese Cobalt \\
+    PWF & Product Waste Footprint \\
+    RCP & Representative Concentration Pathway \\
+    ReCiPe & A standard LCIA method set \\
+    SSPs & Shared Socioeconomic Pathways \\
+    UBP & Umweltbelastungspünkte (Environmental Impact Points) \\
+    WF & Waste Footprint \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
 
@@
 \label{sec:conclusions}
-\cbox{Section word count: 250}
+\cbox{Section word count: 150}
 We have created the T-reX tool, an extension to the brightway LCA ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database.
@@
 These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
-The T-reX tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.
+The T-reX tool can be easily customised by the user to calculate the footprints of other supply chain flows such as water, gas, and critical raw materials.
 
-This paper extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply-chain risk and potential environmental damage.
+This work extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply chain risk and potential environmental damage.
 
@@
 All data used in this analysis are publicly available online under the noted sources.
-The T-reX tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T-reX}.
+The T-reX tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/Tool-reX}.
 The full source code for the T-reX tool is available at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.
+A user guide and comprehensive documentation are available at \url{https://Tool-reX.readthedocs.io}.
 
+% \section*{CRediT authorship contribution statement}
+% \cbox{Co-authors, please check this and change as necessary.}
+% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
+% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
+% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+
+% \subsection*{Alternative CRediT statement}
 \section*{CRediT authorship contribution statement}
-\cbox{Co-authors, please check this and change as necessary.}
-\textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-\textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-\textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-\textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-
-\subsection*{Alternative CRediT statement}
 

### 2024-03-07 latex_files/methodology.tex
update V2 of the manuscript

@@
 This section is divided into two parts.
-In \autoref{sec:method-T-reX_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of 5 Li-ion batteries.
+In \autoref{sec:method-T-reX_tool}, we describe the T-reX, and in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
 
-\subsection{The T-reX tool}\label{sec:method-T-reX_tool}
+\subsection{T-reX}\label{sec:method-T-reX_tool}
 
@@
 Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the 'pseudo' LCIA methods created by T-reX can be used in the accustomed way.
+T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the 'pseudo- LCIA' methods created by T-reX can be used in the accustomed way.
 The T-reX tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
@@
 T-reX is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
-Currently, it has been tested with all available system models of ecoinvent 3.5--3.10~\citep{ecoinvent2016version3}.
+Currently, it has been tested with all available system models of ecoinvent versions 3.5--3.10~\citep{ecoinvent2016version3}.
 
@@
 In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the ecoinvent database.
-The user can also customise T-reX to calculate the waste and material inventory footprints of a custom database, or a prospective database based on future scenarios by implementing \texttt{premise} with an included module.
+The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or to use a prospective database based on future scenarios by implementing \texttt{premise} with an included module.
 
-The supplementary information [ADD LINK] contains the full metadata of T-reX tool, along with a list of the constituent modules of the T-reX tool, a description of their functions and a detailed computational workflow.
-Further details can be found in the user guide and documentation~\citep{mcdowall2023T-reXdocs}.
+The supplementary information [ADD LINK] contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
+Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
 
@@
 
-T-reX is a Python package that allows one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same was as standard LCIA methods.
+T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated such as to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
 
-If desired, prospective databases can be custom-defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} that can be paired with a variety of mitigation scenarios.
+If desired, prospective databases can be defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} and can be paired with a variety of mitigation scenarios.
 
-Subsequent expansion of the databases into lists of exchanges allows relevent material and waste flows to be identified and categorised.
-The search queries are tailored to the specific database and the user can easily modify them to suit their needs.
+The deconstruction of the databases by T-reX into lists of exchanges allows the relevant material and waste flows to be identified and categorised by search functions.
+The queries facilitating this are tailored to the specific database and the user can easily modify them to suit their needs.
 The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand.
-The exchange editing function of T-reX then takes each list of exchanges and appends to the relevent activity a copy of the technosphere exchange as `pseudo-biosphere' exchange that matches the `pseudo LCIA' method.
+The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
 
@@
 The waste categories include incineration, recycling, and total waste, and are listed in the supplementary information [ADD LINK].
-One advantage of T-reX is that is is able to identify of waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
+One advantage of T-reX is that it is able to identify waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
 Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
 
-In addition to the waste categories, the \texttt{queries\_materials} module defines the material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
-The CRM list is a list of 30 materials that are considered critical to the EU economy and are at risk of supply disruption.
-Further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
+In addition to the waste categories, T-reX includes numerous material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
+The CRM list contains 30 materials that are considered critical to the EU economy and are at risk of supply disruption.
+Several further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
 The identity of the materials considered and their categorical groupings are easily customisable by the user.
@@
 
-The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
-A useful feature of the T-reX tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
+The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+A useful feature T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
 For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
-Similarly, the total demand for all critical raw materials (CRMs) can be easily calculated in the same manner.
+Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.
 
-As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the `crustal scarcity indicator' (which provides only an aggregated, abstracted endpoint)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
-In the T-reX tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
-This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply-chain impacts of the demand for this metal.
+T-reX, in contrast, uses accounting for material demand (and waste generation) based on exchanges solely within the technosphere.
+This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply chain impacts of the demand for this metal.
 If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
@@
 
-% \subsubsection{Theoretical basis}
+\subsubsection{Mathematical basis}
 
-% XXX
+\cbox{very nice formulae coming soon}
 
-\subsubsection{T-reX tool workflow}
+\subsubsection{The workflow of T-reX}
 
-The workflow of the T-reX tool is divided into several modules, each of which performs a specific function.
-The modules are designed to be used in a specific order, but the user can also use them individually to perform specific tasks.
+The workflow of T-reX is divided into several modules, each performing a seperate function.
+The modules are designed to be used in a particular order, but the user can also use them individually to perform specific tasks.
 The standard workflow is as follows:
@@
     \item Exchange editing---whereby the technosphere exchange is mirrored as a `pseudo-biosphere' exchange
-    \item Database verification
+    \item Database verification---to ensure that T-reX has manipulated the database correctly
 \end{enumerate}
 
-This workflow creates one or more manipulated LCA databases that can then be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+This workflow creates copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX 'pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
 
-A generic flowchart of the T-reX tool workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material [ADD LINK] contains a more detailed computational flowchart.
+A generic flowchart of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material [ADD LINK] contains a more detailed computational flowchart.
 
@@
     \centering
-    \caption{Workflow of the T-reX tool.
-Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate the waste and material inventory footprints.}
+    \caption{Workflow of the T-reX.
+Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
     \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}
@@
 
-\subsection{Case Study Methodology}\label{sec:method-case_study}
+\subsection{Case study methodology}\label{sec:method-case_study}
 
-We investigated five types of Li-ion batteries, each represented by specific market activities:
+We investigated five types of Li-ion batteries, each represented by their unamended global market activities:
 \begin{itemize}[itemsep=0pt]
+    \item Li-ion, LFP, rechargeable, prismatic
+    \item Li-ion, LiMn2O4, rechargeable, prismatic
     \item Li-ion, NMC111, rechargeable, prismatic
-    \item Li-ion, LiMn2O4, rechargeable, prismatic
+    \item Li-ion, NMC811, rechargeable, prismatic
     \item Li-ion, NCA, rechargeable, prismatic
-    \item Li-ion, NMC811, rechargeable, prismatic
-    \item Li-ion, LFP, rechargeable, prismatic
 \end{itemize}
@@
     \item EDIP 2003 no LT
-    \item Crustal Scarcity
+    \item CSI 2020
 \end{itemize}
 
-The primary source of life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff.
+The source of the life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff.
 Additionally, the T-reX tool was used to create prospective database sets using the \mbox{REMIND} model with the following Representative Concentration Pathways (RCPs):
@@
 
-For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with the T-reX tool over the time series: 2020, 2040, 2060, 2080, 2100.
+For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with T-reX over the time series: 2020, 2040, 2060, 2080, 2100.
 
-\subsubsection{Calculations}
-For each combination of activity, method, and database, a single score `LCIA' was calculated along with details of the top contributing processes.
-Additionally, for the Waste and Material Footprint methods, a contribution analysis was performed.
-This involved utilizing the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, an additional component of the Brightway2 LCA framework.
-This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their CPC codes.
-This provides insight into the products and sectors in the supply chain of the activity that carry the most responsibility for the final footprint.
+\subsubsection{LCA calculations}
+For each combination of activity, method, and database, a single score was calculated along with details of the top contributing processes.
+Additionally, for the T-reX methods, a contribution analysis was performed.
+This involved utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
+This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
+This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand.

### 2024-03-07 latex_files/results.tex
V2 almost
update V2 of the manuscript

@@
 
-An example of the output from the application of the T-reX tool has been included in the supplementary material. The manipulated ecoinvent databases 
-can be recreated using the code and instructions available in the package documentation~\citep{mcdowall2023T-reXdocs}.
+An example of the output from the application of the T-reX tool has been included in the supplementary material.
+The manipulated ecoinvent databases can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 
-As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (as well as a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at market.
+As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (with variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the T-reX tool.
-This section includes some highlights of the results and the full results are available in the supplementary material.
-Because the T-reX methods are integrated into the brightway project as if they were LCIA methods, the results can be visualised in the same way.
+This section includes some highlights of the results with full results included in the supplementary material [ADD LINK].
+Because the `pseudo LCIA' T-reX methods are integrated into the brightway project as if they were LCIA methods, the LCA calculations can be performed in the customary way.
 In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
@@
 \subsubsection{Sankey visualisation of flows in waste and material inventory footprints}\label{sec:results-case_study-sankey}
-\
 
-\autoref{fig:sankey} Sankey diagram of the total solid waste footprint flows for the NMC 811 battery in the case study
+\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' for the NMC 811 Li-ion battery at the global market in the ecoinvent database version 3.9.1
 \cbox{I won't write about these until I know what figures we want to show. We can explore them in ActivityBrowser and decide.}
@@
     \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{Sankey diagram of the total solid waste footprint flows for the NMC 811 battery in the case study.}\label{fig:sankey_waste}
+    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the ecoinvent database version 3.9.1}\label{fig:sankey_waste}
 \end{figure}
 
-\autoref{fig:sankey_material} shows the Sankey diagram of the chromium inventory demand footprint flows for the LFP battery in the case study.
+\autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1.
 
@@
     \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{Sankey diagram of the chromium inventory demand footprint flows for the LFP battery in the case study.}\label{fig:sankey_material}
+    \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1}\label{fig:sankey_material}
 \end{figure}
@@
 
-\autoref{fig:waste_total} shows the total solid waste footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+\autoref{fig:waste_total} shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
 The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced.
@@
 
-\subsubsection{Contribution of sectors in the supply chain}\label{sec:results-case_study-topsectors}
+\subsubsection{Contribution of industrial sectors in the supply chain}\label{sec:results-case_study-topsectors}
 
@@
 For the material demand methods especially, this distinction is critical.
-For example, application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the T-reX method provides an aggregation of the exchanges with the market for that metal.
-The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply-chain pressures that are not captured by the standard LCIA methods.
+For example, the application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the T-reX method provides an aggregation of the exchanges with the market for that metal.
+The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply chain pressures that are not captured by the standard LCIA methods.
 This makes the T-reX methods more sensitive to the modeling choices (e.g., allocation) that are generally embedded in LCA databases.
@@
 
-\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in ecoinvent version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
+\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in the ecoinvent database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
 The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
@@
 The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
-In the T-reX method, the source database is exploded into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+In the T-reX method, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
 

### 2024-03-07 latex_files/supplementary_material.tex
add sankeys, etc.
update V2 of the manuscript

@@
 
-\section{Detailled description of the T-reX tool workflow}
+\section{Description of the computational workflow}
 
@@
 In the T-reX tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
-This offers a different perspective, allowing for the estimation of overall supply-chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply-chain impacts of the demand for this metal.
+This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply chain impacts of the demand for this metal.
 If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.

### 2024-03-08 latex_files/introduction.tex
small update to V2

@@
 
-To facilitate the quantification of waste and material flows in LCA, we have developed a Python package built to extend the \texttt{brightway framework}~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway framework}~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX means for `Tool for re-X' (reduce, reuse, recycle, etc.).
 T-reX enables LCA practitioners to easily manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.

### 2024-03-08 latex_files/main.tex
small update to V2

@@
         \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
-        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use.
-        \item[Bullet 4: The significance of the results to the environment, economics, or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
-        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
+        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use
+        \item[Bullet 4: The significance of the results to the environment, economics, or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains
+        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting
     \end{description}
@@
     \begin{highlights}
-        \item T-reX, a new tool for quantifying waste and material flows in LCA.
-        \item Assesses supply risks by calculating demand for critical materials.
-        \item Simplifies quantification of user-specified waste and material categories.
-        \item Rapidly identifies waste and material demand hotspots.
-        \item Presents a case study of five Li-ion battery supply chains`'.
+        \item T-reX, a new tool for quantifying waste and material flows in LCA
+        \item Assesses supply risks by calculating demand for critical materials
+        \item Simplifies quantification of user-specified waste and material categories
+        \item Rapidly identifies waste and material demand hotspots
+        \item Presents a case study of five Li-ion battery supply chains
     \end{highlights}
@@
 
-        A case study of 5 Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of five Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
 \cbox{Total word count: ca. 5200, Limit: 5000}
+\cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
+\cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2_SupplementaryMaterial.pdf}{here}}
+\cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}
 
 % list of abbreviations
-\begin{table}[H]extension
+\begin{table}[H]
     \centering
@@
 \cbox{Section word count: 1700}
-\cbox{To be split into two (sub) sections: Introduction and Background, can be easily condensed by 200 words or so.}
 \input{introduction.tex}
@@
 \label{sec:discussion}
+\cbox{Section word count: 450}
 \input{discussion.tex}
-\cbox{Section word count: 450}
 
@@
 These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
-The T-reX tool can be easily customised by the user to calculate the footprints of other supply chain flows such as water, gas, and critical raw materials.
+T-reX can be easily customised by the user to calculate the footprints of other supply chain flows such as water, gas, and critical raw materials.
 
@@
 \section*{Data availability} 
-All data used in this analysis are publicly available online under the noted sources.
-The T-reX tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/Tool-reX}.
+All data used in this analysis are publicly available online under the noted sources.T-reX is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T_reX_LCA}.
 The full source code for the T-reX tool is available at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide and comprehensive documentation are available at \url{https://Tool-reX.readthedocs.io}.
+A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.
 
@@
 \section*{CRediT authorship contribution statement}
+
+\begin{itemize}[noitemsep]
+    \item \textbf{SCM}: Stewart Charles McDowall
+    \item \textbf{EL}: Elizabeth Lanphear
+    \item \textbf{SC}: Stefano Cucurachi
+    \item \textbf{CFB}: Carlos Felipe Blanco
+\end{itemize}
 

### 2024-03-08 latex_files/methodology.tex
small update to V2

@@
 
-This workflow creates copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX 'pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+This workflow creates copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
 
@@
     \caption{Workflow of the T-reX.
-Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
+Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.
+[TO BE IMPROVED!]}
     \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}
@@
 
-The source of the life cycle inventory data for this case study was ecoinvent 3.9.1 cutoff.
-Additionally, the T-reX tool was used to create prospective database sets using the \mbox{REMIND} model with the following Representative Concentration Pathways (RCPs):
+The source of the life cycle inventory data for this case study was the database ecoinvent, version 3.9.1 with the `cutoff' attributional system model.
+Additionally, T-reX was used to create prospective database sets using the \mbox{REMIND} model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}

### 2024-03-08 latex_files/results.tex
small update to V2

@@
     \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the ecoinvent database version 3.9.1}\label{fig:sankey_waste}
+    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
 \end{figure}
@@
     \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1}\label{fig:sankey_material}
+    \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
 \end{figure}
@@
     \includegraphics[width=0.7\linewidth]{figures/carbondioxide.png}
-    \caption{Carbon dioxide waste (from carbon capture and storage) footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:carbondioxide}
+    \caption{Carbon dioxide waste (from carbon capture and storage) footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+[TO BE IMPROVED!]}\label{fig:carbondioxide}
 \end{figure}
@@
     \includegraphics[width=0.7\linewidth]{figures/phosphate.png}
-    \caption{Phosphate material demand footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:phosphate}
+    \caption{Phosphate material demand footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+[TO BE IMPROVED!]}\label{fig:phosphate}
 \end{figure}
@@
     \includegraphics[width=0.6\linewidth]{figures/top-processes.png}
-    \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:top_contribution}
+    \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+[TO BE IMPROVED!]}\label{fig:top_contribution}
 \end{figure}
@@
     }
-    \caption{Contribution of industrial sectors to the liquid waste footprint of the NCA battery from 2020 to 2100 under the SSP2 scenario using the PkBudg500 RCP of the REMIND model.}\label{fig:cpc_contribution}
+    \caption{Contribution of industrial sectors to the liquid waste footprint of the NCA battery from 2020 to 2100 under the SSP2 scenario using the PkBudg500 RCP of the REMIND model.
+[TO BE IMPROVED!]}\label{fig:cpc_contribution}
 \end{figure}
@@
     \end{subfigure}
-    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `T-reX - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:comparison_methods}
+    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `T-reX - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.[TO BE IMPROVED!]}\label{fig:comparison_methods}
 \end{figure}

### 2024-03-08 latex_files/latex_compile.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+echo "================================================================================"
+echo "Compiling the T-reX manuscript..."
+echo "-------------------------------------------------------------------------------"
+# Clean the project
+echo "Cleaning the project..."
+echo "-------------------------------------------------------------------------------"
+latexmk -C
+echo "-------------------------------------------------------------------------------"
+# Compile the project
+echo "Compiling the project..."
+echo "-------------------------------------------------------------------------------"
+latexmk -pdf -quiet main.tex
+echo
+echo "-------------------------------------------------------------------------------"
+echo "Done compiling the T-reX manuscript."
+echo "-------------------------------------------------------------------------------"
+# Copy the output PDF to the specified directory
+cp main.pdf ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf
+
+echo "Copied the output PDF to ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf"
+echo "Opening the PDF..."
+echo "================================================================================"
+echo " ----------------------           FIN!               ---------------------------"           
+echo "================================================================================"
+echo 
+echo
+# Open the PDF with the default viewer in a detached process
+# nohup xdg-open ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf > /dev/null 2>&1 &

### 2024-03-15 latex_files/main.tex
update

@@
 
-        A case study of five Li-ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of five lithium ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
     LFP & Lithium Iron Phosphate \\
+    Li-ion & Lithium ion \\
     LiMn$_2$O$_4$ & Lithium Manganese Oxide \\
@@
 
+The authors would like to acknowledge XXXX, XXXX, XXXX, XXX, who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive opensource software framework for the study and development of secondary raw material systems.
+The authors hope to continue the development of T-reX toward this vision.
+
 \section*{Supplementary material}

### 2024-03-25 origins/LL_originals/DB_explode.py
add

@@ -1,0 +1,64 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Explode DB
+Created on Wed Nov 16 11:31:38 2022
+@author: stew
+"""
+import bw2io as bi
+import bw2data as bd
+import wurst as w
+import pandas as pd
+import os
+
+#%% 
+project = "EOL-WDI-SCM-1"
+db = "ei38_co"
+
+#%% Introduction and project/db choice (if not given in function call)
+
+print("** DBexplode returns a list of all exchanges in given BW2 database **")
+print("\n * Using packages: ", 
+      "\n\t bw2io" , bi.__version__, 
+      "\n\t bw2data" ,bd.__version__,
+      "\n\t wurst" , w.__version__)
+
+try: bd.projects.set_current(project)
+except NameError:
+    print("\n* Available projects: \n\t" , [i[0] for i in bd.projects.report()])
+    project = input("\n* Enter name of project to copy: ")
+    bd.projects.set_current(project)
+    project = input("\n* Enter name of project new project: ")
+    bd.projects.copy_project(project)
+
+try: db = bd.Database(db)
+except NameError: 
+    print("\n* Availabile databases in "+ bd.projects.current() +": " + str(list(bd.databases)))
+    db = input("\n* Enter name of database to be processed: ")
+    db = bd.Database(db)
+
+print("\n* db:", db.name, "in project:", bd.projects.current, "will be processed")
+
+#%% Explode database
+print("\n Opening the sausage... \n")
+guts = w.extract_brightway2_databases(db.name)
+
+print("\n Extracting activities from db...")
+df = pd.DataFrame(guts, columns =['code', 'exchanges'])
+
+print("\n Exploding exchanges from activities...")
+df = df.explode("exchanges", ignore_index=False)
+
+print("\n Pickling...")
+tmp = os.getcwd() + "/tmp"
+if not os.path.isdir(tmp): os.mkdir(tmp)
+pickle_path = os.path.join(tmp, "db.pickle")
+df.to_pickle(pickle_path)
+print(db.name, "was exploded and pickled. yay.")
+
+
+
+
+
+
+

### 2024-03-25 origins/LL_originals/WDI_LL_original.py

@@ -1,0 +1,1250 @@
+#!/usr/bin/env python
+# coding: utf-8
+"""
+## This function searches for and aggregates all wastes (category is flexible) 
+## by total mass or volume and provides this as an indicator in the Impact Assessment stage 
+## of a given LCA in the Activity Browser.
+"""
+
+#%% Section 1: Setting Up Variables and Importing Ecoinvent, Biosphere, and Custom Databases
+
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import brightway2 as bw
+from brightway2 import *
+from bw2data.parameters import ActivityParameter, DatabaseParameter, ProjectParameter, Group
+import bw2data
+
+
+# Set this input to your specific project
+projects.set_current("ecoinvent-import2")
+
+create_core_migrations() #???? is that needed?
+
+# Import Custom Waste Biosphere Flows (Liquid and Solid Waste)
+sp = bw.ExcelImporter("wasteBiosphere_final.xlsx")
+sp.apply_strategies()
+sp.statistics()
+sp.write_database()
+
+# Import Ecoinvent 3.8 Cutoff
+fpei38 = "/Users/girk/Desktop/ecoinvent_3.8_cutoff_ecoSpold02/datasets"
+ei38 = bw.SingleOutputEcospold2Importer(fpei38, 'ecoinvent 3.8 cutoff')
+ei38.apply_strategies()
+ei38.statistics()
+ei38.write_database()
+
+eidb = bw.Database('ecoinvent 3.8 cutoff')
+bio3 = bw.Database('biosphere3')
+wd = bw.Database('cstmBiosphere')
+
+database_names = ('biosphere3', 'cstmBiosphere','ecoinvent 3.8 cutoff', 'example', 'waste demand')
+
+#%% Section 2: Applying Wurst to Format Biosphere, Ecoinvent, and Custom Database Structure in Python
+import wurst as w
+
+custom = w.extract_brightway2_databases(["cstmBiosphere"])
+data = w.extract_brightway2_databases(["ecoinvent 3.8 cutoff"])
+biodata = w.extract_brightway2_databases(["biosphere3"])
+
+
+#%% Section 3: Using Pandas Dataframe to Access Waste Exchanges Without Losing Ability to Connect Exchange to
+## Original Process
+df = pd.DataFrame(data,
[…]

### 2024-03-25 origins/LL_originals/methodtest.py

@@ -1,0 +1,77 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Nov 20 11:45:29 2022
+
+@author: stew
+"""
+
+import os
+import brightway2 as bw
+import pandas as pd
+
+# USER VARIABLES
+# the name of the impact category file:
+file_name = 'TWDliquid.xlsx'
+# the name of the Brightway project to add the impact category to:
+project_name = 'cutoff38'
+# the name of the biosphere database in above project:
+biosphere_name = 'db_waste'
+
+
+# find the right file and read it
+cwd = os.getcwd()
+file_path = os.path.join(cwd, file_name)
+mthd = pd.read_excel(file_path)
+
+# open the right BW project and biosphere database
+bw.projects.set_current(project_name)
+biosphere = bw.Database(biosphere_name)
+bs = pd.DataFrame(biosphere)
+
+#check if there are matching problems
+print('+++ Checking if the flows of the file match the biosphere...')
+new_method = []
+errors = 0
+for row in mthd.iterrows():
+    matches = bs[(bs['name'] == row[1]['name'])]
+    if matches.shape[0] == 0:
+        # this TMR flow does not exist in the biosphere database
+        print('    > no matches for:', row[1]['name'])
+        errors += 1
+    elif matches.shape[0] > 1:
+        # this TMR flow is not well defined and could match multiple flows in the biosphere database
+        print('    > too many matches for:', row[1]['name'])
+        errors += 1
+    else:
+        # this TMR flow has 1 match, we add a tuple of the key and its flow value to our method list
+        key = (biosphere_name, matches['code'].tolist()[0])
+        value = row[1]['meanValue']
+        new_method.append((key, value))
+
+# If there were errors, ask user if they still want to add the impact category
+if errors > 0:
+    m = '\n+++ There were problems matching the TWD flows to the biosphere database'
+    q = '\n+++ Would you still like to add the impact category? [y/n]: '
+    cont = confirmation(question=q, message=m)
+    if not cont:
+        print("+++ There is likely a problem with the file or this version of the biosphere database"
+              "    Check the file or contact the authors mentioning the names of the above flows")
+        raise Exception('USER EXIT')
+        exit()
+else:
+    print("+++ No problems found in matching the flows.")
+
+# Add the data to Brightway:
+try:
+    pass
+    method_name = ('Waste Footprint', 'Total Waste Demand', 'Liquid Waste')
[…]

### 2024-03-27 latex_files/discussion.tex
update-text

@@
-Given that both waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints}, we consider it of high importance that they are included in their LCA.
-Although are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they generally contain convoluted formulae or subjective weighting, and their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}).
+Given that waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints}, we consider it of high importance that they are included LCA accounting.
+Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they generally contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
 
@@
 
-One limitation of the T-reX tool is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated.
+One current limitation of T-reX is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated.
 This is the information that would be needed to thoroughly assess the potential environmental impacts of this waste.
@@
 
-The utility of the T-reX tool in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
 As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adaption of the future background inventories.
-In particular, the inclusion of scenarios with future waste processing technology would greatly improve our predictions of waste and material flows\citep{bisinella2024wastelca}.
+In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts\citep{bisinella2024wastelca}.
 A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.

### 2024-03-27 latex_files/introduction.tex
update-text

--- 2024-03-07 latex_files/introduction.tex
@@
 % 268 these materials from the biosphere into the technosphere."
-\subsection{Background}
+\subsection{Background}\label{sec:intro-background}
 
 The development of a `circular economy' has become a critical area of focus in the imperative pursuit of sustainability objectives and curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{eu2022ecodesign, eu2022repair,eu2015reman}.
+Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair,eu2015reman}.
 In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
-While some material demands are apparent in the final product and the waste generated may be inferred from knowledge of the use and end-of-life (EOL) phases, a significant proportion of these can be `hidden' in the supply chain and thus not reported directly in the final results~\citep{laurenti2016wastefootprint,salviulo2021supplychain}.
-The concept of the 'footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} being since popularised by the Carbon Footprint (CF) and more recently extended to include the Material Footprint (MF) and Waste Footprint (WF)~\citep{cucek2015environmentalfootprints}.
-It has been found that these material footprints can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and that waste footprints have a `strong association' with environmental damage~\citep{laurenti2023wastefootprint}.
-Thus, to reduce the negative externalities of consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible.
+The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' developed strongly and has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
+More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is more often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
+Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
+It has been shown that MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
+
+The incipient Waste Footprint (WF) metric, in contrast, though emerging steadily, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities~\citep{laurenti2016wastefootprint}.
+Despite the fact that reducing life cycle waste is clearly critical to the development of the circular economy~\cite{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent evironmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
+Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that waste footprints can have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Alas, in frequent, tragic failures of evironmental justice, it is often the most vunerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,pellow2023envjusticewaste}.
+
+WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders.
+Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint}.
 
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'---, often identifying critical hotspots and guiding prioritisation of actions.
+LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
 These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
+Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct prospective background databases in an effort to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
 
 Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP) in the case of the Swiss Eco-Factors).
-[ADD CITATION HERE?]
+These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP) in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
 
-\subsubsection{Waste in LCA}
+\subsubsection{Waste in LCA}\label{sec:intro-waste}
 
@@
 Moreover, from a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-Returning from the abstract, the economic viability of the waste processing sector depends on precise knowledge (or reasonable predictions) of material/waste flows, thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be `physically based'.
+Returning from the abstract and etherial to the economic viability of the waste processing sector.
+While critical here are energy costs and the relative prices of virgin and secondary materials, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows.
+Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be physically based'.
+The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.
 
@@
 
-Conventional LCAs consider waste as a `service'~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
+Conventional LCA database models consider waste as a `service' (accounting for the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
 In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy.
@@
 Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\cite{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints}[ADD ONE MORE].
+The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
 Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
@@
 
-\cbox{Laurenti's latest. I'm not sure how to move it earlier...}
-
 \cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and `circularity'.
 In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health.
-The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting (as the authors themselves admit).
-The T-reX tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA.
-Moreover, the T-reX tool is not limited to waste but can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
+The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting (as the authors themselves acknowledge).
+The `T-reX' Python software tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA\@.
+Moreover, the utility of T-reX is not limited to waste, it can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
 
-\subsubsection{Material demand in LCA}
+\subsubsection{Material demand in LCA}\label{sec:intro-material}
 
@@
 
-A relatively new (2020) LCA method, the CSI, was developed to assess long-term global scarcity of minerals.
+A relatively new (2020) LCIA method, the Crustal Scarcity (CSI), was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
 The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg silicon equivalents per kg element and derived from crustal concentrations.
@@
 
-While useful for its stated purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
+While useful for its intended purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
 Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
 
-\subsubsection{Introduction to the T-reX tool}
+\subsection{Introduction to T-reX}\label{sec:intro-trex}
 
-To facilitate the quantification of waste and material flows in LCA, we have developed a Python package built to extend the \texttt{brightway framework}~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX enables LCA practitioners to easily manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} LCA framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX, in full, stands for `Tool for re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases such as to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
 While some methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
 
-\cbox{Prospective databases}
-
-Integration with the \texttt{premise} python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
+Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
 Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
@@
 
-\cbox{Purpose of the T-reX, conclusion}
-
-The purpose of the T-reX tool is not to quantify the environmental impacts of material consumption and waste production, but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
+The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requsite data and impact modelling remains lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
 It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
@@
 In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.
+By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.

### 2024-03-27 latex_files/main.tex
update-text

@@
     \cbox{to be improved :D}
-        \includegraphics{grabs}
+    \includegraphics{grabs.jpg}
 \end{graphicalabstract}
@@
 \begin{highlights}
-        \item T-reX, a new tool for quantifying waste and material flows in LCA
+    \item T-reX, a new software tool for quantifying waste and material flows in LCA
     \item Assesses supply risks by calculating demand for critical materials
@@
     \begin{abstract}
-        \cbox{Abstract word count: 153, Limit: 150}
+        \cbox{Abstract word count: 155, Limit: 150}
         The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
@@
 
-        A case study of five lithium ion batteries demonstrated T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of five lithium ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
 \cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
-\cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2_SupplementaryMaterial.pdf}{here}}
+\cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V3/T-reX_manuscript_V3_SupplementaryMaterial.pdf}{here}}
 \cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}
@@
         EDIP          & Environmental Design of Industrial Products                         \\
+        EEIOA         & Environmentally Extended Input-Output Analysis                      \\
         EOL           & End-of-Life                                                         \\
@@
         IAM           & Integrated Assessment Model                                         \\
+        IE            & Industrial Ecology                                                  \\
         LCA           & Life Cycle Assessment                                               \\
@@
         MF            & Material Footprint                                                  \\
+        MFA           & Material Flow Analysis                                              \\
         NCA           & Nickel Cobalt Aluminum                                              \\
@@
         SSPs          & Shared Socioeconomic Pathways                                       \\
+        T-reX         & Tool for re-X in LCA                                                \\
         UBP           & Umweltbelastungspünkte (Environmental Impact Points)                \\
+        UN            & United Nations                                                      \\
         WF            & Waste Footprint                                                     \\
@@
 \cbox{Section word count: 150}
-We have created the T-reX tool, an extension to the brightway LCA ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database.
+
+We have created `T-reX', an extension to the \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database.
 It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom `pseudo' LCIA methods.
@@
 All data used in this analysis are publicly available online under the noted sources.T-reX is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for the T-reX tool is available at \url{https://www.github.com/Stew-McD/T-reX}.
+The full source code for T-reX is available at \url{https://www.github.com/Stew-McD/T-reX}.
 A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.
@@
 
-\section*{Supplementary material}
+\section*{Supplementary material}\label{sec:supplementary}
 The supplementary material contains the following:
 \begin{enumerate}
-        \item Metadata of the T-reX python package
+    \item Metadata of the T-reX Python package
     \item Details of the modules and user configuration
     \item Description of the computational workflow
-        \item Modular flowchart of the T-reX tool
-        \item Example of the terminal output of the T-reX tool
+    \item Modular flowchart of T-reX
+    \item Example of the terminal output of T-reX
     \item Complete tabulated results of the case study

### 2024-03-27 latex_files/methodology.tex
update-text

@@
 T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the 'pseudo- LCIA' methods created by T-reX can be used in the accustomed way.
-The T-reX tool is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
-The full source code for the T-reX tool is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
+T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
+The full source code for T-reX is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
 T-reX is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
@@
 
-The supplementary information [ADD LINK] contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
+The supplementary material\autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
 Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
@@
 T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated such as to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
+We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of a activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
 
@@
 In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include incineration, recycling, and total waste, and are listed in the supplementary information [ADD LINK].
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material\autoref{sec:supplementary}.
 One advantage of T-reX is that it is able to identify waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
@@
 The identity of the materials considered and their categorical groupings are easily customisable by the user.
-The full list of 59 materials included in the default configuration is provided in the supplementary material [ADD LINK].
+The full list of 59 materials included in the default configuration is provided in the supplementary material\autoref{sec:supplementary}.
 
-The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest rather than keywords in the exchange's name.
 A useful feature T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
@@
 
-As discussed in the introduction~\ref{sec:introduction} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+As presented in the introduction of this study~\autoref{sec:intro-background,sec:intro-waste,sec:intro-material} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
+
 T-reX, in contrast, uses accounting for material demand (and waste generation) based on exchanges solely within the technosphere.
 This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply chain impacts of the demand for this metal.
-If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
+If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
+In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models)despite the presence of nickel in the final products.
 
@@
 
-The workflow of T-reX is divided into several modules, each performing a seperate function.
+The workflow of T-reX is divided into several modules, each performing a separate function.
 The modules are designed to be used in a particular order, but the user can also use them individually to perform specific tasks.
@@
 
-This workflow creates copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
 
-A generic flowchart of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material [ADD LINK] contains a more detailed computational flowchart.
+A generic flowchart of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material\autoref{sec:supplementary} contains a more detailed computational flowchart.
 
@@
     \centering
-    \caption{Workflow of the T-reX.
+    \caption{Workflow of T-reX.
 Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.
@@
 
-In addition to the waste and material inventory footprint methods created by the T-reX tool, the following standard LCIA methods were applied for comparison:
+In addition to the waste and material inventory footprint methods created by T-reX, the following standard LCIA methods were applied for comparison:
 
@@
 
-For each pathway, databases were created with texttt{premise}~\citep{sacchi2022premise} and processed with T-reX over the time series: 2020, 2040, 2060, 2080, 2100.
+For each pathway, databases (at five year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.
 

### 2024-03-27 latex_files/results.tex
update-text

@@
 
-An example of the output from the application of the T-reX tool has been included in the supplementary material.
+An example of the output from the application of T-reX has been included in the supplementary material.
 The manipulated ecoinvent databases can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
@@
 As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (with variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
-The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of the T-reX tool.
-This section includes some highlights of the results with full results included in the supplementary material [ADD LINK].
+The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
+This section includes some highlights of the results with full results included in the supplementary material\autoref{sec:supplementary}.
 Because the `pseudo LCIA' T-reX methods are integrated into the brightway project as if they were LCIA methods, the LCA calculations can be performed in the customary way.
@@
 In each case there was a slight downward trend in the waste footprints between 2020 and 2100.
-This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is included in the models over this time.
+This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is included in the models over this time (which can large amounts of waste in its extraction and combustion phases).
 For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
@@
 
-A comparison of the results from the T-reX tool's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
+A comparison of the results from T-reX's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
 In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
@@
 For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste, Laurenti et al.
-reported 95\% of the total waste, whereas the T-reX tool reported only 3\%.
+reported 95\% of the total waste, whereas T-reX reported only 3\%.
 The reason for this discrepancy is explained by the fact that in the method of Laurenti et al., a ``waste flow was regarded as hazardous when the hazardousness

### 2024-03-27 latex_files/supplementary_material.tex
update-text

@@
 
-\title{\textbf{Supplementary Information for the T-reX Manuscript}}
+\title{\textbf{supplementary material for the T-reX Manuscript}}
 \author{Stewart Charles McDowall, Elizabeth Lanphere, Carlos Felipe Blanco, Stefano Cucurachi \\
@@
 
-\section{Metadata of the T-reX tool}
+\section{Metadata of T-reX}
 
@@
 
-\section{List of modules in the T-reX python package}
+\section{List of modules in the T-reX Python package}
 
@@
 The \texttt{search\_waste} and \texttt{search\_material} modules apply user-defined search parameters from \texttt{queries\_waste} and \texttt{queries\_materials} to identify relevant waste and material flows in the list of technosphere exchanges generated by \texttt{explode\_database} and categorises them accordingly.
-The results of the search functions are stored in \texttt{.csv} files for subsequent use in the T-reX tool's workflow.
+The results of the search functions are stored in \texttt{.csv} files for subsequent use in T-reX's workflow.
 
@@
     \centering
-    \caption{Waste categories in the default configuration of the T-reX tool.}\label{tab:waste_categories}
+    \caption{Waste categories in the default configuration of T-reX.}\label{tab:waste_categories}
     \begin{tabularx}{\textwidth}{p{2.5cm}p{1cm}p{1.5cm}XX}
@@
 
-The logic for the identification of material exchanges with the T-reX tool differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
+The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest.
 That is, for material $x$, all exchanges with the name `market for material $x$' are identified and subsequently apportioned a (`pseudo-biosphere') material demand exchange of the same sign and magnitude as the original exchange.
-A useful feature of the T-reX tool is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
+A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can easily aggregate these exchanges.
 For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
@@
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
-In the T-reX tool, however, the accounting for material demand is based on exchanges solely within the technosphere.
+In T-reX, however, the accounting for material demand is based on exchanges solely within the technosphere.
 This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, the T-reX tool can provide insight into the broader supply chain impacts of the demand for this metal.
+Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
 If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
@@
 \subsection{LCIA method management}
-The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in the T-reX tool.
+The \texttt{method\_editor} module manages the addition, deletion, and verification of the custom LCIA methods used in T-reX.
 This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange.
@@
 The \texttt{exchange\_editor} module loads the \texttt{.csv} files created by the search functions and appends `pseudo-biosphere' exchanges to the matching activities in the LCA database.
-This is the most computationally intensive part of the T-reX tool, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
+This is the most computationally intensive part of T-reX, as (depending on the search configuration) there are generally more than 100,000 exchanges to be appended to the database.
 
 \subsection{Database Verification}
-The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that the T-reX tool has processed the database correctly.
+The \texttt{verify\_database} module calculates LCA scores for randomly selected activities using Waste Footprint and Material Demand Footprint methods to confirm that T-reX has processed the database correctly.
 
-\section{Modular flowchart of the T-reX tool}
+\section{Modular flowchart of T-reX}
 
@@
     \includegraphics[height=0.8\textheight]{figures/T-reX_flowchart.pdf}
-    \caption{Computational flowchart of the T-reX tool.
+    \caption{Computational flowchart of T-reX.
 In green are the individual modules, in blue are their outputs.
@@
 
-\section{Example terminal output of the T-reX tool}
+\section{Example terminal output of T-reX}
 
-Due to its length, an example of the terminal output of the T-reX tool is included as a separate file in the supplementary material, named `T-reX\_example\_output.html', which can be opened in an internet browser.
-The example output is from the execution of the T-reX tool on the `ecoinvent 3.9.1' database with the IAM model `REMIND' and the RCP `PkBudg500' in the years 2065 and 2100.
+Due to its length, an example of the terminal output of T-reX is included as a separate file in the supplementary material, named `T-reX\_example\_output.html', which can be opened in an internet browser.
+The example output is from the execution of T-reX on the `ecoinvent 3.9.1' database with the IAM model `REMIND' and the RCP `PkBudg500' in the years 2065 and 2100.
 The example output includes the execution of the main module which in turn executes the \texttt{future\_scenarios}, \texttt{explode\_database}, \texttt{search\_waste}, \texttt{search\_material}, \texttt{make\_custom\_database}, \texttt{method\_editor}, \texttt{exchange\_editor}, and \texttt{verify\_database} modules.
 
-Due to licensing restrictions, the manipulated ecoinvent databases produced by the T-reX tool are not included in the supplementary material.
-However, with a licence to the ecoinvent database, the user can easily reproduce the results of the T-reX tool by following the instructions in the documentation.
+Due to licensing restrictions, the manipulated ecoinvent databases produced by T-reX are not included in the supplementary material.
+However, with a licence to the ecoinvent database, the user can easily reproduce the results of T-reX by following the instructions in the documentation.
 

### 2024-03-30 latex_files/discussion.tex
auto-update

@@
-Given that waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints}, we consider it of high importance that they are included LCA accounting.
+Given that waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, we consider it of high importance that they are included LCA accounting.
 Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they generally contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
@@
 As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adaption of the future background inventories.
-In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts\citep{bisinella2024wastelca}.
+In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
 A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.

### 2024-03-30 latex_files/introduction.tex
auto-update

@@
-% "As discussed in the introduction 1 there are some existing material demand methods in the standard LCIA method
-% 264 sets, including the ‘crustal scarcity indicator’ (which provides only an aggregated, abstracted endpoint) (Arvidsson
-% 265 et al., 2020) and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental
-% 266 units) (Hauschild and Potting, 2004). In these methods, the material demand is calculated based on the total mass that
-% 267 is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring
-% 268 these materials from the biosphere into the technosphere."
 \subsection{Background}\label{sec:intro-background}
@@
 Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that waste footprints can have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Alas, in frequent, tragic failures of evironmental justice, it is often the most vunerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,pellow2023envjusticewaste}.
+Alas, in frequent, tragic, failures of evironmental justice, it is often the most vunerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
@@
 This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\cite{bisinella2024wastelca, laurenti2023wastefootprint}.
+Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}.
 The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
@@
 
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
+Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
 Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.

### 2024-03-30 latex_files/methodology.tex
auto-update

@@
 
-Developed in the Python programming language, T-reX extends the brightway LCA framework, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
+Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
 Additionally, the \texttt{wurst} package---which can expand entire databases into a list of exchanges---is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
@@
 T-reX is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
-Currently, it has been tested with all available system models of ecoinvent versions 3.5--3.10~\citep{ecoinvent2016version3}.
+Currently, it has been tested with all available system models of ecoinvent versions 3.5--3.10.
 
-T-reX can be used directly from the command line, or imported as a Python module, in which case, the user can access the individual functions and modules.
+T-reX can be used directly from the command line, or imported as a Python package, in which case, the user can access the individual functions and modules.
 In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the ecoinvent database.
-The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or to use a prospective database based on future scenarios by implementing \texttt{premise} with an included module.
+The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or to use a prospective database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.
 
@@
 
-T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated such as to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
+T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
 We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of a activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
@@
 In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material\autoref{sec:supplementary}.
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material~\autoref{sec:supplementary}.
 One advantage of T-reX is that it is able to identify waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
@@
 The identity of the materials considered and their categorical groupings are easily customisable by the user.
-The full list of 59 materials included in the default configuration is provided in the supplementary material\autoref{sec:supplementary}.
+The full list of 59 materials included in the default configuration is provided in the supplementary material~\autoref{sec:supplementary}.
 
@@
 
-As presented in the introduction of this study~\autoref{sec:intro-background,sec:intro-waste,sec:intro-material} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+As presented in the introduction of this study~\autoref{sec:intro-background},~\autoref{sec:intro-waste},~\autoref{sec:intro-material} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
@@
     \caption{Workflow of T-reX.
-Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.
-[TO BE IMPROVED!]}
+Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
     \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}
@@
 The source of the life cycle inventory data for this case study was the database ecoinvent, version 3.9.1 with the `cutoff' attributional system model.
-Additionally, T-reX was used to create prospective database sets using the \mbox{REMIND} model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
+Additionally, T-reX was used to create prospective database sets using the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}

### 2024-03-30 latex_files/results.tex
auto-update

@@
 An example of the output from the application of T-reX has been included in the supplementary material.
-The manipulated ecoinvent databases can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+The manipulated ecoinvent databases, which are the main product of T-reX can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 
-As described in \autoref{sec:method-case_study}, this case study calculated the waste and material footprints (with variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
+As described in~\autoref{sec:method-case_study}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
@@
 
+%!@@!!!!!!!!!!
 \autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' for the NMC 811 Li-ion battery at the global market in the ecoinvent database version 3.9.1
-\cbox{I won't write about these until I know what figures we want to show. We can explore them in ActivityBrowser and decide.}
 
-\textcolor{red}{[and decide which ones are most interesting (or will fit on a page.
-The most interesting are the ones that look like the flying spaghetti monster).
-I would probably make a new sankey (back to coding!), or edit the AB one in inkscape to make it understandable on the page.]}
-
+%!!!!!!!!!!!
 \begin{figure}[H]
@@
 
-\autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1.
+% \autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1.
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
-\end{figure}
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
+%     \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
+% \end{figure}
 
@@
 
+%!!!!
 \begin{figure}[H]
@@
 \end{figure}
+%!!!!
 
-The inclusion of carbon capture and storage (CCS) in the prospective databases using the PkBudg500 RCP is evident in \autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
-This result highlights the fact that (frequently) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
-The actual deployment of these technologies---approximately 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}---falls far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
+The inclusion of carbon capture and storage (CCS) in the prospective databases using the PkBudg500 RCP is evident in~\autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
+This result highlights the fact that the (often) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
+The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
 
+%!!!!!!
 \begin{figure}[H]
@@
 
-For the phosphate demand footprints that are depicted in \autoref{fig:phosphate}, the LFP (lithium iron phosphate) battery has a much larger footprint than the other batteries, consistent with its composition.
+For the phosphate demand footprints that are depicted in~\autoref{fig:phosphate}, the LFP (lithium iron phosphate) battery has a much larger footprint than the other batteries, consistent with its composition.
 In this case, the phosphate footprint of all batteries is shown to decrease over the period from 2020--2100, and the RCP scenarios are seen to converge between 2020 and 2040.
@@
 
+%!!!!
+
 \subsubsection{Contribution of industrial sectors in the supply chain}\label{sec:results-case_study-topsectors}
@@
 
-\subsubsection{Comparison with `similar' methods}\label{sec:results-case_study-methodcomparison}
+\subsubsection{Comparison with `similar' LCIA methods}\label{sec:results-case_study-methodcomparison}
 
-A comparison of the results from T-reX's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in \autoref{fig:comparison_methods}.
+A comparison of the results from T-reX's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in~\autoref{fig:comparison_methods}.
 In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
@@
 \autoref{tab:results-case_study-comparison} presents a comparison of the results from the two studies, where possible.
-For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste, Laurenti et al.
-reported 95\% of the total waste, whereas T-reX reported only 3\%.
+For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste fraction, Laurenti et al.
+reported 95\%, whereas T-reX reported only 3\%.
 The reason for this discrepancy is explained by the fact that in the method of Laurenti et al., a ``waste flow was regarded as hazardous when the hazardousness
@@
 The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
-In the T-reX method, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+In T-reX , the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
 

### 2024-04-01 latex_files/cover_letter.tex
update

@@
 
-We are pleased to submit our original research article titled ``T-reX: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
+We are pleased to submit our original research article titled ``T-reX\: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
 This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.
@@
 
-The manuscript complies with the journal's requirements for an original research article, having a word count of [Insert Word Count], including [Insert Number of Figures] figures and [Insert Number of Tables] tables.
+The manuscript complies with the journal's requirements for an original research article, having a word count of 4969, including four figures and two tables.
 
@@
     \item {Valentina Bisinella,  DTU --- Technical University of Denmark, valenb@dtu.dk}
-    \item {Reviewer 3 Name, Affiliation, Email Address}
+    \item {Chris Mutel, Cauldron Solutions and brightway, cmutel@gmail.com}
 \end{enumerate}

### 2024-04-01 latex_files/introduction.tex
update

@@
 
-\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and `circularity'.
+~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and `circularity'.
 In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health.

### 2024-04-01 latex_files/main.tex
auto-update
update

@@
 
-\linenumbers
-% add something about the future stuffg
+\linenumbers{}
+
 % SPOTLIGHTS (5 bullet points, max 120 characters each)
+
 {\Large Spotlights}
 \vspace{1em}
-\begin{description}[style=nextline]
-    \item[Bullet 1: Critical context and background information on the problem addressed] Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
-    \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
-    \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use
-    \item[Bullet 4: The significance of the results to the environment, economics, or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains
-    \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting
-\end{description}
+\begin{itemize}
+    % Bullet 1: Critical context and background information on the problem addressed
+    \item Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
+    % Bullet 2: A brief overview of the key finding of the study (or findings if necessary)
+    \item A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
+    % Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript
+    \item Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use
+    % Bullet 4: The significance of the results to the environment, economics, or society
+    \item The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains
+    % Bullet 5: Future vision or the most important implications for continued research
+    \item Improved data availability and quality would enable more detailed and accurate waste and material footprinting
+\end{itemize}
 
@@
 \begin{graphicalabstract}
-    \cbox{to be improved :D}
     \includegraphics{grabs.jpg}
@@
     \begin{abstract}
-        \cbox{Abstract word count: 155, Limit: 150}
+        % \cbox{Abstract word count: 155, Limit: 150}
         The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
@@
     \begin{keyword}
-        circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
+        circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
     \end{keyword}
@@
     \centering
-    \caption{List of abbreviations}
-    \label{tab:abbreviations}
+    \caption{List of abbreviations}\label{tab:abbreviations}
     \begin{tabular}{ll}
@@
         Li-ion        & Lithium ion                                                         \\
-        LiMn$_2$O$_4$ & Lithium Manganese Oxide                                             \\
+        LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
         MF            & Material Footprint                                                  \\
@@
 %% main text
-\section{Introduction}
-\label{sec:introduction}
-\cbox{Section word count: 1700}
+\section{Introduction}\label{sec:introduction}
+%\cbox{Section word count: 2000}
 \input{introduction.tex}
 
-\section{Methodology}
-\label{sec:methodology}
-\cbox{Section word count: 1600}
+\section{Methodology}\label{sec:methodology}
+%\cbox{Section word count: 1700}
 \input{methodology.tex}
 
-\section{Results}
-\cbox{Section word count: 1300}
-\label{sec:results}
+\section{Results}\label{sec:results}
+% \cbox{Section word count: 1200}
 \input{results.tex}
 
-\section{Discussion}
-\label{sec:discussion}
-\cbox{Section word count: 450}
+\section{Discussion}\label{sec:discussion}
+% \cbox{Section word count: 450}
 \input{discussion.tex}
 
-\section{Conclusions}
-\label{sec:conclusions}
-\cbox{Section word count: 150}
+\section{Conclusions}\label{sec:conclusions}
+% \cbox{Section word count: 150}
 
@@
 
-The authors would like to acknowledge XXXX, XXXX, XXXX, XXX, who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive opensource software framework for the study and development of secondary raw material systems.
+The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive opensource software framework for the study and development of secondary raw material systems.
 The authors hope to continue the development of T-reX toward this vision.

### 2024-04-01 latex_files/methodology.tex
update

@@
 This section is divided into two parts.
-In \autoref{sec:method-T-reX_tool}, we describe the T-reX, and in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
+In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-case-study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
 
-\subsection{T-reX}\label{sec:method-T-reX_tool}
+\subsection{T-reX}\label{sec:method-TreX}
 
@@
 Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the 'pseudo- LCIA' methods created by T-reX can be used in the accustomed way.
+T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the `pseudo- LCIA' methods created by T-reX can be used in the accustomed way.
 T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
@@
 If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models)despite the presence of nickel in the final products.
+In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
 
@@
 
-This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere database(s) that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere databases that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
 
@@
 Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}
-    \label{fig:methods-flowchart}
+    \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure}
 
-\subsection{Case study methodology}\label{sec:method-case_study}
+\subsection{Case study methodology}\label{sec:method-case-study}
 

### 2024-04-01 latex_files/results.tex
update

@@
 
-As described in~\autoref{sec:method-case_study}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
+As described in~\autoref{sec:method-case-study}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.

### 2024-04-01 latex_files/introduction.tex
update

@@
 The development of a `circular economy' has become a critical area of focus in the imperative pursuit of sustainability objectives and curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair,eu2015reman}.
+Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
 In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
@@
 
-The incipient Waste Footprint (WF) metric, in contrast, though emerging steadily, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities~\citep{laurenti2016wastefootprint}.
+The incipient Waste Footprint (WF) metric, in contrast, though emerging steadily, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities---~\citep{laurenti2016wastefootprint,demirer2019wastefootprint}.
 Despite the fact that reducing life cycle waste is clearly critical to the development of the circular economy~\cite{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent evironmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.

### 2024-04-02 latex_files/discussion.tex
add casestudy data

@@
 T-reX advances the state-of-the-art in LCA providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
-Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
+Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
 
@@
 
-First, by adjusting the user configuration for future scenarios and waste/material categorisation we produced a set of customised versions of current and prospective ecoinvent databases along with a `pseudo-biosphere' T-reX database.
-Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for a number of present and future supply chains were trivial to calculate.
+First, by adjusting the user configuration for future scenarios and waste/material categorisation we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
+Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate.
 Additionally, visual exploration in \texttt{ActivityBrowser} was possible because the T-reX `pseudo-LCIA' methods are integrated within the \texttt{brightway} project in the same way as standard LCIA methods.
@@
 Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
-A methodic classification of waste exchanges and the end-of-life fates is expected to be facilitated by the ever more detailed and disaggregated data that is seen in each successive release of ecoinvent~\citep{fitzgerald2023ecoinventdocumentation}.
+A methodic classification of waste exchanges and the end-of-life fates is expected to be facilitated by the ever more detailed and disaggregated data that is seen in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
 

### 2024-04-02 latex_files/introduction.tex
add casestudy data

@@
 
-The development of a `circular economy' has become a critical area of focus in the imperative pursuit of sustainability objectives and curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+The development of a `circular economy' has become a central focus in the frantic pursuit of sustainability objectives and the imperative curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
@@
 Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
-It has been shown that MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
+It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
 
-The incipient Waste Footprint (WF) metric, in contrast, though emerging steadily, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities---~\citep{laurenti2016wastefootprint,demirer2019wastefootprint}.
-Despite the fact that reducing life cycle waste is clearly critical to the development of the circular economy~\cite{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent evironmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that waste footprints can have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Alas, in frequent, tragic, failures of evironmental justice, it is often the most vunerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the incipient `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
+While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
+Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs can have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Alas, in frequent, tragic, failures of environmental justice, it is often the most vulnerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
 WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders.
-Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint}.
+Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
 
@@
 These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
-Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct prospective background databases in an effort to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
+Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
 
 Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP) in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
+These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
 
@@
 Moreover, from a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-Returning from the abstract and etherial to the economic viability of the waste processing sector.
+Returning from the abstract and ethereal to the economic viability of the waste processing sector.
 While critical here are energy costs and the relative prices of virgin and secondary materials, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows.
@@
 
-Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and predicting its life cycle externalities, but there remains a conspicuous knowledge gap regarding the waste footprints of human activities and their environmental impacts~\citep{laurenti2023wastefootprint}.
+Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities.
+There remains, however, a conspicuous knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
 
@@
 In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling the long-term impacts of disposal.
+In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling of the long-term impacts of disposal.
 
@@
 
-~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and `circularity'.
-In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health.
-The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting (as the authors themselves acknowledge).
+\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
+In that study, it was shown that the waste footprint correlates well with many other LCIA methods, particularly human health.
+The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, and not easily reproducible.
+As the authors acknowledge in their article, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
+In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
 The `T-reX' Python software tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA\@.
@@
 
-In the context of a mineral-hungry renewable energy transition and constant geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}.
+In the context of geo-political tension and a mineral-hungry `renewable energy transition', ever greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
 While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
@@
 A relatively new (2020) LCIA method, the Crustal Scarcity (CSI), was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
-The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg silicon equivalents per kg element and derived from crustal concentrations.
+The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations.
 CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
@@
 
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} LCA framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX, in full, stands for `Tool for re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases such as to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases such as to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
 While some methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
@@
 
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requsite data and impact modelling remains lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
+The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
 It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.

### 2024-04-02 latex_files/main.tex
add casestudy data

@@
 \usepackage{xurl}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{pdfpages}
@@
 
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
     % ------------------------------------------------------
@@
     % Bullet 1: Critical context and background information on the problem addressed
-    \item Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
+    \item Tracking waste and material flows is essential for developing supply chain resilience and the circular economy
           % Bullet 2: A brief overview of the key finding of the study (or findings if necessary)
-    \item A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
+    \item T-reX was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
           % Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript
-    \item Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use
+    \item T-reX is an adaptable and extensible open-source Python tool complementing the \texttt{brightway} LCA framework and \texttt{ActivityBrowser},
           % Bullet 4: The significance of the results to the environment, economics, or society
-    \item The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains
+    \item T-reX can be used to identify pre-consumer waste generation and material demand hotspots often hidden in supply chains
           % Bullet 5: Future vision or the most important implications for continued research
@@
 \begin{highlights}
-    \item T-reX, a new software tool for quantifying waste and material flows in LCA
-    \item Assesses supply risks by calculating demand for critical materials
-    \item Simplifies quantification of user-specified waste and material categories
-    \item Rapidly identifies waste and material demand hotspots
-    \item Presents a case study of five Li-ion battery supply chains
+    \item T-reX, a new open-source tool for quantifying waste and material flows in Life Cycle Assessment (LCA)
+    \item Assesses supply risks by calculating demand inventories for critical materials
+    \item Simplifies quantification of user-specified waste and material footprint categories
+    \item Rapidly identifies hotspots of waste generation and material demand in supply chains
+    \item Presents a case study of five lithium-ion batteries to demonstrate T-reX's utility
 \end{highlights}
@@
 
-    \title{T-reX: A tool to quantify waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
+    \title{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
 
-        A case study of five lithium ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
+        A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
 T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
@@
 \end{frontmatter}
-\cbox{Total word count: ca. 5200, Limit: 5000}
-\cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
-\cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V3/T-reX_manuscript_V3_SupplementaryMaterial.pdf}{here}}
-\cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}
+% \cbox{Total word count: ca. 5200, Limit: 5000}
+% \cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
+% \cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V3/T-reX_manuscript_V3_SupplementaryMaterial.pdf}{here}}
+% \cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}
 
@@
     \centering
-    \caption{List of abbreviations}\label{tab:abbreviations}
+    \caption{List of abbreviations and terminology}\label{tab:abbreviations}
     \begin{tabular}{ll}
@@
         LFP               & Lithium Iron Phosphate                                              \\
-        Li-ion        & Lithium ion                                                         \\
+        Li-ion            & Lithium-ion                                                         \\
         LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
@@
         ReCiPe            & A standard LCIA method set                                          \\
+        SDG               & Sustainable Development Goals                                       \\
         SSPs              & Shared Socioeconomic Pathways                                       \\
-        T-reX         & Tool for re-X in LCA                                                \\
+        T-reX             & The Tool for analysing re-X in LCA                                  \\
         UBP               & Umweltbelastungspünkte (Environmental Impact Points)                \\
@@
 \section{Methodology}\label{sec:methodology}
-%\cbox{Section word count: 1700}
+%\cbox{Section word count: 1600}
 \input{methodology.tex}
@@
 \section{Discussion}\label{sec:discussion}
-% \cbox{Section word count: 450}
+% \cbox{Section word count: 500}
 \input{discussion.tex}
@@
 
-We have created `T-reX', an extension to the \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database.
-It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom `pseudo' LCIA methods.
-These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods.
-T-reX can be easily customised by the user to calculate the footprints of other supply chain flows such as water, gas, and critical raw materials.
+We have created `T-reX', an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
+It `explodes' and reconstructs the database to identify waste and material exchanges, edits these, and writes matching custom `pseudo' LCIA methods.
+These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints can be calculated like existing LCIA methods (such as carbon dioxide emissions).
+T-reX can be easily customised by the user to calculate the footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials (CRMs).
 
-This work extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply chain risk and potential environmental damage.
+This work advances the state-of-the-art by introducing an instrument to help elucidate the complex relationships between material demand, waste generation, and environmental impacts.
+T-reX enables the detailed exploration of 'footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
+T-reX hopes to provide a step towards a better understanding of these factors that are so crucial for the attainment of `sustainability' and the aspired `circular economy'.
 
@@
 
-The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive opensource software framework for the study and development of secondary raw material systems.
-The authors hope to continue the development of T-reX toward this vision.
+The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems.
+The authors hope to continue the development of T-reX in the service of our shared ideals.
 
@@
     \item Example of the terminal output of T-reX
+    \item Example of the terminal output of the case study
     \item Complete tabulated results of the case study
-    \item Complete visualisations of the case study
+    \item Complete visualisations from the case study
 \end{enumerate}

### 2024-04-02 latex_files/methodology.tex
add casestudy data

@@
 This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-case-study}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
+In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
 
@@
 Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
-Additionally, the \texttt{wurst} package---which can expand entire databases into a list of exchanges---is used to facilitate database searching and data transformation at the exchange level~\citep{mutel2017wurst}.
+Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}.
 Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---after running T-reX, the manipulated databases and the `pseudo- LCIA' methods created by T-reX can be used in the accustomed way.
-T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is open source under the CC-0 licence.
+T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---and after running T-reX, the manipulated databases and the `pseudo-LCIA' methods created by T-reX can be used in the project in the accustomed way.
+T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is fully open-source under the CC-0 licence.
 The full source code for T-reX is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
-T-reX is designed to be used with ecoinvent databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
-Currently, it has been tested with all available system models of ecoinvent versions 3.5--3.10.
+T-reX is designed to be used with \texttt{ecoinvent} databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
+Currently, it has been tested with all available system models of \texttt{ecoinvent}versions 3.5--3.10.
 
 T-reX can be used directly from the command line, or imported as a Python package, in which case, the user can access the individual functions and modules.
-In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the ecoinvent database.
-The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or to use a prospective database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.
+In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database.
+The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or a `prospective' database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.
 
-The supplementary material\autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
+The supplementary material in \autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
 Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
@@
 T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
-We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of a activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
+We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of an activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
 
@@
 
-The deconstruction of the databases by T-reX into lists of exchanges allows the relevant material and waste flows to be identified and categorised by search functions.
+The deconstruction of the databases by T-reX into lists of exchanges allows the relevant flows of material and waste to be identified and categorised by the search functions.
 The queries facilitating this are tailored to the specific database and the user can easily modify them to suit their needs.
 The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand.
-The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
+The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a mirrored `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
 
 In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material~\autoref{sec:supplementary}.
-One advantage of T-reX is that it is able to identify waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
-Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary}.
+One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
+Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (except CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
 
@@
 The identity of the materials considered and their categorical groupings are easily customisable by the user.
-The full list of 59 materials included in the default configuration is provided in the supplementary material~\autoref{sec:supplementary}.
+The full list of 59 materials included in the default configuration is provided in the supplementary material in ~\autoref{sec:supplementary}.
 
 The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest rather than keywords in the exchange's name.
-A useful feature T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
+A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
 For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
@@
 
-As presented in the introduction of this study~\autoref{sec:intro-background},~\autoref{sec:intro-waste},~\autoref{sec:intro-material} there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+% As presented in~\autoref{sec:intro-background},~\autoref{sec:intro-waste}, and~\autoref{sec:intro-material} of the introduction, there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
+
 In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
 
-T-reX, in contrast, uses accounting for material demand (and waste generation) based on exchanges solely within the technosphere.
+T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere.
 This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
@@
 If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-In the results of the Li-ion battery case study in \autoref{sec:results-case_study}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
+In the results of the Li-ion battery case study in \autoref{sec:results-casestudy}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
 
-\subsubsection{Mathematical basis}
+% \subsubsection{Mathematical basis}
 
-\cbox{very nice formulae coming soon}
+% \cbox{very nice formulae coming soon from LL hopefully}
 
@@
 
-A generic flowchart of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material\autoref{sec:supplementary} contains a more detailed computational flowchart.
+An overview of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material in \autoref{sec:supplementary} contains a more detailed computational flowchart.
 
@@
     \caption{Workflow of T-reX.
-Application of T-reX to one or more LCA databases creates a brightway project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
+Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
     \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
@@
 
-\subsection{Case study methodology}\label{sec:method-case-study}
+\subsection{Case study methodology}\label{sec:method-casestudy}
 
@@
 
-The source of the life cycle inventory data for this case study was the database ecoinvent, version 3.9.1 with the `cutoff' attributional system model.
-Additionally, T-reX was used to create prospective database sets using the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
+The source of the life cycle inventory data for this case study was \texttt{ecoinvent}, version 3.9.1, with the `cutoff' attributional system model.
+Additionally, T-reX was used to create prospective database sets using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}
@@
 
-For each pathway, databases (at five year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.
+For each pathway, databases (at five-year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.
 
@@
 This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
-This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand.
+This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

### 2024-04-02 latex_files/results.tex
add casestudy data

@@
 
-An example of the output from the application of T-reX has been included in the supplementary material.
-The manipulated ecoinvent databases, which are the main product of T-reX can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+An example of the output from the application of T-reX has been included in the supplementary material\autoref{sec:supplementary}.
+The manipulated \texttt{ecoinvent} databases, which are the main product of T-reX can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
-\subsection{Case study: Li-ion batteries}\label{sec:results-case_study}
+\subsection{Case study: Li-ion batteries}\label{sec:results-casestudy}
 
-As described in~\autoref{sec:method-case-study}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
+As described in~\autoref{sec:method-casestudy}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
-This section includes some highlights of the results with full results included in the supplementary material\autoref{sec:supplementary}.
-Because the `pseudo LCIA' T-reX methods are integrated into the brightway project as if they were LCIA methods, the LCA calculations can be performed in the customary way.
+This section includes some highlights of the results with full results included in the supplementary material (\autoref{sec:supplementary}).
+Because the `pseudo-LCIA' methods created by T-reX are integrated into the user's \texttt{brightway} project as if they were standard LCIA methods, the footprint calculations can be performed in the customary way.
 In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
@@
 %!@@!!!!!!!!!!
-\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' for the NMC 811 Li-ion battery at the global market in the ecoinvent database version 3.9.1
+\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' for the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1
 
@@
     \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
+    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
 \end{figure}
 
-% \autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1.
+% \autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the \texttt{ecoinvent} database version 3.9.1.
 
@@
 %     \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-%     \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the ecoinvent database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
+%     \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
 % \end{figure}
 
-\subsubsection{Temporal and scenario variation in waste and material inventory footprints}\label{sec:results-case_study-total_footprints}
+\subsubsection{Temporal and scenario variation in waste and material inventory footprints}
+%\label{sec:results-case_study-total_footprints}
 
-\autoref{fig:waste_total} shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+%! total
+\autoref{fig:waste_total} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
 The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced.
 The LiMn2O4 battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery.
-In each case there was a slight downward trend in the waste footprints between 2020 and 2100.
-This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is included in the models over this time (which can large amounts of waste in its extraction and combustion phases).
+In each case, there was a slight downward trend in the waste footprints between 2020 and 2100.
+This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
 For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
 
+%! CO2
+The inclusion of carbon capture and storage (CCS) in the prospective databases is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
+This result highlights the fact that the (often) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
+The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
 %!!!!
+
 \begin{figure}[H]
     \centering
-    \includegraphics[width=0.7\linewidth]{figures/total_waste.png}
-    \caption{Total solid waste footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_total}
+    \includegraphics[width=0.7\linewidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
+The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
 \end{figure}
@@
 
-The inclusion of carbon capture and storage (CCS) in the prospective databases using the PkBudg500 RCP is evident in~\autoref{fig:carbondioxide}, which shows a rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
-This result highlights the fact that the (often) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
-The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
-
-%!!!!!!
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.7\linewidth]{figures/carbondioxide.png}
-    \caption{Carbon dioxide waste (from carbon capture and storage) footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
-[TO BE IMPROVED!]}\label{fig:carbondioxide}
-\end{figure}
-
-For the phosphate demand footprints that are depicted in~\autoref{fig:phosphate}, the LFP (lithium iron phosphate) battery has a much larger footprint than the other batteries, consistent with its composition.
-In this case, the phosphate footprint of all batteries is shown to decrease over the period from 2020--2100, and the RCP scenarios are seen to converge between 2020 and 2040.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.7\linewidth]{figures/phosphate.png}
-    \caption{Phosphate material demand footprints for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
-[TO BE IMPROVED!]}\label{fig:phosphate}
-\end{figure}
-
-\subsubsection{Contribution of `top-processes' in the supply chain}\label{sec:results-case_study-topprocesses}
+\subsubsection{Contribution of `top-processes' in the supply chain}%\label{sec:results-case_study-topprocesses}
 
@@
 This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model.
-The fractional contributions of the top processes remains relatively steady over the coming century in this case.
+The fractional contributions of the top processes remain relatively steady over the coming century in this case.
 
@@
 
-\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in the ecoinvent database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
+\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in the \texttt{ecoinvent} database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
 The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
 \autoref{tab:results-case_study-comparison} presents a comparison of the results from the two studies, where possible.
-For liquid, solid, and recycled waste, the results were closely aligned, however for hazardous waste fraction, Laurenti et al.
+For liquid, solid, and recycled waste, the results were closely aligned, however, for the hazardous waste fraction, Laurenti et al.
 reported 95\%, whereas T-reX reported only 3\%.
@@
 The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
-In T-reX , the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
 
@@
     \centering
-\caption{Comparison of the results from the T-reX battery case study (database: ecoinvent 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: ecoinvent 3.5, activity:`Li-ion').}
+    \caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}
     \label{tab:results-case_study-comparison}
@@
         \midrule
-Total solid waste (kg/kg)& 50.9 & 62.5 \\
+        Total solid waste (kg/kg)  & 50.9           & 62.5                     \\
         Total liquid waste (kg/kg) & 3.53           & 3.63                     \\

### 2024-04-02 00_JRR-submission_folder/T-reX_supplementary-material/T-reX_case-study/_05d_visualisation_bar_combine_top-processes.py

@@ -1,0 +1,109 @@
+import os
+import xml.etree.ElementTree as ET
+from tqdm import tqdm
+from svglib.svglib import svg2rlg
+from reportlab.graphics import renderPDF
+from reportlab.pdfgen import canvas
+from reportlab.pdfbase.ttfonts import TTFont
+from reportlab.pdfbase import pdfmetrics
+
+from _00_main import DIR_VISUALISATION
+
+# Register Libertine font
+font_path = "misc/LinLibertine_Rah.ttf"  # Change as needed
+pdfmetrics.registerFont(TTFont("Libertine", font_path))
+
+# Directory containing SVG files
+svg_dir = DIR_VISUALISATION / "individual-methods_top-processes"
+print(f'{"*"*50}')
+print(f"Reading SVG files from directory: {svg_dir}")
+print(f'{"*"*50}')
+
+# Define the layout and title
+filename = str(DIR_VISUALISATION / "combined_methods_top-processes.pdf")
+title = "T-reX: Contributions to total LCIA score from top five processes"
+columns = 5
+figures = sorted([f for f in os.listdir(svg_dir) if f.endswith(".svg")], reverse=True)
+figures_count = len(figures)
+rows = figures_count // columns + (1 if figures_count % columns else 0)
+
+if figures_count == 0:
+    raise ValueError("No SVG files found in the directory.")
+else:
+    print(f"Found {figures_count} SVG files.")
+
+# # Initialize variables to store the maximum dimensions
+# max_width = 205
+# max_height = 532
+
+max_width = 0
+max_height = 0
+
+# Define margin size
+margin = 2  # You can adjust this value
+
+# Iterate through each SVG file to find max dimensions
+progress_bar = tqdm(total=len(figures), desc="Calculating maximum dimensions")
+for svg_file in figures:
+    file_path = os.path.join(svg_dir, svg_file)
+    svg_obj = svg2rlg(file_path)
+    max_width = max(max_width, svg_obj.width)
+    max_height = max(max_height, svg_obj.height)
+    min_width = min(max_width, svg_obj.width)
+    min_height = min(max_height, svg_obj.height)
+    progress_bar.update(1)
+progress_bar.close()
+
+print(f'{"*"*50}')
+print(f"Maximum width: {round(max_width,1)}, Maximum height: {round(max_height,1)}")
[…]

### 2024-04-02 latex_files/compile_pdf.sh

@@ -1,0 +1,43 @@
+#!/bin/bash
+
+# Define paths and filenames
+source_folder="." # Assuming your LaTeX files are in the current directory
+target_folder="../00_JRR-submission_folder"
+main_tex_file="main.tex" # Name of your main LaTeX file
+bib_file="references.bib" # Name of your bibliography file
+output_pdf_name="00_T-reX_manuscript.pdf" # Desired name for the output PDF
+echo "======================================================="
+echo "Compiling $main_tex_file to $output_pdf_name"
+# Create target folder if it doesn't exist
+mkdir -p "$target_folder"
+echo "======================================================="
+latexmk -c $main_tex_file || exit 1
+latexmk -pdf $main_tex_file || exit 1
+
+# Clean up before compiling
+# pdflatex -c "$main_tex_file" || exit 1
+
+# # Compile LaTeX document
+# pdflatex -interaction=batchmode -file-line-error "$main_tex_file"
+# bibtex "${main_tex_file%.*}" # Run bibtex if needed
+# pdflatex -interaction=batchmode -file-line-error "$main_tex_file"
+# pdflatex -interaction=batchmode -file-line-error "$main_tex_file" # Run twice for references
+
+# Check if PDF was generated
+if [ ! -f "${main_tex_file%.*}.pdf" ]; then
+    echo "PDF generation failed."
+    exit 1
+fi
+
+# Move the PDF and .bib to the target folder
+cp "${main_tex_file%.*}.pdf" "$target_folder/$output_pdf_name"
+cp "$source_folder/$bib_file" "$target_folder/T-reX_references.bib"
+
+# Clean up temporary files
+rm -f *.aux *.log *.bbl *.blg *.idx *.ind *.lof *.lot *.out *.toc *.acn *.acr *.alg *.glg *.glo *.gls *.ist
+
+echo "Process completed successfully."
+echo "PDF saved to $target_folder/$output_pdf_name"
+echo "Bibliography saved to $target_folder/T-reX_references.bib"
+open "$target_folder/$output_pdf_name" # Open the PDF
+echo "======================================================="

### 2024-04-02 latex_files/methodology.tex
update

@@
 
-In addition to the waste categories, T-reX includes numerous material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}.
-The CRM list contains 30 materials that are considered critical to the EU economy and are at risk of supply disruption.
-Several further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas.
-The identity of the materials considered and their categorical groupings are easily customisable by the user.
+In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
+Especially valuable in the case of CRMs is T-reX ability to uncover and quantify these threads of background consumption, which can be inconspiciusly embededded in products or consumed in seemingly distant supply chain activities.
+The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.
 The full list of 59 materials included in the default configuration is provided in the supplementary material in ~\autoref{sec:supplementary}.
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+    \includegraphics[width=14cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure}
@@
     \item Li-ion, LFP, rechargeable, prismatic
-    \item Li-ion, LiMn2O4, rechargeable, prismatic
+    \item Li-ion, LiMn\(_2\)O\(_4\), rechargeable, prismatic
     \item Li-ion, NMC111, rechargeable, prismatic

### 2024-04-02 latex_files/results.tex
update

@@
     \centering
-    \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
+    \includegraphics[width=19cm]{figures/T-reX_LFP_chromium.png}
     \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
@@
 The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced.
-The LiMn2O4 battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery.
+The  LiMn\(_2\)O\(_4\) battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery.
 In each case, there was a slight downward trend in the waste footprints between 2020 and 2100.
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+    \includegraphics[width=19cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
     \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
@@
 
-\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the cobalt footprint of the LiMn2O4 battery under the baseline scenario from 2020--2100.
+\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the graphite footprint of the  LiMn\(_2\)O\(_4\) battery under the baseline scenario from 2020--2100.
 The total footprint is seen to almost triple, from 2.2~kg/kg in 2020 to 6.2~kg/kg in 2100.
@@
     \centering
-    \includegraphics[width=0.6\linewidth]{figures/top-processes.png}
-    \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
-[TO BE IMPROVED!]}\label{fig:top_contribution}
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
+    \caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.6\linewidth]{figures/cpc_contribution.png
-    }
-    \caption{Contribution of industrial sectors to the liquid waste footprint of the NCA battery from 2020 to 2100 under the SSP2 scenario using the PkBudg500 RCP of the REMIND model.
-[TO BE IMPROVED!]}\label{fig:cpc_contribution}
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
+    \caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
 \end{figure}
@@
 \centering
-    \begin{subfigure}[b]{0.8\linewidth}
-        \includegraphics[width=\linewidth]{figures/comparison_methods2.png}
-        \caption{}\label{fig:comparison_methods_a}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.8\linewidth}
-        \includegraphics[width=\linewidth]{figures/comparison_methods1.png}
-        \caption{}\label{fig:comparison_methods_b}
-    \end{subfigure}
-    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `T-reX - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.[TO BE IMPROVED!]}\label{fig:comparison_methods}
+\includegraphics[width=18cm]{figures/T-reX-coalANDedip.pdf}
+\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
 \end{figure}

### 2024-04-02 latex_files/main.tex
update

@@
 This work advances the state-of-the-art by introducing an instrument to help elucidate the complex relationships between material demand, waste generation, and environmental impacts.
-T-reX enables the detailed exploration of 'footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
+T-reX enables the detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
 T-reX hopes to provide a step towards a better understanding of these factors that are so crucial for the attainment of `sustainability' and the aspired `circular economy'.

### 2024-04-02 latex_files/methodology.tex
update

@@
 In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
-Especially valuable in the case of CRMs is T-reX ability to uncover and quantify these threads of background consumption, which can be inconspiciusly embededded in products or consumed in seemingly distant supply chain activities.
+Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspiciusly embededded in products or consumed in seemingly distant supply chain activities.
 The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.

### 2024-04-02 latex_files/results.tex
update

@@
     \centering
-    \includegraphics[width=19cm]{figures/T-reX_LFP_chromium.png}
+    \includegraphics[width=16cm]{figures/T-reX_LFP_chromium.png}
     \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
@@
 %! total
-\autoref{fig:waste_total} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+\autoref{fig:waste_totalANDcarbon} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
 The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced.
@@
     \centering
-    \includegraphics[width=19cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+    \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
     \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
@@
     \centering
-\includegraphics[width=18cm]{figures/T-reX-coalANDedip.pdf}
+    \includegraphics[width=16cm]{figures/T-reX-coalANDedip.pdf}
     \caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
@@
 
-\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of ~1400 activities in the \texttt{ecoinvent} database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
+\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of \~1400 activities in the \texttt{ecoinvent} database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
 The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
@@
 reported 95\%, whereas T-reX reported only 3\%.
-The reason for this discrepancy is explained by the fact that in the method of Laurenti et al., a ``waste flow was regarded as hazardous when the hazardousness
+The reason for this discrepancy is explained by the fact that in the method of Laurenti et al.\, a ``waste flow was regarded as hazardous when the hazardousness
 was clearly stated in its subsequent waste treatment activity''.

### 2024-04-08 latex_files/discussion.tex
word_slash_LL

@@
-Given that waste generation and material demand are often strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, we consider it of high importance that they are included LCA accounting.
-Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they generally contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
+Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting.
+Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
 
-T-reX advances the state-of-the-art in LCA providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
+T-reX advances the state-of-the-art in LCA by providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
 Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
 
-The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility as well as some limitations of T-reX.
+The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility, and limitations of T-reX.
 
-First, by adjusting the user configuration for future scenarios and waste/material categorisation we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
+By adjusting the user configuration for future scenarios and waste/material categorisation, we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
 Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate.
-Additionally, visual exploration in \texttt{ActivityBrowser} was possible because the T-reX `pseudo-LCIA' methods are integrated within the \texttt{brightway} project in the same way as standard LCIA methods.
+Additionally, visual exploration in \texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are structurally identical to standard LCIA methods when integrated within the \texttt{brightway} project.
 
-One current limitation of T-reX is that it does not yet provide specific information (in a readily accessible format) on the composition of the waste generated.
-This is the information that would be needed to thoroughly assess the potential environmental impacts of this waste.
+One limitation of T-reX is that it does not yet provide specific information in a readily accessible format on the composition of the waste generated.
+This information is required to thoroughly assess the potential environmental impacts of this waste.
 Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
-A methodic classification of waste exchanges and the end-of-life fates is expected to be facilitated by the ever more detailed and disaggregated data that is seen in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
+A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
 
 The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
-As demonstrated in the results of the case study---where there was often very little scenario-temporal change in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adaption of the future background inventories.
+As demonstrated in the results of the case study---where there were often very few scenario-temporal changes in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adoption of future background inventories.
 In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.

### 2024-04-08 latex_files/introduction.tex
word_slash_LL

@@
 
-The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' developed strongly and has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
-More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is more often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
+The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
+More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
 Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
@@
 
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the incipient `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
+Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
 While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs can have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Alas, in frequent, tragic, failures of environmental justice, it is often the most vulnerable communities who suffer disproportionately from the social and ecological impacts that can be created by wasteful behaviours and sub-standard end-of-life treatment practices~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment ~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
@@
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate these impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
+LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
@@
 Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time.
-Moreover, from a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-Returning from the abstract and ethereal to the economic viability of the waste processing sector.
-While critical here are energy costs and the relative prices of virgin and secondary materials, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows.
-Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and that `modelling [of its management] must be physically based'.
+From a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
+From a practical economic viewpoint, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows, in addition to energy costs and the relative prices of virgin and secondary materials.
+Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and `modelling [of its management] must be physically based'.
 The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.
@@
 Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities.
-There remains, however, a conspicuous knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
+There remains, however, a knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
 
@@
 Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research and studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
+The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
+Studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
 Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
-The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
+The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
 
@@
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible ways to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
+Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible methods to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
 
 \cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-In that study, it was shown that the waste footprint correlates well with many other LCIA methods, particularly human health.
 The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, and not easily reproducible.
-As the authors acknowledge in their article, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
+As the authors acknowledge, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
 In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
@@
 
-A relatively new (2020) LCIA method, the Crustal Scarcity (CSI), was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
+The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
 The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations.
@@
 
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases such as to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
-While some methods with similar aims exist, they lack customisability and specificity~\citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting~\citep{laurenti2023wastefootprint}.
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
 

### 2024-04-08 latex_files/main.tex
word_slash_LL

@@
     \begin{abstract}
-        % \cbox{Abstract word count: 155, Limit: 150}
-        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
-        Thus, identifying and quantifying waste and material flows are of fundamental importance.
-Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of activities, where the implementation of circular principles could be most effective.
+        % \cbox{Abstract word count: 150, Limit: 150}
+        A fundamental principle of ‘circular economy’ is to minimise material consumption and waste generation.
+Thus, identifying and quantifying waste and material flows is critical.
+Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity’s life cycle, where the implementation of circular principles could be most effective.
 
-        Introducing T-reX, a Python tool extending the brightway ecosystem, allowing facile quantification of user-defined supply chain demands in current and prospective scenarios.
+        We introduce T-reX, a Python tool extending the brightway ecosystem, to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
 
-        A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
+        A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
+T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
 
@@
 \section{Introduction}\label{sec:introduction}
-%\cbox{Section word count: 2000}
+%\cbox{Section word count: 2000 ; current: 1990}
 \input{introduction.tex}
@@
 \section{Methodology}\label{sec:methodology}
-%\cbox{Section word count: 1600}
+%\cbox{Section word count: 1600 ; current: 1593}
 \input{methodology.tex}
@@
 \section{Results}\label{sec:results}
-% \cbox{Section word count: 1200}
+% \cbox{Section word count: 1200 ; current: 1151}
 \input{results.tex}
@@
 \section{Discussion}\label{sec:discussion}
-% \cbox{Section word count: 500}
+% \cbox{Section word count: 500 ; current: 471}
 \input{discussion.tex}
@@
 \section{Conclusions}\label{sec:conclusions}
-% \cbox{Section word count: 150}
+% \cbox{Section word count: 150 ; current: 150}
 
-We have created `T-reX', an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It `explodes' and reconstructs the database to identify waste and material exchanges, edits these, and writes matching custom `pseudo' LCIA methods.
-These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints can be calculated like existing LCIA methods (such as carbon dioxide emissions).
-T-reX can be easily customised by the user to calculate the footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials (CRMs).
+T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
+It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
+These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
 
-This work advances the state-of-the-art by introducing an instrument to help elucidate the complex relationships between material demand, waste generation, and environmental impacts.
-T-reX enables the detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX hopes to provide a step towards a better understanding of these factors that are so crucial for the attainment of `sustainability' and the aspired `circular economy'.
+T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts.
+This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
+T-reX is a step towards a better understanding of these interconnections for the practical realization of `sustainability' and the `circular economy'.
 

### 2024-04-08 latex_files/methodology.tex
word_slash_LL

--- 2024-04-02 latex_files/methodology.tex
@@
 
-T-reX can be used directly from the command line, or imported as a Python package, in which case, the user can access the individual functions and modules.
+T-reX can be used directly from the command line, or imported as a Python package where the user can access the individual functions and modules.
 In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database.
@@
 In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
-Especially valuable in the case of CRMs is T-reX ability to uncover and quantify these threads of background consumption, which can be inconspiciusly embededded in products or consumed in seemingly distant supply chain activities.
+Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
 The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.
@@
 
-The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the so-called relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
 A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.

### 2024-04-08 latex_files/results.tex
word_slash_LL

@@
 Such comparability was also observed for other fossil-fuel-related methods (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
-Correlation between standard LCIA methods and T-reX methods, is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
+Correlation between standard LCIA methods and T-reX methods is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
 In standard LCIA methods, impact scores are derived exclusively from the magnitude of the exchanges between the biosphere and the technosphere, that is, extraction and emission.

### 2024-04-10 latex_files/introduction.tex
accept LL changes

@@
 
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
+Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the incipient `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
 While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment ~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 

### 2024-04-10 latex_files/main.tex
accept LL changes

@@
         % \cbox{Abstract word count: 150, Limit: 150}
-        A fundamental principle of ‘circular economy’ is to minimise material consumption and waste generation.
+        The quintessential principle of the ‘circular economy’ is to minimise material consumption and waste generation.
 Thus, identifying and quantifying waste and material flows is critical.
@@
 This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realization of `sustainability' and the `circular economy'.
+T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
 

### 2024-04-10 latex_files/main.tex
update-pdf

@@
 %% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
 %% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
@@
 \begin{document}
+\vspace{20m}
+\includegraphics[width=10cm]{graphical-abstract.pdf}
 
+\clearpage
+% \newpage
 \linenumbers{}
@@
 
-\newpage
 %% GRAPHICAL ABSTRACT
-\begin{graphicalabstract}
-    \includegraphics{grabs.jpg}
-\end{graphicalabstract}
+% \begin{graphicalabstract}
+%     \includegraphics[width=8cm]{graphical-abstract.pdf}
+% \end{graphicalabstract}
+%% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
 
-%% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
+\vspace{4em}
+
 \begin{highlights}
-    \item T-reX, a new open-source tool for quantifying waste and material flows in Life Cycle Assessment (LCA)
+    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
     \item Assesses supply risks by calculating demand inventories for critical materials
@@
         % \cbox{Abstract word count: 150, Limit: 150}
-        The quintessential principle of the ‘circular economy’ is to minimise material consumption and waste generation.
+        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
 Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity’s life cycle, where the implementation of circular principles could be most effective.
+Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
 
@@
 \end{frontmatter}
-% \cbox{Total word count: ca. 5200, Limit: 5000}
-% \cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
-% \cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V3/T-reX_manuscript_V3_SupplementaryMaterial.pdf}{here}}
-% \cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}
 
@@
     \centering
-    \caption{List of abbreviations and terminology}\label{tab:abbreviations}
+    \caption*{List of abbreviations and terminology}\label{tab:abbreviations}
     \begin{tabular}{ll}
@@
 \section{Introduction}\label{sec:introduction}
-%\cbox{Section word count: 2000 ; current: 1990}
 \input{introduction.tex}
@@
 \section{Methodology}\label{sec:methodology}
-%\cbox{Section word count: 1600 ; current: 1593}
 \input{methodology.tex}
@@
 \section{Results}\label{sec:results}
-% \cbox{Section word count: 1200 ; current: 1151}
 \input{results.tex}
@@
 \section{Discussion}\label{sec:discussion}
-% \cbox{Section word count: 500 ; current: 471}
 \input{discussion.tex}
@@
 \section{Conclusions}\label{sec:conclusions}
-% \cbox{Section word count: 150 ; current: 150}
 
@@
 \section*{Data availability}
-All data used in this analysis are publicly available online under the noted sources.T-reX is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for T-reX is available at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.
+All data used is publically available online under the noted sources.
+T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
+It installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
+The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}.
+A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
 
@@
 
-\section*{Supplementary material}\label{sec:supplementary}
-The supplementary material contains the following:
+\section{Supplementary material}\label{sec:supplementary}
+The supplementary material supplied with this manuscript contains the following:
 \begin{enumerate}
@@
 % \appendix
-
 % \section{Example output}

### 2024-04-10 latex_files/.compile_pdf.sh
README update

@@ -3,41 +3,46 @@
 # Define paths and filenames
 source_folder="." # Assuming your LaTeX files are in the current directory
 target_folder="../00_JRR-submission_folder"
-main_tex_file="main.tex" # Name of your main LaTeX file
-bib_file="references.bib" # Name of your bibliography file
-output_pdf_name="00_T-reX_manuscript.pdf" # Desired name for the output PDF
+main_tex_file="main.tex"                    # Name of your main LaTeX file
+bib_file="references.bib"                   # Name of your bibliography file
+output_pdf_name="00_T-reX_manuscript.pdf"   # Desired name for the output PDF
+output_docx_name="00_T-reX_manuscript.docx" # Desired name for the output DOCX
+output_html_name="00_T-reX_manuscript.html" # Desired name for the output HTML
+
 echo "======================================================="
-echo "Compiling $main_tex_file to $output_pdf_name"
+echo "Compiling $main_tex_file to $output_pdf_name, $output_docx_name, and $output_html_name"
+
 # Create target folder if it doesn't exist
 mkdir -p "$target_folder"
-echo "======================================================="
-latexmk -c $main_tex_file || exit 1
-latexmk -pdf $main_tex_file || exit 1
 
 # Clean up before compiling
-# pdflatex -c "$main_tex_file" || exit 1
+latexmk -c $main_tex_file || exit 1
 
-# # Compile LaTeX document
-# pdflatex -interaction=batchmode -file-line-error "$main_tex_file"
-# bibtex "${main_tex_file%.*}" # Run bibtex if needed
-# pdflatex -interaction=batchmode -file-line-error "$main_tex_file"
-# pdflatex -interaction=batchmode -file-line-error "$main_tex_file" # Run twice for references
+# Compile LaTeX document to PDF
+latexmk -pdf $main_tex_file || exit 1
 
 # Check if PDF was generated
 if [ ! -f "${main_tex_file%.*}.pdf" ]; then
-    echo "PDF generation failed."
-    exit 1
+	echo "PDF generation failed."
+	exit 1
 fi
+
+# Convert PDF to DOCX
+pandoc -s "${main_tex_file%.*}.pdf" -o "$target_folder/$output_docx_name" || exit 1
+
+# Convert PDF to HTML
[…]

### 2024-05-02 latex_files/introduction.tex
commit text changes

--- 2024-04-08 latex_files/introduction.tex
@@
-\subsection{Background}\label{sec:intro-background}
+\subsection{Introduction to T-reX}\label{sec:intro-trex}
 
-The development of a `circular economy' has become a central focus in the frantic pursuit of sustainability objectives and the imperative curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+Development of a `circular economy' has become a central focus in the frantic pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
@@
 More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
-Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
+Whether at the level of products, entire industries, nations, or even continents, the material footprint aims to quantify the total material consumed in supply chains.
 It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
 
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
+Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed.
 While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
 Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment ~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
@@
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
+LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
@@
 
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+
+Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
+Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
+Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.@\texttt{Premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
+The authors believe that there is an urgent need for research investment in this area.
+
+The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
+It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
+By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
+In any case, minimising primary material consumption and the generation of waste is of paramount importance.
+By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.
+
+\subsection{Background and need}\label{sec:intro-background}
 \subsubsection{Waste in LCA}\label{sec:intro-waste}
@@
 
-Existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste.
+Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA\citep{lauran2020abioticdepletion}.
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
@@
 \cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, and not easily reproducible.
-As the authors acknowledge, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
+The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' .
 In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
@@
 
-In the context of geo-political tension and a mineral-hungry `renewable energy transition', ever greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
+In the context of geopolitical tension and a mineral-hungry `renewable energy transition', greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
+While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system rather than the primary material flows themselves.
 
@@
 Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
-
-\subsection{Introduction to T-reX}\label{sec:intro-trex}
-
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
-
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
-Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Indeed, despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
-
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
-In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.

### 2024-05-02 latex_files/main.tex
update-pdf
commit text changes

@@
 \begin{document}
-\vspace{20m}
+\vspace{20em}
 \includegraphics[width=10cm]{graphical-abstract.pdf}
@@
 
-    \title{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
+    \title{T-reX\@: Quantifying waste and material footprints in current and future \\ Life Cycle Assessment (LCA) databases}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
 
-        We introduce T-reX, a Python tool extending the brightway ecosystem, to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
+        We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
@@
         A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
+T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
 
@@
 T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
-It installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
+It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
 The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}.

### 2024-05-02 latex_files/methodology.tex
accept LL changes
commit text changes

@@
 This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
+In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries, highlighting inconsistencies in LCA methodology and probing departures from projections in prospective models.
 
@@
 
-T-reX can be used directly from the command line, or imported as a Python package where the user can access the individual functions and modules.
+T-reX can be used directly from the command line, or imported as a Python package with which the user can access the individual functions and modules.
 In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database.
@@
 Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
-The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.
-The full list of 59 materials included in the default configuration is provided in the supplementary material in ~\autoref{sec:supplementary}.
+The identity of the materials considered, and their categorical groupings, are easily customisable by the user.
+The full list of 59 materials included in the default configuration is provided in the supplementary material in~\autoref{sec:supplementary}.
 
@@
 Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.
-
-% As presented in~\autoref{sec:intro-background},~\autoref{sec:intro-waste}, and~\autoref{sec:intro-material} of the introduction, there are some existing material demand methods in the standard LCIA method sets, including the CSI (which provides only an aggregated, abstracted endpoint in units of kg~Si~eq)~\citep{arvidsson2020csi} and the (deprecated) EDIP 2003 material use indicators (which provide endpoints in fundamental units)~\citep{hauschild2003edip}.
 
@@
     \centering
-    \caption{Workflow of T-reX.
+    \caption{The workflow of T-reX presented in a generalised (a) and computational (b) format.
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=14cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+    \includegraphics[width=9cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure}
@@
 
-We investigated five types of Li-ion batteries, each represented by their unamended global market activities:
+We investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
 \begin{itemize}[itemsep=0pt]
@@
 \begin{itemize}[itemsep=0pt]
-    \item ReCiPe 2016 v1.03, midpoint (I)
-    \item EF v3.0 no LT
-    \item EDIP 2003 no LT
-    \item CSI 2020
+    \item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
+    \item EF v3.0 no LT~\citep{eu2023ef}
+    \item EDIP 2003 no LT~\citep{hauschild2003edip}
+    \item CSI 2020~\citep{arvidsson2020csi}
 \end{itemize}
 
-The source of the life cycle inventory data for this case study was \texttt{ecoinvent}, version 3.9.1, with the `cutoff' attributional system model.
-Additionally, T-reX was used to create prospective database sets using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
+Additional to the present-day databases, T-reX created prospective database sets based on \texttt{ecoinvent} 3.9.1 using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}

### 2024-05-02 latex_files/results.tex
commit text changes

@@
 
-An example of the output from the application of T-reX has been included in the supplementary material\autoref{sec:supplementary}.
-The manipulated \texttt{ecoinvent} databases, which are the main product of T-reX can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}.
+The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 
-As described in~\autoref{sec:method-casestudy}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
+As described in~\autoref{sec:method-casestudy}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered background inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
 The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
-This section includes some highlights of the results with full results included in the supplementary material (\autoref{sec:supplementary}).
+This section includes some highlights of the results, with full results included in the supplementary material (\autoref{sec:supplementary}).
 Because the `pseudo-LCIA' methods created by T-reX are integrated into the user's \texttt{brightway} project as if they were standard LCIA methods, the footprint calculations can be performed in the customary way.
@@
 
-%!@@!!!!!!!!!!
-\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' for the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1
+\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+As the supply chains are expansive, this visualisation has been greatly simplified for print, the full version is included in the supplementary material~\autoref{sec:supplementary}.
+As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
+Due to the nature of LCA, the waste flows do not represent the `real' physical flows directly, but rather an abstraction thereof based on, in this case, economic allocation---the market price of the functional unit produced by the activity~\citep{guinee2004economicallocation}.
+If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and be more directly related to the physical flows of the activities.
 
-%!!!!!!!!!!!
 \begin{figure}[H]
     \centering
-    \includegraphics[width=16cm]{figures/T-reX_LFP_chromium.png}
-    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
+    \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
+    \caption{A simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
 \end{figure}
-
-% \autoref{fig:sankey_material} presents a Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the \texttt{ecoinvent} database version 3.9.1.
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=0.7\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
-%     \caption{Sankey diagram depicting the supply chain flows of elemental chromium for 1~kg of the LFP (lithium iron phosphate) battery in the case study at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_material}
-% \end{figure}
 
@@
 The  LiMn\(_2\)O\(_4\) battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery.
-In each case, there was a slight downward trend in the waste footprints between 2020 and 2100.
-This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
+In each case, there was only a slight downward trend in the waste footprints between 2020 and 2100.
+This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in fossil fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
 For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
@@
 The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
+The authors consider that the over-representation of CCS and the under-representation of waste processing technologies in the prospective databases is one significant limitation to the validity of prospective LCA databases.
+Significant work is needed to improve the representation of these technologies in the databases to ensure that the results are more realistic and useful for decision-making.
 %!!!!
@@
 \autoref{fig:top_contribution} shows the contribution of the `top-processes' to the graphite footprint of the  LiMn\(_2\)O\(_4\) battery under the baseline scenario from 2020--2100.
-The total footprint is seen to almost triple, from 2.2~kg/kg in 2020 to 6.2~kg/kg in 2100.
-This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model.
-The fractional contributions of the top processes remain relatively steady over the coming century in this case.
+The total footprint is seen to more than triple, from 0.16~kg/kg in 2020 to 0.52~kg/kg in 2050 and onward to 2100.
+The largest contributor is not, in this case, the battery-embedded graphite electrode (which is the second top process), but `silicon production' from further up in the supply chain.
+This result is likely a reflection of the electrification of the transport and energy sectors that is included in the REMIND model.
+The results of this kind of analysis with T-reX can be used to identify the most significant processes in the supply chain and to target interventions to reduce the potential environmental impact of production.
 
@@
 
-\autoref{fig:cpc_contribution} shows the contribution of sectors (grouped by CPC) to the total natural gas footprint of the LFP battery under the PkBudg500 pathway.
-In this example, the relative contribution of the sector `47160: Electronic integrated circuits' is seen to decrease from 11\% in 2020 to 6\% in 2100, while over the same period the contribution of the sector `46430: Parts of primary cells, primary batteries and electrodes' increases from 29\% to 38\%.
+\autoref{fig:cpc_contribution} shows the contribution of sectors (grouped by CPC codes) to the total natural gas footprint of the LFP battery under the PkBudg500 pathway.
+In this example, the relative contribution of the sector `47160: Electronic integrated circuits' is seen to decrease from 11\% in 2020 to 8\% in 2100, while over the same period the contribution of the sector `46430: Parts of primary cells, primary batteries and electrodes' increases from 32\% to 38\%.
 The method used to calculate these contributions involves traversing the supply chain branches to a certain level (max.
@@
 The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply chain pressures that are not captured by the standard LCIA methods.
-This makes the T-reX methods more sensitive to the modeling choices (e.g., allocation) that are generally embedded in LCA databases.
+This makes the T-reX methods more sensitive to the modeling choices (e.g., economic allocation) that are generally embedded in LCA databases.
 
@@
 
-\cite{laurenti2023wastefootprint} used an alternative method for calculating the waste footprint of \~1400 activities in the \texttt{ecoinvent} database version 3.5, which contains only a generic `market for battery, Li-ion, rechargeable, prismatic'.
+In the most similar published method, \cite{laurenti2023wastefootprint} used an alternative procedure for calculating the waste footprint of \~1400 activities in the \texttt{ecoinvent} database version 3.5.
+Direct comparison is not possible, as this database contains only one generic Li-ion battery, `market for battery, Li-ion, rechargeable, prismatic'.
 The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
@@
 reported 95\%, whereas T-reX reported only 3\%.
-The reason for this discrepancy is explained by the fact that in the method of Laurenti et al.\, a ``waste flow was regarded as hazardous when the hazardousness
-was clearly stated in its subsequent waste treatment activity''.
+The reason for this discrepancy is explained by the fact that in the method of Laurenti et al.\, a ``waste flow was regarded as hazardous when the hazardousness was clearly stated in its subsequent waste treatment activity''.
 The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
@@
     \centering
-    \caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}
-    \label{tab:results-case_study-comparison}
+    \caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}\label{tab:results-case_study-comparison}
     \begin{tabular}{lll}

### 2024-09-22 latex_files/.main.tex
update-manuscript-reflecting-elsevier_reviewer_suggestions.

--- 2024-04-10 latex_files/main.tex
@@
 \begin{document}
-\vspace{20m}
-\includegraphics[width=10cm]{graphical-abstract.pdf}
 
-\clearpage
-% \newpage
 \linenumbers{}
@@
 
-%% GRAPHICAL ABSTRACT
+% \newpage
+% %% GRAPHICAL ABSTRACT
 % \begin{graphicalabstract}
-%     \includegraphics[width=8cm]{graphical-abstract.pdf}
+%     \includegraphics{graphical-abstract.pdf}
 % \end{graphicalabstract}
+
 %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
@@
 \begin{highlights}
-    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
+    \item 'T-reX', a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
     \item Assesses supply risks by calculating demand inventories for critical materials
@@
 Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
+Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle.
+Allowing the implementation of circular principles where could be most effective.
 
-        We introduce T-reX, a Python tool extending the brightway ecosystem, to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
+        In this spirit, we introduce T-reX, a Python tool extending the brightway ecosystem that allows for facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
 
-        A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
+        A simple example of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints.
+In a context where the quantification of potential environmental burdens is uncertain and subjective, T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating the analysis of material consumption and waste generation in LCA.
 
@@
         SSPs              & Shared Socioeconomic Pathways                                       \\
-        T-reX             & The Tool for analysing re-X in LCA                                  \\
+        T-reX             & Tool for analysing re-X in LCA                                  \\
         UBP               & Umweltbelastungspünkte (Environmental Impact Points)                \\
@@
 T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
+It can take apart and reconstruct databases in order to identify and edit waste and material exchanges and write matching custom `pseudo' LCIA methods.
 These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
@@
 T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts.
-This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
+This instrument enables detailed exploration of footprints and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
+T-reX is a step towards a better understanding of these interconnections for the practical realisation of sustainability and the circular economy.
 
@@
 \section*{Data availability}
-All data used is publically available online under the noted sources.
+All data used in this analysis are publicly available online under the noted sources.
 T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
@@
 \end{enumerate}
+Further information and extended data is available in the T-reX package's GitHub repository at \url{https://www.github.com/Stew-McD/T-reX}.
 

### 2024-09-22 latex_files/introduction.tex
update-manuscript-reflecting-elsevier_reviewer_suggestions.

@@
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
-These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
+These scores can then be aggregated into one metric for each impact category that can be compared across products and processes.
 Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
@@
 To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
 
@@
 Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.@\texttt{Premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
+Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
+\texttt{premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
 The authors believe that there is an urgent need for research investment in this area.
@@
 In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.
+By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and potentially highlight opportunities for waste reduction and material efficiency.
 
@@
 
-Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA\citep{lauran2020abioticdepletion}.
+Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA~\citep{lauran2020abioticdepletion}.
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
@@
 
-\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
+~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
 The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' .

### 2024-09-22 latex_files/methodology.tex
update-manuscript-reflecting-elsevier_reviewer_suggestions.

@@
 
-Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
+Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising especially  the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
 Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}.
@@
 
-In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
+In the default configuration of T-reX, there are 10 waste categories which are then further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
 The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary}.
@@
 
-In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
+In addition to these waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
 Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
@@
 
-The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
 A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
@@
 
-T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere.
+T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere, a material basis.
 This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
@@
 
-We investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
+To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
 \begin{itemize}[itemsep=0pt]
@@
     \item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
-    \item EF v3.0 no LT~\citep{eu2023ef}
-    \item EDIP 2003 no LT~\citep{hauschild2003edip}
-    \item CSI 2020~\citep{arvidsson2020csi}
+    \item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
+    \item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
+    \item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
 \end{itemize}
@@
 \subsubsection{LCA calculations}
-For each combination of activity, method, and database, a single score was calculated along with details of the top contributing processes.
+For each combination of activity, method, and database, a unified metric was calculated along with details of the top contributing processes.
 Additionally, for the T-reX methods, a contribution analysis was performed.
-This involved utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
-This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
-This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
+Utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem, this function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
+The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
+T-reX is, an sich, agnostic to interests of its user, providing a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

### 2024-09-22 latex_files/results.tex
update-manuscript-reflecting-elsevier_reviewer_suggestions.

@@
 
-An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}.
+An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, subsection s.
 The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
@@
 
-\subsubsection{Sankey visualisation of flows in waste and material inventory footprints}\label{sec:results-case_study-sankey}
+\subsubsection{Sankey visualisation of flows inventory footprints}\label{sec:results-case_study-sankey}
 
-\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
-As the supply chains are expansive, this visualisation has been greatly simplified for print, the full version is included in the supplementary material~\autoref{sec:supplementary}.
+\autoref{fig:sankey_waste} presents a heavily simplified Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material~\autoref{sec:supplementary}.
 As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
 Due to the nature of LCA, the waste flows do not represent the `real' physical flows directly, but rather an abstraction thereof based on, in this case, economic allocation---the market price of the functional unit produced by the activity~\citep{guinee2004economicallocation}.
-If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and be more directly related to the physical flows of the activities.
+If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and would be more directly related to the physical flows of the activities.
 
@@
     \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
-    \caption{A simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
+    \caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
 \end{figure}

### 2024-09-22 reviews/external/JRC-review/T-reX_response-to-review_external_JRC.md
it-builds!

Dear Reviewers and Editor,

We would like to extend our sincere thanks for the time and effort you have invested in reviewing our manuscript on T-reX. We are particularly grateful for the detailed and constructive feedback provided. We have carefully addressed each of your comments in the responses below and have made the necessary revisions to the manuscript. We hope that you find the revised version satisfactory.

### Reviewer One

#### Comment:

“I think that this manuscript needs to be ameliorated.”

**Summary:**The reviewer requested more detail for the case study, as well as a sensitivity or uncertainty analysis.Key points raised:

- Comment 1: System boundary
- Comment 2: Input-output data
- Comment 3: Material footprint data (with more discussion, and a call for an ‘overall material footprint’)
- Comment 4: Sensitivity or uncertainty analysis

#### Author Responses:

##### Comment 1.1 – System Boundary:

_“The system boundary of the case study should be defined more clearly, possibly with diagrams.”_

**Response 1.1:**

(Methodology section 2.2)
We have used the most basic system boundary for this case study, since we are mearly trying to show what the programme can do, not really to provide particular results. We have added another sentence to define the system boundaries more clearly. That being only the Li-ion battery markets, as modelled in ecoinvent version 3.9.1.

##### Comment 1.2 – Input-Output Data:

_“No input-output table provided for the case studies. Please provide this inventory data in the manuscript.”_

**Response 1.2:**
For this example, described in the Methodology section 2.2, the inventory data is taken directly from the relevant activities in the ecoinvent database (version 3.9.1). No adaption was made by the authors to this data. This has been stated more clearly in the text.

##### Comment 1.3 – Material Footprint Data:

_“Results and Discussion section: the study presented some material footprint data. I recommend further discussion and interpretation of the results. Additionally, it is necessary to present the overall material footprint of the battery.”_

**Response 1.3:**
As detailed in the Methodology (section 2.1.1 and 2.1.2) Results (section 3.2), and Discussion sections (first two paragraphs), the material footprint results generated by T-reX are aggregations of the exchanges in the model system targeted by the user’s search criteria. In this case, the default selection includes XX materials and their flows as represented in ecoinvent. The authors believe that presenting a ‘total material footprint’ based on the sum of these individual materials would be misleading, as the LCA models represent only specific aspects of the system. A ‘total material footprint’ without consideration of the material type is not meaningful. One of T-reX’s strengths is its ability to allow users to focus on materials of interest. We have added clarification in the Discussion section (lines X-Y) to make this context clearer to the reader.

##### Comment 1.4 – Sensitivity and Uncertainty Analysis:

_“The authors should also perform a sensitivity analysis or uncertainty analysis.”_

**Response 1.4:**
The battery data used in this case study is taken directly from ecoinvent activities and serves primarily as a basic example to demonstrate the programme’s capabilities. Therefore, we believe that a sensitivity or uncertainty analysis on the case study results is unnecessary and could distract from the overall message of the paper. However, to ensure clarity, we have expanded the discussion on the limitations of the case study results mainto make this point more explicit to the reader.

### 2024-11-26 reviews/external/JRC-review/T-reX_response-to-review_external_JRC.md
working on revision

@@
 We have used the most basic system boundary for this case study, since we are mearly trying to show what the programme can do, not really to provide particular results.
-We have added another sentence to define the system boundaries more clearly.
-That being only the Li-ion battery markets, as modelled in ecoinvent version 3.9.1.
+We have added another sentence to define the system boundaries more clearly (XXX).
+That being that it encompasses exclusively the Li-ion battery markets, as modelled in ecoinvent version 3.9.1.
 
@@
 No adaption was made by the authors to this data.
-This has been stated more clearly in the text.
+However, this has been stated more clearly in the text (XXXXXXX).
 

### 2024-11-26 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
working on revision

%% 
%% Copyright 2019-2024 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.3c of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3c or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

\documentclass[a4paper,fleqn]{cas-dc}

% If the frontmatter runs over more than one page
% use the longmktitle option.

%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}

%\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[authoryear,longnamesfirst]{natbib}

%%%Author macros
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{SCM}
\tsc{EL}
\tsc{SC}
\tsc{CFB}
%%%

% Uncomment and use as if needed
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdefinition{rmk}{Remark}
%\newproof{pf}{Proof}
%\newproof{pot}{Proof of Theorem \ref{thm}}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

% Short title
\shorttitle{T-reX\@: Quantifying waste and material footprints in LCA}

% Short author
\shortauthors{S.C. McDowall et~al.}

% Main title of the paper
\title [mode = title]{T-reX\@: Quantifying waste and material footprints in current and future \\ Life Cycle Assessment (LCA) databases}

% Authors
\author[1]{Stewart Charles McDowall}
\ead{s.c.mcdowall@cml.leidenuniv.nl}
\cormark[1]
\cortext[cor1]{Corresponding author:}

\author[1]{Elizabeth Lanphear}

\author[1]{Stefano Cucurachi}

\author[1]{Carlos Felipe Blanco}

\affiliation[1]{
    organization={Institute of Environmental Sciences (CML), Leiden University},
    addressline={P.O. Box 9518},
    city={Leiden},
    postcode={2300 RA},
    state={South Holland},
    country={The Netherlands}}

% Abstract
\begin{abstract}

%? reviewer 2 - comment 2 - simplify language
The quintessential principle of the circular economy is to minimise material consumption and waste generation. Thus, identifying and quantifying waste and material flows is critical. Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.

We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios. T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens. T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.

\end{abstract}

% Keywords
\verb+\begin{keyword}+ \verb+circular economy+ \verb+waste+ \verb+material+ \verb+life cycle assessment+ \verb+critical raw material+ \verb+supply chain+ \end{keyword}

% Graphical abstract
\begin{graphicalabstract}
  \includegraphics{figs/cas-grabs-TREX.pdf}
\end{graphicalabstract}

% Highlights
\begin{highlights}
    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
    \item Assesses supply risks by calculating demand inventories for critical materials
    \item Simplifies quantification of user-specified waste and material footprint categories
    \item Rapidly identifies hotspots of waste generation and material demand in supply chains
    \item Presents a case study of five lithium-ion batteries to demonstrate T-reX's utility
\end{highlights}

% Keywords
\begin{keyword}
  circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
\end{keyword}

%\nocite{*}

\maketitle

% list of abbreviations
\begin{table}[H]
  \centering
  \caption*{List of abbreviations and terminology}\label{tab:abbreviations}
  \begin{tabular}{ll}
      \toprule
      %? reviewer 3 - RE: CCS definition
      CCS              & Carbon Capture and Storage                                           \\
      CF                & Carbon Footprint                                                    \\
      EF                & Ecological Footprint                                                \\
      CML               & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
      CRM               & Critical Raw Material                                               \\
      CPC               & Cooperative Patent Classification                                   \\
      CSI               & Crustal Scarcity Indicator                                          \\
      CSP               & Crustal Scarcity Potential                                          \\
      EDIP              & Environmental Design of Industrial Products                         \\
      EEIOA             & Environmentally Extended Input-Output Analysis                      \\
      EOL               & End-of-Life                                                         \\
      FOEN              & Swiss Federal Office for the Environment                            \\
      IAM               & Integrated Assessment Model                                         \\
      IE                & Industrial Ecology                                                  \\
      LCA               & Life Cycle Assessment                                               \\
      LCI               & Life Cycle Inventory                                                \\
      LCIA              & Life Cycle Impact Assessment                                        \\
      LFP               & Lithium Iron Phosphate                                              \\
      Li-ion            & Lithium-ion                                                         \\
      LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
      MF                & Material Footprint                                                  \\
      MFA               & Material Flow Analysis                                              \\
      NCA               & Nickel Cobalt Aluminum                                              \\
      NMC               & Nickel Manganese Cobalt                                             \\
      PWF               & Product Waste Footprint                                             \\
      RCP               & Representative Concentration Pathway                                \\
      ReCiPe            & A standard LCIA method set                                          \\
      SDG               & Sustainable Development Goals                                       \\
      SSPs              & Shared Socioeconomic Pathways                                       \\
      T-reX             & The Tool for analysing re-X in LCA                                  \\
      UBP               & Umweltbelastungspünkte (Environmental Impact Points)                \\
      UN                & United Nations                                                      \\
      WF                & Waste Footprint                                                     \\
      \bottomrule
  \end{tabular}
\end{table}

% Main text

% Introduction
\section{Introduction}\label{sec:introduction}
\subsection{Introduction to T-reX}\label{sec:intro-trex}

%? reviewer 2 - comment 2 - simplify language
Development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}. Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}. In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.

The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}. More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}. Whether at the level of products, entire industries, nations, or even continents, the material footprint aims to quantify the total material consumed in supply chains. It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.

Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed. While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}. Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}. Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.

WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders. Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.

%? reviewer 2 - comment 1 - COMPARISON OF T-REX TO OTHER TOOLS
%! fix
Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes. LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}. The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere). These scores can then be aggregated into one metric for each impact category that can be compared across products and processes. Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.

Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}). These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.

To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories. T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.

Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases. Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}. Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field. \texttt{premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming. The authors believe that there is an urgent need for research investment in this area.

%! reviewer 3 - comment 1 - ADDITIONAL INFORMATION ON T-REX's differentiating features comparet to LCI
The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes. It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows. By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others). In any case, minimising primary material consumption and the generation of waste is of paramount importance. By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and potentially highlight opportunities for waste reduction and material efficiency.

%? reviewer 2 - comment 3 - ADDITIONAL INFORMATION ON T-REX's FUTURE DEVELOPMENT
With the continual improvement of LCA databases and the development of prospective databases, T-reX can produce ever more realistic estimates, enhancing its actionable utility derived from the exploration of the potential impacts of future waste management systems and the material demands of emerging technologies. The better the data, the more specific and accurate the results of T-reX analysis, and, thus, the more useful the tool becomes. The authors hope that T-reX will be a valuable asset to the LCA community and contribute to the development of a more sustainable and circular economy.

\subsection{Background and need}\label{sec:intro-background}

\subsubsection{Waste in LCA}\label{sec:intro-waste}

Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time. From a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist. From a practical economic viewpoint, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows, in addition to energy costs and the relative prices of virgin and secondary materials. Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and `modelling [of its management] must be physically based'. The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.

Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities. There remains, however, a knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.

Conventional LCA database models consider waste as a `service' (accounting for the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts. In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy. Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}. There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill. In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments. In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling of the long-term impacts of disposal.

A considerable portion of a product's total waste is generated during earlier stages of its supply chain such as resource extraction, transportation, and manufacturing, thus, often remaining `invisible' to conventional LCA accounting practices~\citep{laurenti2016wastefootprint}. This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators. Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}. The environmental significance of waste and its correlation with other indicators has been the subject of extensive research. Studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}. Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}. The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.

Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA~\citep{lauran2020abioticdepletion}. This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle. Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible methods to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.

~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data. The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' . In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such. The `T-reX' Python software tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA\@. Moreover, the utility of T-reX is not limited to waste, it can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.

\subsubsection{Material demand in LCA}\label{sec:intro-material}

In the context of geopolitical tension and a mineral-hungry `renewable energy transition', greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}. While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system rather than the primary material flows themselves.

The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}. The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations. CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies. The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.

While useful for its intended purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories. Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.

% Methodology

\section{Methodology}\label{sec:methodology}

This section is divided into two parts. In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries, highlighting inconsistencies in LCA methodology and probing departures from projections in prospective models.

\subsection{T-reX}\label{sec:method-TreX}

\subsubsection{Computational framework}

Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising especially  the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}. Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}. Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases. T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---and after running T-reX, the manipulated databases and the `pseudo-LCIA' methods created by T-reX can be used in the project in the accustomed way. T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is fully open-source under the CC-0 licence. The full source code for T-reX is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}. T-reX is designed to be used with \texttt{ecoinvent} databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria. Currently, it has been tested with all available system models of \texttt{ecoinvent}versions 3.5--3.10.

T-reX can be used directly from the command line, or imported as a Python package with which the user can access the individual functions and modules. In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database. The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or a `prospective' database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.

The supplementary material in \autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow. Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.

\subsubsection{Functionality and purpose}

T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods. We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of an activity's technosphere inventory without reference to impact modelling, as in standard LCIA.

If desired, prospective databases can be defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} and can be paired with a variety of mitigation scenarios.

The deconstruction of the databases by T-reX into lists of exchanges allows the relevant flows of material and waste to be identified and categorised by the search functions. The queries facilitating this are tailored to the specific database and the user can easily modify them to suit their needs. The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand. The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a mirrored `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.

In the default configuration of T-reX, there are 10 waste categories which are then further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods. The waste categories include aggregations such as 'total waste', 'hazardous waste' as well as, those for waste sent to treatment by, for example, incineration or recycling. The full details are listed in the supplementary material in~\autoref{sec:supplementary} [section 3 ]. One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.

In addition to these waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption. Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities. The identity of the materials considered, and their categorical groupings, are easily customisable by the user. The full list of 59 materials included in the default configuration is provided in the supplementary material in~\autoref{sec:supplementary}.

Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name. A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows. For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs. Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.

In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.

T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere, a material basis. This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct effects on the market, such as a 'negative consumption' of other materials due their co-production with the functional unit. Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal. If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another. For example, in the results of the Li-ion battery case study , this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.

\subsubsection{The workflow of T-reX}

The workflow of T-reX is divided into several modules, each performing a separate function. The modules are designed to be used in a particular order, but the user can also use them individually to perform specific tasks. The standard workflow is as follows:

\begin{enumerate}
    \item Configuration of waste and material exchange categorisation (optional)
    \item Generation of prospective LCA databases (optional)
    \item Database expansion---to create a list of all exchanges in the database
    \item Identification and categorisation of exchanges
    \item Creation of `pseudo-biosphere' databases
    \item Creation of `pseudo-LCIA' methods to calculate waste and material inventory footprints
    \item Exchange editing---whereby the technosphere exchange is mirrored as a `pseudo-biosphere' exchange
    \item Database verification---to ensure that T-reX has manipulated the database correctly
\end{enumerate}

This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere databases that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.

An overview of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material in \autoref{sec:supplementary} contains a more detailed computational flowchart.

\begin{figure}[H]
    \centering
    \caption{The workflow of T-reX presented in a generalised (a) and computational (b) format. Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
    \includegraphics[width=9cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
\end{figure}

\subsection{Case study methodology}\label{sec:method-casestudy}

To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories. The system boundaies. These following markets were used for the example calculations and unaltered by the authors:
\begin{itemize}[itemsep=0pt]
    \item Li-ion, LFP, rechargeable, prismatic
    \item Li-ion, LiMn\(_2\)O\(_4\), rechargeable, prismatic
    \item Li-ion, NMC111, rechargeable, prismatic
    \item Li-ion, NMC811, rechargeable, prismatic
    \item Li-ion, NCA, rechargeable, prismatic
\end{itemize}

In addition to the waste and material inventory footprint methods created by T-reX, the following standard LCIA methods were applied for comparison:

\begin{itemize}[itemsep=0pt]
    \item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
    \item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
    \item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
    \item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
\end{itemize}

Additional to the present-day databases, T-reX created prospective database sets based on \texttt{ecoinvent} 3.9.1 using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
\begin{itemize}
    \item SSP2-base: representing an approximate 3.5°C increase in global temperatures to 2100
    \item SSP2-PkBudg500: representing the achievement of Paris climate goals, ca. 1.3°C increase to 2100
\end{itemize}

For each pathway, databases (at five-year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.

\subsubsection{LCA calculations}
For each combination of activity, method, and database, a unified metric was calculated along with details of the top contributing processes. Additionally, for the T-reX methods, a contribution analysis was performed. Utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem, this function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff. The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement]. T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

\section{Conclusions}\label{sec:conclusions}

T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database. It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods. These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.

T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts. This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage. T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.

% Back matter
\section*{Data availability}
All data used is publically available online under the noted sources. T-reX is fully open source and dedicated to the public domain under the CC-0 licence. It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.

% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.

\section*{CRediT authorship contribution statement} % VISUAL VERSION

\begin{itemize}[noitemsep]
    \item \textbf{SCM}: Stewart Charles McDowall
    \item \textbf{EL}: Elizabeth Lanphear
    \item \textbf{SC}: Stefano Cucurachi
    \item \textbf{CFB}: Carlos Felipe Blanco
\end{itemize}

\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit-TREX.pdf}

\section*{Declaration of competing interest}
  The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Acknowledgements}
  Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No. 101058522 (project FutuRaM --- futuram.eu). The authors would like to thank the reviewers for their valuable comments and suggestions.

  The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems. The authors hope to continue the development of T-reX in the service of our shared ideals.

  %! replace the sections
\section{Supplementary material}\label{sec:supplementary}
  The supplementary material supplied with this manuscript contains the following:
    \begin{enumerate}
        \item Metadata of the T-reX Python package
        \item Details of the modules and user configuration
        \item Description of the computational workflow
        \item Modular flowchart of T-reX
        \item Example of the terminal output of T-reX
        \item Example of the terminal output of the case study
        \item Complete tabulated results of the case study
        \item Complete visualisations from the case study
    \end{enumerate}

%% Loading bibliography style file
%\bibliographystyle{model1-num-names}
\bibliographystyle{cas-model2-names}

% Loading bibliography database
\bibliography{cas-refs-TREX}

\end{document}

### 2024-11-27 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
update

@@
 \subsubsection{LCA calculations}
-For each combination of activity, method, and database, a unified metric was calculated along with details of the top contributing processes.
-Additionally, for the T-reX methods, a contribution analysis was performed.
-Utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem, this function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
+For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes.
+Additionally, for the T-reX methods, a contribution analysis was performed with the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
+This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
 The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
 T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
+
+\subsection{T-reX tool}\label{sec:results-T-reX}
+
+An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, subsection s.
+The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+
+\subsection{Case study: Li-ion batteries}\label{sec:results-casestudy}
+
+As described in~\autoref{sec:method-casestudy}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered background inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
+The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
+This section includes some highlights of the results, with full results included in the supplementary material (\autoref{sec:supplementary}).
+Because the `pseudo-LCIA' methods created by T-reX are integrated into the user's \texttt{brightway} project as if they were standard LCIA methods, the footprint calculations can be performed in the customary way.
+In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
+
+\subsubsection{Sankey visualisation of flow inventory footprints}\label{sec:results-case_study-sankey}
+
+\autoref{fig:sankey_waste} presents a heavily simplified Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material~\autoref{sec:supplementary}.
+As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
+Due to the nature of LCA, the waste flows do not represent the `real' physical flows directly, but rather an abstraction thereof based on, in this case, economic allocation---the market price of the functional unit produced by the activity~\citep{guinee2004economicallocation}.
+If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and would be more directly related to the physical flows of the activities.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
+    \caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
+\end{figure}
+
+\subsubsection{Temporal and scenario variation in waste and material inventory footprints}
+%\label{sec:results-case_study-total_footprints}
+
+%! total
+\autoref{fig:waste_totalANDcarbon} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
+The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced.
+The  LiMn\(_2\)O\(_4\) battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery.
+In each case, there was only a slight downward trend in the waste footprints between 2020 and 2100.
+This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in fossil fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
+For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
+
+%! CO2
+The inclusion of carbon capture and storage (CCS) in the prospective databases is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
+This result highlights the fact that the (often) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
+The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
+The authors consider that the over-representation of CCS and the under-representation of waste processing technologies in the prospective databases is one significant limitation to the validity of prospective LCA databases.
+Significant work is needed to improve the representation of these technologies in the databases to ensure that the results are more realistic and useful for decision-making.
+%!!!!
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
+The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
+\end{figure}
+%!!!!
+
+\subsubsection{Contribution of `top-processes' in the supply chain}%\label{sec:results-case_study-topprocesses}
+
+\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the graphite footprint of the  LiMn\(_2\)O\(_4\) battery under the baseline scenario from 2020--2100.
+The total footprint is seen to more than triple, from 0.16~kg/kg in 2020 to 0.52~kg/kg in 2050 and onward to 2100.
+The largest contributor is not, in this case, the battery-embedded graphite electrode (which is the second top process), but `silicon production' from further up in the supply chain.
+This result is likely a reflection of the electrification of the transport and energy sectors that is included in the REMIND model.
+The results of this kind of analysis with T-reX can be used to identify the most significant processes in the supply chain and to target interventions to reduce the potential environmental impact of production.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
+    \caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
+\end{figure}
+
+%!!!!
+
+\subsubsection{Contribution of industrial sectors in the supply chain}\label{sec:results-case_study-topsectors}
+
+\autoref{fig:cpc_contribution} shows the contribution of sectors (grouped by CPC codes) to the total natural gas footprint of the LFP battery under the PkBudg500 pathway.
+In this example, the relative contribution of the sector `47160: Electronic integrated circuits' is seen to decrease from 11\% in 2020 to 8\% in 2100, while over the same period the contribution of the sector `46430: Parts of primary cells, primary batteries and electrodes' increases from 32\% to 38\%.
+The method used to calculate these contributions involves traversing the supply chain branches to a certain level (max.
+4, in this case), cutting a specified point (5\% in this case), and grouping the value of the `leaves' by their CPC code.
+The results, therefore, will depend on how deeply the user would like to inspect the supply chain.
+Additionally, the utility of these results is dependent on how well the CPC codes define the processes in the supply chain for the particular case.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
+    \caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
+\end{figure}
+
+\subsubsection{Comparison with `similar' LCIA methods}\label{sec:results-case_study-methodcomparison}
+
+A comparison of the results for `Coal (black)' demand using the T-reX method with the LCIA method `EDIP 2003 - coal no LT' is shown in~\autoref{fig:comparison_methods}.
+In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
+Such comparability was also observed for other fossil-fuel-related demands (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
+Correlation between standard LCIA methods and T-reX methods is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
+In standard LCIA methods, impact scores are derived exclusively from the magnitude of the exchanges between the biosphere and the technosphere, that is, extraction and emission.
+In the T-reX methods, the footprint scores are based on an accounting of either waste generated or material demand, both of which are technosphere-technosphere exchanges in terms of LCA modelling.
+For the material demand methods especially, this distinction is critical.
+For example, the application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the T-reX method provides an aggregation of the exchanges with the market for that metal.
+The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply chain pressures that are not captured by the standard LCIA methods.
+This makes the T-reX methods more sensitive to the modeling choices (e.g., economic allocation) that are generally embedded in LCA databases.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=16cm]{figures/T-reX-coalANDedip.pdf}
+    \caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
+\end{figure}
+
+\subsubsection{Comparison with other studies}\label{sec:results-case_study-comparison}
+
+In the most similar published method, \cite{laurenti2023wastefootprint} used an alternative procedure for calculating the waste footprint of \~1400 activities in the \texttt{ecoinvent} database version 3.5.
+Direct comparison is not possible, as this database contains only one generic Li-ion battery, `market for battery, Li-ion, rechargeable, prismatic'.
+The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
+\autoref{tab:results-case_study-comparison} presents a comparison of the results from the two studies, where possible.
+For liquid, solid, and recycled waste, the results were closely aligned, however, for the hazardous waste fraction, Laurenti et al.
+reported 95\%, whereas T-reX reported only 3\%.
+The reason for this discrepancy is explained by the fact that in the method of Laurenti et al.\, a ``waste flow was regarded as hazardous when the hazardousness was clearly stated in its subsequent waste treatment activity''.
+The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
+In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+
+\begin{table}[H]
+    \centering
+    \caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}\label{tab:results-case_study-comparison}
+    \begin{tabular}{lll}
+        \toprule
+        \textbf{Indicator}         & \textbf{T-reX} & \textbf{Laurenti et al.} \\
+        \midrule
+        Total solid waste (kg/kg)  & 50.9           & 62.5                     \\
+        Total liquid waste (kg/kg) & 3.53           & 3.63                     \\
+        Hazardous waste (kg/kg)    & 1.47           & 62.6                     \\
+        Recycled waste (kg/kg)     & 1.59           & 1.98                     \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting.
+Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
+
+T-reX advances the state-of-the-art in LCA by providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
+Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
+
+The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility, and limitations of T-reX.
+
+By adjusting the user configuration for future scenarios and waste/material categorisation, we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
+Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate.
+Additionally, visual exploration in \texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are structurally identical to standard LCIA methods when integrated within the \texttt{brightway} project.
+
+One limitation of T-reX is that it does not yet provide specific information in a readily accessible format on the composition of the waste generated.
+This information is required to thoroughly assess the potential environmental impacts of this waste.
+Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine the nature of the waste flows.
+For example, if the waste generated could represent an actual loss of resources or  potential environmental risk, or, as is often the case, simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
+
+The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+As demonstrated in the results of the case study---where there were often very few scenario-temporal changes in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adoption of future background inventories.
+In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.
 
@@
 T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
+It deconstructs, manipulates and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
 These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
@@
 This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
+T-reX is a step towards a better understanding of these interconnections for the practical realisation of sustainability and the circular economy.
 

### 2024-11-29 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template.tex
update

--- 2024-11-27 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
@@
 
-We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
+We introduce T-reX, a Python tool extending the Brightway computational ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
@@
 \end{graphicalabstract}
+
+\clearpage
+
+%% GRAPHICAL ABSTRACT
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{figs/grabs_T-reX_graphical-abstract.pdf}
+	\caption{Graphical abstract of the T-reX program for waste and material footprinting in Life Cycle Assessment}
+	\label{fig:graphical_abstract}
+\end{figure}
+
+\clearpage
+
+% Switch to landscape for the included PDF
+\newgeometry{landscape}
+\includepdf[pages=1, width=\textheight]{figs/grabs_T-reX_graphical-abstract.pdf}
+\restoregeometry % Restore the original geometry
 

### 2024-11-29 other/sanitize_filenames.py

@@ -1,18 +1,0 @@
-import os
-from pathvalidate import sanitize_filename, sanitize_filepath
-
-def rename_files_in_directory(root_dir):
-    for dirpath, dirnames, filenames in os.walk(root_dir):
-        for filename in filenames:
-            # Sanitize the filename using pathvalidate
-            sanitized = sanitize_filename(filename, platform="universal")
-            if sanitized != filename:
-                old_filepath = os.path.join(dirpath, filename)
-                new_filepath = os.path.join(dirpath, sanitized)
-                # Rename the file
-                os.rename(old_filepath, new_filepath)
-                print(f'Renamed: {old_filepath} \n\t\t-> {new_filepath}')
-
-# Specify the root directory to start from
-root_directory = "."
-rename_files_in_directory(root_directory)

### 2024-11-30 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template.tex
update

--- 2024-11-27 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
@@
 %% version 1999/12/01 or later.
+%%%
+%
+% T-reX manuscript using the elsevier template for JRC
+%
+% Repository for the T-reX manuscript at https://github.com/Stew-McD/T-reX_Publication
+% Repository for the T-reX python program at https://github.com/Stew-McD/T-reX
+%
+% S.C.McDowall - github/Stew-McD
+%
+%%%%
+
 %% 
@@
 
-\documentclass[a4paper,fleqn]{cas-dc}
+%\documentclass[a4paper,fleqn]{cas-dc}
 
@@
 
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
+\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
 
-%\usepackage[numbers]{natbib}
+\usepackage[numbers]{natbib}
 %\usepackage[authoryear]{natbib}
-\usepackage[authoryear,longnamesfirst]{natbib}
+%\usepackage[authoryear,longnamesfirst]{natbib}
 
@@
 %%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
 
@@
 
-% Abstract
+% Abstract\\
 \begin{abstract}
 
-%? reviewer 2 - comment 2 - simplify language
 The quintessential principle of the circular economy is to minimise material consumption and waste generation.
 Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
+Life Cycle Assessment (LCA) is a powerful quantitative method for this, especially given its capacity to expose the details of an activity's entire life cycle, often guiding the implementation of circular principles to where they could be most effective.
 
-We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
-T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
+Pursuant to these goals the authors have developed 'T-reX', an analytical tool for LCA, written in Python.
+T-reX extends on the existing computational ecosystem centered on Brightway and Premise and can facilitate the quantification of user-defined supply chain demands for activities in current and prospective scenarios.
+T-reX streamlines database manipulation for LCA practitioners, integrating methods to aggregate and analyse demand inventories that could expose hotspots of hidden risk.
 
-A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
+With a simple case study using Lithium-ion batteries, we demonstrate some of T-reX's potential by detailling and exploring aspects of their waste and material footprints.
+Since such footprints are often linked to negative externalities, T-reX could support sustainable decision-making and contribute to the development of the circular economy by promoting the understanding of our material consumption and waste generation.
 
@@
 % Keywords
-\verb+\begin{keyword}+ \verb+circular economy+ \verb+waste+ \verb+material+ \verb+life cycle assessment+ \verb+critical raw material+ \verb+supply chain+ \end{keyword}
+\begin{keywords} 
+circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
+\end{keywords}
 
@@
 \begin{graphicalabstract}
-  \includegraphics{figs/cas-grabs-TREX.pdf}
+  \includegraphics{figs/T-reX_graphical-abstract.png}
 \end{graphicalabstract}
+
+\clearpage
+
+%% GRAPHICAL ABSTRACT FULL SIZE
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{figs/T-reX_graphical-abstract.png}
+	\caption{The graphical abstract of the T-reX program developed for waste and material footprinting in Life Cycle Assessment}
+	\label{fig:graphical_abstract}
+\end{figure}
+
+\clearpage
+
+% Switch to landscape for the included PDF
+\newgeometry{landscape}
+\includepdf[pages=1, width=\textheight]{figs/T-reX_graphical-abstract.pdf}
+\restoregeometry % Restore the original geometry
 
@@
 \end{highlights}
-
-% Keywords
-\begin{keyword}
-  circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
-\end{keyword}
 
@@
 
-%? reviewer 2 - comment 2 - simplify language
-Development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+The development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised world have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
@@
 
-%? reviewer 2 - comment 1 - COMPARISON OF T-REX TO OTHER TOOLS
-%! fix
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
@@
 
-Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
-A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
+Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the identification of material exchanges with T-reX differs from that used to identify waste exchanges.
+That is, the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+A useful feature of T-reX is that, in cases where there are several markets for one material or material grouping, the program can aggregate these flows.
 For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
@@
 This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
-The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
+The groups of the resulting 'leaves' are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
 T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
@@
 \section*{Data availability}
-All data used is publically available online under the noted sources.
-T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
+All data used is publicly available online under the noted sources.
+T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence.
 It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
@@
 
-\section*{CRediT authorship contribution statement} % VISUAL VERSION
+\section*{CRediT authorship contribution visualisation} % IMPROVED VISUAL VERSION
 

### 2024-11-30 00_JCR-submission-folder_SECOND SUBMISSION/cas-dc-template.tex
update

--- 2024-11-30 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template.tex
@@
 
-An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, subsection s.
+An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, appendix 6.
 The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
@@
     \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
-    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
+    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
 The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
@@
 
-T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts.
-This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
+It is essential that we elucidate the complex relationships between material demand and waste generation and environmental and social harms.
+The software we present in this article enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their potential connections threats such as supply chain risks and environmental damage.
 T-reX is a step towards a better understanding of these interconnections for the practical realisation of sustainability and the circular economy.

### 2024-11-30 latex_files-DRAFTING/discussion.tex
update

--- 2024-04-08 latex_files/discussion.tex
@@
 In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
-A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.

### 2024-11-30 02_JCR-submission_folder_FIRST-SUBMISSION-original/T-reX_supplementary-material/T-reX_case-study_data/_02b_process-results_top-processes.py

@@ -1,0 +1,62 @@
+# this is a script to process the contribution analysis results of the case study
+import os
+import numpy as np
+import pandas as pd
+import ast
+from tabulate import tabulate
+
+from _00_main import FILE_RESULTS_PROCESSED, FILE_RESULTS_PROCESSED_TOP_PROCESSES
+
+DATA_FILE_IN = FILE_RESULTS_PROCESSED
+DATA_FILE_OUT = FILE_RESULTS_PROCESSED_TOP_PROCESSES
+
+# make sure of the correct working directory
+os.chdir(os.path.dirname(os.path.abspath(__file__)))
+
+pd.set_option("display.max_columns", 50)
+pd.set_option("display.max_rows", 200)
+
+# load data set
+df_raw = pd.read_csv(DATA_FILE_IN, sep=";", index_col=None)
+df = df_raw.copy().reset_index(drop=True)
+df = df[df.method_0 == "T-reX"].reset_index(drop=True)
+
+df["method_1"] = df["method_1"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+)
+df["method_2"] = df["method_2"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+)
+
+
+df["top_processes"] = df["top_processes"].apply(ast.literal_eval)
+length_top = len(df["top_processes"][0])
+length_top = 5
+print(f"Number of top processes: {length_top}")
+
+for i in range(length_top):
+    df[f"tp_{i+1}"] = df["top_processes"].apply(lambda x: x[i])
+    df[f"tp_{i+1}_score"] = df[f"tp_{i+1}"].apply(lambda x: x[0])
+    # df[f"tp_{i+1}_demand"] = df[f"tp_{i+1}"].apply(lambda x: x[1])
+    df[f"tp_{i+1}_name"] = df[f"tp_{i+1}"].apply(lambda x: x[2])
+    # df[f"tp_{i+1}_location"] = df[f"tp_{i+1}"].apply(lambda x: x[3])
+    df = df.drop(f"tp_{i+1}", axis=1)
+
+df = df.drop("top_processes", axis=1)
+
+cols_tp = [x for x in df.columns if "tp_" in x]
+cols_tp_score = [x for x in cols_tp if "score" in x]
+cols_tp_demand = [x for x in cols_tp if "demand" in x]
+cols_tp_name = [x for x in cols_tp if "name" in x]
+
[…]

### 2024-11-30 02_JCR-submission_folder_FIRST-SUBMISSION-original/T-reX_supplementary-material/T-reX_case-study_data/_03a_process-supply_chain_results.py

@@ -1,0 +1,94 @@
+# this is a script to process the contribution analysis results of the case study
+import os
+import numpy as np
+import pandas as pd
+
+from utils import replace_string_in_df, term_replacements
+from _00_main import FILE_SUPPLYCHAIN_RESULTS, FILE_RESULTS_PROCESSED_SC_1, FILE_RESULTS_PROCESSED_SC_2
+# make sure of the correct working directory
+os.chdir(os.path.dirname(os.path.abspath(__file__)))
+
+pd.set_option("display.max_columns", 50)
+pd.set_option("display.max_rows", 200)
+
+# load data set
+DATA_FILE_IN = FILE_SUPPLYCHAIN_RESULTS
+DATA_FILE_OUT_1 = FILE_RESULTS_PROCESSED_SC_1
+DATA_FILE_OUT_2 = FILE_RESULTS_PROCESSED_SC_2
+
+# load data set
+df_raw = pd.read_csv(DATA_FILE_IN, sep=";", index_col=None)
+df = df_raw.copy().reset_index(drop=True)
+
+# Call the function with the DataFrame and the replacements dictionary
+for term, replacement in term_replacements.items():
+    df_raw = replace_string_in_df(df, term_replacements)
+
+# filter out unnecessary columns
+df = df[df.database != "ecoinvent-3.9.1-cutoff"]
+drop = [
+    "db_model",
+    "db_version",
+    "db_system",
+    "db_base",
+    "code",
+    "db_ssp",
+    "direct emissions",
+    "location",
+    "key",
+    "database",
+    "product",
+]
+
+for d in drop:
+    df = df.drop(columns=[d], errors="ignore")
+
+
+# simplify the names of the activities
+df.name = df.name.apply(lambda x: x.split(", ")[2])
+print(df.name.unique())
+df["method_1"] = df["method_1"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+)
+df["method_2"] = df["method_2"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+)
+
+# check for signigicant methods
+grouped = df["total"].groupby(df["method_2"]).mean().sort_values(ascending=False)
+
+# # filter out methods where the maximum score < 0.01
+df_sig = df.groupby("method_2").filter(lambda x: x["total"].abs().max() > 0.01)
+grouped_sig = (
[…]

### 2024-11-30 04_T-reX_manuscript-review-documents/external/JRC-peer-review/T-reX_manuscript_JRC-submission_response-to-reviewers.md
update

# T-reX manuscript revision

----

## Response to peer review

- Author:
  - S.C. McDowall

- Date:
  - 2024-11-30

- Journal:
  - Resources, Conservation and Recycling

- Manuscript Number:
  - RECYCL-D-24-01859R1

- Manuscript title:
  - *T-reX: Quantifying waste and material footprints in current and
future Life Cycle Assessment (LCA) databases*

-----------

Dear reviewers, editor,

We the authors of T-reX would like to thank you for the time and attention that you have generously contributed to the review of our manuscript. We are especially grateful for the extensive and constructive commentary. We have addressed each of your comments in the following text, and have made the necessary changes to the manuscript. We hope that you find the revised manuscript to be satisfactory.

------------
------------

## 1.  Reviewer One

*“I think that this manuscript needs to be ameliorated.”*

### Summary of review

- More detail for case study requested, also sensitivity/uncertainty analysis

- Comment 1.1: system boundary
- Comment 1.2: input-output data
- Comment 1.3: material footprint data (more discussion, and ‘overall material footprint’)
- Comment 1.4: sensitivity analysis or uncertainty analysis

### Author responses

#### General response

As the reviewer points out, the case study could, indeed, benefit from more detail. Additional explanation has been added to the main text and there is detailed discussion in the supplementary information (especially appendices 3 and 4). Furthermore, the complete datasets are publicly available on Github and Zenodo. The case study was included only to demonstrate some potential use cases of T-reX and not to provide a comprehensive waste and material footprint analysis of the Li-ion battery market. This is detailed in the abstract, as well as subsection 2.2 of the methods and 3.2 of the results. The case study is not intended to be intensive or exhaustive and the results are not intended to be used for decision-making purposes. The authors intent is for the focus of the paper to be on the T-reX program itself.

#### Specific responses

##### Reviewer comment 1.1 - Case study system boundary

"*The system boundary of the case study should define, and it can be shown more clearly by adding diagrams.*"

##### Author response 1.1 - Case study system boundary

A figure could be added to the supplementary information to illustrate the system boundary of the case study. However, given the limitations on space, and that the authors feel it would not add any pertinent information. The system boundary is very simple, as defined in the text (methodology section 2.2). That is, exactly the inventory of the Li-ion battery markets as modelled in ecoinvent version 3.9.1. Essentially, a single market activity for each battery, to serve as a basic example.

##### Reviewer comment 1.2 - Input-output data

"*No input-output table provided for the case studies. Please provide this inventory data in the manuscript.”*

##### Author response 1.2 - Input-output data

As described in the methodology section (subsection 2.2), the inventory data is taken directly from the respective activities in the ecoinvent database (version 3.9.1). The input-output data used in the case study is directly that given by the ecoinvent database, without amendment. Full details are provided in the supplementary information of this manuscript (appendix 3 and 6) and complete data is available in the [Github](https://github.com/Stew-McD/T-reX_Publication) and [Zenodo](https://zenodo.org/records/10925359) repositories.

##### Reviewer comment 1.3 - Material footprint data

*“Result and Discussion section; this study was present some part of material footprint. I recommend further discussions and interpretation the result. In addition, it is necessary to present the overall material footprint of battery.”*

##### Author Response 1.3 - Material footprint data

As is mentioned throughout the manuscript, the material footprint results calculated by T-reX are aggregations of the exchanges in the model system that are targeted by the user’s search criteria (in this case, a default selection of 59 materials and their flows as presented by the ecoinvent database and listed in appendix 3, sub section 2 and 3). The authors believe that it would be misleading to present the reader with a ‘total material footprint’ based on the sum of the material footprints of the individual materials, as the LCA models only represent certain aspects of the system. A ‘total material footprint’ without respect for the nature of the material is meaningless, and indeed, one of the strengths of T-reX is to allow the user to pinpoint the materials of interest to them.

##### Reviewer comment 1.4 - Sensitivity analysis and uncertainty analysis

*“Authors should also perform the sensitivity analysis or uncertainty analysis.”*

#### Author response 1.4 - Sensitivity analysis and uncertainty analysis

Indeed, this is something that also could be done, and may yet make and interesting extension for the next study. However, given that the data for the batteries in this case study is taken directly from the ecoinvent activities and offered as a rudimentary example of the programme’s utility, the authors believe that a sensitivity or uncertainty analysis on the case study results is not called for. The authors believe that the presentation of this could be distracting from the message of the paper. For clarity, however, the authors have added further to the discussion on the limitations of the case study results (paragraphs 4-6 and subsection 4) to make this more apparent to the reader.

-------

## Reviewer Two

*“The manuscript is well-written and presents a clear argument. I had a few minor suggestions.”*

### 2. Reviewer Two

#### Summary of this review

- Expand comparison with existing methods
- Simplify some of the language
- Expand on the future developments and applications of T-reX
- Include higher resolution version of graphical abstract

#### Specific responses

##### Reviewer comment 2.1 - To expand on the comparison with existing methods

*“To clearly communicate the novelty of this work, the author may consider emphasizing explicit Comparison with Existing Methods. Although section 3 has comparisons, it’s hard to tell the differences between T-rex with the existing LCIA methods (i.e CO2 emissions/waste). Providing a detailed comparison of T-reX with existing LCA and LCIA methods will be helpful to justify the novelty.”*

##### Author response 2.1 - To expand on the comparison with existing methods

As the reviewer correctly points out, there are many similarities between the inventory footprints calculated by T-reX, and that of some exiting LCIA methods. Indeed, the choice of the term 'pseudo-LCIA' is derived from this, because the calculation strategy is very similar to LCIA methods in general. Apart from the added functionality of T-reX, one main difference lies in directness of T-reX aggregations, with comparison to the abstracted methods such as Crustal Scarcity Potential, for example. The authors have expanded the comparison with existing methods by adding more specific details in the relevant sections. In the introduction, this is subsection 1.1 (paragraphs 3-5) and 1.2.1 (paragraph 3). In the methodology, this is 2.1.2, paragraphs 6-7 and section 3.3.2 paragraph 3. In the results section, this is subsection 3.2.5 and 3.2.6. In the discussion, this is first two paragraphs (lines 389-411). Additionally, appendix 3, section 3.3.2, paragraph 3 and 4 cover this in more detail and some of this material has been brought forward to the methods section for more prominence (subsection 2.1.2 paragraph 7-8).

##### Reviewer comment 2.2 - To simplify some of the language

*"There are sections where the language can be simplified for better clarity without losing technical precision. For example, the abstract succinctly summarizes the study. However, consider adding a sentence on the main findings from the case study to give a complete overview. The phrase ”frantic pursuit of sustainability objectives” in the introduction could be toned down to “dedicated pursuit of sustainability objectives” for a more neutral tone”*

##### Author response 2.2 - To simplify some of the language

The authors have restructured the abstract to include more reference to the case study (paragraph 3). Additionally, the language in the text has been amended to improve clarity and purvey a more neutral tone, particularly the first sentence of the abstract and the first paragraph of the introduction.

##### Reviewer comment 2.3 - To expand on the future developments and applications of T-reX

*”A discussion on potential future developments and applications of T-reX will be helpful”*

##### Author response 2.3 - To expand on the future developments and applications of T-reX

The authors have expanded the discussion on potential future developments and applications of T-reX in the discussion section (paragraphs 4 and 5). The authors hope that this will provide the reader with a better understanding of the potential of T-reX, as well as its limitations and their relation to some of the challenges in our field, most notably, the need for more comprehensive and reliable data on which to build our models.

##### Reviewer comment 2.4 -  The graphical of abstract

*'low resolution figure is hard to read'*

##### Author response 2.4 -  The graphical of abstract

The publication portal rendered the resolution of the graphical abstract in this way. It was an unfortunate oversight from the authors not to correct this in the initial submission. The graphical abstract in is now also available in various formats and resolutions in the portal submission as well in the online material that is available with the software.

----

## 3. Reviewer Three

*“I think that the paper presents an interesting and novel method to assess waste generation and consumption of material in systems, but some aspects need a more extensive explanation.”*

### Review summary

- Clarification on the difference between T-reX and LCI
- Clarification on the terminology ‘pseudo-LCIA’
- Clarification on the single score calculation of T-reX

### Major comments

##### Reviewer comment 3.1 - Clarification on the difference between T-reX and LCI

*“Lines 105-106: I can understand the difference between the T-reX method and the results that one can derive from the LCI analysis (i.e. the list of elementary flows associated with the system under analysis) in the case of waste, because, as you said, T-reX considers also waste flows that are treated, but what about material flows? Why should T-reX results be different from those that one can obtain by the LCI analysis? This point remains unclear to me even after reading the entire paper.”*

##### Author response 3.1 - Clarification on the difference between T-reX and LCI

One significant difference, and a highlight of the method of the T-reX program is how it makes it easy to identify "hotspots" and perform contribution analysis. Furthermore, the strength of T-reX is to make these calculations simple and easily repeatable (in the way that the databases are manipulated, and the footprints calculated as if they were LCIA methods). The as described in the response to reviewer two, the relevant sections have been refined with the aim of more clearly communicating these differences.

##### Reviewer comment 3.2 - Clarification/replacement of the terminology ‘pseudo-LCIA’

*“Line 213: I think that T-reX is a method to perform the interpretation of the LCI results, it does not evaluate impacts; as it is correctly reported in lines 409-410, T-reX method calculates waste and material inventory footprints, so the terminology”pseudo-LCIA” is in my opinion misleading (even if there is the “pseudo”) and it should be replaced with e.g. “LCI interpretation” method.”*

##### Author response 3.2 - Clarification/replacement of the terminology ‘pseudo-LCIA’

Following from the response to the previous comment, the authors agree that the distinction between LCI and LCIA is critical and the text has been refined so as to make this more explicit. The choice of the term ‘pseudo-LCIA method’ derives from the fact that the T-reX methods are integrated into the code in the same way as LCIA methods. This structure allows for a more user-friendly calculation of these LCI results, being customisable and easy to implement in existing frameworks. The authors have rephrased the relevant sections of the text to make this distinction more clear to the reader.

##### Reviewer comment 3.3 - Clarification on the single score calculation of T-reX

*“Line 300: please say how the single score is calculated for T-reX evaluation. Results section: as in a standard LCA, the results should be reported before starting with their interpretation, so please consider adding here for the analysed systems some of the results in absolute terms. This will really help the reader to understand the type of output that can be achieved with T-reX method.”*

##### Author response 3.3 - Clarification on the single score calculation of T-reX

As the reviewer correctly points out, the phrase "single score" is not clear, and could mislead the reader. T-reX gives a "single score" only in the sense of calculating an aggregate of the particular waste or material streams associated with the production of a given functional unit (in the case study, one kg. of various batteries). The language in subsection 2.2.1 ‘LCA calculations’ was amended to more explicitly convey this point, and further, to better explain the results presented in the case study. The phrase “single score”, was replaced with total summation which the authors believe conveys a more specific and appropriate description of the waste and material inventory results. The sentence now reads *"For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes."*

### **Minor comments**

##### Reviewer comment:

*“In general, when you refer to the SM section 6, please indicate also the sub-section.”*

##### Author response:

The reference to the subsection was incorrect, there is no subsection to appendix 6. This has been corrected in the revised manuscript

##### Reviewer comment:

*“In general, try to avoid putting words in ’’ and use this only when it is strictly necessary (e.g. line 433: why did you write ‘circular economy’ and not simply circular economy?)”*

##### Author response:

This has been corrected in several places in the revised manuscript.

##### Reviewer comment:

*“Line 128: what do you mean with ”generic waste processing models”? e.g. models that do not take into consideration the type of waste? Usually, databases include different treatment processes for the different types of waste, so please explain what you mean.”*

##### Author response:

This sentence has been rephrased in the revised manuscript (line X).

##### Reviewer comment:

*“Line 227:”the waste categories include incineration, recycling”: do you mean “the waste categories include waste sent to incineration, recycling”?“*

##### Author response:

Yes, that was the intended meaning. This has been corrected in the revised manuscript (2.1.2 paragraph 4).

##### Reviewer comment:

*“Lines 230-231:”a characterisation factor of -1 is applied to the waste footprint methods”: what do you mean? Usually, a CF is given to a flow.”*

##### Author response:

This sentence has been removed. Being somewhat technical, it has not a general relevance, and could be confusing, perhaps. The previous phrasing stems from the way that methods are constructed in the brightway code, where the characterisation factor is embedded in the method as a parameter. This is then applied to the relevant flow in the LCIA, as you correctly point out. The -1 corrects the perspective to be that of waste produced and not on the treatment services (as in “Waste is not a service”, Guinee 2021).

##### Reviewer comment:

*“Line 231: please explain the acronym CCS”*

##### Author response:

Yes. Thank you for pointing out the oversight. The expanded form of CCS has been moved to the first mention of carbon capture and storage in the text and the acronym has been added to the list of abbreviations in the front matter.

##### Reviewer comment:

*“Table 4 of the SM: there is sand but not gravel: why?”*

##### Author response:

There is no reason that gravel could not be included, the user of T-reX can decide on the material and waste streams of interest to them. The list of materials included in the default configuration of T-reX is based on the list of critical raw materials and then expanded to include other materials commonly deemed strategic or important. Some were added solely based on the interest of the authors (thus, sand and not gravel). The list is not supposed to be exhaustive or comprehensive and can be expanded or condensed by the user at their will. Instructions for configuration are included in the documentation of the software in its GitHub repository.

##### Reviewer comment:

*“Line 247: which are the methods are you referring to?”*

##### Author response:

This paragraph has been changed to make a more specific reference to the existing LCA methods (such as the Crustal Scarcity Indicator) that are described initially in the introduction (subsections 1.2.1 and 1.2.2). The textual connection was lost due to restructuring.

##### Reviewer comment:

*“Lines 251-253: T-reX does not calculate impacts, so what do you mean with ”including non-direct impacts on the market such as co-production of other materials”?“*

##### Author response:

The word “impacts” has been replaced with “effects” to avoid confusion (subsection 2.1.2, paragraph 8).

##### Reviewer comment:

*“Lines 255-257: are you referring to the cases where the mining of one material is done together with the extraction of another material? Please say more explicitly the situations are you referring to.”*

##### Author response:

Covered in the next response.

##### Reviewer comment:

*“Lines 258-260: where exactly can the reader see this aspect in sub-section 3.2?”*

##### Author response:

As the reviewer suggests, co-production is one example of such a situation.
It was previously intended to include a more detailed exploration of this through the example of nickle metal production, however, due to space constraints, this data will not be in the main section of the manuscript, but is discussed in more detail in the supplementary material (appendix subsection 3.2.2, paragraph 2).

##### Reviewer comment:

*“Line 260 and line 382: ”substitution”: substitution of what?“*

##### Author response:

This has been clarified in the text to be more explicit (subsection 3.2.1). The term “substitution” here refers to the subtraction of co-products from the system in question (those which are not the functional unit) using the standard LCA methodology.

##### Reviewer comment:

*“Lines 289 and 290: LT stays for what?”*

##### Author response:

“LT” stands for “long term”. The abbreviation has been replaced by the complete word (subsection 2.2, second list).

##### Reviewer comment:

*“Line 303-305: this sentence is not clear to me. Please consider rephrasing it.”*

##### Author response:

Indeed, the reference to the graph traversal algorithm used in brightway was not clearly explained here. The text has been revised to improve on the description of the calculation methodology (subsection 2.2.1).

##### Reviewer comment:

*“Lines 322 and 332: the title includes both waste and material but then nothing is said about the material inventory footprint in the sub-section.”*

##### Author response:

Yes, only one example was given to illustrate this type of result as the method and the form is identical for both waste and material inventory footprints. This has been made explicit in the text (subsection 3.2.1). The caption now reads *¨Sankey visualisation of flow inventory footprints"*

##### Reviewer comment:

*“Figure 2: if I consider arrows that enter the cobalt production box, I have 12% + 20% that is more than 22% (cobalt box): why?”*

##### Author response:

I have revised this figure (2) and added much more descriptive text. The Sankey diagram shown in  is a greatly abridged and simplified version of that produced by using T-reX with the ActivityBrowser software. The full Sankey diagram is far too large and complex to print. Examples of these diagrams are included in their full forms in the supplementary information (appendix 8).

##### Reviewer comment:

*“Figure 2 caption: here you say that the diagram shows the total solid waste footprint but at the top of the diagram I see the box ”treatment of inert waste, landfill”, so I do not understand how to read the diagram.”*

##### Author response:

*Figure 2 and its caption have been amended to portray a clearer depiction of the result that the authors are trying to present here. As in the response to the previous comment:  
 , the complexity of these flows make it challenging to contain them in a printable figure.*

##### Reviewer comment

*“Line 343: what do you mean with ”carbon dioxide waste”? Is it not better to call it “waste from CCS”?“*

##### Author response:

*As suggested, the text has been revised to replace “carbon dioxide waste” with “carbon dioxide waste from CCS”. And in the caption, to Figure 3(b) "carbon dioxide waste (produced from carbon capture and storage (CCS) only)".*

##### Reviewer comment:

*“Figure 3b: CCS instead of CSS?”*

##### Author response:

Yes, this typographical error has been corrected in the revised manuscript.

##### Reviewer comment:

*“Line 370: I think that ”T-reX Coal (black) demand method” should be replaced with “Coal (black) demand of T-reX method”. The same consideration is valid for the EDIP method (EDIP is the method, “coal no LT” is one of the results of the method).”*

##### Author response:

This phrasing has been amended as suggested by the reviewer (subsection 3.2.5 and  Fig 6)

##### Reviewer comment:

*“Lines 373 and 374: natural gas is given as example of fossil fuel related methods and zinc for metal demand methods, but natural gas and zinc are just elementary flows and not methods. So please use the correct terminology.”*

##### Author response:

The terminology has been corrected to be more clear (subsection 3.2.5).

##### Reviewer comment:

*“Lines 422-423: how can a waste represent a loss of resources or an environmental risk? Please explain it better, maybe adding some examples.”*

##### Author response:

The message here about the importance of knowing the specific nature and context of what is described as a waste flow in the LCA databases is of fundamental importance in the interpretation of the results obtained from T-reX. This paragraph has been revised to better explain this point (lines X-Y).

##### Reviewer comment:

*“Lines 436-437: I didn't understand the second part of this sentence. Please consider rephrasing it.”*

##### Author response:

This sentence has been rephrased in the revised manuscript make the database manipulation process more clear to the reader. "explode" has been replaced with deconstruct. The second sentence of the conclusion now reads *"It deconstructs, manipulates and reconstructs the database to identify and edit waste and material exchanges..."*

##### Reviewer comment:

*“Line 443: it does not seem to me that T-reX ”examines their connections so supply chain risks and potential environmental damage”: T-reX gives simply an overview of waste generated and materials consumed.”*

##### Author response:

Agreed. It was intended to convey that T-reX can aid the user in this examination by facilitating the generation of specific inventories.The final paragraph of the conclusion has been amended to soften this language and be more specific about what T-reX can do.

----
----

On behalf of all of the authors of T-reX, I would like to extend once our gratitude, the review process was very thorough and I think it has resulted in significant improvement to the manuscript.

All the best,

S.C. McDowall

----
----
----

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/00_T-reX_manuscript.tex
update

--- 2024-11-30 00_JCR-submission-folder_SECOND SUBMISSION/cas-dc-template.tex
@@
 
-%% 
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for cas-dc documentclass for 
-%% double column output.
-
-%\documentclass[a4paper,fleqn]{cas-dc}
+\documentclass[a4paper,fleqn]{cas-dc}
 
@@
 % use the longmktitle option.
+%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
 
-\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
-
-\usepackage[numbers]{natbib}
-%\usepackage[authoryear]{natbib}
+%\usepackage[numbers]{natbib}
+\usepackage[authoryear]{natbib}
 %\usepackage[authoryear,longnamesfirst]{natbib}
 
+\usepackage{booktabs}
 %%%Author macros
@@
 
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage[british]{babel}
+\usepackage{microtype}
+\usepackage{listings}
+\usepackage{xcolor}
+\definecolor{blue}{RGB}{33,150,209}
+\definecolor{red}{RGB}{255,0,0}
+\usepackage{xurl}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{lineno}
+\usepackage{float}
+
+\setlength{\linenumbersep}{5pt}
+% Save the old \texttt command
+\let\oldtexttt\texttt
+% Redefine the \UrlFont to use a smaller typewriter font
+\renewcommand{\UrlFont}{\ttfamily\small}
+% Redefine the \texttt command to use a smaller font size and allow line breaks
+\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
+\usepackage{hyperref}
+
+\usepackage{graphicx}
+
+% ------------------------------------------------------
+
 \begin{document}
+
 \let\WriteBookmarks\relax
@@
 % Short title
-\shorttitle{T-reX\@: Quantifying waste and material footprints in LCA}
+\shorttitle{T-reX: Quantifying waste and material footprints in LCA}
 
@@
 % Main title of the paper
-\title [mode = title]{T-reX\@: Quantifying waste and material footprints in current and future \\ Life Cycle Assessment (LCA) databases}
+\title[mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
 
 % Authors
-\author[1]{Stewart Charles McDowall}
+\author[1]{Stewart Charles McDowall}[orcid=0000-0001-0000-0000]
 \ead{s.c.mcdowall@cml.leidenuniv.nl}
@@
 \cortext[cor1]{Corresponding author:}
+\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.}
 
 \author[1]{Elizabeth Lanphear}
+\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing.}
 
 \author[1]{Stefano Cucurachi}
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}
 
 \author[1]{Carlos Felipe Blanco}
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}
 
@@
 \begin{abstract}
-
 	The quintessential principle of the circular economy is to minimise material consumption and waste generation.
@@
 
-Pursuant to these goals the authors have developed 'T-reX', an analytical tool for LCA, written in Python.
-T-reX extends on the existing computational ecosystem centered on Brightway and Premise and can facilitate the quantification of user-defined supply chain demands for activities in current and prospective scenarios.
-T-reX streamlines database manipulation for LCA practitioners, integrating methods to aggregate and analyse demand inventories that could expose hotspots of hidden risk.
+	Pursuant to these goals the authors have developed `T-reX', an analytical tool
+	for LCA, written in Python. T-reX extends on the existing computational
+	ecosystem centered on Brightway and Premise and can facilitate the
+	quantification of user-defined supply chain demands for activities in current
+	and prospective scenarios. T-reX streamlines database manipulation for LCA
+	practitioners, integrating methods to aggregate and analyse demand inventories
+	that could expose hotspots of hidden risk.
 
-With a simple case study using Lithium-ion batteries, we demonstrate some of T-reX's potential by detailling and exploring aspects of their waste and material footprints.
-Since such footprints are often linked to negative externalities, T-reX could support sustainable decision-making and contribute to the development of the circular economy by promoting the understanding of our material consumption and waste generation.
+	With a simple case study using Lithium-ion batteries, we demonstrate some of
+	T-reX's potential by detailing and exploring aspects of their waste and
+	material footprints. Since such footprints are often linked to negative
+	externalities, T-reX could support sustainable decision-making and contribute
+	to the development of the circular economy by promoting the understanding of
+	our material consumption and waste generation.
+
+\verb+\begin{keyword}+ \verb+\sep+circular economy \verb+\sep+waste \verb+\sep+material footprint \verb+\sep+life cycle assessment \verb+\sep+critical raw materials \verb+sep+supply chain \verb+\end{keyword}+
 
@@
 
-% Keywords
-\begin{keywords} 
-circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
-\end{keywords}
-
 % Graphical abstract
 \begin{graphicalabstract}
-  \includegraphics{figs/T-reX_graphical-abstract.png}
+	\includegraphics[width=\textwidth]{figures/T-reX_graphical-abstract.pdf}
 \end{graphicalabstract}
-
-\clearpage
-
-%% GRAPHICAL ABSTRACT FULL SIZE
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\textwidth]{figs/T-reX_graphical-abstract.png}
-	\caption{The graphical abstract of the T-reX program developed for waste and material footprinting in Life Cycle Assessment}
-	\label{fig:graphical_abstract}
-\end{figure}
-
-\clearpage
-
-% Switch to landscape for the included PDF
-\newgeometry{landscape}
-\includepdf[pages=1, width=\textheight]{figs/T-reX_graphical-abstract.pdf}
-\restoregeometry % Restore the original geometry
 
@@
 \begin{highlights}
-    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
+	\item T-reX, a new open-source program for quantifying waste and material flows in
+	Life Cycle Assessment (LCA)
 	\item Assesses supply risks by calculating demand inventories for critical materials
-    \item Simplifies quantification of user-specified waste and material footprint categories
-    \item Rapidly identifies hotspots of waste generation and material demand in supply chains
-    \item Presents a case study of five lithium-ion batteries to demonstrate T-reX's utility
+	\item Simplifies quantification of user-specified waste and material footprint
+	categories
+	\item Rapidly identifies hotspots of waste generation and material demand in supply
+	chains
+	\item Presents a case study of five lithium-ion batteries to demonstrate T-reX's
+	potential
 \end{highlights}
+
+% Keywords
+\begin{keywords}
+	circular economy \sep waste \sep material footprint \sep life cycle assessment \sep critical raw materials \sep supply chain
+\end{keywords}
+
+\maketitle
 
@@
 
-\maketitle
-
-% list of abbreviations
-\begin{table}[H]
+\begin{table*}
 	\centering
@@
 		\toprule
-      %? reviewer 3 - RE: CCS definition
 		CCS	              & Carbon Capture and Storage                                          \\
@@
 	\end{tabular}
-\end{table}
+\end{table*}
 
@@
 
-The development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised world have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+The development of a circular economy has become a central focus in the pursuit
+of sustainability objectives toward curtailment of our environmental footprint
+within planetary boundaries~\citep{eu2019greendeal,
+	eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+Fundamental to this development is a decrease in primary material consumption
+and a reduction of life cycle waste through the implementation of `re-X'
+strategies (e.g., refuse, rethink, design for---and implementation of---repair,
+remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign,
+	eu2022repair, eu2015reman}. In addition to circular economy goals, contemporary
+geopolitical tensions in an ever more globalised world have highlighted the
+vulnerability of many advanced economies to intentional supply
+disruptions---wrought as an act of competition or even outright
+hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
-The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
-More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
-Whether at the level of products, entire industries, nations, or even continents, the material footprint aims to quantify the total material consumed in supply chains.
-It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
+The concept of the `footprint' as an environmental sustainability indicator
+began with the Ecological Footprint
+(EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by
+the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the
+`footprint family' has adopted many additional metrics---albeit without yet
+coalescing into a coherent or consistent
+framework~\citep{giampietro2014footprintstonowhere,
+	vanham2019footprints,ridoutt2013footprints}. More recently, the footprint
+collection has been extended to include the Material Footprint
+(MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the
+`macro methods' of Industrial Ecology (IE), such as environmentally Extended
+Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material
+Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}. Whether at the level
+of products, entire industries, nations, or even continents, the material
+footprint aims to quantify the total material consumed in supply chains. It has
+been shown that the MF can be `highly representative of damage to human health
+and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this
+metric was recently beatified by the United Nations (UN), becoming the `core
+official indicator' for targets 8.4 and 12.2 of the Sustainable Development
+Goals (SDGs)~\citep{lenzen2022materialfootprint}.
 
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed.
-While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Though steadily
+emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint},
+the `Waste Footprint' (WF) metric is less well-developed. While it seems
+obvious that reducing life cycle waste is critical to the development of the
+circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF
+remains largely overlooked---especially in LCA models where waste itself is
+seldom apportioned any inherent environmental significance aside from the
+emissions related to its treatment~\citep{laurenti2023wastefootprint}. Such
+neglect strikes the authors as unjustified, given that it has been repeatedly
+demonstrated that WFs have a strong association with environmental
+damage~\citep{laurenti2023wastefootprint,doka2024publications,
+	ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}. Additionally, it has
+been shown that the most vulnerable communities suffer disproportionately from
+the social and ecological impacts created by wasteful behaviours and
+substandard end-of-life
+treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
-WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders.
-Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
+WF and MF, the two metrics of focus in the study, can provide a comprehensive
+assessment of potential environmental impacts across the supply chain,
+encapsulating both resource use and pollution/waste generation, while offering
+insights at diverse scales, from individual activities to global systems,
+facilitating communication with a diverse range of stakeholders. Thus, to
+reduce the negative externalities of human consumption and improve supply chain
+resilience, it is essential to uncover, dis-aggregate, and quantify the material
+and waste footprints of human activities in as much detail as
+possible~\citep{bisinella2024wastelca,
+	towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
 
-Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
-The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
-These scores can then be aggregated into one metric for each impact category that can be compared across products and processes.
-Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
+Life Cycle Assessment (LCA) is a useful method for the holistic estimation of
+the environmental impacts of products and processes. LCA can comprehensively
+evaluate impacts across the entire life cycle---from `cradle to grave'--- often
+identifying critical hotspots and guiding prioritisation of
+action~\citep{guinee2011lca}. The standard approach is to apply Life Cycle
+Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe}
+and CML~\citep{guinee2002cml}), which convert the inventory data into a set of
+impact scores based on the sum of the elementary flows (those between the
+technosphere and biosphere). These scores can then be aggregated into one
+metric for each impact category that can be compared across products and
+processes. Extending on standard LCA is \textit{ex-ante} LCA, which employs
+future scenarios to construct `prospective background databases' to predict the
+impacts of supply chains that have not yet been (and may never be)
+realised~\citep{cucurachi2018exante,blanco2020exante}.
 
-Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
+Several LCIA methods include, to some extent, waste generation (Swiss
+Eco-Factors, EDIP and
+EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and
+material consumption (Crustal Scarcity Indicator (CSI) and Swiss
+Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}). These methods,
+however, are generally limited in their scope (especially for waste), do not
+allow for flexible quantification of specific waste and material types, and
+often provide results in characterised units that are abstract or difficult to
+interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental
+	impact points] in the case of the Swiss
+Eco-Factors)~\citep{su2020sustainableproddev}.
 
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
+To facilitate the quantification of waste and material flows in LCA, we
+developed a Python package to extend the \texttt{brightway} open-source LCA
+framework~\citep{mutel2017brightway} and track these exchanges by translating
+them into inventory indicators and `pseudo' LCA impact (LCIA) categories. T-reX
+is `T(ool for) re-X' (reduce, reuse, recycle, etc.\@ \ldots) and enables LCA
+practitioners to manipulate their databases to allow them to easily aggregate
+the mass and volume of any desired exchange, and to create flexible categories
+that differentiate between material categories, waste types, and treatment or
+EOL handling.
 
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
-Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
-\texttt{premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
-The authors believe that there is an urgent need for research investment in this area.
+Integration with the \texttt{premise} Python
+package~\citep{sacchi2022premise}---which connects the projections of
+integrated assessment models (IAMs) with current LCA databases---enables the
+user to create and manipulate prospective LCA databases. Frustratingly, the
+current utility of prospective databases---in general, and in particular for
+the waste sector---is constrained by the fact that, to date, the sectoral
+coverage of future life cycle inventories (LCIs) is largely confined to energy,
+steel, cement, and transport~\citep{sacchi2023premisedocs}. Despite the ever
+more critical need to reliably model and future waste management
+systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent
+development in this field. \texttt{premise} offers an excellent structure for
+the integration of future waste and material flow modelling in LCA databases,
+but its potential contribution to the development of the circular economy
+cannot be realised if the scientific resources for data collection and
+validation are not forthcoming. The authors believe that there is an urgent
+need for research investment in this area.
 
-%! reviewer 3 - comment 1 - ADDITIONAL INFORMATION ON T-REX's differentiating features comparet to LCI
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
-In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and potentially highlight opportunities for waste reduction and material efficiency.
+The purpose of T-reX is not to quantify the environmental impacts of material
+consumption and waste production (the requisite data and impact modelling
+remain lacking), but rather to quantify the material and waste flows
+themselves, even those that are finally consumed by waste treatment processes.
+It provides, thus, not an impact assessment in the traditional sense, but an
+accounting of the material consumed and waste generated by a product or service
+inside of the technosphere, regardless of the end-of-life fate of these flows.
+By definition, the development of the `circular economy' necessitates the
+reduction and ultimate elimination of waste (though whether this objective is
+thermodynamically impossible has long been the subject of debate
+by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many
+others). In any case, minimising primary material consumption and the
+generation of waste is of paramount importance. By allowing LCA practitioners
+to easily classify and quantify these exchanges, T-reX provides a practical
+means to identify hotspots and potentially highlight opportunities for waste
+reduction and material efficiency.
 
-%? reviewer 2 - comment 3 - ADDITIONAL INFORMATION ON T-REX's FUTURE DEVELOPMENT
-With the continual improvement of LCA databases and the development of prospective databases, T-reX can produce ever more realistic estimates, enhancing its actionable utility derived from the exploration of the potential impacts of future waste management systems and the material demands of emerging technologies.
-The better the data, the more specific and accurate the results of T-reX analysis, and, thus, the more useful the tool becomes.
-The authors hope that T-reX will be a valuable asset to the LCA community and contribute to the development of a more sustainable and circular economy.
+With the continual improvement of LCA databases and the development of
+prospective databases, T-reX can produce ever more realistic estimates,
+enhancing its actionable utility derived from the exploration of the potential
+impacts of future waste management systems and the material demands of emerging
+technologies. The better the data, the more specific and accurate the results
+of T-reX analysis, and, thus, the more useful the tool becomes. The authors
+hope that T-reX will be a valuable asset to the LCA community and contribute to
+the development of a more sustainable and circular economy.
 
@@
 
-Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time.
-From a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-From a practical economic viewpoint, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows, in addition to energy costs and the relative prices of virgin and secondary materials.
-Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and `modelling [of its management] must be physically based'.
-The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.
+Though often described simply as a `material with a negative economic
+value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and
+one whose definition is poorly delineated and highly variable across space and
+time. From a systems perspective, the notion of waste is anathema to the
+`circular economy', they cannot co-exist. From a practical economic viewpoint,
+the viability of an extensive waste processing system is highly dependent on
+precise knowledge (or at least, reasonable predictions) of material and waste
+flows, in addition to energy costs and the relative prices of virgin and
+secondary materials. Thus, as~\cite{bisinella2024wastelca} argues, waste must
+be `thoroughly characterised' and `modelling [of its management] must be
+physically based'. The reliable, robust models needed to guide the development
+of the circular economy can be built only on a foundation of high-quality data.
 
-Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities.
-There remains, however, a knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
+Accurate and detailed information about waste and waste systems is, by
+definition, essential for understanding the `circularity' of a human activity
+and for predicting its life cycle externalities. There remains, however, a
+knowledge gap regarding WFs and their associated environmental
+impacts~\citep{laurenti2023wastefootprint}.
 
-Conventional LCA database models consider waste as a `service' (accounting for the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
-In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy.
-Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
-There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill.
-In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling of the long-term impacts of disposal.
+Conventional LCA database models consider waste as a `service' (accounting for
+the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and
+typically use generic waste processing models~\citep{beylot2018} that break the
+causal link between the functional unit and the waste-associated impacts. In
+LCA, waste flows are (almost exclusively) not considered as fundamental
+biosphere exchanges, but rather, as technosphere flows within the human
+economy. Waste produced by an activity is transferred to a relevant waste
+treatment activity where it is accepted `burden-free' and transformed into a
+combination of emissions and other waste
+`products'~\citep{guinee2021wasteisnotaservice}. There can be several treatment
+steps in this pathway leading, ultimately, to a mass of material being
+deposited in a landfill. In this system of waste accounting, the impacts
+apportioned to the waste-producing activity are a sum of those incurred by the
+transport, treatment, and final disposal of the waste into terrestrial or
+aquatic environments. In particular, the extensive work
+of~\cite{doka2024publications} has contributed significantly to understanding
+the environmental impacts of waste treatment processes and the modelling of the
+long-term impacts of disposal.
 
-A considerable portion of a product's total waste is generated during earlier stages of its supply chain such as resource extraction, transportation, and manufacturing, thus, often remaining `invisible' to conventional LCA accounting practices~\citep{laurenti2016wastefootprint}.
-This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
-Studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
-Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
-The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
+A considerable portion of a product's total waste is generated during earlier
+stages of its supply chain such as resource extraction, transportation, and
+manufacturing, thus, often remaining `invisible' to conventional LCA accounting
+practices~\citep{laurenti2016wastefootprint}. This oversight in measuring and
+communicating a cradle-to-grave product waste footprint (PWF) highlights a gap
+in circular economy indicators. Traditional LCA does not typically view waste
+as having environmental significance \textit{per se}, considering instead only
+emissions and resource use resulting from waste
+treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}. The
+environmental significance of waste and its correlation with other indicators
+has been the subject of extensive research. Studies have shown that popular
+resource footprints can cover a significant portion of environmental impact
+variance between
+activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
+Correlations between various environmental indicators are not always
+consistent, however, as seen with the carbon footprint, which often does not
+correlate well with other impact assessment
+scores~\citep{laurenti2012carbonfootprint}. The aggregation of waste in PWFs
+raises concerns among LCA experts, regarding the uncertainties introduced, as
+well as the potential misrepresentation of environmental performance due to
+differences in waste
+types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
 
-Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA~\citep{lauran2020abioticdepletion}.
-This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible methods to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
+Some existing LCA methodologies offer direct indicators at the impact
+assessment level, but generally, there is little attention given to the impacts
+of waste in LCA~\citep{lauran2020abioticdepletion}. This limitation becomes
+particularly evident when attempting to identify waste generation hotspots
+within a product's life cycle. Addressing these hotspots is crucial for
+advancing `circularity', but there is a lack of convenient and flexible methods
+to calculate waste flows in LCA and a pressing need for more comprehensive
+methods that can effectively quantify waste flows and, therefore, contribute to
+a greater understanding of a product's total environmental footprint.
 
@@
 
-In the context of geopolitical tension and a mineral-hungry `renewable energy transition', greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system rather than the primary material flows themselves.
+In the context of geopolitical tension and a mineral-hungry `renewable energy
+transition', greater attention is being given to the security of material
+supply---especially for those considered `critical raw materials'
+(CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
+While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its
+exchanges with the biosphere, its focus is often on the environmental impacts
+of the system rather than the primary material flows themselves.
 
-The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
-The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations.
-CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
-The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.
+The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an
+assessment of the long-term global scarcity of minerals in
+LCA~\citep{arvidsson2020csi}. The CSI introduced crustal scarcity potentials
+(CSPs), which are measured in kg–silicon–equivalents per kg–element and
+derived from crustal concentrations. CSPs, provided for 76 elements, reflect
+the long-term global elemental scarcity based on crustal concentration proxies.
+The CSI, calculated by multiplying CSPs with extracted masses, effectively
+gauges the impact of elemental extraction.
 
-While useful for its intended purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
-Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
+While useful for its intended purpose, the CSI presents its results in an
+abstract unit (kg-Si eq.) that is difficult to interpret and compare with other
+impact categories. Furthermore, the CSPs are not available for all elements (or
+more complex materials), and the method does not allow for the quantification
+of material demands in terms of mass or volume.
 
@@
 
-This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries, highlighting inconsistencies in LCA methodology and probing departures from projections in prospective models.
+This section is divided into two parts. In \autoref{sec:method-TreX}, we
+describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe
+the methodology used to calculate the waste and material inventory footprints
+in a case study of five Li-ion batteries, highlighting inconsistencies in LCA
+methodology and probing departures from projections in prospective models.
 
@@
 
-Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising especially  the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
-Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}.
-Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---and after running T-reX, the manipulated databases and the `pseudo-LCIA' methods created by T-reX can be used in the project in the accustomed way.
-T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is fully open-source under the CC-0 licence.
-The full source code for T-reX is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
-T-reX is designed to be used with \texttt{ecoinvent} databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
-Currently, it has been tested with all available system models of \texttt{ecoinvent}versions 3.5--3.10.
+Developed in the Python programming language, T-reX extends the
+\texttt{brightway LCA framework}, utilising especially the components
+\texttt{bw2data}, \texttt{bw2calc}, and
+\texttt{bw2io}~\citep{mutel2017brightway}. Additionally, the \texttt{wurst}
+package---which can deconstruct databases into a list of exchanges---is used to
+facilitate searching and data transformation at the exchange level (the
+individual supply chain flows)~\citep{mutel2017wurst}. Integration with the
+\texttt{premise} package~\citep{sacchi2022premise}---which integrates the
+projections of integrated assessment models (IAMs) into current LCA
+databases---enables the user to easily create and manipulate prospective LCA
+databases. T-reX is also compatible with
+\texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source
+graphical user interface for LCA---and after running T-reX, the manipulated
+databases and the `pseudo-LCIA' methods created by T-reX can be used in the
+project in the accustomed way. T-reX is installable via the Python Package
+Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is fully open-source under the
+CC-0 licence. The full source code for T-reX is indexed on
+Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the
+GitHub repository~\citep{mcdowall2024T-reXgithub}. T-reX is designed to be used
+with \texttt{ecoinvent} databases~\citep{ecoinvent2016version3}, but could be
+adapted to other databases by changing the search criteria. Currently, it has
+been tested with all available system models of \texttt{ecoinvent}versions
+3.5--3.10.
 
-T-reX can be used directly from the command line, or imported as a Python package with which the user can access the individual functions and modules.
-In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database.
-The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or a `prospective' database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.
+T-reX can be used directly from the command line, or imported as a Python
+package with which the user can access the individual functions and modules. In
+the simplest case, the user can run the program with the default settings,
+which will calculate the waste and material footprint of the
+\texttt{ecoinvent}database. The user can adapt the settings of T-reX as
+desired, to calculate alternative or additional waste and material inventory
+footprints, use a custom database, or a `prospective' database based on future
+scenarios by implementing the premise \texttt{premise} with one of T-reX's
+internal modules.
 
-The supplementary material in \autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
-Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
+The supplementary material in \autoref{sec:supplementary} contains metadata of
+T-reX, along with a list of the constituent modules, a description of their
+functions, and a detailed computational workflow. Further details can be found
+in the GitHub repository and package
+documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
 
@@
 
-T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
-We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of an activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
+T-reX is a Python package that enables one to produce LCA databases---both
+current and prospective---that are manipulated to facilitate the calculation of
+waste and material inventory footprints in the supply chain of any activity in
+the same way as one would apply standard LCIA methods. We refer to the T-reX
+methods as `pseudo-LCIA' methods, as they represent an accounting of an
+activity's technosphere inventory without reference to impact modelling, as in
+standard LCIA.
 
-If desired, prospective databases can be defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} and can be paired with a variety of mitigation scenarios.
+If desired, prospective databases can be defined by the user or constructed
+with the projections of the integrated assessment models such as
+IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer
+a range of options aligned with the Shared Socioeconomic Pathways
+(SSPs)~\citep{ssp2020ghg} and can be paired with a variety of mitigation
+scenarios.
 
-The deconstruction of the databases by T-reX into lists of exchanges allows the relevant flows of material and waste to be identified and categorised by the search functions.
-The queries facilitating this are tailored to the specific database and the user can easily modify them to suit their needs.
-The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand.
-The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a mirrored `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
+The deconstruction of the databases by T-reX into lists of exchanges allows the
+relevant flows of material and waste to be identified and categorised by the
+search functions. The queries facilitating this are tailored to the specific
+database and the user can easily modify them to suit their needs. The
+categories defined in the configuration are used to create T-reX's `pseudo
+LCIA' methods that are indicators of aggregated technosphere demand. The
+exchange editing function of T-reX then takes each list of exchanges and
+appends to the relevant activity a copy of the technosphere exchange as a
+mirrored `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
 
-In the default configuration of T-reX, there are 10 waste categories which are then further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include aggregations such as 'total waste', 'hazardous waste' as well as, those for waste sent to treatment by, for example, incineration or recycling.
-The full details are listed in the supplementary material in~\autoref{sec:supplementary} [section 3 ].
-One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
+In the default configuration of T-reX, there are 10 waste categories which are
+then further divided by their unit of measurement (kilograms and cubic meters)
+to create a total of 20 waste methods. The waste categories include
+aggregations such as 'total waste', 'hazardous waste' as well as, those for
+waste sent to treatment by, for example, incineration or recycling. The full
+details are listed in the supplementary material in~\autoref{sec:supplementary}
+[section 3 ]. One advantage of T-reX is that it can identify `hidden' waste
+exchanges that would otherwise be `consumed' by a treatment process and not
+leave the technosphere.
 
-In addition to these waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
-Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
-The identity of the materials considered, and their categorical groupings, are easily customisable by the user.
-The full list of 59 materials included in the default configuration is provided in the supplementary material in~\autoref{sec:supplementary}.
+In addition to these waste categories, T-reX can incorporate any number of
+material demand categories, the defaults being based on the EU Critical Raw
+Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30
+materials considered essential to the EU economy and also at risk of supply
+disruption. Especially valuable in the case of CRMs is T-reX's ability to
+uncover and quantify these threads of background consumption, which can be
+inconspicuously embedded in products or consumed in seemingly distant supply
+chain activities. The identity of the materials considered, and their
+categorical groupings, are easily customisable by the user. The full list of 59
+materials included in the default configuration is provided in the
+supplementary material in~\autoref{sec:supplementary}.
 
-Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the identification of material exchanges with T-reX differs from that used to identify waste exchanges.
-That is, the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
-A useful feature of T-reX is that, in cases where there are several markets for one material or material grouping, the program can aggregate these flows.
-For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
-Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.
+Following the logic of the standard LCA models~\cite{ecoinvent2016version3,
+	guinee2004economicallocation}, the identification of material exchanges with
+T-reX differs from that used to identify waste exchanges. That is, the search
+queries are based on the names of the relevant `market activities' for the
+material of interest rather than keywords in the exchange's name. A useful
+feature of T-reX is that, in cases where there are several markets for one
+material or material grouping, the program can aggregate these flows. For
+example, exchanges with markets for the rare-earth-elements (REEs) `market for
+cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated
+into a single indicator category for REEs. Similarly, the total demand for all
+critical raw materials (CRMs) can be calculated in the same manner.
 
-In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
+In these methods, the material demand is calculated based on the total mass
+that is extracted from the environment, thus, their focus is essentially solely
+on the mining-related exchanges that bring these materials from the biosphere
+into the technosphere.
 
-T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere, a material basis.
-This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct effects on the market, such as a 'negative consumption' of other materials due their co-production with the functional unit.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
-If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-For example, in the results of the Li-ion battery case study , this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
+T-reX, in contrast, uses accounting for material demand and waste generation
+based on exchanges solely within the technosphere, a material basis. This
+offers a different perspective, allowing for the estimation of overall supply
+chain material demands that consider the entire life cycle of an activity,
+including non-direct effects on the market, such as a 'negative consumption' of
+other materials due their co-production with the functional unit. Consider a
+demand for an activity containing a metal, for example; while the existing
+material use methods allow one to calculate the total mass of that metal that
+is extracted from the environment, T-reX can provide insight into the broader
+supply chain impacts of the demand for this metal. If the production of other
+materials is attributed to the production of this metal, these would appear as
+negative material demands in the T-reX results---supply chain pressure for one
+material can result in lessening of supply chain pressure for another. For
+example, in the results of the Li-ion battery case study , this is indeed the
+case for nickel demand, which, in the final inventory because of such
+co-production, is counter-intuitively negative (due to co-production and
+substitution in the LCA system models) despite the presence of nickel in the
+final products.
 
@@
 
-The workflow of T-reX is divided into several modules, each performing a separate function.
-The modules are designed to be used in a particular order, but the user can also use them individually to perform specific tasks.
-The standard workflow is as follows:
+The workflow of T-reX is divided into several modules, each performing a
+separate function. The modules are designed to be used in a particular order,
+but the user can also use them individually to perform specific tasks. The
+standard workflow is as follows:
 
@@
 	\item Creation of `pseudo-biosphere' databases
-    \item Creation of `pseudo-LCIA' methods to calculate waste and material inventory footprints
-    \item Exchange editing---whereby the technosphere exchange is mirrored as a `pseudo-biosphere' exchange
-    \item Database verification---to ensure that T-reX has manipulated the database correctly
+	\item Creation of `pseudo-LCIA' methods to calculate waste and material inventory
+	      footprints
+	\item Exchange editing---whereby the technosphere exchange is mirrored as a
+	      `pseudo-biosphere' exchange
+	\item Database verification---to ensure that T-reX has manipulated the database
+	      correctly
 \end{enumerate}
 
-This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere databases that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
+This workflow creates a copy of the original \texttt{brightway} project
+containing the original biosphere database, a T-reX `pseudo-biosphere' database
+along with one or more manipulated technosphere databases that can be used to
+calculate the waste and material inventory footprints of activities in the same
+way as standard LCIA calculations.
 
-An overview of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material in \autoref{sec:supplementary} contains a more detailed computational flowchart.
+An overview of the T-reX workflow is presented in
+\autoref{fig:methods-flowchart}. The supplementary material in
+\autoref{sec:supplementary} (part 4) contains a more detailed computational flowchart.
 
-\begin{figure}[H]
+\begin{figure}
 	\centering
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=9cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+	\includegraphics[width=0.9\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure}
@@
 
-To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories.
-The system boundaies.
-These following markets were used for the example calculations and unaltered by the authors:
-\begin{itemize}[itemsep=0pt]
+To provide an example, we investigated five types of Li-ion batteries in the
+\texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3},
+each represented by their unamended global market inventories. The system
+boundaries. These following markets were used for the example calculations and
+unaltered by the authors:
+\begin{itemize}
+	\itemsep=0pt
 	\item Li-ion, LFP, rechargeable, prismatic
@@
 
-In addition to the waste and material inventory footprint methods created by T-reX, the following standard LCIA methods were applied for comparison:
+In addition to the waste and material inventory footprint methods created by
+T-reX, the following standard LCIA methods were applied for comparison:
 
-\begin{itemize}[itemsep=0pt]
+\begin{itemize}
+	\itemsep=0pt
 	\item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
 	\item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
-    \item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
+	\item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long
+	      term emissions method~\citep{hauschild2003edip}
 	\item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
@@
 
-Additional to the present-day databases, T-reX created prospective database sets based on \texttt{ecoinvent} 3.9.1 using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
+Additional to the present-day databases, T-reX created prospective database
+sets based on \texttt{ecoinvent} 3.9.1 using the functionality of
+\texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle
+of the road' with the following Representative Concentration Pathways (RCPs):
 \begin{itemize}
-    \item SSP2-base: representing an approximate 3.5°C increase in global temperatures to 2100
-    \item SSP2-PkBudg500: representing the achievement of Paris climate goals, ca. 1.3°C increase to 2100
+	\item SSP2-base: representing an approximate 3.5°C increase in global temperatures to
+	      2100
+	\item SSP2-PkBudg500: representing the achievement of Paris climate goals, ca. 1.3°C
+	      increase to 2100
 \end{itemize}
 
-For each pathway, databases (at five-year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.
+For each pathway, databases (at five-year time intervals) were created with
+\texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the
+period between 2020 and 2100.
 
 \subsubsection{LCA calculations}
-For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes.
-Additionally, for the T-reX methods, a contribution analysis was performed with the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
-This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
-The groups of the resulting 'leaves' are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
-T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
+For each combination of activity, method, and database, a total summation was
+calculated along with details of the top contributing processes. Additionally,
+for the T-reX methods, a contribution analysis was performed with the
+\texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the
+\texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a
+component of the \texttt{brightway} ecosystem. This function performs one
+subjective method of graph traversal on the impact matrix of the LCA object to
+a specified cutoff. The groups of the resulting `leaves' are collected by their
+Cooperative Patent Classification (CPC) codes [giving a representation of
+		industry sector involvement]. T-reX provides, essentially, a tabulation of
+internal exchanges and an insight into the products and sectors in the supply
+chain of the activity that carry the most substantial responsibility for the
+final waste generation or material demand footprints.
 
+\section{Results}\label{sec:results}
 \subsection{T-reX tool}\label{sec:results-T-reX}
 
-An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, appendix 6.
-The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+An example of the output from the application of T-reX has been included in the
+supplementary material in~\autoref{sec:supplementary}, appendix 6. The
+manipulated \texttt{ecoinvent} databases (which are the main product of T-reX)
+can be recreated by following the instructions in the package
+documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 
-As described in~\autoref{sec:method-casestudy}, this case study calculated the waste and material footprints (with a variety of other indicators) for the unaltered background inventories of five Li-ion batteries with the functional unit being 1~kg of the battery at the global market.
-The purpose of this simple case study was to test, verify, and demonstrate the functionality and limitations of T-reX.
-This section includes some highlights of the results, with full results included in the supplementary material (\autoref{sec:supplementary}).
-Because the `pseudo-LCIA' methods created by T-reX are integrated into the user's \texttt{brightway} project as if they were standard LCIA methods, the footprint calculations can be performed in the customary way.
-In the supplementary material, there are screenshots of selected results obtained using the \texttt{ActivityBrowser} software, including contribution analysis and a Sankey diagram that disaggregates the final footprint result over the activities in the supply chain.
+As described in~\autoref{sec:method-casestudy}, this case study calculated the
+waste and material footprints (with a variety of other indicators) for the
+unaltered background inventories of five Li-ion batteries with the functional
+unit being 1~kg of the battery at the global market. The purpose of this simple
+case study was to test, verify, and demonstrate the functionality and
+limitations of T-reX. This section includes some highlights of the results,
+with full results included in the supplementary material
+(\autoref{sec:supplementary}). Because the `pseudo-LCIA' methods created by
+T-reX are integrated into the user's \texttt{brightway} project as if they were
+standard LCIA methods, the footprint calculations can be performed in the
+customary way. In the supplementary material, there are screenshots of selected
+results obtained using the \texttt{ActivityBrowser} software, including
+contribution analysis and a Sankey diagram that dis-aggregates the final
+footprint result over the activities in the supply chain.
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\centering
-    \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
-    \caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
+	\includegraphics[width=0.9\columnwidth]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
+	\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}\label{fig:sankey_waste}
 \end{figure}
@@
 %! CO2
-The inclusion of carbon capture and storage (CCS) in the prospective databases is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase in the production of carbon dioxide `waste' over the period from 2020--2040 that is not seen in the baseline scenario.
-This result highlights the fact that the (often) downward trends in global warming impacts calculated with prospective databases using standard LCIA methods are dependent on the assumptions made about the introduction of CCS technology.
-The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
-The authors consider that the over-representation of CCS and the under-representation of waste processing technologies in the prospective databases is one significant limitation to the validity of prospective LCA databases.
-Significant work is needed to improve the representation of these technologies in the databases to ensure that the results are more realistic and useful for decision-making.
+The inclusion of carbon capture and storage (CCS) in the prospective databases
+is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase
+in the production of carbon dioxide `waste' over the period from 2020--2040
+that is not seen in the baseline scenario. This result highlights the fact that
+the (often) downward trends in global warming impacts calculated with
+prospective databases using standard LCIA methods are dependent on the
+assumptions made about the introduction of CCS technology. The actual
+deployment of these technologies was only around 37~Mt CO$_2$/yr as of
+2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of
+the RCP scenarios~\citep{sacchi2023premisedocs}. The authors consider that the
+over-representation of CCS and the under-representation of waste processing
+technologies in the prospective databases is one significant limitation to the
+validity of prospective LCA databases. Significant work is needed to improve
+the representation of these technologies in the databases to ensure that the
+results are more realistic and useful for decision-making.
 %!!!!
 
-\begin{figure}[H]
+\begin{figure*}[]
 	\centering
-    \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+	\includegraphics[width=1.8\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
 	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
 The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
-\end{figure}
+\end{figure*}
 %!!!!
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\centering
-    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
-    \caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
+	\includegraphics[width=0.9\columnwidth]{figures/T-reX-wastefootprint-processcontributions.pdf}
+	\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model}\label{fig:top_contribution}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\centering
-    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
-    \caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
+	\includegraphics[width=0.9\columnwidth]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
+	\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model}\label{fig:cpc_contribution}
 \end{figure}
@@
 
-A comparison of the results for `Coal (black)' demand using the T-reX method with the LCIA method `EDIP 2003 - coal no LT' is shown in~\autoref{fig:comparison_methods}.
-In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
-Such comparability was also observed for other fossil-fuel-related demands (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
-Correlation between standard LCIA methods and T-reX methods is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
-In standard LCIA methods, impact scores are derived exclusively from the magnitude of the exchanges between the biosphere and the technosphere, that is, extraction and emission.
-In the T-reX methods, the footprint scores are based on an accounting of either waste generated or material demand, both of which are technosphere-technosphere exchanges in terms of LCA modelling.
-For the material demand methods especially, this distinction is critical.
-For example, the application of the EDIP 2003 or the CSP methods for a given metal will provide a score that is proportional to the amount extracted by mining, whereas the T-reX method provides an aggregation of the exchanges with the market for that metal.
-The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply chain pressures that are not captured by the standard LCIA methods.
-This makes the T-reX methods more sensitive to the modeling choices (e.g., economic allocation) that are generally embedded in LCA databases.
+A comparison of the results for `Coal (black)' demand using the T-reX method
+with the LCIA method `EDIP 2003 - coal no LT' is shown
+in~\autoref{fig:comparison_methods}. In this case, both the trends and the
+magnitude of the scores were very similar, both demonstrating a general
+decrease in the use of coking coal in the battery's footprints. Such
+comparability was also observed for other fossil-fuel-related demands (e.g.,
+natural gas and petroleum), and to a lesser extent, for some of the metal
+demand methods (e.g., zinc and cobalt). Correlation between standard LCIA
+methods and T-reX methods is not generally to be expected, however, due to
+fundamental differences in the way that the methods are constructed. In
+standard LCIA methods, impact scores are derived exclusively from the magnitude
+of the exchanges between the biosphere and the technosphere, that is,
+extraction and emission. In the T-reX methods, the footprint scores are based
+on an accounting of either waste generated or material demand, both of which
+are technosphere-technosphere exchanges in terms of LCA modelling. For the
+material demand methods especially, this distinction is critical. For example,
+the application of the EDIP 2003 or the CSP methods for a given metal will
+provide a score that is proportional to the amount extracted by mining, whereas
+the T-reX method provides an aggregation of the exchanges with the market for
+that metal. The T-reX method, therefore, considers cases of co-production,
+recycling, and substitution, providing a picture of the supply chain pressures
+that are not captured by the standard LCIA methods. This makes the T-reX
+methods more sensitive to the modeling choices (e.g., economic allocation) that
+are generally embedded in LCA databases.
 
-\begin{figure}[H]
+\begin{figure*}
 	\centering
-    \includegraphics[width=16cm]{figures/T-reX-coalANDedip.pdf}
+	\includegraphics[width=1.8\columnwidth]{figures/T-reX-coalANDedip.pdf}
 	\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
-\end{figure}
+\end{figure*}
 
@@
 
-In the most similar published method, \cite{laurenti2023wastefootprint} used an alternative procedure for calculating the waste footprint of \~1400 activities in the \texttt{ecoinvent} database version 3.5.
-Direct comparison is not possible, as this database contains only one generic Li-ion battery, `market for battery, Li-ion, rechargeable, prismatic'.
-The inventory of this battery most closely resembles that of the NMC 111 battery in this case study.
-\autoref{tab:results-case_study-comparison} presents a comparison of the results from the two studies, where possible.
-For liquid, solid, and recycled waste, the results were closely aligned, however, for the hazardous waste fraction, Laurenti et al.
-reported 95\%, whereas T-reX reported only 3\%.
-The reason for this discrepancy is explained by the fact that in the method of Laurenti et al.\, a ``waste flow was regarded as hazardous when the hazardousness was clearly stated in its subsequent waste treatment activity''.
-The authors continue: ``It should be noted that the high hazardousness ratios for many products might indicate a weakness in the validity of this measure''.
-In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
+In the most similar published method, \cite{laurenti2023wastefootprint} used an
+alternative procedure for calculating the waste footprint of \~1400 activities
+in the \texttt{ecoinvent} database version 3.5. Direct comparison is not
+possible, as this database contains only one generic Li-ion battery, `market
+for battery, Li-ion, rechargeable, prismatic'. The inventory of this battery
+most closely resembles that of the NMC 111 battery in this case study.
+\autoref{tab:results-case_study-comparison} presents a comparison of the
+results from the two studies, where possible. For liquid, solid, and recycled
+waste, the results were closely aligned, however, for the hazardous waste
+fraction, Laurenti et al. reported 95\%, whereas T-reX reported only 3\%. The
+reason for this discrepancy is explained by the fact that in the method of
+Laurenti et al.\, a ``waste flow was regarded as hazardous when the
+hazardousness was clearly stated in its subsequent waste treatment activity''.
+The authors continue: ``It should be noted that the high hazardousness ratios
+for many products might indicate a weakness in the validity of this measure''.
+In T-reX, the source database is deconstructed into a list of separate
+exchanges, and only those explicitly defined as hazardous are marked as such.
 
-\begin{table}[H]
+\begin{table}
 	\centering
@@
 
-Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting.
-Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
+\section{Discussion}\label{sec:discussion}
+Waste generation and material demand are strongly associated with the
+environmental impacts of human
+activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints,
+	demirer2019wastefootprint}, and thus must be represented in LCA accounting.
+Although there are existing LCIA methods that provide endpoint impact scores
+related to material demand and waste generation, they contain convoluted
+formulae or subjective weighting, or their complexity and lack of transparency
+can make them difficult to use and
+interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804,
+	arvidsson2020csi,foen2021ecofactors}.
 
-T-reX advances the state-of-the-art in LCA by providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
-Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
+T-reX advances the state-of-the-art in LCA by providing practitioners with a
+simple, flexible, and transparent way to calculate supply chain waste and
+material footprints, delivering results in standard units and as direct
+aggregations of the relevant demand inventories. Once LCA databases in the
+\texttt{brightway} project have been processed with T-reX, the user can easily
+apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and
+material inventory footprints in the same way that they would with a
+conventional LCIA method.
 
-The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility, and limitations of T-reX.
+The simple case study of five Li-ion batteries presented in this paper
+demonstrated the utility, flexibility, and limitations of T-reX.
 
-By adjusting the user configuration for future scenarios and waste/material categorisation, we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
-Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate.
-Additionally, visual exploration in \texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are structurally identical to standard LCIA methods when integrated within the \texttt{brightway} project.
+By adjusting the user configuration for future scenarios and waste/material
+categorisation, we produced a set of customised versions of current and
+prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX
+database. Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste
+and material inventory footprints for present and future supply chains were
+trivial to calculate. Additionally, visual exploration in
+\texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are
+structurally identical to standard LCIA methods when integrated within the
+\texttt{brightway} project.
 
-One limitation of T-reX is that it does not yet provide specific information in a readily accessible format on the composition of the waste generated.
-This information is required to thoroughly assess the potential environmental impacts of this waste.
-Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine the nature of the waste flows.
-For example, if the waste generated could represent an actual loss of resources or  potential environmental risk, or, as is often the case, simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
-A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
+One limitation of T-reX is that it does not yet provide specific information in
+a readily accessible format on the composition of the waste generated. This
+information is required to thoroughly assess the potential environmental
+impacts of this waste. Currently, the user would need to manually explore the
+waste footprint inventory produced by the application of the T-reX to determine
+the nature of the waste flows. For example, if the waste generated could
+represent an actual loss of resources or potential environmental risk, or, as
+is often the case, simply a transfer of the `overburden' in mining activity,
+which is classified as `inert waste'. A methodical classification of waste
+exchanges and end-of-life fates is expected to be facilitated by the
+increasingly detailed and disaggregated data in each successive release of
+\texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
 
-The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
-As demonstrated in the results of the case study---where there were often very few scenario-temporal changes in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adoption of future background inventories.
-In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
-A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.
+The utility of T-reX in studies of future supply chains is limited by the fact
+that the currently available prospective databases focus largely on changes in
+the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
+As demonstrated in the results of the case study---where there were often very
+few scenario-temporal changes in many waste and material footprint
+indicators---the utility of prospective LCA is restricted if there is
+inadequate adoption of future background inventories. In particular, the
+inclusion of scenarios with future waste processing technology could greatly
+improve our predictions of waste and material flows and offer valuable insight
+into their potential impacts~\citep{bisinella2024wastelca}. A strong focus on
+enhancing these prospective databases is, thus, of critical importance to the
+future of prospective LCA, and by extension, to the development of the circular
+economy.
 
@@
 
-T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It deconstructs, manipulates and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
-These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
+T-reX is an extension to the open-source \texttt{brightway} ecosystem that
+enables users to calculate the waste and material footprints of any product or
+service in an LCA database. It deconstructs, manipulates and reconstructs the
+database to identify and edit waste and material exchanges and writes matching
+custom `pseudo' LCIA methods. These exchanges are mirrored in the
+\texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and
+material footprints of any other supply chain flow of interest, such as water,
+gas, or critical raw materials, can be calculated similarly to existing LCIA
+methods.
 
-It is essential that we elucidate the complex relationships between material demand and waste generation and environmental and social harms.
-The software we present in this article enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their potential connections threats such as supply chain risks and environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of sustainability and the circular economy.
+It is essential that we elucidate the complex relationships between material
+demand and waste generation and environmental and social harms. The software we
+present in this article enables detailed exploration of `footprints' and life
+cycle inventories---modelled externalities of human activities---and examines
+their potential connections threats such as supply chain risks and
+environmental damage. T-reX is a step towards a better understanding of these
+interconnections for the practical realisation of sustainability and the
+circular economy.
 
 % Back matter
-\section*{Data availability}
-All data used is publicly available online under the noted sources.
-T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence.
-It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
-
-% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
-% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-
-\section*{CRediT authorship contribution visualisation} % IMPROVED VISUAL VERSION
-
-\begin{itemize}[noitemsep]
-    \item \textbf{SCM}: Stewart Charles McDowall
-    \item \textbf{EL}: Elizabeth Lanphear
-    \item \textbf{SC}: Stefano Cucurachi
-    \item \textbf{CFB}: Carlos Felipe Blanco
-\end{itemize}
-
-\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit-TREX.pdf}
-
-\section*{Declaration of competing interest}
-  The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-\section*{Acknowledgements}
-  Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
-101058522 (project FutuRaM --- futuram.eu).
-The authors would like to thank the reviewers for their valuable comments and suggestions.
-
-  The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems.
-The authors hope to continue the development of T-reX in the service of our shared ideals.
-
-  %! replace the sections
+%\appendix
 \section{Supplementary material}\label{sec:supplementary}
-  The supplementary material supplied with this manuscript contains the following:
+The supplementary material supplied with this manuscript contains the
+following:
 \begin{enumerate}
@@
 
+\section*{Data availability}
+All data used is publicly available online under the noted sources.
+T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence.
+It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
+The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} 
+and the development server is hosted by GitHub at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
+
+\section*{Declaration of competing interest}
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
+\section*{Acknowledgements}
+Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
+101058522 (project FutuRaM --- futuram.eu).
+The authors would like to thank the reviewers for their valuable comments and suggestions.
+
+The authors would like to acknowledge those who were part of the futures
+thinking small group session at the FutuRaM consortium meeting in Berlin in
+June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision
+presented by Stewart Charles McDowall of a comprehensive open-source software
+framework for the study and development of secondary raw material systems. The
+authors hope to continue the development of T-reX in the service of our shared
+ideals.
+
+% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
+% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
+% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
+% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+
+\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
+
+\begin{itemize}
+	\item \textbf{SCM}: Stewart Charles McDowall
+	\item \textbf{EL}: Elizabeth Lanphear
+	\item \textbf{SC}: Stefano Cucurachi
+	\item \textbf{CFB}: Carlos Felipe Blanco
+\end{itemize}
+
+\includegraphics[width=0.9\columnwidth]{figures/T-reX_credit_heatmap.pdf}
+
++\verb+\printcredits +\verb+\credit+ \printcredits
+
 %% Loading bibliography style file
@@
 % Loading bibliography database
-\bibliography{cas-refs-TREX}
+\bibliography{01_T-reX_manuscript.bib}
+
+% insert supplementary information
+%\clearpage
+%\includepdf{T-reX_supplementary-material/00_T-reX_supplementary-material_main.pdf}
 

### 2024-11-30 03_T-reX-latex_files-DRAFTING/cover_letter.tex
update

--- 2024-04-01 latex_files/cover_letter.tex
@@
 \documentclass[10pt]{letter}
-\usepackage[letterpaper,margin=2.2cm]{geometry}
-\usepackage[T1]{fontenc}
+\usepackage[letterpaper,margin=2cm]{geometry}
+% \usepackage[T1]{fontenc}
 \usepackage{libertine}
@@
 
-We are pleased to submit our original research article titled ``T-reX\: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
-This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.
+    We are pleased to submit our original research article titled ``T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases'' for consideration by \textit{Resources, Conservation and Recycling}.
+This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows, an essential aspect of progressing towards sustainable resource management and circular economy objectives.
 
-Our submission provides a comprehensive description of the tool, which integrates seamlessly with the Brightway2 framework to offer a user-friendly and adaptable solution for LCA practitioners.
+    Our submission provides a comprehensive description of the tool, which integrates seamlessly with the \texttt{brightway} framework to offer a user-friendly and adaptable solution for LCA practitioners.
 By automating data analysis and aggregation processes, our tool expedites the identification of environmental hotspots within supply chains, thus facilitating more informed decision-making.
@@
 
-\vspace{1em}
+    \vspace{0.5em}
 
@@
     % Signature block with names in a row using dummy images for alignment
-\begin{center}
-    \begin{minipage}[t]{0.25\textwidth}
-      \centering
+    % \begin{center}
+    \begin{minipage}[t]{0.9\textwidth}
+        % \centering
         \vspace{-2cm}
+        \hspace{0.1cm}
         \includegraphics[height=2cm]{signature_scm.png}\\ % Replace with your actual signature file name.
-      Stewart Charles McDowall\\
+        Stewart Charles McDowall   ---   on behalf of Elizabeth Lanphear, Stefano Cucurachi, and Carlos Felipe Blanco\\
         %   \textbf{Corresponding author}
     \end{minipage}%
-    \begin{minipage}[t]{0.25\textwidth}
-      \centering
-      \vspace{-2cm}
-      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
-      Elizabeth Lanphear
-    \end{minipage}%
-    \begin{minipage}[t]{0.25\textwidth}
-      \centering
-      \vspace{-2cm}
-      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
-      Stefano Cucurachi
-    \end{minipage}%
-    \begin{minipage}[t]{0.25\textwidth}
-      \centering
-      \vspace{-2cm}
-      \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
-      Carlos Felipe Rocha Blanco
-    \end{minipage}
-  \end{center}
+    % \begin{minipage}[t]{0.20\textwidth}
+    %     \centering
+    %     \vspace{-2cm}
+    %     \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
+
+    % \end{minipage}%
+    % \begin{minipage}[t]{0.20\textwidth}
+    %     \centering
+    %     \vspace{-2cm}
+    %     \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
+
+    % \end{minipage}%
+    % \begin{minipage}[t]{0.20\textwidth}
+    %     \centering
+    %     \vspace{-2cm}
+    %     \includegraphics[height=2cm]{dummy_signature.png}\\ % Invisible dummy image
+
+    % \end{minipage}
+    % \end{center}
 

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/T-reX_supplementary-material/T-reX_case-study_data/_01a_calculate-batteries.py

@@ -1,0 +1,376 @@
+import bw2data as bd
+import pandas as pd
+import bw2calc as bc
+from tqdm import tqdm
+from pathlib import Path
+import os
+import ast
+import contextlib
+import bw2analyzer as bwa
+
+ca = bwa.ContributionAnalysis()
+
+# Import user settings
+
+from _00_main import (
+    PROJECT,
+    KEY,
+    KEYWORDS_METHODS,
+    SEARCH_ACTIVITIES,
+    SEARCH_METHODS,
+    GET_RESULTS,
+    GET_SUPPLY_CHAIN_RESULTS,
+    FILE_ACTIVITIES,
+    FILE_METHODS,
+    FILE_RESULTS,
+    FILE_SUPPLYCHAIN_RESULTS,
+)
+
+
+if PROJECT in bd.projects:
+    bd.projects.set_current(PROJECT)
+else:
+    print(f"## Project {PROJECT} not found... ##")
+    print("Choose from:")
+    print(*bd.projects.report(), sep="\n")
+    exit_code = 1
+    raise SystemExit
+
+print(f'\n\n{"="*80}\n')
+print(f"\t\t\t LCA calculations with Brightway ")
+print(f'\n{"="*80}\n')
+
+print(f"Project: \n\t{PROJECT} (containing {len(bd.databases)} databases)\n")
+print(f"Activity search phrase:\n\t{KEY}\n")
+print(f"Method keywords: ", end="\n\t")
+print(*KEYWORDS_METHODS, sep="\n\t", end="\n")
+print(f'\n{"-"*80}\n')
+
+
+def parse_db_name(db_name):
+    # hacky way to deal with the standard ecoinvent names not matching the premise names
+    if db_name == "ecoinvent-3.9.1-cutoff":
+        db_name = "ecoinvent_cutoff_3.9_none_none0-none_2020"
+    if db_name == "ecoinvent-3.9.1-consequential":
+        db_name = "ecoinvent_consequential_3.9_none_none0-none_2020"
+
+    parts = db_name.split("_")
+    db_base = parts[0]
+    db_system = parts[1]
+    db_version = float(parts[2])
+    db_model = parts[3]
+    db_ssp = int(parts[4].split("-")[0][-1])
+    db_target = parts[4].split("-")[1]
+    db_year = int(parts[5])
+
+    return pd.Series(
+        {
+            "db_base": db_base,
+            "db_system": db_system,
+            "db_version": db_version,
+            "db_model": db_model,
[…]

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/T-reX_supplementary-material/T-reX_case-study_data/_05b_visualisation_bar_combine.py

@@ -1,0 +1,110 @@
+import os
+import xml.etree.ElementTree as ET
+from tqdm import tqdm
+from svglib.svglib import svg2rlg
+from reportlab.graphics import renderPDF
+from reportlab.pdfgen import canvas
+from reportlab.pdfbase.ttfonts import TTFont
+from reportlab.pdfbase import pdfmetrics
+
+from _00_main import DIR_VISUALISATION
+
+# Register Libertine font
+font_path = "misc/LinLibertine_Rah.ttf"  # Change as needed
+pdfmetrics.registerFont(TTFont("Libertine", font_path))
+
+# Directory containing SVG files
+svg_dir = DIR_VISUALISATION / "individual-methods_supply-chain"
+if not os.path.isdir(svg_dir):
+    raise ValueError(f"Directory not found: {svg_dir}")
+
+print(f'{"*"*50}')
+print(f"Reading SVG files from directory: {svg_dir}")
+print(f'{"*"*50}')
+
+# Define the layout and title
+filename = str(DIR_VISUALISATION / "combined_methods_supply-chain.pdf")
+title = "T-reX: Individual Contributions by CPC sector"
+columns = 5
+figures = sorted([f for f in os.listdir(svg_dir) if f.endswith(".svg")], reverse=True)
+figures_count = len(figures)
+rows = figures_count // columns + (1 if figures_count % columns else 0)
+
+if figures_count == 0:
+    raise ValueError("No SVG files found in the directory.")
+else:
+    print(f"Found {figures_count} SVG files.")
+
+# Initialize variables to store the maximum dimensions
+max_width = 205
+max_height = 532
+
+# Define margin size
+margin = 2  # You can adjust this value
+
+# # Iterate through each SVG file to find max dimensions
+# progress_bar = tqdm(total=len(figures), desc="Calculating maximum dimensions")
+# for svg_file in figures:
+#     file_path = os.path.join(svg_dir, svg_file)
+#     svg_obj = svg2rlg(file_path)
+#     max_width = max(max_width, svg_obj.width)
+#     max_height = max(max_height, svg_obj.height)
+#     min_width = min(max_width, svg_obj.width)
+#     min_height = min(max_height, svg_obj.height)
+#     progress_bar.update(1)
+# progress_bar.close()
+
+# print(f'{"*"*50}')
[…]

### 2024-11-30 latex_files-DRAFTING/.compile_pdf.sh

@@ -1,99 +1,0 @@
-#!/bin/bash
-
-# Define paths and filenames
-source_folder="."
- # Assuming your LaTeX files are in the current directory
-target_folder="../00_JRR-submission_folder"
-main_tex_file="main.tex"                   
-bib_file="references.bib"                  
-output_pdf_name="00_T-reX_manuscript.pdf"  
-output_docx_name="00_T-reX_manuscript.docx"
-output_html_name="00_T-reX_manuscript.html"
-
-# Optional: Specify Pandoc templates for DOCX and HTML
-docx_template="path/to/docx_template.docx"
-html_template="path/to/html_template.html"
-
-echo "======================================================="
-echo "Compiling $main_tex_file to PDF, DOCX, and HTML"
-
-# Create target folder if it doesn't exist
-mkdir -p "$target_folder"
-
-# Clean up before compiling
-latexmk -c -outdir="$target_folder" "$source_folder/$main_tex_file" || exit 1
-
-# Compile LaTeX document to PDF, output directly to target folder
-latexmk -pdf -outdir="$target_folder" "$source_folder/$main_tex_file" || exit 1
-
-# Conversion commands for DOCX and HTML with template usage if available
-if [ -f "$docx_template" ]; then
-    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" --reference-doc="$docx_template" -o "$target_folder/$output_docx_name" || exit 1
-else
-    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" -o "$target_folder/$output_docx_name" || exit 1
-fi
-
-if [ -f "$html_template" ]; then
-    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" --template="$html_template" -o "$target_folder/$output_html_name" || exit 1
-else
-    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" -o "$target_folder/$output_html_name" || exit 1
-fi
-
-echo "Process completed successfully."
-echo "PDF saved to $target_folder/$output_pdf_name"
-echo "DOCX saved to $target_folder/$output_docx_name"
-echo "HTML saved to $target_folder/$output_html_name"
-echo "======================================================="
-
-
-
-##OLDVERSION
-
-# #!/bin/bash
-# 
-# # Define paths and filenames
-# source_folder="." # Assuming your LaTeX files are in the current directory
[…]

### 2024-11-30 latex_files-DRAFTING/.other/latex_compile.sh

@@ -1,29 +1,0 @@
-#!/bin/bash
-echo "================================================================================"
-echo "Compiling the T-reX manuscript..."
-echo "-------------------------------------------------------------------------------"
-# Clean the project
-echo "Cleaning the project..."
-echo "-------------------------------------------------------------------------------"
-latexmk -C
-echo "-------------------------------------------------------------------------------"
-# Compile the project
-echo "Compiling the project..."
-echo "-------------------------------------------------------------------------------"
-latexmk -pdf -quiet main.tex
-echo
-echo "-------------------------------------------------------------------------------"
-echo "Done compiling the T-reX manuscript."
-echo "-------------------------------------------------------------------------------"
-# Copy the output PDF to the specified directory
-cp main.pdf ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf
-
-echo "Copied the output PDF to ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf"
-echo "Opening the PDF..."
-echo "================================================================================"
-echo " ----------------------           FIN!               ---------------------------"           
-echo "================================================================================"
-echo 
-echo
-# Open the PDF with the default viewer in a detached process
-# nohup xdg-open ../reviews/T-reX_Manuscript_V2/T-reX_manuscript_V2.pdf > /dev/null 2>&1 &

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/00_T-reX_manuscript.tex
commit\

@@
 \definecolor{blue}{RGB}{33,150,209}
-\definecolor{red}{RGB}{255,0,0}
+\definecolor{red}{RGB}{255,0,0}% \usepackage{listings}
 \usepackage{xurl}
-\usepackage{pdfpages}
+% \usepackage{pdfpages}
 \usepackage{enumitem}
-\usepackage{lineno}
+% \usepackage{lineno}
 \usepackage{float}
-
-\setlength{\linenumbersep}{5pt}
+% 
+% \setlength{\linenumbersep}{5pt}
 % Save the old \texttt command
@@
 % Short author
-\shortauthors{S.C. McDowall et~al.}
+\shortauthors{S2. McDowall et~al.}
 
@@
 % Authors
-\author[1]{Stewart Charles McDowall}[orcid=0000-0001-0000-0000]
+\author[1]{Stewart Charles McDowall}[orcid=0000-0002-5688-1279]
 \ead{s.c.mcdowall@cml.leidenuniv.nl}
@@
 \cortext[cor1]{Corresponding author:}
-\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.}
-
-\author[1]{Elizabeth Lanphear}
-\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing.}
-
-\author[1]{Stefano Cucurachi}
-\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}
-
-\author[1]{Carlos Felipe Blanco}
-\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}
+\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization}
+
+\author[1]{Elizabeth Lanphear}[orcid=0009-0003-6364-0873]
+\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing}
+
+\author[1]{Stefano Cucurachi}[orcid=0000-0003-2748-655X]
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
+
+\author[1]{Carlos Felipe Blanco}[orcid=0000-0001-8199-8420]
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
 
@@
 
-\begin{figure}
+\begin{figure*}
 	\centering
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-	\includegraphics[width=0.9\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
-\end{figure}
+	\includegraphics[width=1.8\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+\end{figure*}
 
@@
 assumptions made about the introduction of CCS technology. The actual
-deployment of these technologies was only around 37~Mt CO$_2$/yr as of
+deployment of these technologies was only around 37~Mt CO\(_2\)/yr as of
 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of
@@
 	\centering
-	\includegraphics[width=1.8\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+	\includegraphics[width=2.1\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
 	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
@@
 A comparison of the results for `Coal (black)' demand using the T-reX method
-with the LCIA method `EDIP 2003 - coal no LT' is shown
+with the LCIA method `EDIP 2003 – coal no LT' is shown
 in~\autoref{fig:comparison_methods}. In this case, both the trends and the
@@
 	\centering
-	\includegraphics[width=1.8\columnwidth]{figures/T-reX-coalANDedip.pdf}
-	\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
+	\includegraphics[width=2.0\columnwidth]{figures/T-reX-coalANDedip.pdf}
+	\caption{Comparison of the (a) `T-reX  	\caption{Comparison of the (a) `T-reX  Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
+Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
 \end{figure*}
@@
 	\centering
-	\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}\label{tab:results-case_study-comparison}
+	\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion')}\label{tab:results-case_study-comparison}
 	\begin{tabular}{lll}

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/00_T-reX_manuscript.tex
update

@@
 
-\usepackage{booktabs}
-%%%Author macros
+%Author macros
 \def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
@@
 
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[T1]{fontenc}
-\usepackage[british]{babel}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{xcolor}
-\definecolor{blue}{RGB}{33,150,209}
-\definecolor{red}{RGB}{255,0,0}% \usepackage{listings}
-\usepackage{xurl}
-% \usepackage{pdfpages}
-\usepackage{enumitem}
-% \usepackage{lineno}
-\usepackage{float}
+% Essential packages
+\usepackage{graphicx}
+\usepackage{float} % For precise figure placement
+\usepackage[authoryear]{natbib} % Use natbib for citations
+
+% Optional packages (uncomment only if necessary)
+%\usepackage{subcaption} % For subfigures
+%\usepackage{xcolor} % If custom colours are needed
+%\usepackage{lineno} % For line numbering (use \linenumbers to activate)
+
+% Ensure compatibility
+\usepackage{booktabs} % For professional-quality tables
+
+% \setlength{\linenumbersep}{5pt}
 %
-% \setlength{\linenumbersep}{5pt}
-% Save the old \texttt command
-\let\oldtexttt\texttt
-% Redefine the \UrlFont to use a smaller typewriter font
-\renewcommand{\UrlFont}{\ttfamily\small}
-% Redefine the \texttt command to use a smaller font size and allow line breaks
-\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
-\usepackage{hyperref}
-
 \usepackage{graphicx}
@@
 Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful quantitative method for this, especially given its capacity to expose the details of an activity's entire life cycle, often guiding the implementation of circular principles to where they could be most effective.
+ Life Cycle Assessment (LCA) is a powerful quantitative method for this,
+especially given its capacity to expose the details of an activity's entire life cycle, 
+often guiding the implementation of circular principles to where they could be most effective.
 
@@
 our material consumption and waste generation.
-
-\verb+\begin{keyword}+ \verb+\sep+circular economy \verb+\sep+waste \verb+\sep+material footprint \verb+\sep+life cycle assessment \verb+\sep+critical raw materials \verb+sep+supply chain \verb+\end{keyword}+
-
 \end{abstract}
@@
 
+\begin{keyword}
+circular economy \sep wste \sep material footprint \sep life cycle assessment \sep critical raw materials \sep supply chain
+\end{keyword}
+
 % Highlights
 \begin{highlights}
-	\item T-reX, a new open-source program for quantifying waste and material flows in
-	Life Cycle Assessment (LCA)
+	\item T-reX, a new open-source program for quantifying waste and material flows in	Life Cycle Assessment (LCA)
 	\item Assesses supply risks by calculating demand inventories for critical materials
-	\item Simplifies quantification of user-specified waste and material footprint
-	categories
-	\item Rapidly identifies hotspots of waste generation and material demand in supply
-	chains
-	\item Presents a case study of five lithium-ion batteries to demonstrate T-reX's
-	potential
+	\item Simplifies quantification of user-specified waste and material footprint categories
+	\item Rapidly identifies hotspots of waste generation and material demand in supply	chains
+	\item Presents a case study of five lithium-ion batteries to demonstrate T-reX's potential
 \end{highlights}
@@
 
-\maketitle
+%\maketitle
 
@@
 \section{Introduction}\label{sec:introduction}
+
 \subsection{Introduction to T-reX}\label{sec:intro-trex}
@@
 of sustainability objectives toward curtailment of our environmental footprint
-within planetary boundaries~\citep{eu2019greendeal,
-	eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+within planetary boundaries~\citep{eu2019greendeal,	eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+
 Fundamental to this development is a decrease in primary material consumption
@@
 often provide results in characterised units that are abstract or difficult to
-interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental
-	impact points] in the case of the Swiss
-Eco-Factors)~\citep{su2020sustainableproddev}.
+interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the SwissEco-Factors)~\citep{su2020sustainableproddev}.
 
@@
 resource footprints can cover a significant portion of environmental impact
-variance between
-activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
+variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
 Correlations between various environmental indicators are not always
@@
 of material demands in terms of mass or volume.
-
-% Methodology
 
@@
 functions, and a detailed computational workflow. Further details can be found
-in the GitHub repository and package
-documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
+in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
 
@@
 to create a total of 20 waste methods. The waste categories include
-aggregations such as 'total waste', 'hazardous waste' as well as, those for
+aggregations such as `total waste' `hazardous waste' as well as, those for`'
 waste sent to treatment by, for example, incineration or recycling. The full
 details are listed in the supplementary material in~\autoref{sec:supplementary}
-[section 3 ]. One advantage of T-reX is that it can identify `hidden' waste
+[section 3]. One advantage of T-reX is that it can identify `hidden' waste
 exchanges that would otherwise be `consumed' by a treatment process and not
@@
 
-Following the logic of the standard LCA models~\cite{ecoinvent2016version3,
-	guinee2004economicallocation}, the identification of material exchanges with
+Following the logic of the standard LCA models~\cite{ecoinvent2016version3, guinee2004economicallocation},
+ the identification of material exchanges with
 T-reX differs from that used to identify waste exchanges. That is, the search
@@
 An overview of the T-reX workflow is presented in
-\autoref{fig:methods-flowchart}. The supplementary material in
-\autoref{sec:supplementary} (part 4) contains a more detailed computational flowchart.
+\autoref{fig:methods-flowchart}. Section four of the supplementary material in
+\autoref{sec:supplementary} contains a more detailed computational flowchart.
 
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-	\includegraphics[width=1.8\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+	\includegraphics[width=1.7\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure*}
@@
 unaltered by the authors:
+
 \begin{itemize}
-	\itemsep=0pt
 	\item Li-ion, LFP, rechargeable, prismatic
@@
 \begin{itemize}
-	\itemsep=0pt
 	\item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
@@
 An example of the output from the application of T-reX has been included in the
-supplementary material in~\autoref{sec:supplementary}, appendix 6. The
+supplementary material in section 6 of the ~\autoref{sec:supplementary}. The
 manipulated \texttt{ecoinvent} databases (which are the main product of T-reX)
@@
 
-\autoref{fig:sankey_waste} presents a heavily simplified Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+\autoref{fig:sankey_waste} presents a heavily abridged Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
 As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material~\autoref{sec:supplementary}.
@@
 
-\begin{figure}
+\begin{figure}[H!]
 	\centering
-	\includegraphics[width=0.9\columnwidth]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
-	\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}\label{fig:sankey_waste}
+	\includegraphics[width=0.8\columnwidth]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
+	\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser showing the total solid waste footprint flows for the
+	 NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}label{fig:sankey_waste}
 \end{figure}
@@
 \subsubsection{Temporal and scenario variation in waste and material inventory footprints}
+
 %\label{sec:results-case_study-total_footprints}
@@
 
-\begin{figure*}[]
+\begin{figure*}[H]
 	\centering
-	\includegraphics[width=2.1\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+	\includegraphics[width=/textwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
 	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
@@
 \end{figure*}
-%!!!!
 
@@
 	\centering
-	\includegraphics[width=2.0\columnwidth]{figures/T-reX-coalANDedip.pdf}
-	\caption{Comparison of the (a) `T-reX  	\caption{Comparison of the (a) `T-reX  Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
-Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
+	\includegraphics[width=1.9\columnwidth]{figures/T-reX-coalANDedip.pdf}
+	\caption{Comparison of the (a) `T-reX – Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 – coal no longterm'.
+In the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
 \end{figure*}
@@
 
-In the most similar published method, \cite{laurenti2023wastefootprint} used an
+In the most similar published method,~\cite{laurenti2023wastefootprint} used an
 alternative procedure for calculating the waste footprint of \~1400 activities
@@
 The supplementary material supplied with this manuscript contains the
-following:
+following sections:
 \begin{enumerate}
@@
 
-\section*{Declaration of competing interest}
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
 \section*{Acknowledgements}
@@
 
+\section*{Declaration of competing interest}
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
 % \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
@@
 
-\includegraphics[width=0.9\columnwidth]{figures/T-reX_credit_heatmap.pdf}
-
-+\verb+\printcredits +\verb+\credit+ \printcredits
+\includegraphics[width=0.8\columnwidth]{figures/T-reX_credit_heatmap.pdf}
 
@@
 %\clearpage
-%\includepdf{T-reX_supplementary-material/00_T-reX_supplementary-material_main.pdf}
+%#\includepdf{T-reX_supplementary-material/00_T-reX_supplementary-material_main.pdf}
 

### 2024-12-02 00_T-reX_journal-upload-folder/T-reX_latex-files/T-reX_supplementary-material/T-reX_supplementary-material.tex
always one more...

--- 2024-03-27 latex_files/supplementary_material.tex
@@
 \usepackage{microtype}
-\usepackage[tt=false]{libertine}
+\usepackage{fontspec}
 \usepackage{xcolor}
 \usepackage{xurl}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{pdfpages}
@@
 \usepackage{tabularx}
+\setmainfont[Mapping=tex-text,Ligatures={Common,Rare,Discretionary}]{Linux Libertine}
 % Define colors
@@
 % Line numbers settings
-\setlength{\linenumbersep}{5pt}
-\linenumbers
+%\setlength{\linenumbersep}{5pt}
+%\linenumbers
 
@@
 
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \begin{document}
 
-\title{\textbf{supplementary material for the T-reX Manuscript}}
-\author{Stewart Charles McDowall, Elizabeth Lanphere, Carlos Felipe Blanco, Stefano Cucurachi \\
+\title{\textbf{Supplementary material for the T-reX Manuscript}}
+\author{Stewart Charles McDowall, Elizabeth Lanphear, Stefano Cucurachi, Carlos Felipe Blanco  \\
     \textit{Institute of Environmental Sciences (CML), Leiden University, Leiden, The Netherlands}}
-\date{\today}
+\date{1 December 2024}
 
@@
 \clearpage
+
+% \section*{Supplementary material}\label{sec:supplementary}
+% The supplementary material contains the following:
+% \begin{enumerate}
+%     \item Metadata of the T-reX Python package
+%     \item Details of the modules and user configuration
+%     \item Description of the computational workflow
+%     \item Modular flowchart of T-reX
+%     \item Example of the terminal output of T-reX
+%    \item{Python scripts used in the case study}
+%     \item Complete tabulated results of the case study
+%     \item Complete visualisations from the case study
+% \end{enumerate}
 
@@
         Documentation     & \url{T-reX.readthedocs.io}                   \\
-    Main dependencies & brightway2, premise, wurst \\
+        Main dependencies & \texttt{brightway, premise, wurst}           \\
         \bottomrule
@@
     \item \texttt{future\_scenarios}: Creates prospective LCA databases based on future scenarios.
-    \item \texttt{explode\_database}: Responsible for expanding a Brightway2 database into detailed exchange lists.
+    \item \texttt{explode\_database}: Responsible for expanding a \texttt{brightway} database into detailed exchange lists.
     \item \texttt{search\_waste}: Provides functions for searching and categorising waste generation-related exchange data.
@@
     \centering
-    \caption{Waste categories in the default configuration of T-reX.}\label{tab:waste_categories}
+    \caption{Waste categories in the default configuration of T-reX}\label{tab:waste_categories}
     \begin{tabularx}{\textwidth}{p{2.5cm}p{1cm}p{1.5cm}XX}
@@
 
-As an example, \autoref{tab:wf_results} presents a list of the number of waste exchanges identified in the prospective database built from `ecoinvent 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
+As an example, \autoref{tab:wf_results} presents a list of the number of waste exchanges identified in the prospective database built from `\texttt{ecoinvent} 3.9.1' according to the IAM model `REMIND' with the RCP `PkBudg500' in the year 2100.
 Note that the carbon dioxide waste category does not include emissions to the atmosphere, which is a typical focus of LCIA studies.
@@
     \centering
-    \caption{T-reX waste search result counts for the database `ecoinvent cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'.}\label{tab:wf_results}
+    \caption{T-reX waste search result counts for the database `\texttt{ecoinvent} cutoff 3.9.1, REMIND, SSP2, PkBudg500, 2100'}\label{tab:wf_results}
     \begin{tabular}{llr}
@@
 Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
-If the production other materials are attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
+If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
 In the results of the Li-ion battery case study in the paper to which this material is attached, it was, indeed, the case for the demand for nickel, which, because of such effects, is counter-intuitively negative despite the presence of nickel in the final products.
@@
 \begin{longtable}{ll}
-    \caption{A comprehensive list of various materials and their groupings in the default configuration.} \label{tab:list_materials} \\
+    \caption{A comprehensive list of various materials and their groupings in the default configuration}\label{tab:list_materials} \\
     \toprule
@@
 Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module.
-This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `T-reX'.
+This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the \texttt{brightway} project as a biosphere database named `T-reX'.
 
@@
 This module uses the custom `pseudo-biosphere' databases created by \texttt{make\_custom\_database} to create these waste and material footprint LCIA methods that have the same unit as the respective technosphere exchange.
-The methods are stored in the Brightway2 project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods.
+The methods are stored in the \texttt{brightway} project and can be used for calculating the waste and material footprints of activities in the LCA database in the same way as with other LCIA methods.
 Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (with the exception of CCS exchanges), changing the perspective from waste consumed by treatment to waste generated by the activity.
@@
 
-\autoref{fig:flowchart} shows the logical flow through the individual modules of T-reX tool.
+\autoref{fig:flowchart} shows the logical flow through the individual modules of T-reX tool in a generalised (a) and module-centric (b) format.
 
@@
     \centering
-    \includegraphics[height=0.8\textheight]{figures/T-reX_flowchart.pdf}
-    \caption{Computational flowchart of T-reX.
+    \includegraphics[height=0.8\textheight]{T-reX_flowchart-supplementary.pdf}
+    \caption{The generalised (a) and computational flowcharts of T-reX.
 In green are the individual modules, in blue are their outputs.
@@
 
-Due to its length, an example of the terminal output of T-reX is included as a separate file in the supplementary material, named `T-reX\_example\_output.html', which can be opened in an internet browser.
-The example output is from the execution of T-reX on the `ecoinvent 3.9.1' database with the IAM model `REMIND' and the RCP `PkBudg500' in the years 2065 and 2100.
+Due to its length, an example of the terminal output of T-reX is included as a separate file in the supplementary material, named `T-reX\_example\_terminal\_output.pdf'.
+The example output is from the execution of T-reX on the `\texttt{ecoinvent} 3.9.1' database with the IAM model `REMIND' and the RCP `PkBudg500' in the years 2065 and 2100.
 The example output includes the execution of the main module which in turn executes the \texttt{future\_scenarios}, \texttt{explode\_database}, \texttt{search\_waste}, \texttt{search\_material}, \texttt{make\_custom\_database}, \texttt{method\_editor}, \texttt{exchange\_editor}, and \texttt{verify\_database} modules.
 
-Due to licensing restrictions, the manipulated ecoinvent databases produced by T-reX are not included in the supplementary material.
-However, with a licence to the ecoinvent database, the user can easily reproduce the results of T-reX by following the instructions in the documentation.
+Due to licensing restrictions, the manipulated \texttt{ecoinvent} databases produced by T-reX are not included in the supplementary material.
+However, with a licence to the \texttt{ecoinvent} database, the user can easily reproduce the results of T-reX by following the instructions in the documentation.
+
+\clearpage
+
+\section{Python scripts used in the case study}
+
+The Python scripts used in the case study are included as separate \texttt{.py} files in the `case-study' folder of the supplementary material.
+The terminal output of the case-study calculations is included in the `case-study' folder as a \texttt{.html} file, which can be opened in an internet browser and as a plain text file.
+
+\clearpage
+\section{Complete tabulated results of the case study}
+
+Due to their length, the complete tabulated results of the case study are included as separate files in the `case-study/data\_batteries' folder of the supplementary material.
+The results are presented in the form of \texttt{.csv} files.
+More are available in the GitHub repositories \url{https://github.com/Stew-McD/T-reX_Publication} (publication) and \url{https://github.com/Stew-McD/T-reX} (codebase).
+
+\clearpage
+
+\section{Complete visualisations from the case study}
+
+Due to their size, the complete visualisations from the case study are included as separate files in the `case-study/visualisation\_batteries' folder of the supplementary material.
+The visualisations are presented in the form of merged \texttt{.pdf} files and separate \texttt{.svg} file.
+More are available in the GitHub repositories \url{https://github.com/Stew-McD/T-reX_Publication} (publication) and \url{https://github.com/Stew-McD/T-reX} (codebase).
 

### 2024-12-02 20240201_T-reX_JournalUploadFolder/main.tex
always one more...

--- 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/00_T-reX_manuscript.tex
@@
-%% 
-%% Copyright 2019-2024 Elsevier Ltd
-%% 
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3c of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3c or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%%
-%
-% T-reX manuscript using the elsevier template for JRC
-%
-% Repository for the T-reX manuscript at https://github.com/Stew-McD/T-reX_Publication
-% Repository for the T-reX python program at https://github.com/Stew-McD/T-reX
-%
-% S.C.McDowall - github/Stew-McD
-%
-%%%%
-
 \documentclass[a4paper,fleqn]{cas-dc}
@@
 
-%\usepackage[numbers]{natbib}
-\usepackage[authoryear]{natbib}
-%\usepackage[authoryear,longnamesfirst]{natbib}
-
-%Author macros
 \def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
@@
 \tsc{CFB}
+
+%\usepackage[numbers]{natbib}
+\usepackage[authoryear]{natbib}
+% \usepackage[authoryear,longnamesfirst]{natbib}
+%\usepackage{booktabs}
+%\usepackage[tt=false]{libertine}
+%\usepackage{xurl}
+
 %%%
-
-% Essential packages
-\usepackage{graphicx}
-\usepackage{float} % For precise figure placement
-\usepackage[authoryear]{natbib} % Use natbib for citations
-
-% Optional packages (uncomment only if necessary)
-%\usepackage{subcaption} % For subfigures
-%\usepackage{xcolor} % If custom colours are needed
-%\usepackage{lineno} % For line numbering (use \linenumbers to activate)
-
-% Ensure compatibility
-\usepackage{booktabs} % For professional-quality tables
-
-% \setlength{\linenumbersep}{5pt}
-%
-\usepackage{graphicx}
-
-% ------------------------------------------------------
-
 \begin{document}
-
 	\let\WriteBookmarks\relax
@@
 	% Short author
-\shortauthors{S2. McDowall et~al.}
+	\shortauthors{McDowall, S2. et~al.}
 
 	% Main title of the paper
-\title[mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
+	\title [mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
 
-% Authors
+	% First author
+	%
+	% Options: Use if required
+	% eg: \author[1,3]{Author Name}[type=editor,
+	%       style=chinese,
+	%       auid=000,
+	%       bioid=1,
+	%       prefix=Sir,
+	%       orcid=0000-0000-0000-0000,
+	%       facebook=<facebook id>,
+	%       twitter=<twitter id>,
+	%       linkedin=<linkedin id>,
+	%       gplus=<gplus id>]
+
 	\author[1]{Stewart Charles McDowall}[orcid=0000-0002-5688-1279]
+
+	% Corresponding author indication
+	\cormark[1]
+	% Email id of the first author
 	\ead{s.c.mcdowall@cml.leidenuniv.nl}
-\cormark[1]
-\cortext[cor1]{Corresponding author:}
-\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization}
+	% URL of the first author
+	\ead[url]{scmcd.ch}
+	% Credit authorship
+	\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing: original draft, Writing: review \& editing, Visualization}
+
+	\affiliation[1]{
+		organization={Institute of Environmental Sciences (CML), Leiden University},
+		addressline={P.O. Box 9518},
+		city={Leiden},
+		postcode={2300RA},
+		state={South Holland},
+		country={The Netherlands}}
 
@@
 
-\affiliation[1]{
-	organization={Institute of Environmental Sciences (CML), Leiden University},
-	addressline={P.O. Box 9518},
-	city={Leiden},
-	postcode={2300 RA},
-	state={South Holland},
-	country={The Netherlands}}
+	% Corresponding author text
+	\cortext[1]{Corresponding author}
 
-% Abstract\\
+	% Here goes the abstract
 	\begin{abstract}
-The quintessential principle of the circular economy is to minimise material consumption and waste generation.
-Thus, identifying and quantifying waste and material flows is critical.
- Life Cycle Assessment (LCA) is a powerful quantitative method for this,
-especially given its capacity to expose the details of an activity's entire life cycle, 
-often guiding the implementation of circular principles to where they could be most effective.
+		The quintessential principle of the circular economy is to minimise material consumption and waste generation. Thus, identifying and quantifying waste and material flows is critical.
 
-Pursuant to these goals the authors have developed `T-reX', an analytical tool
-for LCA, written in Python. T-reX extends on the existing computational
-ecosystem centered on Brightway and Premise and can facilitate the
-quantification of user-defined supply chain demands for activities in current
-and prospective scenarios. T-reX streamlines database manipulation for LCA
-practitioners, integrating methods to aggregate and analyse demand inventories
-that could expose hotspots of hidden risk.
+		Life Cycle Assessment (LCA) is a powerful quantitative method for this, especially given its capacity to expose the details of an activity's entire life cycle, often guiding the implementation of circular principles to where they could be most effective.
 
-With a simple case study using Lithium-ion batteries, we demonstrate some of
-T-reX's potential by detailing and exploring aspects of their waste and
-material footprints. Since such footprints are often linked to negative
-externalities, T-reX could support sustainable decision-making and contribute
-to the development of the circular economy by promoting the understanding of
-our material consumption and waste generation.
+		Pursuant to these goals the authors have developed `T-reX', an analytical tool for LCA, written in Python.
+T-reX extends on the existing computational ecosystem centered on \texttt{Brightway} and \texttt{Premise} and can facilitate the quantification of user-defined supply chain demands for activities in current and prospective scenarios.
+T-reX streamlines database manipulation for LCA practitioners, integrating methods to aggregate and analyse demand inventories that could expose hotspots of hidden risk.
+
+		With a simple case study using lithium-ion batteries, we demonstrate some of T-reX's potential by detailing and exploring aspects of their waste and material footprints.
+Since such footprints are often linked to negative externalities, T-reX could support sustainable decision-making and contribute to the development of the circular economy by promoting the understanding of our material consumption and waste generation.
 	\end{abstract}
 
-% Graphical abstract
+	% Use if graphical abstract is present
 	\begin{graphicalabstract}
-	\includegraphics[width=\textwidth]{figures/T-reX_graphical-abstract.pdf}
+		\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract.pdf}
+		%\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract_MID-RES.png}
 	\end{graphicalabstract}
-
-\begin{keyword}
-circular economy \sep wste \sep material footprint \sep life cycle assessment \sep critical raw materials \sep supply chain
-\end{keyword}
 
@@
 	\begin{highlights}
-	\item T-reX, a new open-source program for quantifying waste and material flows in	Life Cycle Assessment (LCA)
+		\item T-reX, a new open-source program for waste \& material footprints in Life Cycle Assessment (LCA)
 		\item Assesses supply risks by calculating demand inventories for critical materials
@@
 	% Keywords
+	% Each keyword is seperated by \sep
+
 	\begin{keywords}
-	circular economy \sep waste \sep material footprint \sep life cycle assessment \sep critical raw materials \sep supply chain
+		circular economy, waste footprint, material footprint, life cycle assessment, critical raw materials, supply chain
 	\end{keywords}
 
-%\maketitle
+	\maketitle
 
-%\nocite{*}
-
-\begin{table*}
+	% Main text
+	\begin{table*}[!htbp]
 		\centering
-	\caption*{List of abbreviations and terminology}\label{tab:abbreviations}
+		\caption{List of abbreviations}\label{tab:abbreviations}
 		\begin{tabular}{ll}
@@
 
-% Main text
-
-% Introduction
 	\section{Introduction}\label{sec:introduction}  
-
 	\subsection{Introduction to T-reX}\label{sec:intro-trex}
 
-The development of a circular economy has become a central focus in the pursuit
-of sustainability objectives toward curtailment of our environmental footprint
-within planetary boundaries~\citep{eu2019greendeal,	eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+	The development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal,eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 
-Fundamental to this development is a decrease in primary material consumption
-and a reduction of life cycle waste through the implementation of `re-X'
-strategies (e.g., refuse, rethink, design for---and implementation of---repair,
-remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign,
-	eu2022repair, eu2015reman}. In addition to circular economy goals, contemporary
-geopolitical tensions in an ever more globalised world have highlighted the
-vulnerability of many advanced economies to intentional supply
-disruptions---wrought as an act of competition or even outright
-hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+	Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X'v strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
+In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised world have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outrighthostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
-The concept of the `footprint' as an environmental sustainability indicator
-began with the Ecological Footprint
-(EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by
-the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the
-`footprint family' has adopted many additional metrics---albeit without yet
-coalescing into a coherent or consistent
-framework~\citep{giampietro2014footprintstonowhere,
-	vanham2019footprints,ridoutt2013footprints}. More recently, the footprint
-collection has been extended to include the Material Footprint
-(MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the
-`macro methods' of Industrial Ecology (IE), such as environmentally Extended
-Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material
-Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}. Whether at the level
-of products, entire industries, nations, or even continents, the material
-footprint aims to quantify the total material consumed in supply chains. It has
-been shown that the MF can be `highly representative of damage to human health
-and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this
-metric was recently beatified by the United Nations (UN), becoming the `core
-official indicator' for targets 8.4 and 12.2 of the Sustainable Development
-Goals (SDGs)~\citep{lenzen2022materialfootprint}.
+	The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet
+	coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere,vanham2019footprints,ridoutt2013footprints}. More recently, the footprint
+	collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the
+	`macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material
+	Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}. Whether at the level of products, entire industries, nations, or even continents, the material
+	footprint aims to quantify the total material consumed in supply chains. It has been shown that the MF can be `highly representative of damage to human health
+	and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core
+	official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
 
-Though steadily
-emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint},
-the `Waste Footprint' (WF) metric is less well-developed. While it seems
-obvious that reducing life cycle waste is critical to the development of the
-circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF
-remains largely overlooked---especially in LCA models where waste itself is
-seldom apportioned any inherent environmental significance aside from the
-emissions related to its treatment~\citep{laurenti2023wastefootprint}. Such
-neglect strikes the authors as unjustified, given that it has been repeatedly
-demonstrated that WFs have a strong association with environmental
-damage~\citep{laurenti2023wastefootprint,doka2024publications,
-	ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}. Additionally, it has
-been shown that the most vulnerable communities suffer disproportionately from
-the social and ecological impacts created by wasteful behaviours and
-substandard end-of-life
-treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+	Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed.
+While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
+Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
-WF and MF, the two metrics of focus in the study, can provide a comprehensive
-assessment of potential environmental impacts across the supply chain,
-encapsulating both resource use and pollution/waste generation, while offering
-insights at diverse scales, from individual activities to global systems,
-facilitating communication with a diverse range of stakeholders. Thus, to
-reduce the negative externalities of human consumption and improve supply chain
-resilience, it is essential to uncover, dis-aggregate, and quantify the material
-and waste footprints of human activities in as much detail as
-possible~\citep{bisinella2024wastelca,
-	towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
+	WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems,facilitating communication with a diverse range of stakeholders.
+Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca,towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
 
-Life Cycle Assessment (LCA) is a useful method for the holistic estimation of
-the environmental impacts of products and processes. LCA can comprehensively
-evaluate impacts across the entire life cycle---from `cradle to grave'--- often
-identifying critical hotspots and guiding prioritisation of
-action~\citep{guinee2011lca}. The standard approach is to apply Life Cycle
-Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe}
-and CML~\citep{guinee2002cml}), which convert the inventory data into a set of
-impact scores based on the sum of the elementary flows (those between the
-technosphere and biosphere). These scores can then be aggregated into one
-metric for each impact category that can be compared across products and
-processes. Extending on standard LCA is \textit{ex-ante} LCA, which employs
-future scenarios to construct `prospective background databases' to predict the
-impacts of supply chains that have not yet been (and may never be)
-realised~\citep{cucurachi2018exante,blanco2020exante}.
+	Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
+LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
+The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
+These scores can then be aggregated into one metric for each impact category that can be compared across products and processes.
+Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
 
-Several LCIA methods include, to some extent, waste generation (Swiss
-Eco-Factors, EDIP and
-EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and
-material consumption (Crustal Scarcity Indicator (CSI) and Swiss
-Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}). These methods,
-however, are generally limited in their scope (especially for waste), do not
-allow for flexible quantification of specific waste and material types, and
-often provide results in characterised units that are abstract or difficult to
+	Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
+These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to
 	interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the SwissEco-Factors)~\citep{su2020sustainableproddev}.
@@
 	them into inventory indicators and `pseudo' LCA impact (LCIA) categories. T-reX
-is `T(ool for) re-X' (reduce, reuse, recycle, etc.\@ \ldots) and enables LCA
+	is `T[ool for] re-X' (reduce, reuse, recycle, etc.) and enables LCA
 	practitioners to manipulate their databases to allow them to easily aggregate
@@
 	~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' .
+The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged'.
 In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
@@
 
-Developed in the Python programming language, T-reX extends the
-\texttt{brightway LCA framework}, utilising especially the components
+	Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising especially the components
 	\texttt{bw2data}, \texttt{bw2calc}, and
@@
 
-The supplementary material in \autoref{sec:supplementary} contains metadata of
+	The supplementary material in \autoref{sec:supplementary-1} contains metadata of
 	T-reX, along with a list of the constituent modules, a description of their
 	functions, and a detailed computational workflow. Further details can be found
-in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
+	in the GitHub repository and in the package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
 
@@
 	to create a total of 20 waste methods. The waste categories include
-aggregations such as `total waste' `hazardous waste' as well as, those for`'
-waste sent to treatment by, for example, incineration or recycling. The full
-details are listed in the supplementary material in~\autoref{sec:supplementary}
-[section 3]. One advantage of T-reX is that it can identify `hidden' waste
+	aggregations such as `total waste' `hazardous waste' as well as, those for
+	`waste sent to treatment' by, for example, incineration or recycling. The full
+	details are listed in the supplementary material in~\autoref{sec:supplementary-3}. One advantage of T-reX is that it can identify `hidden' waste
 	exchanges that would otherwise be `consumed' by a treatment process and not
@@
 	materials included in the default configuration is provided in the
-supplementary material in~\autoref{sec:supplementary}.
+	supplementary material in~\autoref{sec:supplementary-3}.
 
@@
 	chain material demands that consider the entire life cycle of an activity,
-including non-direct effects on the market, such as a 'negative consumption' of
+	including non-direct effects on the market, such as a `negative consumption' of
 	other materials due their co-production with the functional unit. Consider a
@@
 	material can result in lessening of supply chain pressure for another. For
-example, in the results of the Li-ion battery case study , this is indeed the
+	example, in the results of the Li-ion battery case study, this is indeed the
 	case for nickel demand, which, in the final inventory because of such
@@
 	An overview of the T-reX workflow is presented in
-\autoref{fig:methods-flowchart}. Section four of the supplementary material in
-\autoref{sec:supplementary} contains a more detailed computational flowchart.
+	\autoref{fig:methods-flowchart}. The supplementary material in
+	\autoref{sec:supplementary-3} provides more detailed information.
 
-\begin{figure*}
+	\begin{figure*}[]
 		\centering
+		\includegraphics[width=1.8\columnwidth]{T-reX_method.pdf}
 		\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format.
-Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-	\includegraphics[width=1.7\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}\label{fig:methods-flowchart}
 	\end{figure*}
@@
 		\item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
-	\item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long
-	      term emissions method~\citep{hauschild2003edip}
+		\item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
 		\item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
@@
 	An example of the output from the application of T-reX has been included in the
-supplementary material in section 6 of the ~\autoref{sec:supplementary}. The
+	supplementary material~\autoref{sec:supplementary-5}. The
 	manipulated \texttt{ecoinvent} databases (which are the main product of T-reX)
@@
 	with full results included in the supplementary material
-(\autoref{sec:supplementary}). Because the `pseudo-LCIA' methods created by
+	(\autoref{sec:supplementary-7} and \autoref{sec:supplementary-8}). Because the `pseudo-LCIA' methods created by
 	T-reX are integrated into the user's \texttt{brightway} project as if they were
@@
 	\autoref{fig:sankey_waste} presents a heavily abridged Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
-As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material~\autoref{sec:supplementary}.
+As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material in~\autoref{sec:supplementary-8}.
 As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
@@
 
-\begin{figure}[H!]
+	\begin{figure}[!htbp]
 		\centering
-	\includegraphics[width=0.8\columnwidth]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
-	\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser showing the total solid waste footprint flows for the
-	 NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}label{fig:sankey_waste}
+		\includegraphics[width=0.8\columnwidth]{T-reX_NMC811_WasteTotalSolid.pdf}
+		\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the
+			NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}\label{fig:sankey_waste}
 	\end{figure}
@@
 
-\begin{figure*}[H]
+	\begin{figure*}[!htbp]
 		\centering
-	\includegraphics[width=/textwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+		\includegraphics[width=1.8\columnwidth]{T-reX-wastefootprint-totalANDcarbon.pdf}
 		\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
@@
 
-\subsubsection{Contribution of `top-processes' in the supply chain}%\label{sec:results-case_study-topprocesses}
+	\subsubsection{Contribution of `top-processes' in the supply chain}\label{sec:results-case_study-topprocesses}
 
@@
 
-\begin{figure}
+	\begin{figure}[!htbp]
 		\centering
-	\includegraphics[width=0.9\columnwidth]{figures/T-reX-wastefootprint-processcontributions.pdf}
+		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-processcontributions.pdf}
 		\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model}\label{fig:top_contribution}
 	\end{figure}
-
-%!!!!
 
@@
 		\centering
-	\includegraphics[width=0.9\columnwidth]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
+		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-sectorcontributions.pdf}
 		\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model}\label{fig:cpc_contribution}
@@
 
-\begin{figure*}
+	\begin{figure*}[!htbp]
 		\centering
-	\includegraphics[width=1.9\columnwidth]{figures/T-reX-coalANDedip.pdf}
+		\includegraphics[width=1.8\columnwidth]{T-reX-coalANDedip.pdf}
 		\caption{Comparison of the (a) `T-reX – Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 – coal no longterm'.
@@
 	% Back matter
-%\appendix
-\section{Supplementary material}\label{sec:supplementary}
+	\appendix
 	The supplementary material supplied with this manuscript contains the
@@
 	\begin{enumerate}
-	\item Metadata of the T-reX Python package
-	\item Details of the modules and user configuration
-	\item Description of the computational workflow
-	\item Modular flowchart of T-reX
-	\item Example of the terminal output of T-reX
-	\item Example of the terminal output of the case study
-	\item Complete tabulated results of the case study
-	\item Complete visualisations from the case study
+		\item Metadata of the T-reX Python package\label{sec:supplementary-1}
+		\item Details of the modules and user configuration\label{sec:supplementary-2}
+		\item Description of the computational workflow\label{sec:supplementary-3}
+		\item Modular flowchart of T-reX\label{sec:supplementary-4}
+		\item Example of the terminal output of T-reX\label{sec:supplementary-5}
+		\item Example of the terminal output of the case study\label{sec:supplementary-6}
+		\item Complete tabulated results of the case study\label{sec:supplementary-7}
+		\item Complete visualisations from the case study\label{sec:supplementary-8}
 	\end{enumerate}
@@
 
-% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
-% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+	% To print the credit authorship contribution details
+	\printcredits
 
-\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
+	%\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
 
-\begin{itemize}
-	\item \textbf{SCM}: Stewart Charles McDowall
-	\item \textbf{EL}: Elizabeth Lanphear
-	\item \textbf{SC}: Stefano Cucurachi
-	\item \textbf{CFB}: Carlos Felipe Blanco
-\end{itemize}
+	%\begin{itemize}
+	%	\item \textbf{SCM}: Stewart Charles McDowall
+	%	\item \textbf{EL}: Elizabeth Lanphear
+	%	\item \textbf{SC}: Stefano Cucurachi
+	%	\item \textbf{CFB}: Carlos Felipe Blanco
+	%\end{itemize}
 
-\includegraphics[width=0.8\columnwidth]{figures/T-reX_credit_heatmap.pdf}
+	\begin{figure}[!htbp]
+		\centering
+		\includegraphics[width=0.8\columnwidth]{T-reX_credit_heatmap.pdf}
+		\caption{CRediT authorship visualisation}\label{fig:credit_heatmap}
+	\end{figure}
 
@@
 	% Loading bibliography database
-\bibliography{01_T-reX_manuscript.bib}
-
-% insert supplementary information
-%\clearpage
-%#\includepdf{T-reX_supplementary-material/00_T-reX_supplementary-material_main.pdf}
+	\bibliography{01_T-reX_manuscript}
 

### 2024-12-02 00_T-reX_journal-upload-folder/T-reX_latex-files/T-reX_supplementary-material/T-reX_case-study_data/_05c_visualisation_bar_top-processes.py

@@ -1,0 +1,281 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+import matplotlib as mpl
+from matplotlib.lines import Line2D
+import matplotlib.ticker as mticker
+import scienceplots
+from tqdm import tqdm
+import seaborn as sns
+import random
+import numpy as np
+
+from utils import replace_string_in_df, term_replacements, replace_strings_in_string
+from _00_main import DIR_VISUALISATION, FILE_RESULTS_PROCESSED_TOP_PROCESSES
+
+print("Generating figures for top processes...")
+
+
+# Create a custom formatter
+class CustomFormatter(mticker.ScalarFormatter):
+    def __init__(self, fmt="%1.1f"):
+        super().__init__(useOffset=False, useMathText=True)
+        self.fmt = fmt
+
+    def _set_format(self):
+        self.format = self.fmt
+
+
+# Set a fixed figure size to match your largest plot (274x343 pixels)
+# fixed_figsize = (250 / 100, 500 / 100)  # Convert pixels to inches
+
+# plt.rcParams["figure.figsize"] = fixed_figsize
+plt.rcParams["figure.autolayout"] = False
+
+# set global formatting options
+rc_fonts = {
+    "font.family": "serif",
+    "text.usetex": True,
+    "text.latex.preamble": "\n".join(
+        [
+            "\\usepackage{libertine}",
+            "\\usepackage[libertine]{newtxmath}",
+        ]
+    ),
+}
+mpl.style.use(["science"])
+mpl.rcParams.update(rc_fonts)
+
+
+# Read in the data
+df_raw = pd.read_csv(FILE_RESULTS_PROCESSED_TOP_PROCESSES, sep=";", index_col=None)
+
+for term, replacement in term_replacements.items():
+    df_raw = replace_string_in_df(df_raw, term_replacements)
+
+# Aggregate the methods into a single column
+method_full = list(zip(df_raw["method_0"], df_raw["method_1"], df_raw["method_2"]))
+df_raw.insert(0, "method_full", method_full)
+
+
+activities = df_raw["name"].unique()
+methods_full = df_raw["method_full"].unique()
+scenarios = df_raw["db_target"].unique()
+
+df = df_raw.copy().reset_index(drop=True)
+# filter out unnecessary columns
+
+cols_tp = [x for x in df.columns if "tp_" in x]
+cols_tp_score = [x for x in cols_tp if "score" in x]
[…]

### 2024-12-03 00_T-reX_journal-upload-folder/T-reX_latex-files/T-reX_manuscript-review_changes.tex
add revision comparisons

--- 2024-12-02 20240201_T-reX_JournalUploadFolder/main.tex
@@
 \documentclass[a4paper,fleqn]{cas-dc}
-
-% If the frontmatter runs over more than one page
-% use the longmktitle option.
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
 
@@
 
-%\usepackage[numbers]{natbib}
 \usepackage[authoryear]{natbib}
-% \usepackage[authoryear,longnamesfirst]{natbib}
-%\usepackage{booktabs}
-%\usepackage[tt=false]{libertine}
-%\usepackage{xurl}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage[british]{babel}
+\usepackage{microtype}
+\usepackage{xcolor}
+\usepackage{xurl}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{lineno}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{geometry}
+\usepackage{tabularx}
+\usepackage{fontspec}
+\setmainfont[Mapping=tex-text,Ligatures={Common,Rare,Discretionary}]{Linux Libertine}
 
-%%%
 \begin{document}
@@
 
-	% Short title
 \shorttitle{T-reX: Quantifying waste and material footprints in LCA}
 
-	% Short author
 \shortauthors{McDowall, S2. et~al.}
 
-	% Main title of the paper
 \title [mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
-
-	% First author
-	%
-	% Options: Use if required
-	% eg: \author[1,3]{Author Name}[type=editor,
-	%       style=chinese,
-	%       auid=000,
-	%       bioid=1,
-	%       prefix=Sir,
-	%       orcid=0000-0000-0000-0000,
-	%       facebook=<facebook id>,
-	%       twitter=<twitter id>,
-	%       linkedin=<linkedin id>,
-	%       gplus=<gplus id>]
 
@@
 
-	% Corresponding author indication
 \cormark[1]
-	% Email id of the first author
 \ead{s.c.mcdowall@cml.leidenuniv.nl}
-	% URL of the first author
 \ead[url]{scmcd.ch}
-	% Credit authorship
 \credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing: original draft, Writing: review \& editing, Visualization}
@@
 
-	% Corresponding author text
 \cortext[1]{Corresponding author}
 
-	% Here goes the abstract
 \begin{abstract}
-		The quintessential principle of the circular economy is to minimise material consumption and waste generation. Thus, identifying and quantifying waste and material flows is critical.
+    The quintessential principle of the \protect\TLSdel{`circular economy'} \protect\TLSins{circular economy} is to minimise material consumption and waste generation.
+Thus, identifying and quantifying waste and material flows is critical.
 
-		Life Cycle Assessment (LCA) is a powerful quantitative method for this, especially given its capacity to expose the details of an activity's entire life cycle, often guiding the implementation of circular principles to where they could be most effective.
+    Life Cycle Assessment (LCA) is a powerful \protect\TLSins{quantitative} method for this, \protect\TLSins{especially} given its capacity to \protect\TLSdel{pinpoint hotspots} \protect\TLSins{expose the details} of \protect\TLSdel{environmental impact throughout} an activity's \protect\TLSins{entire} life cycle, \protect\TLSdel{where} \protect\TLSins{often guiding} the implementation of circular principles \protect\TLSins{to where they} could be most effective.
 
-		Pursuant to these goals the authors have developed `T-reX', an analytical tool for LCA, written in Python.
-T-reX extends on the existing computational ecosystem centered on \texttt{Brightway} and \texttt{Premise} and can facilitate the quantification of user-defined supply chain demands for activities in current and prospective scenarios.
-T-reX streamlines database manipulation for LCA practitioners, integrating methods to aggregate and analyse demand inventories that could expose hotspots of hidden risk.
+\protect\TLSdel{We introduce T-reX, a Python}
 
-		With a simple case study using lithium-ion batteries, we demonstrate some of T-reX's potential by detailing and exploring aspects of their waste and material footprints.
-Since such footprints are often linked to negative externalities, T-reX could support sustainable decision-making and contribute to the development of the circular economy by promoting the understanding of our material consumption and waste generation.
+    \protect\TLSins{Pursuant to these goals the authors have developed `T-reX', an analytical} tool \protect\TLSdel{extending} \protect\TLSins{for LCA, written in Python.
+T-reX extends on the existing computational ecosystem centered on \texttt{Brightway} and \texttt{Premise} and can facilitate} the \protect\TLSdel{brightway ecosystem, to allow facile} quantification of user-defined supply chain demands \protect\TLSins{for activities} in current and prospective scenarios.
+T-reX streamlines database manipulation for LCA \protect\TLSdel{practitioners and integrates} \protect\TLSins{practitioners, integrating} methods to aggregate and analyse \protect\TLSdel{waste and material flows, facilitating rapid hotspot identification.
+
+A} \protect\TLSins{demand inventories that could expose hotspots of hidden risk.
+
+    With a simple} case study \protect\TLSdel{of five} \protect\TLSins{using} lithium-ion \protect\TLSdel{batteries demonstrates} \protect\TLSins{batteries, we demonstrate some of} T-reX's \protect\TLSdel{utility, quantifying categorised} \protect\TLSins{potential by detailing and exploring aspects of their} waste and material \protect\TLSdel{inventory} \protect\TLSins{footprints.
+Since such} footprints \protect\TLSdel{and their potential environmental burdens.} \protect\TLSins{are often linked to negative externalities,} T-reX \protect\TLSdel{can aid} \protect\TLSins{could support} sustainable decision-making and contribute to the development of \protect\TLSdel{`circular economy'} \protect\TLSins{the circular economy} by \protect\TLSdel{facilitating analysis} \protect\TLSins{promoting the understanding} of \protect\TLSins{our} material consumption and waste \protect\TLSdel{generation in LCA.} \protect\TLSins{generation.}
 \end{abstract}
 
-	% Use if graphical abstract is present
 \begin{graphicalabstract}
-		\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract.pdf}
-		%\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract_MID-RES.png}
+	\includegraphics[width=1\columnwidth]{figs/grabs_T-reX_graphical-abstract.pdf}
 	\end{graphicalabstract}
 
-	% Highlights
 \begin{highlights}
@@
 
-	% Keywords
-	% Each keyword is seperated by \sep
-
 \begin{keywords}
-		circular economy, waste footprint, material footprint, life cycle assessment, critical raw materials, supply chain
+circular economy\sep waste footprint\sep material footprint\sep life cycle assessment\sep critical raw materials \sep supply chain
 \end{keywords}
@@
 
-	% Main text
 \begin{table*}[!htbp]
@@
 \section{Introduction}\label{sec:introduction}  
-	\subsection{Introduction to T-reX}\label{sec:intro-trex}
+\subsection[[Introduction to T-reX]]{Introduction to T-reX}\label{sec:intro-trex}
 
@@
 
-	\subsection{Background and need}\label{sec:intro-background}
+\subsection[[Background and need]]{Background and need}\label{sec:intro-background}
 
@@
 of material demands in terms of mass or volume.
+
+\subsection*{}
 
@@
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX_method.pdf}
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX_method.pdf}
 	\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format.
-Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}\label{fig:methods-flowchart}
+Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply \protect\TLSdel{chain.
+	\label{fig:methods-flowchart}} chain.}\label{fig:methods-flowchart}
 \end{figure*}
@@
 
-	To provide an example, we investigated five types of Li-ion batteries in the
+\protect\TLSdel{We}
+\protect\TLSins{To provide an example, we} investigated five types of Li-ion batteries in the
 \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3},
-	each represented by their unamended global market inventories. The system
+each represented by their unamended global market \protect\TLSdel{inventories:
+[itemsep=0pt]} \protect\TLSins{inventories. The system
 boundaries. These following markets were used for the example calculations and
@@
 
-	\begin{itemize}
+}\begin{itemize}
 	\item Li-ion, LFP, rechargeable, prismatic
@@
 
+\protect\TLSdel{[itemsep=0pt]}
 \begin{itemize}
 	\item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
-		\item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
-		\item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
-		\item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
+	\item \protect\TLSdel{EF v3.0 no LT~} \protect\TLSins{Environmental Footprint (EF) v3.1 (no long-term emissions)~}\citep{eu2023ef}
+	\item \protect\TLSins{Spatial differentiation in life cycle impact assessment---the} EDIP 2003 no \protect\TLSdel{LT~} \protect\TLSins{long term emissions method~}\citep{hauschild2003edip}
+	\item \protect\TLSdel{CSI} \protect\TLSins{The crustal scarcity indicator method---CSI} 2020~\citep{arvidsson2020csi}
 \end{itemize}
@@
 \subsubsection{LCA calculations}
-	For each combination of activity, method, and database, a total summation was
+For each combination of activity, method, and database, a \protect\TLSdel{single score} \protect\TLSins{total summation} was
 calculated along with details of the top contributing processes. Additionally,
-	for the T-reX methods, a contribution analysis was performed with the
+for the T-reX methods, a contribution analysis was \protect\TLSdel{performed. This involved utilising} \protect\TLSins{performed with} the
 \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the
 \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a
-	component of the \texttt{brightway} ecosystem. This function performs one
-	subjective method of graph traversal on the impact matrix of the LCA object to
-	a specified cutoff. The groups of the resulting `leaves' are collected by their
-	Cooperative Patent Classification (CPC) codes [giving a representation of
+component of the \texttt{brightway} ecosystem. This function performs \protect\TLSins{one
+subjective method of} graph traversal on the impact matrix of the LCA object to
+a specified \protect\TLSdel{cutoff and} \protect\TLSins{cutoff. The} groups \protect\TLSins{of} the resulting \protect\TLSdel{leaves} \protect\TLSins{`leaves' are collected} by their
+Cooperative Patent Classification (CPC) \protect\TLSdel{codes. This provides} \protect\TLSins{codes [giving a representation of
 industry sector involvement]. T-reX provides, essentially, a tabulation of
-	internal exchanges and an insight into the products and sectors in the supply
+internal exchanges and an} insight into the products and sectors in the supply
 chain of the activity that carry the most substantial responsibility for the
@@
 An example of the output from the application of T-reX has been included in the
-	supplementary material~\autoref{sec:supplementary-5}. The
-	manipulated \texttt{ecoinvent} databases (which are the main product of T-reX)
+supplementary \protect\TLSdel{material in~\autoref{sec:supplementary}.} \protect\TLSins{material~\autoref{sec:supplementary-5}.} The
+manipulated \texttt{ecoinvent} \protect\TLSdel{databases, which} \protect\TLSins{databases (which} are the main product of \protect\TLSdel{T-reX,} \protect\TLSins{T-reX)}
 can be recreated by following the instructions in the package
@@
 limitations of T-reX. This section includes some highlights of the results,
-	with full results included in the supplementary material
-	(\autoref{sec:supplementary-7} and \autoref{sec:supplementary-8}). Because the `pseudo-LCIA' methods created by
+with full results included in the supplementary material \protect\TLSdel{(\autoref{sec:supplementary}).}
+\protect\TLSins{(\autoref{sec:supplementary-7} and \autoref{sec:supplementary-8}).} Because the `pseudo-LCIA' methods created by
 T-reX are integrated into the user's \texttt{brightway} project as if they were
@@
 results obtained using the \texttt{ActivityBrowser} software, including
-	contribution analysis and a Sankey diagram that dis-aggregates the final
+contribution analysis and a Sankey diagram that \protect\TLSdel{disaggregates} \protect\TLSins{dis-aggregates} the final
 footprint result over the activities in the supply chain.
 
-	\subsubsection{Sankey visualisation of flow inventory footprints}\label{sec:results-case_study-sankey}
+\subsubsection[Sankey visualisation of  flow inventory footprints]{Sankey visualisation of \protect\TLSdel{flows in waste and material} \protect\TLSins{flow} inventory footprints}\label{sec:results-case_study-sankey}
 
-	\autoref{fig:sankey_waste} presents a heavily abridged Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
-As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material in~\autoref{sec:supplementary-8}.
-As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
+\autoref{fig:sankey_waste} presents a \protect\TLSins{heavily abridged} Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+As the supply chains are expansive, this visualisation has been greatly simplified \protect\TLSins{to be legible} for print, \protect\TLSdel{the full version is} \protect\TLSins{multiple versions are} included in the supplementary \protect\TLSdel{material~\autoref{sec:supplementary}.} \protect\TLSins{material in~\autoref{sec:supplementary-8}.} As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
 Due to the nature of LCA, the waste flows do not represent the `real' physical flows directly, but rather an abstraction thereof based on, in this case, economic allocation---the market price of the functional unit produced by the activity~\citep{guinee2004economicallocation}.
-If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and would be more directly related to the physical flows of the activities.
+If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and \protect\TLSins{would} be more directly related to the physical flows of the activities.
 
+\protect\TLSdel{[htbp]}
 \begin{figure}[!htbp]
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX_NMC811_WasteTotalSolid.pdf}
-		\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the
-			NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1}\label{fig:sankey_waste}
+
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX_NMC811_WasteTotalSolid.pdf}
+	\caption{A \protect\TLSins{greatly} simplified Sankey diagram \protect\TLSins{from \texttt{ActivityBrowser}} showing the total solid waste footprint flows for the
+	 NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version \protect\TLSdel{3.9.1.}\label{fig:sankey_waste}} 3.9.1}\label{fig:sankey_waste}
 \end{figure}
@@
 
-	%\label{sec:results-case_study-total_footprints}
-
-	%! total
 \autoref{fig:waste_totalANDcarbon} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.
@@
 In each case, there was only a slight downward trend in the waste footprints between 2020 and 2100.
-This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in fossil fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
+This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in \protect\TLSdel{fossil-fuel} \protect\TLSins{fossil fuel} use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
 For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
 
-	%! CO2
 The inclusion of carbon capture and storage (CCS) in the prospective databases
@@
 assumptions made about the introduction of CCS technology. The actual
-	deployment of these technologies was only around 37~Mt CO\(_2\)/yr as of
+deployment of these technologies was only around 37~Mt \protect\TLSdel{CO$_2$/yr} \protect\TLSins{CO\(_2\)/yr} as of
 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of
-	the RCP scenarios~\citep{sacchi2023premisedocs}. The authors consider that the
-	over-representation of CCS and the under-representation of waste processing
+the RCP scenarios~\citep{sacchi2023premisedocs}. The authors consider that the \protect\TLSdel{over representation}
+\protect\TLSins{over-representation} of CCS and the \protect\TLSdel{under representation} \protect\TLSins{under-representation} of waste processing
 technologies in the prospective databases is one significant limitation to the
-	validity of prospective LCA databases. Significant work is needed to improve
+validity of prospective LCA databases.
+
+\protect\TLSdel{[htbp]} \protect\TLSins{Significant work is needed to improve
 the representation of these technologies in the databases to ensure that the
 results are more realistic and useful for decision-making.
-	%!!!!
 
-	\begin{figure*}[!htbp]
+}\begin{figure*}[!htbp]
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX-wastefootprint-totalANDcarbon.pdf}
-		\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
+
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX-wastefootprint-totalANDcarbon.pdf}
+	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste \protect\TLSdel{(from} \protect\TLSins{(produced from} carbon capture and storage (CCS) only).
 The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
@@
 
-	\subsubsection{Contribution of `top-processes' in the supply chain}\label{sec:results-case_study-topprocesses}
+\subsubsection[Contribution of `top-processes' in the supply  chain}\label{sec:results-case_study-topprocesses}
+
+\autoref{fig:top_contribution]{Contribution of `top-processes' in the supply \protect\TLSdel{chain} \protect\TLSins{chain}\label{sec:results-case_study-topprocesses}}
 
@@
 
+\protect\TLSdel{[htbp]}
 \begin{figure}[!htbp]
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-processcontributions.pdf}
-		\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model}\label{fig:top_contribution}
+
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX-wastefootprint-processcontributions.pdf}
+	\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND \protect\TLSdel{model.}\label{fig:top_contribution}} model}\label{fig:top_contribution}
 \end{figure}
@@
 
+\protect\TLSdel{[htbp]}
 \begin{figure}
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-sectorcontributions.pdf}
-		\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model}\label{fig:cpc_contribution}
+
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX-wastefootprint-sectorcontributions.pdf}
+	\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND \protect\TLSdel{model.}\label{fig:cpc_contribution}} model}\label{fig:cpc_contribution}
 \end{figure}
@@
 
-	A comparison of the results for `Coal (black)' demand using the T-reX method
-	with the LCIA method `EDIP 2003 – coal no LT' is shown
+A comparison of the results \protect\TLSdel{from T-reX's} \protect\TLSins{for} `Coal (black)' demand \protect\TLSins{using the T-reX} method
+with the LCIA method `EDIP 2003 \protect\TLSdel{-} \protect\TLSins{–} coal no LT' is shown
 in~\autoref{fig:comparison_methods}. In this case, both the trends and the
@@
 decrease in the use of coking coal in the battery's footprints. Such
-	comparability was also observed for other fossil-fuel-related demands (e.g.,
+comparability was also observed for other fossil-fuel-related \protect\TLSdel{methods} \protect\TLSins{demands} (e.g.,
 natural gas and petroleum), and to a lesser extent, for some of the metal
@@
 that are not captured by the standard LCIA methods. This makes the T-reX
-	methods more sensitive to the modeling choices (e.g., economic allocation) that
+methods more sensitive to the modeling choices \protect\TLSdel{(e.g.,economic} \protect\TLSins{(e.g., economic} allocation) that
 are generally embedded in LCA databases.
 
+\protect\TLSdel{[htbp]}
 \begin{figure*}[!htbp]
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX-coalANDedip.pdf}
-		\caption{Comparison of the (a) `T-reX – Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 – coal no longterm'.
-In the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
+
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX-coalANDedip.pdf}
+	\caption{Comparison of the (a) `T-reX \protect\TLSdel{-} \protect\TLSins{–} Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 \protect\TLSdel{-} \protect\TLSins{–} coal no \protect\TLSdel{LT'  in} \protect\TLSins{longterm'.
+In} the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
 \end{figure*}
@@
 
-	In the most similar published method,~\cite{laurenti2023wastefootprint} used an
+In the most similar published \protect\TLSdel{method, \cite{laurenti2023wastefootprint}} \protect\TLSins{method,~\cite{laurenti2023wastefootprint}} used an
 alternative procedure for calculating the waste footprint of \~1400 activities
@@
 
+\protect\TLSdel{[htbp]}
 \begin{table}
 	\centering
-		\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion')}\label{tab:results-case_study-comparison}
+	\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, \protect\TLSdel{activity:`Li-ion').}\label{tab:results-case_study-comparison}} activity:`Li-ion')}\label{tab:results-case_study-comparison}
 	\begin{tabular}{lll}
@@
 
-	% Back matter
 \appendix
@@
 
-	% To print the credit authorship contribution details
 \printcredits
-
-	%\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
-
-	%\begin{itemize}
-	%	\item \textbf{SCM}: Stewart Charles McDowall
-	%	\item \textbf{EL}: Elizabeth Lanphear
-	%	\item \textbf{SC}: Stefano Cucurachi
-	%	\item \textbf{CFB}: Carlos Felipe Blanco
-	%\end{itemize}
 
@@
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX_credit_heatmap.pdf}
-		\caption{CRediT authorship visualisation}\label{fig:credit_heatmap}
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX_credit_heatmap.pdf}
+	\caption{\protect\TLSins{CRediT} authorship \protect\TLSdel{contribution statement
+
+[noitemsep]
+	 \textbf{SCM}: Stewart Charles McDowall
+	\item \textbf{EL}: Elizabeth Lanphear
+	\item \textbf{SC}: Stefano Cucurachi
+	\item \textbf{CFB}: Carlos Felipe Blanco
+\end{itemize}
+
+\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{figs/T-reX_credit_heatmap.pdf}
+
+\section*{Declaration of competing interest}
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
+\section*{Acknowledgements}
+Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
+101058522 (project FutuRaM --- futuram.eu).
+The authors would like to thank the reviewers for their valuable comments and suggestions.
+
+The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems.
+The authors hope to continue the development of T-reX in the service of our shared ideals.
+
+}\section*{} visualisation}\label{fig:credit_heatmap}
 \end{figure}
 
-	%% Loading bibliography style file
-	%\bibliographystyle{model1-num-names}
 \bibliographystyle{cas-model2-names}
 
-	% Loading bibliography database
-	\bibliography{01_T-reX_manuscript}
+\bibliography{cas-refs}
 

### 2024-12-03 00_T-reX_journal-upload-folder/T-reX_latex-files/cas-dc-template_copyfordiff.tex
add revision comparisons

--- 2024-11-27 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
@@
-%% 
-%% Copyright 2019-2024 Elsevier Ltd
-%% 
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3c of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3c or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for cas-dc documentclass for 
-%% double column output.
-
 \documentclass[a4paper,fleqn]{cas-dc}
@@
 % use the longmktitle option.
-
 %\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
 
-%\usepackage[numbers]{natbib}
-%\usepackage[authoryear]{natbib}
-\usepackage[authoryear,longnamesfirst]{natbib}
-
-%%%Author macros
 \def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
@@
 \tsc{CFB}
+
+%\usepackage[numbers]{natbib}
+\usepackage[authoryear]{natbib}
+% \usepackage[authoryear,longnamesfirst]{natbib}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage[british]{babel}
+\usepackage{microtype}
+% \usepackage[tt=false]{libertine}
+\usepackage{xcolor}
+\usepackage{xurl}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{lineno}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{geometry}
+\usepackage{tabularx}
+\usepackage{fontspec}
+\setmainfont[Mapping=tex-text,Ligatures={Common,Rare,Discretionary}]{Linux Libertine}
+
 %%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
-
 \begin{document}
@@
 % Short title
-\shorttitle{T-reX\@: Quantifying waste and material footprints in LCA}
+\shorttitle{T-reX: Quantifying waste and material footprints in LCA}
 
 % Short author
-\shortauthors{S.C. McDowall et~al.}
+\shortauthors{McDowall, S2. et~al.}
 
 % Main title of the paper
-\title [mode = title]{T-reX\@: Quantifying waste and material footprints in current and future \\ Life Cycle Assessment (LCA) databases}
+\title [mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
 
-% Authors
-\author[1]{Stewart Charles McDowall}
+% First author
+%
+% Options: Use if required
+% eg: \author[1,3]{Author Name}[type=editor,
+%       style=chinese,
+%       auid=000,
+%       bioid=1,
+%       prefix=Sir,
+%       orcid=0000-0000-0000-0000,
+%       facebook=<facebook id>,
+%       twitter=<twitter id>,
+%       linkedin=<linkedin id>,
+%       gplus=<gplus id>]
+
+\author[1]{Stewart Charles McDowall}[orcid=0000-0002-5688-1279]
+
+% Corresponding author indication
+\cormark[1]
+% Email id of the first author
 \ead{s.c.mcdowall@cml.leidenuniv.nl}
-\cormark[1]
-\cortext[cor1]{Corresponding author:}
-
-\author[1]{Elizabeth Lanphear}
-
-\author[1]{Stefano Cucurachi}
-
-\author[1]{Carlos Felipe Blanco}
+% URL of the first author
+\ead[url]{scmcd.ch}
+% Credit authorship
+\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing: original draft, Writing: review \& editing, Visualization}
 
@@
   city={Leiden},
-    postcode={2300 RA},
+  postcode={2300RA},
   state={South Holland},
@@
 
-% Abstract
+\author[1]{Elizabeth Lanphear}[orcid=0009-0003-6364-0873]
+\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing}
+
+\author[1]{Stefano Cucurachi}[orcid=0000-0003-2748-655X]
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
+
+\author[1]{Carlos Felipe Blanco}[orcid=0000-0001-8199-8420]
+\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
+
+% Corresponding author text
+\cortext[1]{Corresponding author}
+
+% Here goes the abstract
 \begin{abstract}
+        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
+Thus, identifying and quantifying waste and material flows is critical.
+Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
 
-%? reviewer 2 - comment 2 - simplify language
-The quintessential principle of the circular economy is to minimise material consumption and waste generation.
-Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
-
-We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
+We introduce T-reX, a Python tool extending the brightway ecosystem, to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
 T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
@@
 A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
-
+T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
 \end{abstract}
 
-% Keywords
-\verb+\begin{keyword}+ \verb+circular economy+ \verb+waste+ \verb+material+ \verb+life cycle assessment+ \verb+critical raw material+ \verb+supply chain+ \end{keyword}
-
-% Graphical abstract
+% Use if graphical abstract is present
 \begin{graphicalabstract}
-  \includegraphics{figs/cas-grabs-TREX.pdf}
+	\includegraphics[width=1\columnwidth]{figs/grabs_T-reX_graphical-abstract.pdf}
+	%\includegraphics[width=\columnwidth]{figs/grabs_T-reX_graphical-abstract_MID-RES.png}
 \end{graphicalabstract}
@@
 \begin{highlights}
-    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
+	\item T-reX, a new open-source program for waste \& material footprints in Life Cycle Assessment (LCA)
 	\item Assesses supply risks by calculating demand inventories for critical materials
@@
 	\item Rapidly identifies hotspots of waste generation and material demand in supply	chains
-    \item Presents a case study of five lithium-ion batteries to demonstrate T-reX's utility
+	\item Presents a case study of five lithium-ion batteries to demonstrate T-reX's potential
 \end{highlights}
@@
 % Keywords
-\begin{keyword}
-  circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
-\end{keyword}
+% Each keyword is seperated by \sep
 
-%\nocite{*}
+\begin{keywords}
+circular economy\sep waste footprint\sep material footprint\sep life cycle assessment\sep critical raw materials \sep supply chain
+\end{keywords}
 
@@
 
-% list of abbreviations
-\begin{table}[H]
+% Main text
+\begin{table*}[!htbp]
 	\centering
-  \caption*{List of abbreviations and terminology}\label{tab:abbreviations}
+	\caption{List of abbreviations}\label{tab:abbreviations}
 	\begin{tabular}{ll}
 		\toprule
-      %? reviewer 3 - RE: CCS definition
 		CCS	              & Carbon Capture and Storage                                          \\
@@
 	\end{tabular}
-\end{table}
+\end{table*}
 
-% Main text
-
-% Introduction
 \section{Introduction}\label{sec:introduction}  
-\subsection{Introduction to T-reX}\label{sec:intro-trex}
-
-%? reviewer 2 - comment 2 - simplify language
-Development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
+The development of a `circular economy' has become a central focus in the frantic pursuit of sustainability objectives and the imperative curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
 Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
+In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
 
@@
 More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
-Whether at the level of products, entire industries, nations, or even continents, the material footprint aims to quantify the total material consumed in supply chains.
+Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
 It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
 
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed.
+Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the incipient `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
 While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
+Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
+Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
 
@@
 
-%? reviewer 2 - comment 1 - COMPARISON OF T-REX TO OTHER TOOLS
-%! fix
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
+LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
-These scores can then be aggregated into one metric for each impact category that can be compared across products and processes.
+These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
 Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
@@
 These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
-
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
-
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
-Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
-\texttt{premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
-The authors believe that there is an urgent need for research investment in this area.
-
-%! reviewer 3 - comment 1 - ADDITIONAL INFORMATION ON T-REX's differentiating features comparet to LCI
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
-In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and potentially highlight opportunities for waste reduction and material efficiency.
-
-%? reviewer 2 - comment 3 - ADDITIONAL INFORMATION ON T-REX's FUTURE DEVELOPMENT
-With the continual improvement of LCA databases and the development of prospective databases, T-reX can produce ever more realistic estimates, enhancing its actionable utility derived from the exploration of the potential impacts of future waste management systems and the material demands of emerging technologies.
-The better the data, the more specific and accurate the results of T-reX analysis, and, thus, the more useful the tool becomes.
-The authors hope that T-reX will be a valuable asset to the LCA community and contribute to the development of a more sustainable and circular economy.
-
-\subsection{Background and need}\label{sec:intro-background}
 
@@
 
-Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA~\citep{lauran2020abioticdepletion}.
+Existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste.
 This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
@@
 
-~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' .
+\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
+The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, and not easily reproducible.
+As the authors acknowledge, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
 In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
@@
 
-In the context of geopolitical tension and a mineral-hungry `renewable energy transition', greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system rather than the primary material flows themselves.
+In the context of geo-political tension and a mineral-hungry `renewable energy transition', ever greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
+While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
 
@@
 
-% Methodology
+\subsection{Introduction to T-reX}\label{sec:intro-trex}
+
+To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
+
+Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
+Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
+Indeed, despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
+
+The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
+It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
+By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
+In any case, minimising primary material consumption and the generation of waste is of paramount importance.
+By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.
 
@@
 This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries, highlighting inconsistencies in LCA methodology and probing departures from projections in prospective models.
+In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
 
@@
 
-Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising especially  the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
+Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
 Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}.
@@
 
-In the default configuration of T-reX, there are 10 waste categories which are then further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include aggregations such as 'total waste', 'hazardous waste' as well as, those for waste sent to treatment by, for example, incineration or recycling.
-The full details are listed in the supplementary material in~\autoref{sec:supplementary} [section 3 ].
+In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary}.
 One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
+Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (except CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
 
-In addition to these waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
+In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
 Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
-The identity of the materials considered, and their categorical groupings, are easily customisable by the user.
+The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.
 The full list of 59 materials included in the default configuration is provided in the supplementary material in~\autoref{sec:supplementary}.
 
-Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
+The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
 A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
@@
 
-T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere, a material basis.
-This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct effects on the market, such as a 'negative consumption' of other materials due their co-production with the functional unit.
+T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere.
+This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
 Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
 If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-For example, in the results of the Li-ion battery case study , this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
+In the results of the Li-ion battery case study in \autoref{sec:results-casestudy}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
+
+% \subsubsection{Mathematical basis}
+
+% \cbox{very nice formulae coming soon from LL hopefully}
 
@@
 
-\begin{figure}[H]
+\begin{figure*}[!htbp]
 	\centering
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=9cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
-\end{figure}
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX_method.pdf}\label{fig:methods-flowchart}
+\end{figure*}
 
@@
 
-To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories.
-The system boundaies.
-These following markets were used for the example calculations and unaltered by the authors:
+We investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
 \begin{itemize}[itemsep=0pt]
@@
 	\item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
-    \item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
-    \item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
-    \item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
+	\item EF v3.0 no LT~\citep{eu2023ef}
+	\item EDIP 2003 no LT~\citep{hauschild2003edip}
+	\item CSI 2020~\citep{arvidsson2020csi}
 \end{itemize}
@@
 \subsubsection{LCA calculations}
-For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes.
-Additionally, for the T-reX methods, a contribution analysis was performed with the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
-This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
-The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
-T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
+For each combination of activity, method, and database, a single score was calculated along with details of the top contributing processes.
+Additionally, for the T-reX methods, a contribution analysis was performed.
+This involved utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
+This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
+This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
 
+\section{Results}\label{sec:results}
 \subsection{T-reX tool}\label{sec:results-T-reX}
 
-An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}, subsection s.
-The manipulated \texttt{ecoinvent} databases (which are the main product of T-reX) can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
+An example of the output from the application of T-reX has been included in the supplementary material in~\autoref{sec:supplementary}.
+The manipulated \texttt{ecoinvent} databases, which are the main product of T-reX, can be recreated by following the instructions in the package documentation~\citep{mcdowall2023T-reXdocs}.
 
@@
 
-\subsubsection{Sankey visualisation of flow inventory footprints}\label{sec:results-case_study-sankey}
+\subsubsection{Sankey visualisation of flows in waste and material inventory footprints}\label{sec:results-case_study-sankey}
 
-\autoref{fig:sankey_waste} presents a heavily simplified Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
-As the supply chains are expansive, this visualisation has been greatly simplified to be legible for print, multiple versions are included in the supplementary material~\autoref{sec:supplementary}.
+\autoref{fig:sankey_waste} presents a Sankey diagram depicting the supply chain flows of `total solid waste' attributed to the NMC 811 Li-ion battery at the global market in the \texttt{ecoinvent} database version 3.9.1.
+As the supply chains are expansive, this visualisation has been greatly simplified for print, the full version is included in the supplementary material~\autoref{sec:supplementary}.
 As is evident in the figure, the largest contributors to the waste footprint are the extraction and processing of the raw materials, with the largest flows coming from the extraction of cobalt, lithium, and nickel.
 Due to the nature of LCA, the waste flows do not represent the `real' physical flows directly, but rather an abstraction thereof based on, in this case, economic allocation---the market price of the functional unit produced by the activity~\citep{guinee2004economicallocation}.
-If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and would be more directly related to the physical flows of the activities.
+If an unallocated database were available, the waste flows could be calculated with T-reX in the same way and be more directly related to the physical flows of the activities.
 
-\begin{figure}[H]
+\begin{figure}[htbp]
 	\centering
-    \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
-    \caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
+	\includegraphics[width=8cm]{figs/T-reX_NMC811_WasteTotalSolid.pdf}
+	\caption{A simplified Sankey diagram showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
 \end{figure}
@@
 In each case, there was only a slight downward trend in the waste footprints between 2020 and 2100.
-This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in fossil fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
+This is mostly in the period between 2020 and 2040 and could be attributable to the relatively rapid decrease in fossil-fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion).
 For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.
@@
 The actual deployment of these technologies was only around 37~Mt CO$_2$/yr as of 2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of the RCP scenarios~\citep{sacchi2023premisedocs}.
-The authors consider that the over-representation of CCS and the under-representation of waste processing technologies in the prospective databases is one significant limitation to the validity of prospective LCA databases.
-Significant work is needed to improve the representation of these technologies in the databases to ensure that the results are more realistic and useful for decision-making.
+The authors consider that the over representation of CCS and the under representation of waste processing technologies in the prospective databases is one significant limitation to the validity of prospective LCA databases.
 %!!!!
 
-\begin{figure}[H]
+\begin{figure*}[htbp]
 	\centering
-    \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
+	\includegraphics[width=16cm]{figs/T-reX-wastefootprint-totalANDcarbon.pdf}
 	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
 The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
-\end{figure}
+\end{figure*}
 %!!!!
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
 	\centering
-    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
+	\includegraphics[width=9cm]{figs/T-reX-wastefootprint-processcontributions.pdf}
 	\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
 	\centering
-    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
+	\includegraphics[width=9cm]{figs/T-reX-wastefootprint-sectorcontributions.pdf}
 	\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
@@
 
-A comparison of the results for `Coal (black)' demand using the T-reX method with the LCIA method `EDIP 2003 - coal no LT' is shown in~\autoref{fig:comparison_methods}.
+A comparison of the results from T-reX's `Coal (black)' demand method with the LCIA method `EDIP 2003 - coal no LT' is shown in~\autoref{fig:comparison_methods}.
 In this case, both the trends and the magnitude of the scores were very similar, both demonstrating a general decrease in the use of coking coal in the battery's footprints.
-Such comparability was also observed for other fossil-fuel-related demands (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
+Such comparability was also observed for other fossil-fuel-related methods (e.g., natural gas and petroleum), and to a lesser extent, for some of the metal demand methods (e.g., zinc and cobalt).
 Correlation between standard LCIA methods and T-reX methods is not generally to be expected, however, due to fundamental differences in the way that the methods are constructed.
@@
 The T-reX method, therefore, considers cases of co-production, recycling, and substitution, providing a picture of the supply chain pressures that are not captured by the standard LCIA methods.
-This makes the T-reX methods more sensitive to the modeling choices (e.g., economic allocation) that are generally embedded in LCA databases.
+This makes the T-reX methods more sensitive to the modeling choices (e.g.,economic allocation) that are generally embedded in LCA databases.
 
-\begin{figure}[H]
+\begin{figure*}[htbp]
 	\centering
-    \includegraphics[width=16cm]{figures/T-reX-coalANDedip.pdf}
+	\includegraphics[width=16cm]{figs/T-reX-coalANDedip.pdf}
 	\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}[H]
+\begin{table}[htbp]
 	\centering
@@
 
+\section{Discussion}\label{sec:discussion}
 Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting.
@@
 This information is required to thoroughly assess the potential environmental impacts of this waste.
-Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine the nature of the waste flows.
-For example, if the waste generated could represent an actual loss of resources or  potential environmental risk, or, as is often the case, simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
+Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
 A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
@@
 In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
-A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the circular economy.
+A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.
 
 \section{Conclusions}\label{sec:conclusions}
-
 T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It deconstructs, manipulates and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
+It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
 These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
@@
 This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of sustainability and the circular economy.
+T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
 
 % Back matter
+\appendix
+The supplementary material supplied with this manuscript contains the
+following sections:
+\begin{enumerate}
+	\item Metadata of the T-reX Python package\label{sec:supplementary-1}
+	\item Details of the modules and user configuration\label{sec:supplementary-2}
+	\item Description of the computational workflow\label{sec:supplementary-3}
+	\item Modular flowchart of T-reX\label{sec:supplementary-4}
+	\item Example of the terminal output of T-reX\label{sec:supplementary-5}
+	\item Example of the terminal output of the case study\label{sec:supplementary-6}
+	\item Complete tabulated results of the case study\label{sec:supplementary-7}
+	\item Complete visualisations from the case study\label{sec:supplementary-8}
+\end{enumerate}
+
 \section*{Data availability}
-All data used is publically available online under the noted sources.
-T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
+All data used is publicly available online under the noted sources.
+T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence.
 It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
+The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} 
+and the development server is hosted by GitHub at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
 
-% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
-% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
+\section*{Acknowledgements}
+Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
+101058522 (project FutuRaM --- futuram.eu).
+The authors would like to thank the reviewers for their valuable comments and suggestions.
 
-\section*{CRediT authorship contribution statement} % VISUAL VERSION
+The authors would like to acknowledge those who were part of the futures
+thinking small group session at the FutuRaM consortium meeting in Berlin in
+June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision
+presented by Stewart Charles McDowall of a comprehensive open-source software
+framework for the study and development of secondary raw material systems. The
+authors hope to continue the development of T-reX in the service of our shared
+ideals.
+
+\section*{Declaration of competing interest}
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
+% To print the credit authorship contribution details
+\printcredits
+
+%\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
+
+%\begin{itemize}
+%	\item \textbf{SCM}: Stewart Charles McDowall
+%	\item \textbf{EL}: Elizabeth Lanphear
+%	\item \textbf{SC}: Stefano Cucurachi
+%	\item \textbf{CFB}: Carlos Felipe Blanco
+%\end{itemize}
+% \subsection*{Alternative CRediT statement}
+\section*{CRediT authorship contribution statement}
 
@@
 
-\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit-TREX.pdf}
+\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{figs/T-reX_credit_heatmap.pdf}
 
@@
 
-  %! replace the sections
 \section{Supplementary material}\label{sec:supplementary}
@@
 % Loading bibliography database
-\bibliography{cas-refs-TREX}
+\bibliography{cas-refs}
+% \appendix
+% \section{Example output}
+% \label{app:example_output}
 

### 2024-12-03 03_T-reX-latex_files-atfirstsubmission/latex_files-atfirstsubmission/methodology.tex
add revision comparisons

--- 2024-11-26 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template-TREX.tex
@@
-%% 
-%% Copyright 2019-2024 Elsevier Ltd
-%% 
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3c of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3c or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for cas-dc documentclass for 
-%% double column output.
-
-\documentclass[a4paper,fleqn]{cas-dc}
-
-% If the frontmatter runs over more than one page
-% use the longmktitle option.
-
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
-
-%\usepackage[numbers]{natbib}
-%\usepackage[authoryear]{natbib}
-\usepackage[authoryear,longnamesfirst]{natbib}
-
-%%%Author macros
-\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
-\tsc{SCM}
-\tsc{EL}
-\tsc{SC}
-\tsc{CFB}
-%%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
-
-\begin{document}
-\let\WriteBookmarks\relax
-\def\floatpagepagefraction{1}
-\def\textpagefraction{.001}
-
-% Short title
-\shorttitle{T-reX\@: Quantifying waste and material footprints in LCA}
-
-% Short author
-\shortauthors{S.C. McDowall et~al.}
-
-% Main title of the paper
-\title [mode = title]{T-reX\@: Quantifying waste and material footprints in current and future \\ Life Cycle Assessment (LCA) databases}
-
-% Authors
-\author[1]{Stewart Charles McDowall}
-\ead{s.c.mcdowall@cml.leidenuniv.nl}
-\cormark[1]
-\cortext[cor1]{Corresponding author:}
-
-\author[1]{Elizabeth Lanphear}
-
-\author[1]{Stefano Cucurachi}
-
-\author[1]{Carlos Felipe Blanco}
-
-\affiliation[1]{
-    organization={Institute of Environmental Sciences (CML), Leiden University},
-    addressline={P.O. Box 9518},
-    city={Leiden},
-    postcode={2300 RA},
-    state={South Holland},
-    country={The Netherlands}}
-
-% Abstract
-\begin{abstract}
-
-%? reviewer 2 - comment 2 - simplify language
-The quintessential principle of the circular economy is to minimise material consumption and waste generation.
-Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
-
-We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
-T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
-
-A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
-
-\end{abstract}
-
-% Keywords
-\verb+\begin{keyword}+ \verb+circular economy+ \verb+waste+ \verb+material+ \verb+life cycle assessment+ \verb+critical raw material+ \verb+supply chain+ \end{keyword}
-
-% Graphical abstract
-\begin{graphicalabstract}
-  \includegraphics{figs/cas-grabs-TREX.pdf}
-\end{graphicalabstract}
-
-% Highlights
-\begin{highlights}
-    \item T-reX, a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
-    \item Assesses supply risks by calculating demand inventories for critical materials
-    \item Simplifies quantification of user-specified waste and material footprint categories
-    \item Rapidly identifies hotspots of waste generation and material demand in supply chains
-    \item Presents a case study of five lithium-ion batteries to demonstrate T-reX's utility
-\end{highlights}
-
-% Keywords
-\begin{keyword}
-  circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
-\end{keyword}
-
-%\nocite{*}
-
-\maketitle
-
-% list of abbreviations
-\begin{table}[H]
-  \centering
-  \caption*{List of abbreviations and terminology}\label{tab:abbreviations}
-  \begin{tabular}{ll}
-      \toprule
-      %? reviewer 3 - RE: CCS definition
-      CCS              & Carbon Capture and Storage                                           \\
-      CF                & Carbon Footprint                                                    \\
-      EF                & Ecological Footprint                                                \\
-      CML               & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
-      CRM               & Critical Raw Material                                               \\
-      CPC               & Cooperative Patent Classification                                   \\
-      CSI               & Crustal Scarcity Indicator                                          \\
-      CSP               & Crustal Scarcity Potential                                          \\
-      EDIP              & Environmental Design of Industrial Products                         \\
-      EEIOA             & Environmentally Extended Input-Output Analysis                      \\
-      EOL               & End-of-Life                                                         \\
-      FOEN              & Swiss Federal Office for the Environment                            \\
-      IAM               & Integrated Assessment Model                                         \\
-      IE                & Industrial Ecology                                                  \\
-      LCA               & Life Cycle Assessment                                               \\
-      LCI               & Life Cycle Inventory                                                \\
-      LCIA              & Life Cycle Impact Assessment                                        \\
-      LFP               & Lithium Iron Phosphate                                              \\
-      Li-ion            & Lithium-ion                                                         \\
-      LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
-      MF                & Material Footprint                                                  \\
-      MFA               & Material Flow Analysis                                              \\
-      NCA               & Nickel Cobalt Aluminum                                              \\
-      NMC               & Nickel Manganese Cobalt                                             \\
-      PWF               & Product Waste Footprint                                             \\
-      RCP               & Representative Concentration Pathway                                \\
-      ReCiPe            & A standard LCIA method set                                          \\
-      SDG               & Sustainable Development Goals                                       \\
-      SSPs              & Shared Socioeconomic Pathways                                       \\
-      T-reX             & The Tool for analysing re-X in LCA                                  \\
-      UBP               & Umweltbelastungspünkte (Environmental Impact Points)                \\
-      UN                & United Nations                                                      \\
-      WF                & Waste Footprint                                                     \\
-      \bottomrule
-  \end{tabular}
-\end{table}
-
-% Main text
-
-% Introduction
-\section{Introduction}\label{sec:introduction}
-\subsection{Introduction to T-reX}\label{sec:intro-trex}
-
-%? reviewer 2 - comment 2 - simplify language
-Development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
-
-The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
-More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
-Whether at the level of products, entire industries, nations, or even continents, the material footprint aims to quantify the total material consumed in supply chains.
-It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
-
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the `Waste Footprint' (WF) metric is less well-developed.
-While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and substandard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
-
-WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders.
-Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
-
-%? reviewer 2 - comment 1 - COMPARISON OF T-REX TO OTHER TOOLS
-%! fix
-Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
-The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
-These scores can then be aggregated into one metric for each impact category that can be compared across products and processes.
-Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
-
-Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
-
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
-
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
-Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
-\texttt{premise} offers an excellent structure for the integration of future waste and material flow modelling in LCA databases, but its potential contribution to the development of the circular economy cannot be realised if the scientific resources for data collection and validation are not forthcoming.
-The authors believe that there is an urgent need for research investment in this area.
-
-%! reviewer 3 - comment 1 - ADDITIONAL INFORMATION ON T-REX's differentiating features comparet to LCI
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
-In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and potentially highlight opportunities for waste reduction and material efficiency.
-
-%? reviewer 2 - comment 3 - ADDITIONAL INFORMATION ON T-REX's FUTURE DEVELOPMENT
-With the continual improvement of LCA databases and the development of prospective databases, T-reX can produce ever more realistic estimates, enhancing its actionable utility derived from the exploration of the potential impacts of future waste management systems and the material demands of emerging technologies.
-The better the data, the more specific and accurate the results of T-reX analysis, and, thus, the more useful the tool becomes.
-The authors hope that T-reX will be a valuable asset to the LCA community and contribute to the development of a more sustainable and circular economy.
-
-\subsection{Background and need}\label{sec:intro-background}
-
-\subsubsection{Waste in LCA}\label{sec:intro-waste}
-
-Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time.
-From a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-From a practical economic viewpoint, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows, in addition to energy costs and the relative prices of virgin and secondary materials.
-Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and `modelling [of its management] must be physically based'.
-The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.
-
-Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities.
-There remains, however, a knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
-
-Conventional LCA database models consider waste as a `service' (accounting for the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
-In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy.
-Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
-There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill.
-In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling of the long-term impacts of disposal.
-
-A considerable portion of a product's total waste is generated during earlier stages of its supply chain such as resource extraction, transportation, and manufacturing, thus, often remaining `invisible' to conventional LCA accounting practices~\citep{laurenti2016wastefootprint}.
-This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
-Studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
-Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
-The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
-
-Some existing LCA methodologies offer direct indicators at the impact assessment level, but generally, there is little attention given to the impacts of waste in LCA~\citep{lauran2020abioticdepletion}.
-This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible methods to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
-
-~\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-The method presented, as the authors explain, suffers from errors due to double counting due to the way that exchanges are identified and `tagged' .
-In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
-The `T-reX' Python software tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA\@.
-Moreover, the utility of T-reX is not limited to waste, it can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
-
-\subsubsection{Material demand in LCA}\label{sec:intro-material}
-
-In the context of geopolitical tension and a mineral-hungry `renewable energy transition', greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system rather than the primary material flows themselves.
-
-The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
-The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations.
-CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
-The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.
-
-While useful for its intended purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
-Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
-
-% Methodology
-
-\section{Methodology}\label{sec:methodology}
-
 This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries, highlighting inconsistencies in LCA methodology and probing departures from projections in prospective models.
+In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
 
@@
 Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-    \includegraphics[width=9cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
+    \includegraphics[width=12cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
 \end{figure}
@@
     \item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
-    \item Environmental Footprint (EF) v3.1 (no long-term emissions)~\citep{eu2023ef}
-    \item Spatial differentiation in life cycle impact assessment---the EDIP 2003 no long term emissions method~\citep{hauschild2003edip}
-    \item The crustal scarcity indicator method---CSI 2020~\citep{arvidsson2020csi}
+    \item EF v3.0 no LT~\citep{eu2023ef}
+    \item EDIP 2003 no LT~\citep{hauschild2003edip}
+    \item CSI 2020~\citep{arvidsson2020csi}
 \end{itemize}
@@
 \subsubsection{LCA calculations}
-For each combination of activity, method, and database, a unified metric was calculated along with details of the top contributing processes.
+For each combination of activity, method, and database, a single score was calculated along with details of the top contributing processes.
 Additionally, for the T-reX methods, a contribution analysis was performed.
-Utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem, this function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff.
-The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
-T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
-
-\section{Conclusions}\label{sec:conclusions}
-
-T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
-These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
-
-T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts.
-This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
-
-% Back matter
-\section*{Data availability}
-All data used is publically available online under the noted sources.
-T-reX is fully open source and dedicated to the public domain under the CC-0 licence.
-It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}.
-A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
-
-% \section*{CRediT authorship contribution statement} % STANDARD TEXTUAL VERSION
-% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
-% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
-% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
-
-\section*{CRediT authorship contribution statement} % VISUAL VERSION
-
-\begin{itemize}[noitemsep]
-    \item \textbf{SCM}: Stewart Charles McDowall
-    \item \textbf{EL}: Elizabeth Lanphear
-    \item \textbf{SC}: Stefano Cucurachi
-    \item \textbf{CFB}: Carlos Felipe Blanco
-\end{itemize}
-
-\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit-TREX.pdf}
-
-\section*{Declaration of competing interest}
-  The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-\section*{Acknowledgements}
-  Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
-101058522 (project FutuRaM --- futuram.eu).
-The authors would like to thank the reviewers for their valuable comments and suggestions.
-
-  The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems.
-The authors hope to continue the development of T-reX in the service of our shared ideals.
-
-  %! replace the sections
-\section{Supplementary material}\label{sec:supplementary}
-  The supplementary material supplied with this manuscript contains the following:
-    \begin{enumerate}
-        \item Metadata of the T-reX Python package
-        \item Details of the modules and user configuration
-        \item Description of the computational workflow
-        \item Modular flowchart of T-reX
-        \item Example of the terminal output of T-reX
-        \item Example of the terminal output of the case study
-        \item Complete tabulated results of the case study
-        \item Complete visualisations from the case study
-    \end{enumerate}
-
-%% Loading bibliography style file
-%\bibliographystyle{model1-num-names}
-\bibliographystyle{cas-model2-names}
-
-% Loading bibliography database
-\bibliography{cas-refs-TREX}
-
-\end{document}
+This involved utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
+This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
+This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

### 2024-12-03 00_Journal-folder_resubmission/T-reX_supplementary-material/T-reX_case-study_data/_00_main.py

@@ -1,0 +1,159 @@
+import subprocess
+import os
+from pathlib import Path
+import shutil
+
+
+from utils import *
+
+# ************************************ USER SETTINGS ************************************
+PROJECT = "TreX-premise-SSP2-cutoff"
+KEY = "market for battery, Li-ion, rechargeable"
+
+KEYWORDS_METHODS = [
+    "ReCiPe 2016 v1.03, midpoint (I)",
+    "EF v3.1 EN15804",
+    "EDIP 2003 no LT",
+    "Crustal Scarcity Indicator 2020",
+    "T-reX",
+]
+
+DELETE_OLD_RAW_RESULTS = 0
+DELETE_OLD_RAW_RESULTS_SC = 0
+
+DELETE_OLD_PROCESSED_RESULTS = 0
+DELETE_OLD_PROCESSED_RESULTS_SC = 0
+
+DELETE_OLD_FIGURES = 0
+
+SEARCH_ACTIVITIES = 1
+SEARCH_METHODS = 1
+
+GET_RESULTS = 1
+GET_SUPPLY_CHAIN_RESULTS = 1
+
+PROCESS_RESULTS = 1
+PROCESS_SUPPLY_CHAIN_RESULTS = 1
+DO_VISUALISATION = 1
+DO_VISUALISATION_SUPPLY_CHAIN = 1
+# ************************************ END USER SETTINGS ************************************
+# Directory setup
+CWD = Path.cwd()
+DIR_DATA = CWD / "data_batteries"
+DIR_VISUALISATION = CWD / "visualisation_batteries"
+
+os.makedirs(DIR_DATA, exist_ok=True)
+os.makedirs(DIR_VISUALISATION, exist_ok=True)
+
+FILE_ACTIVITIES = DIR_DATA / "activities.csv"
+FILE_METHODS = DIR_DATA / "methods.csv"
+
+FILE_RESULTS = DIR_DATA / "raw_results.csv"
+FILE_SUPPLYCHAIN_RESULTS = DIR_DATA / "raw_supply_chain_results.csv"
+
+FILE_RESULTS_PROCESSED = DIR_DATA / "results_significant.csv"
+FILE_RESULTS_PROCESSED_TOP_PROCESSES = (
+    DIR_DATA / "results_significant_top_processes.csv"
+)
+FILE_RESULTS_PROCESSED_SC_1 = DIR_DATA / "results_significant_supply_chain.csv"
[…]

### 2024-12-03 00_Journal-folder_resubmission/T-reX_supplementary-material/T-reX_case-study_data/_02a_process-results.py

@@ -1,0 +1,90 @@
+# this is a script to process the results of the case study, easy to adapt to other case studies
+
+import os
+import pandas as pd
+
+from _00_main import FILE_RESULTS, FILE_RESULTS_PROCESSED
+from utils import replace_string_in_df, term_replacements
+
+# make sure of the correct working directory
+os.chdir(os.path.dirname(os.path.abspath(__file__)))
+
+pd.set_option("display.max_columns", None)
+pd.set_option("display.max_rows", None)
+
+# load data set
+DATA_FILE_IN = FILE_RESULTS
+DATA_FILE_OUT = FILE_RESULTS_PROCESSED
+
+MIN_SCORE = 10^(-6)
+
+
+print(f"Loading data set: {DATA_FILE_IN}")
+df_raw = pd.read_csv(DATA_FILE_IN, sep=";", index_col=None)
+df = df_raw.copy()
+
+# Call the function with the DataFrame and the replacements dictionary
+for term, replacement in term_replacements.items():
+    df_raw = replace_string_in_df(df, term_replacements)
+
+print(f"Data set loaded. Shape: {df.shape}")
+print(f"Columns: \n{df.columns.values}")
+print("Processing data...")
+# filter out unnecessary columns
+df = df[df.database != "ecoinvent-3.9.1-cutoff"]
+drop = [
+    "db_model",
+    "db_version",
+    "db_system",
+    "db_base",
+    "code",
+    "db_ssp",
+    "database",
+]
+
+
+print(f"Filtering out columns: {drop}")
+for d in drop:
+    df = df.drop(columns=[d], errors="ignore")
+
+
+# simplify the names of the activities
+df.name = df.name.apply(lambda x: x.split(", ")[2])
+
+df = df.assign(method_1=df["method_1"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+))
+df = df.assign(method_2=df["method_2"].apply(
+    lambda x: (x[:32] + "[...]") if len(x) > 35 else x
+))
+
+
+# check for signigicant methods
+grouped = df["score"].groupby(df["method_2"]).mean().sort_values(ascending=False)
+
+# # filter out methods where the maximum score is less than a certain value
+print(f"Filtering out methods where the maximum score < {MIN_SCORE}")
+df_sig = df.groupby("method_2").filter(lambda x: x["score"].abs().max() > MIN_SCORE)
[…]

### 2024-12-03 02_Journal-folder_first-submission_original/T-reX_supplementary-material/T-reX_case-study_data/_05a_visualisation_bar.py

@@ -1,0 +1,332 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+import matplotlib as mpl
+from matplotlib.lines import Line2D
+import scienceplots
+from tqdm import tqdm
+import seaborn as sns
+import random
+import numpy as np
+
+from utils import term_replacements, replace_string_in_df, replace_strings_in_string
+from _00_main import FILE_RESULTS_PROCESSED_SC_2, DIR_VISUALISATION
+
+FILE_IN = FILE_RESULTS_PROCESSED_SC_2
+
+# Set a fixed figure size to match your largest plot (274x343 pixels)
+# fixed_figsize = (250 / 100, 500 / 100)  # Convert pixels to inches
+
+# plt.rcParams["figure.figsize"] = fixed_figsize
+# plt.rcParams["figure.autolayout"] = False
+
+
+# set global formatting options
+rc_fonts = {
+    "font.family": "serif",
+    "text.usetex": True,
+    "text.latex.preamble": "\n".join(
+        [
+            "\\usepackage{libertine}",
+            "\\usepackage[libertine]{newtxmath}",
+        ]
+    ),
+}
+mpl.style.use(["science"])
+mpl.rcParams.update(rc_fonts)
+
+# Read in the data
+df_raw = pd.read_csv(FILE_IN, sep=";", index_col=None)
+
+
+for term, replacement in term_replacements.items():
+    df_raw = replace_string_in_df(df_raw, term_replacements)
+
+# Aggregate the methods into a single column
+method_full = list(zip(df_raw["method_0"], df_raw["method_1"], df_raw["method_2"]))
+df_raw.insert(0, "method_full", method_full)
+
+# # Define the plotting colors
+# color_palette = [
+#     "#FF5733", "#3B9C9C", "#FFC300", "#FF6363", "#4CAF50",
+#     "#9C27B0", "#2196F3", "#F44336", "#FFEB3B", "#673AB7",
+#     "#009688", "#795548", "#00BCD4", "#E91E63", "#FF7043",
+#     "#4CAF50", "#FF9800", "#607D8B", "#9C27B0", "#3F51B5",
+#     "#E91E63", "#FFC107", "#8BC34A", "#2196F3", "#FF5722",
+#     "#00BCD4", "#FF5252", "#536DFE", "#FF9800", "#9E9E9E",
[…]

### 2024-12-03 02_Journal-folder_first-submission_revised/T-reX_supplementary-material/T-reX_case-study_data/_04a_visualisation_scatter.py

@@ -1,0 +1,256 @@
+import matplotlib as mpl
+from matplotlib.lines import Line2D
+import scienceplots
+import pandas as pd
+import numpy as np
+import os
+
+from utils import *
+
+
+# set global formatting options
+rc_fonts = {
+    "font.family": "serif",
+    "text.usetex": True,
+    "text.latex.preamble": "\n".join(
+        [
+            "\\usepackage{libertine}",
+            "\\usepackage[libertine]{newtxmath}",
+        ]
+    ),
+}
+
+title = "Li-ion batteries:"
+mpl.style.use(["science", "scatter"])
+mpl.rcParams.update(rc_fonts)
+
+import matplotlib.pyplot as plt
+
+from _00_main import FILE_RESULTS_PROCESSED, DIR_VISUALISATION
+from utils import term_replacements, replace_strings_in_string
+
+# # for testing
+# x = np.linspace(0, 10, 100)
+# y = np.sin(x)
+# plt.plot(x, y)
+
+# make sure of the correct working directory
+os.chdir(os.path.dirname(os.path.abspath(__file__)))
+
+FILE_IN = FILE_RESULTS_PROCESSED
+
+# read in the data
+df_raw = pd.read_csv(FILE_IN, sep=";", index_col=None)
+
+df_raw.sort_values(
+    by=["method_0", "method_1", "method_2", "db_target", "db_year"], inplace=True
+)
+
+df_raw["method_full"] = list(
+    zip(df_raw["method_0"], df_raw["method_1"], df_raw["method_2"])
+)
+
+total_methods = len(df_raw.method_full.unique())
+
+activities = df_raw.name.unique()  # eg. ["NMC111", "NCA", "NMC811", "LFP", "LiMn2O4"]
+scenarios = df_raw.db_target.unique()  # eg. ["Base", "PkBudg500"]
+years = df_raw.db_year.unique()  # eg. [2020, 2030, 2040, 2050]
+methods_0 = df_raw.method_0.unique()
+methods_1 = df_raw.method_1.unique()
+methods_2 = df_raw.method_2.unique()
+width_1 = 90  # mm
+width_3on2 = 140  # mm
+width_2 = 190  # mm
+
+markers = ["o", "x", "^", "D", "*", "s"]  # Add more markers as needed
+marker_weight = 1
+
+colorblind_friendly_palette = [
+    "#377eb8",
+    "#ff7f00",
+    "#4daf4a",
+    "#e41a1c",
[…]

### 2024-12-03 02_Journal-folder_first-submission_revised/T-reX_supplementary-material/T-reX_case-study_data/_04b_visualisation_scatter_combine.py

@@ -1,0 +1,85 @@
+import os
+from tqdm import tqdm
+from svglib.svglib import svg2rlg
+from reportlab.graphics import renderPDF
+from reportlab.pdfgen import canvas
+from reportlab.pdfbase.ttfonts import TTFont
+from reportlab.pdfbase import pdfmetrics
+
+# Register Libertine font
+font_path = "LinLibertine_Rah.ttf"  # Change as needed
+pdfmetrics.registerFont(TTFont("Libertine", font_path))
+
+from _00_main import DIR_VISUALISATION
+# Directory containing SVG files
+svg_dir = DIR_VISUALISATION / "individual-methods"
+print(f'{"*"*50}')
+print(f"Reading SVG files from directory: {svg_dir}")
+print(f'{"*"*50}')
+
+# Define the layout and title
+title = "T-reX --- Scatter plots of individual methods"
+columns = 3
+figures = [f for f in os.listdir(svg_dir) if f.endswith(".svg")]
+figures = sorted(figures, reverse=True)
+figures_count = len(figures)
+rows = figures_count // columns + (1 if figures_count % columns else 0)
+
+if figures_count == 0:
+    raise ValueError("No SVG files found in the directory.")
+else:
+    print(f"Found {figures_count} SVG files.")
+
+# Get the size of the first SVG to determine the size for all
+first_svg = svg2rlg(os.path.join(svg_dir, figures[0]))
+img_width = first_svg.width
+img_height = first_svg.height
+print(f"Each SVG size: {round(img_width, 2)}x{round(img_height, 2)} units.")
+
+# Calculate the size of the page
+title_height = 50  # Adjust as needed
+page_width = columns * img_width
+page_height = rows * img_height + title_height
+print(
+    f"Creating a single-page PDF with size: {round(page_width,2)}x{round(page_height,2)} units."
+)
+
+# Create a PDF file
+c = canvas.Canvas(
+    str(DIR_VISUALISATION / "scatter-combined_methods.pdf"),
+    pagesize=(page_width, page_height),
+)
+
+# Add the title
+c.setFont("Libertine", 16)  # Set Libertine font and size for the title
+title_width = c.stringWidth(title, "Libertine", 16)
+c.drawString(
+    (page_width - title_width) / 2, page_height - 35, title
+)  # Center the title
+
+# Add metadata
+c.setTitle("T-reX: Scatter Plots of Individual Methods")  # PDF Title
[…]

### 2024-12-03 02_Journal-folder_first-submission_revised/T-reX_supplementary-material/T-reX_case-study_data/_04c_visualisation_scatter_similar_methods.py

@@ -1,0 +1,93 @@
+import os
+from svglib.svglib import svg2rlg
+from reportlab.graphics import renderPDF
+from reportlab.pdfgen import canvas
+from reportlab.pdfbase.ttfonts import TTFont
+from reportlab.pdfbase import pdfmetrics
+from _00_main import DIR_VISUALISATION
+
+pairs = {
+    "EDIP 2003 no LT-non-renewable resources no LT-aluminium no LT": "T-reX-Demand: Aluminium-Aluminium",
+    "EDIP 2003 no LT-non-renewable resources no LT-brown coal no LT": "T-reX-Demand: Coal(brown)-Coal(brown)",
+    "EDIP 2003 no LT-non-renewable resources no LT-coal no LT": "T-reX-Demand: Coal(black)-Coal(black)",
+    "EDIP 2003 no LT-non-renewable resources no LT-cobalt no LT": "T-reX-Demand: Cobalt-Cobalt",
+    "EDIP 2003 no LT-non-renewable resources no LT-copper no LT": "T-reX-Demand: Copper-Copper",
+    "EDIP 2003 no LT-non-renewable resources no LT-natural gas no LT": "T-reX-Demand: Natural gas-Natural gas",
+    "EDIP 2003 no LT-non-renewable resources no LT-nickel no LT": "T-reX-Demand: Nickel-Nickel",
+    "EDIP 2003 no LT-non-renewable resources no LT-oil no LT": "T-reX-Demand: Petroleum-Petroleum",
+    "EDIP 2003 no LT-non-renewable resources no LT-zinc no LT": "T-reX-Demand: Zinc-Zinc",
+}
+
+# Register Libertine font
+font_path = "LinLibertine_Rah.ttf"  # Update the path as needed
+pdfmetrics.registerFont(TTFont("Libertine", font_path))
+
+# Directory containing SVG files
+svg_dir = DIR_VISUALISATION / "individual-methods"
+print(f"Reading SVG files from directory: {svg_dir}")
+
+# Define the layout and title
+title = "T-reX --- Comparison of Similar Methods"
+page_width = 0
+page_height = 0
+title_height = 40  # Space for title
+
+# Calculate total page height based on pairs
+for edip, T_reX in pairs.items():
+    edip_file = next((f for f in os.listdir(svg_dir) if edip in f), None)
+    T_reX_file = next((f for f in os.listdir(svg_dir) if T_reX in f), None)
+
[…]

### 2024-12-03 03_T-reX-latex_files-atfirstsubmission/latex_files-atfirstsubmission/.compile_pdf.sh

@@ -1,0 +1,99 @@
+#!/bin/bash
+
+# Define paths and filenames
+source_folder="."
+ # Assuming your LaTeX files are in the current directory
+target_folder="../00_JRR-submission_folder"
+main_tex_file="main.tex"                   
+bib_file="references.bib"                  
+output_pdf_name="00_T-reX_manuscript.pdf"  
+output_docx_name="00_T-reX_manuscript.docx"
+output_html_name="00_T-reX_manuscript.html"
+
+# Optional: Specify Pandoc templates for DOCX and HTML
+docx_template="path/to/docx_template.docx"
+html_template="path/to/html_template.html"
+
+echo "======================================================="
+echo "Compiling $main_tex_file to PDF, DOCX, and HTML"
+
+# Create target folder if it doesn't exist
+mkdir -p "$target_folder"
+
+# Clean up before compiling
+latexmk -c -outdir="$target_folder" "$source_folder/$main_tex_file" || exit 1
+
+# Compile LaTeX document to PDF, output directly to target folder
+latexmk -pdf -outdir="$target_folder" "$source_folder/$main_tex_file" || exit 1
+
+# Conversion commands for DOCX and HTML with template usage if available
+if [ -f "$docx_template" ]; then
+    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" --reference-doc="$docx_template" -o "$target_folder/$output_docx_name" || exit 1
+else
+    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" -o "$target_folder/$output_docx_name" || exit 1
+fi
+
+if [ -f "$html_template" ]; then
+    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" --template="$html_template" -o "$target_folder/$output_html_name" || exit 1
+else
+    pandoc -s "$target_folder/${main_tex_file%.*}.pdf" -o "$target_folder/$output_html_name" || exit 1
+fi
+
+echo "Process completed successfully."
+echo "PDF saved to $target_folder/$output_pdf_name"
+echo "DOCX saved to $target_folder/$output_docx_name"
+echo "HTML saved to $target_folder/$output_html_name"
+echo "======================================================="
+
+
+
+##OLDVERSION
+
+# #!/bin/bash
+# 
+# # Define paths and filenames
+# source_folder="." # Assuming your LaTeX files are in the current directory
[…]

### 2025-02-09 00_T-reX_journal-upload-folder/T-reX_latex-files/cas-dc-template.tex
auto-update

--- 2024-12-02 20240201_T-reX_JournalUploadFolder/main.tex
@@
 % \usepackage[authoryear,longnamesfirst]{natbib}
-%\usepackage{booktabs}
-%\usepackage[tt=false]{libertine}
-%\usepackage{xurl}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage[british]{babel}
+\usepackage{microtype}
+% \usepackage[tt=false]{libertine}
+\usepackage{xcolor}
+\usepackage{xurl}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{lineno}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{geometry}
+\usepackage{tabularx}
+\usepackage{fontspec}
+\setmainfont[Mapping=tex-text,Ligatures={Common,Rare,Discretionary}]{Linux Libertine}
 
@@
 \begin{graphicalabstract}
-		\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract.pdf}
-		%\includegraphics[width=\columnwidth]{grabs_T-reX_graphical-abstract_MID-RES.png}
+	\includegraphics[width=1\columnwidth]{figs/grabs_T-reX_graphical-abstract.pdf}
+	%\includegraphics[width=\columnwidth]{figs/grabs_T-reX_graphical-abstract_MID-RES.png}
 \end{graphicalabstract}
@@
 \begin{keywords}
-		circular economy, waste footprint, material footprint, life cycle assessment, critical raw materials, supply chain
+circular economy\sep waste footprint\sep material footprint\sep life cycle assessment\sep critical raw materials \sep supply chain
 \end{keywords}
@@
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX_method.pdf}
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX_method.pdf}
 	\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format.
@@
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX_NMC811_WasteTotalSolid.pdf}
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX_NMC811_WasteTotalSolid.pdf}
 	\caption{A greatly simplified Sankey diagram from \texttt{ActivityBrowser} showing the total solid waste footprint flows for the
@@
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX-wastefootprint-totalANDcarbon.pdf}
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX-wastefootprint-totalANDcarbon.pdf}
 	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only).
@@
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-processcontributions.pdf}
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX-wastefootprint-processcontributions.pdf}
 	\caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model}\label{fig:top_contribution}
@@
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX-wastefootprint-sectorcontributions.pdf}
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX-wastefootprint-sectorcontributions.pdf}
 	\caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model}\label{fig:cpc_contribution}
@@
 	\centering
-		\includegraphics[width=1.8\columnwidth]{T-reX-coalANDedip.pdf}
+	\includegraphics[width=1.8\columnwidth]{figs/T-reX-coalANDedip.pdf}
 	\caption{Comparison of the (a) `T-reX – Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 – coal no longterm'.
@@
 	\centering
-		\includegraphics[width=0.8\columnwidth]{T-reX_credit_heatmap.pdf}
+	\includegraphics[width=0.8\columnwidth]{figs/T-reX_credit_heatmap.pdf}
 	\caption{CRediT authorship visualisation}\label{fig:credit_heatmap}
@@
 % Loading bibliography database
-	\bibliography{01_T-reX_manuscript}
+\bibliography{cas-refs}
 

### 2025-02-09 03_T-reX-latex_files-DRAFTING/.main.tex
auto-update

--- 2024-09-22 latex_files/.main.tex
@@
 \begin{highlights}
-    \item 'T-reX', a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
+    \item ``T-reX'', a new open-source program for quantifying waste and material flows in Life Cycle Assessment (LCA)
     \item Assesses supply risks by calculating demand inventories for critical materials
@@
 
-    \title{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
+    \title{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases.\@}
     \author[1]{Stewart Charles McDowall\corref{cor1}}
@@
     \begin{abstract}
-        % \cbox{Abstract word count: 150, Limit: 150}
         The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.

### 2025-02-09 03_T-reX-latex_files-DRAFTING/introduction.tex
auto-update

--- 2024-09-22 latex_files/introduction.tex
@@
 Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
+LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'---often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}.
 The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
@@
 To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc...) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
+T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc\ldots) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and treatment or EOL handling.
 

### 2025-02-09 03_T-reX-latex_files-DRAFTING/main.tex
auto-update

--- 2024-05-02 latex_files/main.tex
@@
 \usepackage[tt=false]{libertine}
+\usepackage{listings}
 \usepackage{xcolor}
@@
 \begin{document}
-\vspace{20em}
-\includegraphics[width=10cm]{graphical-abstract.pdf}
+\vspace{30em}
+\includegraphics[width=\textwidth]{graphical-abstract.pdf}
 
@@
 %% GRAPHICAL ABSTRACT
-% \begin{graphicalabstract}
-%     \includegraphics[width=8cm]{graphical-abstract.pdf}
-% \end{graphicalabstract}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{grabs_T-reX_graphical-abstract.pdf}
+	\caption{Graphical abstract of the T-reX program for waste and material footprinting in Life Cycle Assessment}
+	\label{fig:graphical_abstract}
+\end{figure}
+
+% Switch to landscape for the included PDF
+\newgeometry{landscape}
+\includepdf[pages=1, width=\textheight]{grabs_T-reX_graphical-abstract.pdf}
+\restoregeometry % Restore the original geometry
+
 %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
@@
 % list of abbreviations
-\begin{table}[H]
+\begin{table}[]
     \centering
@@
 \section*{Data availability}
-All data used is publically available online under the noted sources.
+All data used is publicly available online under the noted sources.
 T-reX is fully open source and dedicated to the public domain under the CC-0 licence.

### 2025-02-09 03_T-reX-latex_files-DRAFTING/methodology.tex
auto-update

--- 2024-09-22 latex_files/methodology.tex
@@
 In the default configuration of T-reX, there are 10 waste categories which are then further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary}.
+The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary} [section 3 ].
 One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
@@
 
-To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
+To provide an example, we investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories.
+The system boundaies.
+These following markets were used for the example calculations and unaltered by the authors:
 \begin{itemize}[itemsep=0pt]
@@
 The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement].
-T-reX is, an sich, agnostic to interests of its user, providing a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
+T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

### 2025-02-09 03_T-reX-latex_files-atfirstsubmission/latex_files-atfirstsubmission/results.tex
auto-update

--- 2024-12-03 00_T-reX_journal-upload-folder/T-reX_latex-files/cas-dc-template_copyfordiff.tex
@@
-\documentclass[a4paper,fleqn]{cas-dc}
-
-% If the frontmatter runs over more than one page
-% use the longmktitle option.
-%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
-
-\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
-\tsc{SCM}
-\tsc{EL}
-\tsc{SC}
-\tsc{CFB}
-
-%\usepackage[numbers]{natbib}
-\usepackage[authoryear]{natbib}
-% \usepackage[authoryear,longnamesfirst]{natbib}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[T1]{fontenc}
-\usepackage[british]{babel}
-\usepackage{microtype}
-% \usepackage[tt=false]{libertine}
-\usepackage{xcolor}
-\usepackage{xurl}
-\usepackage{pdfpages}
-\usepackage{enumitem}
-\usepackage{lineno}
-\usepackage{float}
-\usepackage{longtable}
-\usepackage{geometry}
-\usepackage{tabularx}
-\usepackage{fontspec}
-\setmainfont[Mapping=tex-text,Ligatures={Common,Rare,Discretionary}]{Linux Libertine}
-
-%%%
-\begin{document}
-\let\WriteBookmarks\relax
-\def\floatpagepagefraction{1}
-\def\textpagefraction{.001}
-
-% Short title
-\shorttitle{T-reX: Quantifying waste and material footprints in LCA}
-
-% Short author
-\shortauthors{McDowall, S2. et~al.}
-
-% Main title of the paper
-\title [mode = title]{T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
-
-% First author
-%
-% Options: Use if required
-% eg: \author[1,3]{Author Name}[type=editor,
-%       style=chinese,
-%       auid=000,
-%       bioid=1,
-%       prefix=Sir,
-%       orcid=0000-0000-0000-0000,
-%       facebook=<facebook id>,
-%       twitter=<twitter id>,
-%       linkedin=<linkedin id>,
-%       gplus=<gplus id>]
-
-\author[1]{Stewart Charles McDowall}[orcid=0000-0002-5688-1279]
-
-% Corresponding author indication
-\cormark[1]
-% Email id of the first author
-\ead{s.c.mcdowall@cml.leidenuniv.nl}
-% URL of the first author
-\ead[url]{scmcd.ch}
-% Credit authorship
-\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing: original draft, Writing: review \& editing, Visualization}
-
-\affiliation[1]{
-  organization={Institute of Environmental Sciences (CML), Leiden University},
-  addressline={P.O. Box 9518},
-  city={Leiden},
-  postcode={2300RA},
-  state={South Holland},
-  country={The Netherlands}}
-
-\author[1]{Elizabeth Lanphear}[orcid=0009-0003-6364-0873]
-\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing}
-
-\author[1]{Stefano Cucurachi}[orcid=0000-0003-2748-655X]
-\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
-
-\author[1]{Carlos Felipe Blanco}[orcid=0000-0001-8199-8420]
-\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}
-
-% Corresponding author text
-\cortext[1]{Corresponding author}
-
-% Here goes the abstract
-\begin{abstract}
-        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
-Thus, identifying and quantifying waste and material flows is critical.
-Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.
-
-We introduce T-reX, a Python tool extending the brightway ecosystem, to allow facile quantification of user-defined supply chain demands in current and prospective scenarios.
-T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.
-
-A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens.
-T-reX can aid sustainable decision-making and contribute to the development of `circular economy' by facilitating analysis of material consumption and waste generation in LCA.
-\end{abstract}
-
-% Use if graphical abstract is present
-\begin{graphicalabstract}
-	\includegraphics[width=1\columnwidth]{figs/grabs_T-reX_graphical-abstract.pdf}
-	%\includegraphics[width=\columnwidth]{figs/grabs_T-reX_graphical-abstract_MID-RES.png}
-\end{graphicalabstract}
-
-% Highlights
-\begin{highlights}
-	\item T-reX, a new open-source program for waste \& material footprints in Life Cycle Assessment (LCA)
-	\item Assesses supply risks by calculating demand inventories for critical materials
-	\item Simplifies quantification of user-specified waste and material footprint categories
-	\item Rapidly identifies hotspots of waste generation and material demand in supply	chains
-	\item Presents a case study of five lithium-ion batteries to demonstrate T-reX's potential
-\end{highlights}
-
-% Keywords
-% Each keyword is seperated by \sep
-
-\begin{keywords}
-circular economy\sep waste footprint\sep material footprint\sep life cycle assessment\sep critical raw materials \sep supply chain
-\end{keywords}
-
-\maketitle
-
-% Main text
-\begin{table*}[!htbp]
-	\centering
-	\caption{List of abbreviations}\label{tab:abbreviations}
-	\begin{tabular}{ll}
-		\toprule
-		CCS	              & Carbon Capture and Storage                                          \\
-		CF                & Carbon Footprint                                                    \\
-		EF                & Ecological Footprint                                                \\
-		CML               & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
-		CRM               & Critical Raw Material                                               \\
-		CPC               & Cooperative Patent Classification                                   \\
-		CSI               & Crustal Scarcity Indicator                                          \\
-		CSP               & Crustal Scarcity Potential                                          \\
-		EDIP              & Environmental Design of Industrial Products                         \\
-		EEIOA             & Environmentally Extended Input-Output Analysis                      \\
-		EOL               & End-of-Life                                                         \\
-		FOEN              & Swiss Federal Office for the Environment                            \\
-		IAM               & Integrated Assessment Model                                         \\
-		IE                & Industrial Ecology                                                  \\
-		LCA               & Life Cycle Assessment                                               \\
-		LCI               & Life Cycle Inventory                                                \\
-		LCIA              & Life Cycle Impact Assessment                                        \\
-		LFP               & Lithium Iron Phosphate                                              \\
-		Li-ion            & Lithium-ion                                                         \\
-		LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
-		MF                & Material Footprint                                                  \\
-		MFA               & Material Flow Analysis                                              \\
-		NCA               & Nickel Cobalt Aluminum                                              \\
-		NMC               & Nickel Manganese Cobalt                                             \\
-		PWF               & Product Waste Footprint                                             \\
-		RCP               & Representative Concentration Pathway                                \\
-		ReCiPe            & A standard LCIA method set                                          \\
-		SDG               & Sustainable Development Goals                                       \\
-		SSPs              & Shared Socioeconomic Pathways                                       \\
-		T-reX             & The Tool for analysing re-X in LCA                                  \\
-		UBP               & Umweltbelastungspünkte (Environmental Impact Points)                \\
-		UN                & United Nations                                                      \\
-		WF                & Waste Footprint                                                     \\
-		\bottomrule
-	\end{tabular}
-\end{table*}
-
-\section{Introduction}\label{sec:introduction}  
-The development of a `circular economy' has become a central focus in the frantic pursuit of sustainability objectives and the imperative curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}.
-Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}.
-In addition to circular economy goals, contemporary geo-political tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.
-
-The concept of the `footprint' as an environmental sustainability indicator began with the Ecological Footprint (EF)~\citep{wackernagel1994ecologicalfootprint} and after being popularised by the Carbon Footprint (CF)~\citep{cucek2015environmentalfootprints} the `footprint family' has adopted many additional metrics---albeit without yet coalescing into a coherent or consistent framework~\citep{giampietro2014footprintstonowhere, vanham2019footprints,ridoutt2013footprints}.
-More recently, the footprint collection has been extended to include the Material Footprint (MF)~\citep{weidmann2013materialfootprint}, which is often encountered in the `macro methods' of Industrial Ecology (IE), such as environmentally Extended Input-Output Analysis (EEIOA)~\citep{lenzen2022materialfootprint} and Material Flow Analysis (MFA)~\citep{schaffartzik2013mfafootprint}.
-Whether at the level of products, entire industries, nations or even continents, the material footprint aims to quantify the total material consumed in supply chains.
-It has been shown that the MF can be `highly representative of damage to human health and biodiversity'~\citep{steinmann2017resourcefootprints} and indeed, this metric was recently beatified by the United Nations (UN), becoming the `core official indicator' for targets 8.4 and 12.2 of the Sustainable Development Goals (SDGs)~\citep{lenzen2022materialfootprint}.
-
-Though steadily emerging~\citep{laurenti2016wastefootprint,demirer2019wastefootprint,guillotreau2023wastefootprint}, the incipient `Waste Footprint' (WF) metric, in contrast, has yet to enjoy such noble recognition for its attempts to measure and classify the waste generated by human activities.
-While it seems obvious that reducing life cycle waste is critical to the development of the circular economy~\citep{towa2020wastefootprint,ellenmacarthur2015ce}, the WF remains largely overlooked---especially in LCA models where waste itself is seldom apportioned any inherent environmental significance aside from the emissions related to its treatment~\citep{laurenti2023wastefootprint}.
-Such ignominious neglect strikes the authors as unjustified, given that it has been repeatedly demonstrated that WFs have a strong association with environmental damage~\citep{laurenti2023wastefootprint,doka2024publications, ridoutt2010wasteimpacts,jaio2013wasteabsorbtionfootprint}.
-Additionally, it has been shown that the most vulnerable communities suffer disproportionately from the social and ecological impacts created by wasteful behaviours and sub-standard end-of-life treatment~\citep{pellow2023envjusticewaste,akese2018envjustice}.
-
-WF and MF, the two metrics of focus in the study, can provide a comprehensive assessment of potential environmental impacts across the supply chain, encapsulating both resource use and pollution/waste generation, while offering insights at diverse scales, from individual activities to global systems, facilitating communication with a diverse range of stakeholders.
-Thus, to reduce the negative externalities of human consumption and improve supply chain resilience, it is essential to uncover, disaggregate, and quantify the material and waste footprints of human activities in as much detail as possible~\citep{bisinella2024wastelca, towa2020wastefootprint,berger2020mineralsinlca,sonderegger2020mineralsinlca}.
-
-Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes.
-LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of actions~\citep{guinee2011lca}.
-The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere).
-These scores can then be aggregated into a single score for each impact category, that can be compared across products and processes.
-Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.
-
-Several LCIA methods include, to some extent, waste generation (Swiss Eco-Factors, EDIP and EN15804)~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption (Crustal Scarcity Indicator (CSI) and Swiss Eco-Factors~\citep{arvidsson2020csi,foen2021ecofactors}).
-These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., \textit{Ümweltbelastungspunkte} (UBP) [English: environmental impact points] in the case of the Swiss Eco-Factors)~\citep{su2020sustainableproddev}.
-
-\subsubsection{Waste in LCA}\label{sec:intro-waste}
-
-Though often described simply as a `material with a negative economic value'~\citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and highly variable across space and time.
-From a systems perspective, the notion of waste is anathema to the `circular economy', they cannot co-exist.
-From a practical economic viewpoint, the viability of an extensive waste processing system is highly dependent on precise knowledge (or at least, reasonable predictions) of material and waste flows, in addition to energy costs and the relative prices of virgin and secondary materials.
-Thus, as~\cite{bisinella2024wastelca} argues, waste must be `thoroughly characterised' and `modelling [of its management] must be physically based'.
-The reliable, robust models needed to guide the development of the circular economy can be built only on a foundation of high-quality data.
-
-Accurate and detailed information about waste and waste systems is, by definition, essential for understanding the `circularity' of a human activity and for predicting its life cycle externalities.
-There remains, however, a knowledge gap regarding WFs and their associated environmental impacts~\citep{laurenti2023wastefootprint}.
-
-Conventional LCA database models consider waste as a `service' (accounting for the treatment, not the material)~\citep{guinee2021wasteisnotaservice} and typically use generic waste processing models~\citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.
-In LCA, waste flows are (almost exclusively) not considered as fundamental biosphere exchanges, but rather, as technosphere flows within the human economy.
-Waste produced by an activity is transferred to a relevant waste treatment activity where it is accepted `burden-free' and transformed into a combination of emissions and other waste `products'~\citep{guinee2021wasteisnotaservice}.
-There can be several treatment steps in this pathway leading, ultimately, to a mass of material being deposited in a landfill.
-In this system of waste accounting, the impacts apportioned to the waste-producing activity are a sum of those incurred by the transport, treatment, and final disposal of the waste into terrestrial or aquatic environments.
-In particular, the extensive work of~\cite{doka2024publications} has contributed significantly to understanding the environmental impacts of waste treatment processes and the modelling of the long-term impacts of disposal.
-
-A considerable portion of a product's total waste is generated during earlier stages of its supply chain such as resource extraction, transportation, and manufacturing, thus, often remaining `invisible' to conventional LCA accounting practices~\citep{laurenti2016wastefootprint}.
-This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators.
-Traditional LCA does not typically view waste as having environmental significance \textit{per se}, considering instead only emissions and resource use resulting from waste treatment~\citep{bisinella2024wastelca, laurenti2023wastefootprint}.
-The environmental significance of waste and its correlation with other indicators has been the subject of extensive research.
-Studies have shown that popular resource footprints can cover a significant portion of environmental impact variance between activities~\citep{steinmann2017resourcefootprints,laurenti2023wastefootprint}.
-Correlations between various environmental indicators are not always consistent, however, as seen with the carbon footprint, which often does not correlate well with other impact assessment scores~\citep{laurenti2012carbonfootprint}.
-The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.
-
-Existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste.
-This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle.
-Addressing these hotspots is crucial for advancing `circularity', but there is a lack of convenient and flexible methods to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a greater understanding of a product's total environmental footprint.
-
-\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, while also proposing simple measures to quantify waste `hazardousness' and `circularity' with this data.
-The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, and not easily reproducible.
-As the authors acknowledge, due to the way that exchanges are identified and `tagged' this method also suffers from errors due to double counting.
-In T-reX, the source database is deconstructed into a list of separate exchanges, and only those explicitly defined as hazardous are marked as such.
-The `T-reX' Python software tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA\@.
-Moreover, the utility of T-reX is not limited to waste, it can be used to categorise and aggregate any technosphere exchange of interest (or customised grouping thereof), such as water, gas, and critical raw materials.
-
-\subsubsection{Material demand in LCA}\label{sec:intro-material}
-
-In the context of geo-political tension and a mineral-hungry `renewable energy transition', ever greater attention is being given to the security of material supply---especially for those considered `critical raw materials' (CRMs)~\citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain,iea2023crm,iea2023energytechperspectives}.
-While LCA seeks to model the technosphere (a.k.a.\ the anthroposphere) and its exchanges with the biosphere, its focus is often on the environmental impacts of the system---the endpoints---rather than the primary material flows themselves.
-
-The Crustal Scarcity (CSI) LCIA method was developed in order to introduce an assessment of the long-term global scarcity of minerals in LCA~\citep{arvidsson2020csi}.
-The CSI introduced crustal scarcity potentials (CSPs), which are measured in kg--silicon--equivalents per kg--element and derived from crustal concentrations.
-CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies.
-The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.
-
-While useful for its intended purpose, the CSI presents its results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories.
-Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.
-
-\subsection{Introduction to T-reX}\label{sec:intro-trex}
-
-To facilitate the quantification of waste and material flows in LCA, we developed a Python package to extend the \texttt{brightway} open-source LCA framework~\citep{mutel2017brightway} and track these exchanges by translating them into inventory indicators and `pseudo' LCA impact (LCIA) categories.
-T-reX is `T(ool for) re-X' (reduce, reuse, recycle, etc.) and enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.
-
-Integration with the \texttt{premise} Python package~\citep{sacchi2022premise}---which connects the projections of integrated assessment models (IAMs) with current LCA databases---enables the user to create and manipulate prospective LCA databases.
-Frustratingly, the current utility of prospective databases---in general, and in particular for the waste sector---is constrained by the fact that, to date, the sectoral coverage of future life cycle inventories (LCIs) is largely confined to energy, steel, cement, and transport~\citep{sacchi2023premisedocs}.
-Indeed, despite the ever more critical need to reliably model and future waste management systems,~\cite{bisinella2024wastelca} reports an alarming lack of coherent development in this field.
-
-The purpose of T-reX is not to quantify the environmental impacts of material consumption and waste production (the requisite data and impact modelling remain lacking), but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes.
-It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows.
-By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste (though whether this objective is thermodynamically impossible has long been the subject of debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others).
-In any case, minimising primary material consumption and the generation of waste is of paramount importance.
-By allowing LCA practitioners to easily classify and quantify these exchanges, T-reX provides a practical means to identify hotspots and highlight opportunities for waste reduction and material efficiency.
-
-\section{Methodology}\label{sec:methodology}
-
-This section is divided into two parts.
-In \autoref{sec:method-TreX}, we describe the T-reX program, and in \autoref{sec:method-casestudy}, we describe the methodology used to calculate the waste and material inventory footprints in a case study of five Li-ion batteries.
-
-\subsection{T-reX}\label{sec:method-TreX}
-
-\subsubsection{Computational framework}
-
-Developed in the Python programming language, T-reX extends the \texttt{brightway LCA framework}, utilising the components \texttt{bw2data}, \texttt{bw2calc}, and \texttt{bw2io}~\citep{mutel2017brightway}.
-Additionally, the \texttt{wurst} package---which can deconstruct databases into a list of exchanges---is used to facilitate searching and data transformation at the exchange level (the individual supply chain flows)~\citep{mutel2017wurst}.
-Integration with the \texttt{premise} package~\citep{sacchi2022premise}---which integrates the projections of integrated assessment models (IAMs) into current LCA databases---enables the user to easily create and manipulate prospective LCA databases.
-T-reX is also compatible with \texttt{ActivityBrowser}~\citep{steubing2020activitybrowser}---an open-source graphical user interface for LCA---and after running T-reX, the manipulated databases and the `pseudo-LCIA' methods created by T-reX can be used in the project in the accustomed way.
-T-reX is installable via the Python Package Index (PyPI)~\citep{mcdowall2023T-reXpipy} and is fully open-source under the CC-0 licence.
-The full source code for T-reX is indexed on Zenodo~\citep{mcdowall2023T-reXzenodo} and under further development in the GitHub repository~\citep{mcdowall2024T-reXgithub}.
-T-reX is designed to be used with \texttt{ecoinvent} databases~\citep{ecoinvent2016version3}, but could be adapted to other databases by changing the search criteria.
-Currently, it has been tested with all available system models of \texttt{ecoinvent}versions 3.5--3.10.
-
-T-reX can be used directly from the command line, or imported as a Python package with which the user can access the individual functions and modules.
-In the simplest case, the user can run the program with the default settings, which will calculate the waste and material footprint of the \texttt{ecoinvent}database.
-The user can adapt the settings of T-reX as desired, to calculate alternative or additional waste and material inventory footprints, use a custom database, or a `prospective' database based on future scenarios by implementing the premise \texttt{premise} with one of T-reX's internal modules.
-
-The supplementary material in \autoref{sec:supplementary} contains metadata of T-reX, along with a list of the constituent modules, a description of their functions, and a detailed computational workflow.
-Further details can be found in the GitHub repository and package documentation~\citep{mcdowall2024T-reXgithub, mcdowall2023T-reXdocs}.
-
-\subsubsection{Functionality and purpose}
-
-T-reX is a Python package that enables one to produce LCA databases---both current and prospective---that are manipulated to facilitate the calculation of waste and material inventory footprints in the supply chain of any activity in the same way as one would apply standard LCIA methods.
-We refer to the T-reX methods as `pseudo-LCIA' methods, as they represent an accounting of an activity's technosphere inventory without reference to impact modelling, as in standard LCIA.
-
-If desired, prospective databases can be defined by the user or constructed with the projections of the integrated assessment models such as IMAGE~\citep{stehfest2014image} and REMIND~\citep{remind2020model}, which offer a range of options aligned with the Shared Socioeconomic Pathways (SSPs)~\citep{ssp2020ghg} and can be paired with a variety of mitigation scenarios.
-
-The deconstruction of the databases by T-reX into lists of exchanges allows the relevant flows of material and waste to be identified and categorised by the search functions.
-The queries facilitating this are tailored to the specific database and the user can easily modify them to suit their needs.
-The categories defined in the configuration are used to create T-reX's `pseudo LCIA' methods that are indicators of aggregated technosphere demand.
-The exchange editing function of T-reX then takes each list of exchanges and appends to the relevant activity a copy of the technosphere exchange as a mirrored `pseudo-biosphere' exchange that matches the `pseudo-LCIA' method.
-
-In the default configuration, there are 10 waste categories which are further divided by their unit of measurement (kilograms and cubic meters) to create a total of 20 waste methods.
-The waste categories include incineration, recycling, and total waste, and are listed in the supplementary material in~\autoref{sec:supplementary}.
-One advantage of T-reX is that it can identify `hidden' waste exchanges that would otherwise be `consumed' by a treatment process and not leave the technosphere.
-Since `waste is not a service'~\citep{guinee2021wasteisnotaservice}, a characterisation factor of -1 is applied to the waste footprint methods (except CCS exchanges), changing the perspective from `waste consumed by treatment' to `waste generated by the activity'.
-
-In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption.
-Especially valuable in the case of CRMs is T-reX's ability to uncover and quantify these threads of background consumption, which can be inconspicuously embedded in products or consumed in seemingly distant supply chain activities.
-The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas.
-The full list of 59 materials included in the default configuration is provided in the supplementary material in~\autoref{sec:supplementary}.
-
-The logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name.
-A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows.
-For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs.
-Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.
-
-In these methods, the material demand is calculated based on the total mass that is extracted from the environment, thus, their focus is essentially solely on the mining-related exchanges that bring these materials from the biosphere into the technosphere.
-
-T-reX, in contrast, uses accounting for material demand and waste generation based on exchanges solely within the technosphere.
-This offers a different perspective, allowing for the estimation of overall supply chain material demands that consider the entire life cycle of an activity, including non-direct impacts on the market such as co-production of other materials.
-Consider a demand for an activity containing a metal, for example; while the existing material use methods allow one to calculate the total mass of that metal that is extracted from the environment, T-reX can provide insight into the broader supply chain impacts of the demand for this metal.
-If the production of other materials is attributed to the production of this metal, these would appear as negative material demands in the T-reX results---supply chain pressure for one material can result in lessening of supply chain pressure for another.
-In the results of the Li-ion battery case study in \autoref{sec:results-casestudy}, we will see that this is indeed the case for nickel demand, which, in the final inventory because of such co-production, is counter-intuitively negative (due to co-production and substitution in the LCA system models) despite the presence of nickel in the final products.
-
-% \subsubsection{Mathematical basis}
-
-% \cbox{very nice formulae coming soon from LL hopefully}
-
-\subsubsection{The workflow of T-reX}
-
-The workflow of T-reX is divided into several modules, each performing a separate function.
-The modules are designed to be used in a particular order, but the user can also use them individually to perform specific tasks.
-The standard workflow is as follows:
-
-\begin{enumerate}
-	\item Configuration of waste and material exchange categorisation (optional)
-	\item Generation of prospective LCA databases (optional)
-	\item Database expansion---to create a list of all exchanges in the database
-	\item Identification and categorisation of exchanges
-	\item Creation of `pseudo-biosphere' databases
-	\item Creation of `pseudo-LCIA' methods to calculate waste and material inventory footprints
-	\item Exchange editing---whereby the technosphere exchange is mirrored as a `pseudo-biosphere' exchange
-	\item Database verification---to ensure that T-reX has manipulated the database correctly
-\end{enumerate}
-
-This workflow creates a copy of the original \texttt{brightway} project containing the original biosphere database, a T-reX `pseudo-biosphere' database along with one or more manipulated technosphere databases that can be used to calculate the waste and material inventory footprints of activities in the same way as standard LCIA calculations.
-
-An overview of the T-reX workflow is presented in \autoref{fig:methods-flowchart}. The supplementary material in \autoref{sec:supplementary} contains a more detailed computational flowchart.
-
-\begin{figure*}[!htbp]
-	\centering
-	\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format.
-Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
-	\includegraphics[width=1.8\columnwidth]{figs/T-reX_method.pdf}\label{fig:methods-flowchart}
-\end{figure*}
-
-\subsection{Case study methodology}\label{sec:method-casestudy}
-
-We investigated five types of Li-ion batteries in the \texttt{ecoinvent} 3.9.1 `cutoff' LCA database~\citep{ecoinvent2016version3}, each represented by their unamended global market inventories:
-\begin{itemize}[itemsep=0pt]
-	\item Li-ion, LFP, rechargeable, prismatic
-	\item Li-ion, LiMn\(_2\)O\(_4\), rechargeable, prismatic
-	\item Li-ion, NMC111, rechargeable, prismatic
-	\item Li-ion, NMC811, rechargeable, prismatic
-	\item Li-ion, NCA, rechargeable, prismatic
-\end{itemize}
-
-In addition to the waste and material inventory footprint methods created by T-reX, the following standard LCIA methods were applied for comparison:
-
-\begin{itemize}[itemsep=0pt]
-	\item ReCiPe 2016 v1.03, midpoint (I)~\citep{huijbregts2016recipe}
-	\item EF v3.0 no LT~\citep{eu2023ef}
-	\item EDIP 2003 no LT~\citep{hauschild2003edip}
-	\item CSI 2020~\citep{arvidsson2020csi}
-\end{itemize}
-
-Additional to the present-day databases, T-reX created prospective database sets based on \texttt{ecoinvent} 3.9.1 using the functionality of \texttt{premise} with the REMIND model and the baseline scenario SSP2 `middle of the road' with the following Representative Concentration Pathways (RCPs):
-\begin{itemize}
-	\item SSP2-base: representing an approximate 3.5°C increase in global temperatures to 2100
-	\item SSP2-PkBudg500: representing the achievement of Paris climate goals, ca. 1.3°C increase to 2100
-\end{itemize}
-
-For each pathway, databases (at five-year time intervals) were created with \texttt{premise}~\citep{sacchi2022premise} and processed with T-reX for the period between 2020 and 2100.
-
-\subsubsection{LCA calculations}
-For each combination of activity, method, and database, a single score was calculated along with details of the top contributing processes.
-Additionally, for the T-reX methods, a contribution analysis was performed.
-This involved utilising the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem.
-This function performs graph traversal on the impact matrix of the LCA object to a specified cutoff and groups the resulting leaves by their Cooperative Patent Classification (CPC) codes.
-This provides insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.
-
-\section{Results}\label{sec:results}
 \subsection{T-reX tool}\label{sec:results-T-reX}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
     \centering
-	\includegraphics[width=8cm]{figs/T-reX_NMC811_WasteTotalSolid.pdf}
+    \includegraphics[width=8cm]{figures/T-reX_NMC811_WasteTotalSolid.pdf}
     \caption{A simplified Sankey diagram showing the total solid waste footprint flows for the NMC 811 battery supply chain at the global market in the \texttt{ecoinvent} database version 3.9.1.}\label{fig:sankey_waste}
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[H]
     \centering
-	\includegraphics[width=16cm]{figs/T-reX-wastefootprint-totalANDcarbon.pdf}
+    \includegraphics[width=16cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
     \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only).
 The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
-\end{figure*}
+\end{figure}
 %!!!!
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
     \centering
-	\includegraphics[width=9cm]{figs/T-reX-wastefootprint-processcontributions.pdf}
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
     \caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
     \centering
-	\includegraphics[width=9cm]{figs/T-reX-wastefootprint-sectorcontributions.pdf}
+    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
     \caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[H]
     \centering
-	\includegraphics[width=16cm]{figs/T-reX-coalANDedip.pdf}
+    \includegraphics[width=16cm]{figures/T-reX-coalANDedip.pdf}
     \caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
     \centering
@@
 \end{table}
-
-\section{Discussion}\label{sec:discussion}
-Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting.
-Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.
-
-T-reX advances the state-of-the-art in LCA by providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories.
-Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.
-
-The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility, and limitations of T-reX.
-
-By adjusting the user configuration for future scenarios and waste/material categorisation, we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database.
-Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate.
-Additionally, visual exploration in \texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are structurally identical to standard LCIA methods when integrated within the \texttt{brightway} project.
-
-One limitation of T-reX is that it does not yet provide specific information in a readily accessible format on the composition of the waste generated.
-This information is required to thoroughly assess the potential environmental impacts of this waste.
-Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'.
-A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.
-
-The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}.
-As demonstrated in the results of the case study---where there were often very few scenario-temporal changes in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adoption of future background inventories.
-In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}.
-A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.
-
-\section{Conclusions}\label{sec:conclusions}
-T-reX is an extension to the open-source \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of any product or service in an LCA database.
-It `explodes' and reconstructs the database to identify and edit waste and material exchanges and writes matching custom `pseudo' LCIA methods.
-These exchanges are mirrored in the \texttt{brightway} project as `pseudo-biosphere' flows and thus, the waste and material footprints of any other supply chain flow of interest, such as water, gas, or critical raw materials, can be calculated similarly to existing LCIA methods.
-
-T-reX elucidates the complex relationships between material demand, waste generation, and environmental impacts.
-This instrument enables detailed exploration of `footprints' and life cycle inventories---modelled externalities of human activities---and examines their connections to supply chain risks and potential environmental damage.
-T-reX is a step towards a better understanding of these interconnections for the practical realisation of `sustainability' and the `circular economy'.
-
-% Back matter
-\appendix
-The supplementary material supplied with this manuscript contains the
-following sections:
-\begin{enumerate}
-	\item Metadata of the T-reX Python package\label{sec:supplementary-1}
-	\item Details of the modules and user configuration\label{sec:supplementary-2}
-	\item Description of the computational workflow\label{sec:supplementary-3}
-	\item Modular flowchart of T-reX\label{sec:supplementary-4}
-	\item Example of the terminal output of T-reX\label{sec:supplementary-5}
-	\item Example of the terminal output of the case study\label{sec:supplementary-6}
-	\item Complete tabulated results of the case study\label{sec:supplementary-7}
-	\item Complete visualisations from the case study\label{sec:supplementary-8}
-\end{enumerate}
-
-\section*{Data availability}
-All data used is publicly available online under the noted sources.
-T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence.
-It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
-The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} 
-and the development server is hosted by GitHub at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.
-
-\section*{Acknowledgements}
-Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
-101058522 (project FutuRaM --- futuram.eu).
-The authors would like to thank the reviewers for their valuable comments and suggestions.
-
-The authors would like to acknowledge those who were part of the futures
-thinking small group session at the FutuRaM consortium meeting in Berlin in
-June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision
-presented by Stewart Charles McDowall of a comprehensive open-source software
-framework for the study and development of secondary raw material systems. The
-authors hope to continue the development of T-reX in the service of our shared
-ideals.
-
-\section*{Declaration of competing interest}
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-% To print the credit authorship contribution details
-\printcredits
-
-%\section*{CRediT authorship visualisation} % IMPROVED VISUAL VERSION
-
-%\begin{itemize}
-%	\item \textbf{SCM}: Stewart Charles McDowall
-%	\item \textbf{EL}: Elizabeth Lanphear
-%	\item \textbf{SC}: Stefano Cucurachi
-%	\item \textbf{CFB}: Carlos Felipe Blanco
-%\end{itemize}
-% \subsection*{Alternative CRediT statement}
-\section*{CRediT authorship contribution statement}
-
-\begin{itemize}[noitemsep]
-	\item \textbf{SCM}: Stewart Charles McDowall
-	\item \textbf{EL}: Elizabeth Lanphear
-	\item \textbf{SC}: Stefano Cucurachi
-	\item \textbf{CFB}: Carlos Felipe Blanco
-\end{itemize}
-
-\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{figs/T-reX_credit_heatmap.pdf}
-
-\section*{Declaration of competing interest}
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-\section*{Acknowledgements}
-Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No.
-101058522 (project FutuRaM --- futuram.eu).
-The authors would like to thank the reviewers for their valuable comments and suggestions.
-
-The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive open-source software framework for the study and development of secondary raw material systems.
-The authors hope to continue the development of T-reX in the service of our shared ideals.
-
-\section{Supplementary material}\label{sec:supplementary}
-The supplementary material supplied with this manuscript contains the following:
-\begin{enumerate}
-	\item Metadata of the T-reX Python package
-	\item Details of the modules and user configuration
-	\item Description of the computational workflow
-	\item Modular flowchart of T-reX
-	\item Example of the terminal output of T-reX
-	\item Example of the terminal output of the case study
-	\item Complete tabulated results of the case study
-	\item Complete visualisations from the case study
-\end{enumerate}
-
-%% Loading bibliography style file
-%\bibliographystyle{model1-num-names}
-\bibliographystyle{cas-model2-names}
-
-% Loading bibliography database
-\bibliography{cas-refs}
-% \appendix
-% \section{Example output}
-% \label{app:example_output}
-
-\end{document}

### 2025-02-16 03_T-reX-latex_files-atfirstsubmission/latex_files-atfirstsubmission/cover_letter.tex
auto-update

--- 2024-04-01 latex_files/cover_letter.tex
@@
 
-We are pleased to submit our original research article titled ``T-reX\: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}.
+    We are pleased to submit our original research article titled ``T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases'' for consideration by \textit{Resources, Conservation and Recycling}.
 This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.
 
-Our submission provides a comprehensive description of the tool, which integrates seamlessly with the Brightway2 framework to offer a user-friendly and adaptable solution for LCA practitioners.
+    Our submission provides a comprehensive description of the tool, which integrates seamlessly with the \texttt{brightway} framework to offer a user-friendly and adaptable solution for LCA practitioners.
 By automating data analysis and aggregation processes, our tool expedites the identification of environmental hotspots within supply chains, thus facilitating more informed decision-making.
